### 2019-04-22 fertile.Rmd
initial commit

---
title: Title here

# to produce blinded version set to 1
blinded: 0

authors: 
- name: Author 1
  thanks: The authors gratefully acknowledge ...
  affiliation: Department of YYY, University of XXX

- name: Author 2
  affiliation: Department of ZZZ, University of WWW

keywords:
- 3 to 6 keywords
- that do not appear in the title

abstract: |
  The text of your abstract.  200 or fewer words.

bibliography: bibliography.bib
output: rticles::asa_article
---

# Introduction

This template demonstrates some of the basic latex you'll need to know to create a ASA article.

\section{Verifications}
\label{sec:verify}
This section will be just long enough to illustrate what a full page of
text looks like, for margins and spacing.

@marwick2018packaging

### 2019-04-28 fertile.Rmd
preliminary sources

@@
 
-# Introduction
+# Sources Summaries
 
-This template demonstrates some of the basic latex you'll need to know to create a ASA article.
+Main points of different sources, as well as info about how they might be used for the paper.
 
-\section{Verifications}
+\section{Background Information}
 \label{sec:verify}
-This section will be just long enough to illustrate what a full page of
-text looks like, for margins and spacing.
+
+Here are some sources we might use for motivation behind the project:
+
+### The Reproducibility Crisis
+
+Big idea: most scientific fields are facing a reproducibility crisis and poor statistical use is considered one of the important reasons behind this.
+
+Not sure how useful this is due to the fact that it does not necessarily focus on the same kind of reproducibility we are looking at, which is code reproducibility rather than experimental reproducibility.
+
+@baker20161
+
+### Why is reproducibility important?
+
+From Popper in `The Logic of Scientific Discovery`: "non-reproducible single occurrences are of no significance to science."
+
+@popper2005logic
+
+### What does reproducibility mean in data science?
+
+"The ability of a researcher to duplicate the results of a prior study using the same materials as were used by the original investigator.
+That is, a second researcher might use the same raw data to build the same analysis files and implement the same statistical analysis in an attempt to yield the same results."
+
+@goodman2016does
+
+From "The Practice of Reproducibile Research": Can all of the figures/calculations related to the result be reproduced in a single button press or at least with a reasonably small effort?
+
+@kitzes2017practice
+
+### What makes a data science project reproducibile?
+
+Another way to think about this is: what features does fertile have that match what different sources think needs to be checked to ensure reproducibility?
+
+Some big ideas from the OpenSci discussion page:
+
+* code should use tidy style
+* project should have short vignette files -- most written code should be in an R/ directory
+* code and data should be stored in separate folders
+* a run environment and dependencies should be specified
+* there should be a README file
+* there should be a data dictionary
+* one master script
+* code should only use relative paths
+* code should be well commented and all variables should be defined
+
+Big ideas from "Packaging data analytical work reproducibly using R"
+
+* research should be organized like an R package!
+* clear separation of data, method, and output
+* specify the computational environment that was used for the original analysis (typically in a plain text file)
+* there should be a README describing the project and where to get started
+* script files with reusable functions should go in an R/ directory
+* raw data files should be kept in a data/ directory
+* analysis scripts and report files should go in an analysis/ directory
+* for simpe projects, scripts should be given ascending names. For more complicated ones, the use of some sort of makefile is recommended
+* there should also be a DESCRIPTION file with information about the authors, project license, and software dependencies
 
 @marwick2018packaging
+
+From "The Practice of Reproducible Research"
+
+* Are the data openly accessible? If hosted online, is the web address reliable long-term?
+* Are they in a commonly used and well-documented file format? Avoid spreadsheets and instead use plain text data if possible!
+* Is the raw data available? Is sufficient metadata provided?
+
+* Are dependencies described properly?
+* Is full history of source code available through a public version history
+
+* Is there a README?
+* Are functions documented?
+* Is there narrative documentation explaining how the different pieces work together?
+* Are there usage examples?
+
+Folder setup might look like:
+
+1. Raw Data
+    + Data
+    + README
+2. Clean Data
+    + Data
+3. Results
+    + Results file
+4. Src
+    + Analysis script
+    + Script to clean data
+
+@kitzes2017practice
+
+From R OpenSci's Reproducibility Guide:
+
+http://ropensci.github.io/reproducibility-guide/
+
+* Is it clear where to begin?
+* Can you determinine which files were used as input to create output files?
+* Is there documentation about every result?
+* Are exact versions of external applications noted?
+* If using randomness, are seeds noted?
+* Have you specified a license or noted licenses if you used other people's content?
+* Are files easy to find?
+* Is it clear what the most recent file is?
+* Are there any folders that could be deleted?
+* Is analysis output done hierarchically?
+* Are there lots of manual data manipulation steps?
+
+From "A Guide to Reproducible Code in Ecology and Evolution" (these ideas are pretty universal, though):
+
+A basic project structure:
+
+* The data folder contains all input data (and metadata) used in the analysis.
+* The doc folder contains the manuscript
+* The figs directory contains figures generated by the analysis
+* The output folder contains any type of intermediate or output files (e.g. simulation outputs, models, processed datasets, etc.). You might separate this and also have a cleaned-data folder.
+* The R directory contains R scripts with function definitions.
+* The reports folder contains RMarkdown files that document the analysis or report on results
+
+* Consistent, ordered naming of scripts
+* Use portable paths
+* Write unit tests (only for advanced coding)
+* Show the packages you used
+* Record dependencies and versions of outside things you use
+
+@cooper2017guide
+
+# Why is this package important?
+
+From "Five Concrete Reasons Your Students Should Be Learning to Analyze Data in the Reproducible Paradigm"
+
+* Helps with transparency: make clear data cleaning steps between raw data and final data
+* Helps with collaboration: easier to share code when it takes very few steps to run on a different computer
+
+@bray2014five
+
+The other big ideas on the importance/use of this package (not from sources):
+
+* Make it easier for professors to grade students' code
+* Should be usable by intro level data science students
+* Possible use by reviewers of journal articles (and by those writing the articles)
+
+# Similar R Packages
+
+### rrtools
+
+https://github.com/benmarwick/rrtools
+
+* Creates a basic R package named after your research topic
+* Generates a license file
+* Connects to GitHub and creates a repository
+* Generates a README
+* Generates a reproducible directory structure
+* Creates a dockerfile
+* Creates a minimal travis file
+* Sets up testthat

### 2019-05-21 fertile.Rmd
broman, stodden, why R?

@@
 
+# Karl Broman's Suggestions
+
+* Encapsulate everything within one directory
+* Separate raw data from derived data
+* Separate data from code
+* Use relative paths
+* Choose filenames correctly
+* Write README files
+
+http://kbroman.org/steps2rr/pages/organize.html
+
 # Why is this package important?
@@
 
+# Why focus on R?
+
+* R is the most popular language for statistical programming and is specifically designed for statistics
+* R is great for reprodubility because the code is readable by users, and RMarkdown is a great way to show/explain processes
+* R is easy to install and begin using
+
+--> from https://openresearchsoftware.metajnl.com/articles/10.5334/jors.bu/print/
+
 # Similar R Packages
@@
 * Sets up testthat
+
+# Msc sources to look at:
+
+### Victoria Stodden's "Implementing Reproducible Research" book
+https://books.google.com/books?hl=en&lr=&id=JcmSAwAAQBAJ&oi=fnd&pg=PP1&dq=Victoria+stodden+reproducibility&ots=ym_gvSxMNH&sig=B5N6FlkoocVjIprIZevtvrL33qQ#v=onepage&q=coombes&f=false

### 2019-11-08 fertile.Rmd
abstract - draft #1

@@
 abstract: |
-  The text of your abstract.  200 or fewer words.
+  The production of reproducible research is key for advancing our scientific knowlege, encouraging transparency, collaboration, and understanding.
+However, there is no clear consensus on the standards that constitute research reproducibility.
+In an attempt to build a reproducibility framework in the R community, several packages have been developed to address different aspects of reproducibility, but none are comprehensive.
+In this paper, we present `fertile`, an R package with an extensive selection of functions designed to thoroughly test reproducibility.
+Fertile makes creating a reproducibile project as easy as possible for users of all levels of sophistication, providing a simple and comprehensive framework for testing reproducibility in the R community.
+
+  Note: should we add something about the idea of testing the package in a     classroom?
 
@@
 
-# Sources Summaries
+# Introduction
+
+# The Importance of Ensuring Reproducibility
+
+Practicing reproducible research is key for advancing our scientific knowledge.
+One reason for this is that reproducible research encourages transparency.
+Providing one's audience with well organized, reproducible code and results allows readers to understand the steps taken to generate findings from raw data and determine for themselves whether or not they believe the results to be reliable or trustworthy.
+
+Additionally, reproducibility encourages collaboration and extended research, allows others to easily apply the methods used in one project to their own work with minimal effort. @bray2014five
+
+Fertile is designed to make reproducibility simple, providing fast and easy methods to test an R project for reproducibility.
+The package is intended to be usable by introductory data science students in their first semester of R.
+
+--> make sure this info is included somewhere:
+
+From "Five Concrete Reasons Your Students Should Be Learning to Analyze Data in the Reproducible Paradigm" (@bray2014five)
+
+* Helps with transparency: make clear data cleaning steps between raw data and final data
+* Helps with collaboration: easier to share code when it takes very few steps to run on a different computer
+
+The other big ideas on the importance/use of this package (not from sources):
+
+* Make it easier for professors to grade students' code
+* Should be usable by intro level data science students
+* Possible use by reviewers of journal articles (and by those writing the articles)
+
+# What Defines Reproducibility in Data Science?
+
+# Creating Comprehensive Reproducibility Reports With "fertile"
 
 Main points of different sources, as well as info about how they might be used for the paper.
-
-\section{Background Information}
-\label{sec:verify}
 
@@
 
-# Why is this package important?
-
-From "Five Concrete Reasons Your Students Should Be Learning to Analyze Data in the Reproducible Paradigm"
-
-* Helps with transparency: make clear data cleaning steps between raw data and final data
-* Helps with collaboration: easier to share code when it takes very few steps to run on a different computer
-
-@bray2014five
-
-The other big ideas on the importance/use of this package (not from sources):
-
-* Make it easier for professors to grade students' code
-* Should be usable by intro level data science students
-* Possible use by reviewers of journal articles (and by those writing the articles)
-
 # Why focus on R?

### 2019-11-08 fertile.tex
abstract - draft #1

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

%% load any required packages here

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf Title here}

  \author{
        Author 1 \thanks{The authors gratefully acknowledge \ldots{}} \\
    Department of YYY, University of XXX\\
     and \\     Author 2 \\
    Department of ZZZ, University of WWW\\
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Title here}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
The text of your abstract. 200 or fewer words.
\end{abstract}

\noindent%
{\it Keywords:} 3 to 6 keywords, that do not appear in the title
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\section{Introduction}\label{introduction}

\section{The Importance of Ensuring
Reproducibility}\label{the-importance-of-ensuring-reproducibility}

Practicing reproducible research is key for advancing our scientific
knowledge. One reason for this is that reproducible research encourages
transparency. Providing one's audience with well organized, reproducible
code and results allows readers to understand the steps taken to
generate findings from raw data and determine for themselves whether or
not they believe the results to be reliable or trustworthy.

Additionally, reproducibility encourages collaboration and extended
research, allows others to easily apply the methods used in one project
to their own work with minimal effort. \citet{bray2014five}

Fertile is designed to make reproducibility simple, providing fast and
easy methods to test an R project for reproducibility. The package is
intended to be usable by introductory data science students in their
first semester of R.

--\textgreater{} make sure this info is included somewhere:

From ``Five Concrete Reasons Your Students Should Be Learning to Analyze
Data in the Reproducible Paradigm'' (\citet{bray2014five})

\begin{itemize}
\tightlist
\item
  Helps with transparency: make clear data cleaning steps between raw
  data and final data
\item
  Helps with collaboration: easier to share code when it takes very few
  steps to run on a different computer
\end{itemize}

The other big ideas on the importance/use of this package (not from
sources):

\begin{itemize}
\tightlist
\item
  Make it easier for professors to grade students' code
\item
  Should be usable by intro level data science students
\item
  Possible use by reviewers of journal articles (and by those writing
  the articles)
\end{itemize}

\section{What Defines Reproducibility in Data
Science?}\label{what-defines-reproducibility-in-data-science}

\section{\texorpdfstring{Creating Comprehensive Reproducibility Reports
With
``fertile''}{Creating Comprehensive Reproducibility Reports With fertile}}\label{creating-comprehensive-reproducibility-reports-with-fertile}

Main points of different sources, as well as info about how they might
be used for the paper.

Here are some sources we might use for motivation behind the project:

\subsubsection{The Reproducibility
Crisis}\label{the-reproducibility-crisis}

Big idea: most scientific fields are facing a reproducibility crisis and
poor statistical use is considered one of the important reasons behind
this.

Not sure how useful this is due to the fact that it does not necessarily
focus on the same kind of reproducibility we are looking at, which is
code reproducibility rather than experimental reproducibility.

\citet{baker20161}

\subsubsection{Why is reproducibility
important?}\label{why-is-reproducibility-important}

From Popper in \texttt{The\ Logic\ of\ Scientific\ Discovery}:
``non-reproducible single occurrences are of no significance to
science.''

\citet{popper2005logic}

\subsubsection{What does reproducibility mean in data
science?}\label{what-does-reproducibility-mean-in-data-science}

``The ability of a researcher to duplicate the results of a prior study
using the same materials as were used by the original investigator. That
is, a second researcher might use the same raw data to build the same
analysis files and implement the same statistical analysis in an attempt
to yield the same results.''

\citet{goodman2016does}

From ``The Practice of Reproducibile Research'': Can all of the
figures/calculations related to the result be reproduced in a single
button press or at least with a reasonably small effort?

\citet{kitzes2017practice}

\subsubsection{What makes a data science project
reproducibile?}\label{what-makes-a-data-science-project-reproducibile}

Another way to think about this is: what features does fertile have that
match what different sources think needs to be checked to ensure
reproducibility?

Some big ideas from the OpenSci discussion page:

\begin{itemize}
\tightlist
\item
  code should use tidy style
\item
  project should have short vignette files -- most written code should
  be in an R/ directory
\item
  code and data should be stored in separate folders
\item
  a run environment and dependencies should be specified
\item
  there should be a README file
\item
  there should be a data dictionary
\item
  one master script
\item
  code should only use relative paths
\item
  code should be well commented and all variables should be defined
\end{itemize}

Big ideas from ``Packaging data analytical work reproducibly using R''

\begin{itemize}
\tightlist
\item
  research should be organized like an R package!
\item
  clear separation of data, method, and output
\item
  specify the computational environment that was used for the original
  analysis (typically in a plain text file)
\item
  there should be a README describing the project and where to get
  started
\item
  script files with reusable functions should go in an R/ directory
\item
  raw data files should be kept in a data/ directory
\item
  analysis scripts and report files should go in an analysis/ directory
\item
  for simpe projects, scripts should be given ascending names. For more
  complicated ones, the use of some sort of makefile is recommended
\item
  there should also be a DESCRIPTION file with information about the
  authors, project license, and software dependencies
\end{itemize}

\citet{marwick2018packaging}

From ``The Practice of Reproducible Research''

\begin{itemize}
\item
  Are the data openly accessible? If hosted online, is the web address
  reliable long-term?
\item
  Are they in a commonly used and well-documented file format? Avoid
  spreadsheets and instead use plain text data if possible!
\item
  Is the raw data available? Is sufficient metadata provided?
\item
  Are dependencies described properly?
\item
  Is full history of source code available through a public version
  history
\item
  Is there a README?
\item
  Are functions documented?
\item
  Is there narrative documentation explaining how the different pieces
  work together?
\item
  Are there usage examples?
\end{itemize}

Folder setup might look like:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Raw Data

  \begin{itemize}
  \tightlist
  \item
    Data
  \item
    README
  \end{itemize}
\item
  Clean Data

  \begin{itemize}
  \tightlist
  \item
    Data
  \end{itemize}
\item
  Results

  \begin{itemize}
  \tightlist
  \item
    Results file
  \end{itemize}
\item
  Src

  \begin{itemize}
  \tightlist
  \item
    Analysis script
  \item
    Script to clean data
  \end{itemize}
\end{enumerate}

\citet{kitzes2017practice}

From R OpenSci's Reproducibility Guide:

\url{http://ropensci.github.io/reproducibility-guide/}

\begin{itemize}
\tightlist
\item
  Is it clear where to begin?
\item
  Can you determinine which files were used as input to create output
  files?
\item
  Is there documentation about every result?
\item
  Are exact versions of external applications noted?
\item
  If using randomness, are seeds noted?
\item
  Have you specified a license or noted licenses if you used other
  people's content?
\item
  Are files easy to find?
\item
  Is it clear what the most recent file is?
\item
  Are there any folders that could be deleted?
\item
  Is analysis output done hierarchically?
\item
  Are there lots of manual data manipulation steps?
\end{itemize}

From ``A Guide to Reproducible Code in Ecology and Evolution'' (these
ideas are pretty universal, though):

A basic project structure:

\begin{itemize}
\item
  The data folder contains all input data (and metadata) used in the
  analysis.
\item
  The doc folder contains the manuscript
\item
  The figs directory contains figures generated by the analysis
\item
  The output folder contains any type of intermediate or output files
  (e.g.~simulation outputs, models, processed datasets, etc.). You might
  separate this and also have a cleaned-data folder.
\item
  The R directory contains R scripts with function definitions.
\item
  The reports folder contains RMarkdown files that document the analysis
  or report on results
\item
  Consistent, ordered naming of scripts
\item
  Use portable paths
\item
  Write unit tests (only for advanced coding)
\item
  Show the packages you used
\item
  Record dependencies and versions of outside things you use
\end{itemize}

\citet{cooper2017guide}

\section{Karl Broman's Suggestions}\label{karl-bromans-suggestions}

\begin{itemize}
\tightlist
\item
  Encapsulate everything within one directory
\item
  Separate raw data from derived data
\item
  Separate data from code
\item
  Use relative paths
\item
  Choose filenames correctly
\item
  Write README files
\end{itemize}

\url{http://kbroman.org/steps2rr/pages/organize.html}

\section{Why focus on R?}\label{why-focus-on-r}

\begin{itemize}
\tightlist
\item
  R is the most popular language for statistical programming and is
  specifically designed for statistics
\item
  R is great for reprodubility because the code is readable by users,
  and RMarkdown is a great way to show/explain processes
\item
  R is easy to install and begin using
\end{itemize}

--\textgreater{} from
\url{https://openresearchsoftware.metajnl.com/articles/10.5334/jors.bu/print/}

\section{Similar R Packages}\label{similar-r-packages}

\subsubsection{rrtools}\label{rrtools}

\url{https://github.com/benmarwick/rrtools}

\begin{itemize}
\tightlist
\item
  Creates a basic R package named after your research topic
\item
  Generates a license file
\item
  Connects to GitHub and creates a repository
\item
  Generates a README
\item
  Generates a reproducible directory structure
\item
  Creates a dockerfile
\item
  Creates a minimal travis file
\item
  Sets up testthat
\end{itemize}

\section{Msc sources to look at:}\label{msc-sources-to-look-at}

\subsubsection{\texorpdfstring{Victoria Stodden's ``Implementing
Reproducible Research''
book}{Victoria Stodden's Implementing Reproducible Research book}}\label{victoria-stoddens-implementing-reproducible-research-book}

\url{https://books.google.com/books?hl=en\&lr}=\&id=JcmSAwAAQBAJ\&oi=fnd\&pg=PP1\&dq=Victoria+stodden+reproducibility\&ots=ym\_gvSxMNH\&sig=B5N6FlkoocVjIprIZevtvrL33qQ\#v=onepage\&q=coombes\&f=false

\bibliographystyle{agsm}
\bibliography{bibliography.bib}

\end{document}

### 2019-11-20 fertile.Rmd
revised abstract; #1

@@
 ---
-title: Title here
+title: "Creating optimal conditions for reproducible data analysis in R with `fertile`"
 
@@
 authors: 
-- name: Author 1
-  thanks: The authors gratefully acknowledge ...
-  affiliation: Department of YYY, University of XXX
+- name: Audrey Bertin
+  thanks: The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the `tidyverse` team. 
+  affiliation: Program in Statistical and Data Sciences, Smith College
 
-- name: Author 2
-  affiliation: Department of ZZZ, University of WWW
+- name: Benjamin S. Baumer
+  affiliation: Program in Statistical and Data Sciences, Smith College
 
 keywords:
-- 3 to 6 keywords
-- that do not appear in the title
+- reproducibility
+- statistical software
+- workflow
+- collaboration
 
 abstract: |
-  The production of reproducible research is key for advancing our scientific knowlege, encouraging transparency, collaboration, and understanding.
-However, there is no clear consensus on the standards that constitute research reproducibility.
-In an attempt to build a reproducibility framework in the R community, several packages have been developed to address different aspects of reproducibility, but none are comprehensive.
-In this paper, we present `fertile`, an R package with an extensive selection of functions designed to thoroughly test reproducibility.
-Fertile makes creating a reproducibile project as easy as possible for users of all levels of sophistication, providing a simple and comprehensive framework for testing reproducibility in the R community.
-
-  Note: should we add something about the idea of testing the package in a     classroom?
+  The advancement of scientific knowledge increasingly depends on ensuring that data-driven research is reproducible: that two people with the same data obtain the same results.
+However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation, and no clear consensus on standards of what constitutes reproducibility in published research.
+We focus on a series of common mistakes programmers make while conducting data science projects in R, primarily through the RStudio integrated development environment.
+`fertile` is an R package that operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that is already written for potential problems).
+Furthermore, `fertile` is designed to educate the user about why the mistakes are problematic, and how to fix them.
+We discuss experimental results from testing `fertile` in an introductory data science course.
 

### 2019-11-20 fertile.tex
merge

@@
 \begin{abstract}
-The text of your abstract. 200 or fewer words.
+The production of reproducible research is key for advancing our
+scientific knowlege, encouraging transparency, collaboration, and
+understanding. However, there is no clear consensus on the standards
+that constitute research reproducibility. In an attempt to build a
+reproducibility framework in the R community, several packages have been
+developed to address different aspects of reproducibility, but none are
+comprehensive. In this paper, we present \texttt{fertile}, an R package
+with an extensive selection of functions designed to thoroughly test
+reproducibility. Fertile makes creating a reproducibile project as easy
+as possible for users of all levels of sophistication, providing a
+simple and comprehensive framework for testing reproducibility in the R
+community.
+
+Note: should we add something about the idea of testing the package in a
+classroom?
 \end{abstract}

### 2019-11-20 fertile.Rmd
merge

@@
 abstract: |
+
   The advancement of scientific knowledge increasingly depends on ensuring that data-driven research is reproducible: that two people with the same data obtain the same results.

### 2019-11-24 fertile.Rmd
start of extended abstract draft

@@
 
+# Extended Abstract
+
+In data science, an analyst has created *reproducible* work if their code and data produces identical results when given to another analyst.
+
+This idea of reproducibility is becoming increasingly important to the advancement of our scientific knowledge.
+One reason for this is that reproducible research encourages transparency.
+Providing one's audience with well organized, reproducible code and results allows readers to understand the steps taken to generate findings from raw data and determine for themselves whether or not they believe the results to be reliable or trustworthy.
+
+Additionally, reproducibility encourages collaboration and extended research, allowing others to easily apply the methods used in one project to their own work with minimal effort.
+
+However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation, and no clear consensus on standards of what constitutes reproducibility in published research.
+
+• Note: What are these behavioral and technical challenges, and which ones should we mention?
+
+In defining the steps toward creating reproducible analyses, researchers place their emphasis on different areas, including file structure, documentation, the use of file paths in code, and accessibility.
+Within these categories, the recommendations for what steps to take in ensuring reproducibility also vary.
+
+• Why focus on R?
+
 # Introduction
@@
 
-Practicing reproducible research is key for advancing our scientific knowledge.
 One reason for this is that reproducible research encourages transparency.

### 2019-11-24 fertile.tex
start of extended abstract draft

--- 2019-11-08 fertile.tex
@@
 {
-  \title{\bf Title here}
+  \title{\bf Creating optimal conditions for reproducible data analysis in R with
+\texttt{fertile}}
 
   \author{
-        Author 1 \thanks{The authors gratefully acknowledge \ldots{}} \\
-    Department of YYY, University of XXX\\
-     and \\     Author 2 \\
-    Department of ZZZ, University of WWW\\
+        Audrey Bertin \thanks{The authors gratefully acknowledge contributions from Hadley Wickham,
+Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the
+\texttt{tidyverse} team.} \\
+    Program in Statistical and Data Sciences, Smith College\\
+     and \\     Benjamin S. Baumer \\
+    Program in Statistical and Data Sciences, Smith College\\
       }
@@
   \begin{center}
-    {\LARGE\bf Title here}
+    {\LARGE\bf Creating optimal conditions for reproducible data analysis in R with
+\texttt{fertile}}
   \end{center}
@@
 \begin{abstract}
-The text of your abstract. 200 or fewer words.
+The advancement of scientific knowledge increasingly depends on ensuring
+that data-driven research is reproducible: that two people with the same
+data obtain the same results. However, while the necessity of
+reproducibility is clear, there are significant behavioral and technical
+challenges that impede its widespread implementation, and no clear
+consensus on standards of what constitutes reproducibility in published
+research. We focus on a series of common mistakes programmers make while
+conducting data science projects in R, primarily through the RStudio
+integrated development environment. \texttt{fertile} is an R package
+that operates in two modes: proactively (to prevent reproducibility
+mistakes from happening in the first place), and retroactively
+(analyzing code that is already written for potential problems).
+Furthermore, \texttt{fertile} is designed to educate the user about why
+the mistakes are problematic, and how to fix them. We discuss
+experimental results from testing \texttt{fertile} in an introductory
+data science course.
 \end{abstract}
@@
 \noindent%
-{\it Keywords:} 3 to 6 keywords, that do not appear in the title
+{\it Keywords:} reproducibility, statistical software, workflow, collaboration
 \vfill
@@
 \spacingset{1.45} % DON'T change the spacing!
+
+\section{Extended Abstract}\label{extended-abstract}
+
+In data science, an analyst has created \emph{reproducible} work if
+their code and data produces identical results when given to another
+analyst.
+
+This idea of reproducibility is becoming increasingly important to the
+advancement of our scientific knowledge. One reason for this is that
+reproducible research encourages transparency. Providing one's audience
+with well organized, reproducible code and results allows readers to
+understand the steps taken to generate findings from raw data and
+determine for themselves whether or not they believe the results to be
+reliable or trustworthy.
+
+Additionally, reproducibility encourages collaboration and extended
+research, allowing others to easily apply the methods used in one
+project to their own work with minimal effort.
+
+However, while the necessity of reproducibility is clear, there are
+significant behavioral and technical challenges that impede its
+widespread implementation, and no clear consensus on standards of what
+constitutes reproducibility in published research.
+
+• Note: What are these behavioral and technical challenges, and which
+ones should we mention?
+
+In defining the steps toward creating reproducible analyses, researchers
+place their emphasis on different areas, including file structure,
+documentation, the use of file paths in code, and accessibility. Within
+these categories, the recommendations for what steps to take in ensuring
+reproducibility also vary.
+
+• Why focus on R?
 
@@
 
-Practicing reproducible research is key for advancing our scientific
-knowledge. One reason for this is that reproducible research encourages
+One reason for this is that reproducible research encourages
 transparency. Providing one's audience with well organized, reproducible

### 2019-12-01 fertile.Rmd
more abstract drafting

--- 2019-11-20 fertile.Rmd
@@
 
+# Extended Abstract
+
+In the field of data science, an analyst has created *reproducible* work if: 1) their code and data files have been made publicly available, and 2) these files produce identical results when run by another analyst.
+
+In the age of modern computer technology, where knowledge can be instantly shared worldwide, reproducibility is becoming incredibly important to the advancement of scientific knowledge.
+
+The first reason for this is that reproducible research is transparent.
+When researchers provide the code and data used for their work in a well organized and reproducible format, readers are able to understand the steps taken to generate findings from raw data and determine whether or not they believe the results to be reliable or trustworthy.
+Without reproducibility, research results must be blindly trusted.
+
+The second reason is that reproducibility allows for collaboration and extended research.
+The creators of reproducible research can easily receive feedback and recommendations on their work, and others interested in the research topic can see the code and easily apply the methods and ideas used in one project to their own future work with minimal effort.
+
+However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation, and no clear consensus on standards of what constitutes reproducibility in published research.
+
+In defining the steps toward creating reproducible analyses, researchers place their emphasis on different areas, including file structure, documentation, the use of file paths in code, and accessibility.
+Within these categories, the recommendations for what steps to take in ensuring reproducibility also vary.
+
+Much of the discussion around reproducibility is also generalized, written in a way applicable to users working with a variety of statistical software programs.
+As a result, analysts interested in practicing reproducibility cannot easily access recommendations tailored specifically to their programming language or software of choice.
+
+Additionally, all statistical software programs operate differently.
+Since analysts do not all share the same software preferences, there can be no single set of recommendations or guidelines for reproducibility.
+Instead, recommendations must be specific to certain software programs.
+
+In this paper, we focus on reproducibility in the R programming language, with a specific concentration on the RStudio integrated development environment, due to its popularity for statistical analyses and the ease with which analysts can download and begin using the software.
+
+We propose an R package called `fertile`, which focuses on a series of common reproducibility-harming mistakes programmers make while conducting data science projects in R, warning users of their errors and providing recommendations for how to correct them.
+
+`fertile` operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that is already written for potential problems).
+
+Retroactively, `fertile` is designed to be run on an R Project folder, analyzing the project structure and the analyst's use of file paths and considering randomness, among other areas of interest.
+`fertile` creates reproducibility reports, identifying mistakes that users have made and providing recommendations for remedies.
+For example, fertile might identify an analyst's use of randomness in code and recommend setting a seed.
+
+Proactively, the package works similarly, identifying mistakes as they are made by the user and outputting an educational warning message identifying, and providing a solution for, the mistake.
+In its proactive warning system, `fertile` focuses primarily on the use of file paths.
+As users execute code, `fertile` looks for file paths passed to functions, identifying when absolute paths are provided or when paths point to a location outside of the project directory.
+
+`fertile` is designed to educate the user about why the mistakes are problematic, and how to fix them.
+It is written for a wide audience, simple enough to be used by students in an introductory data science course.
+
 # Introduction

### 2019-12-01 fertile.tex
more abstract drafting (.tex file)

@@
 
-In data science, an analyst has created \emph{reproducible} work if
-their code and data produces identical results when given to another
+In the field of data science, an analyst has created \emph{reproducible}
+work if: 1) their code and data files have been made publicly available,
+and 2) these files produce identical results when run by another
 analyst.
 
-This idea of reproducibility is becoming increasingly important to the
-advancement of our scientific knowledge. One reason for this is that
-reproducible research encourages transparency. Providing one's audience
-with well organized, reproducible code and results allows readers to
-understand the steps taken to generate findings from raw data and
-determine for themselves whether or not they believe the results to be
-reliable or trustworthy.
+In the age of modern computer technology, where knowledge can be
+instantly shared worldwide, reproducibility is becoming incredibly
+important to the advancement of scientific knowledge.
 
-Additionally, reproducibility encourages collaboration and extended
-research, allowing others to easily apply the methods used in one
-project to their own work with minimal effort.
+The first reason for this is that reproducible research is transparent.
+When researchers provide the code and data used for their work in a well
+organized and reproducible format, readers are able to understand the
+steps taken to generate findings from raw data and determine whether or
+not they believe the results to be reliable or trustworthy. Without
+reproducibility, research results must be blindly trusted.
+
+The second reason is that reproducibility allows for collaboration and
+extended research. The creators of reproducible research can easily
+receive feedback and recommendations on their work, and others
+interested in the research topic can see the code and easily apply the
+methods and ideas used in one project to their own future work with
+minimal effort.
 
@@
 constitutes reproducibility in published research.
-
-• Note: What are these behavioral and technical challenges, and which
-ones should we mention?
 
@@
 
-• Why focus on R?
+Much of the discussion around reproducibility is also generalized,
+written in a way applicable to users working with a variety of
+statistical software programs. As a result, analysts interested in
+practicing reproducibility cannot easily access recommendations tailored
+specifically to their programming language or software of choice.
+
+Additionally, all statistical software programs operate differently.
+Since analysts do not all share the same software preferences, there can
+be no single set of recommendations or guidelines for reproducibility.
+Instead, recommendations must be specific to certain software programs.
+
+In this paper, we focus on reproducibility in the R programming
+language, with a specific concentration on the RStudio integrated
+development environment, due to its popularity for statistical analyses
+and the ease with which analysts can download and begin using the
+software.
+
+We propose an R package called \texttt{fertile}, which focuses on a
+series of common reproducibility-harming mistakes programmers make while
+conducting data science projects in R, warning users of their errors and
+providing recommendations for how to correct them.
+
+\texttt{fertile} operates in two modes: proactively (to prevent
+reproducibility mistakes from happening in the first place), and
+retroactively (analyzing code that is already written for potential
+problems).
+
+Retroactively, \texttt{fertile} is designed to be run on an R Project
+folder, analyzing the project structure and the analyst's use of file
+paths and considering randomness, among other areas of interest.
+\texttt{fertile} creates reproducibility reports, identifying mistakes
+that users have made and providing recommendations for remedies. For
+example, fertile might identify an analyst's use of randomness in code
+and recommend setting a seed.
+
+Proactively, the package works similarly, identifying mistakes as they
+are made by the user and outputting an educational warning message
+identifying, and providing a solution for, the mistake. In its proactive
+warning system, \texttt{fertile} focuses primarily on the use of file
+paths. As users execute code, \texttt{fertile} looks for file paths
+passed to functions, identifying when absolute paths are provided or
+when paths point to a location outside of the project directory.
+
+\texttt{fertile} is designed to educate the user about why the mistakes
+are problematic, and how to fix them. It is written for a wide audience,
+simple enough to be used by students in an introductory data science
+course.
 

### 2019-12-08 fertile.Rmd
more edits (still need final section)

@@
 
-In defining the steps toward creating reproducible analyses, researchers place their emphasis on different areas, including file structure, documentation, the use of file paths in code, and accessibility.
+In defining the steps toward creating reproducible analyses, different researchers place their emphasis on different areas, including file structure, documentation, the use of file paths in code, and accessibility.
 Within these categories, the recommendations for what steps to take in ensuring reproducibility also vary.
@@
 Much of the discussion around reproducibility is also generalized, written in a way applicable to users working with a variety of statistical software programs.
-As a result, analysts interested in practicing reproducibility cannot easily access recommendations tailored specifically to their programming language or software of choice.
+Since all statistical software programs operate differently, generalized recommendations on reproducibility often cannot go in depth.
+While they provide some useful guidelines, they can often be relatively unhelpful, particularly to new analysts who are unsure how to apply such recommendations within the software programs they are using.
 
-Additionally, all statistical software programs operate differently.
-Since analysts do not all share the same software preferences, there can be no single set of recommendations or guidelines for reproducibility.
-Instead, recommendations must be specific to certain software programs.
+In order to be most effective, reproducibility recommendations must be tailored to specific software programs.
+However, there has been very little work on reproducibility that is tailored to this level of specificity.
+In this paper, attempt to remedy that, focusing on reproducibility in the R programming language with a concentration on the RStudio integrated development environment.
 
-In this paper, we focus on reproducibility in the R programming language, with a specific concentration on the RStudio integrated development environment, due to its popularity for statistical analyses and the ease with which analysts can download and begin using the software.
+R is an ideal candidate for reproducibility recommendations due to the language's popularity for statistical analyses and the ease with which analysts can download and begin using the software.
 

### 2019-12-08 fertile.tex
more edits (still need final section)

@@
 
-In defining the steps toward creating reproducible analyses, researchers
-place their emphasis on different areas, including file structure,
-documentation, the use of file paths in code, and accessibility. Within
-these categories, the recommendations for what steps to take in ensuring
-reproducibility also vary.
+In defining the steps toward creating reproducible analyses, different
+researchers place their emphasis on different areas, including file
+structure, documentation, the use of file paths in code, and
+accessibility. Within these categories, the recommendations for what
+steps to take in ensuring reproducibility also vary.
 
@@
 written in a way applicable to users working with a variety of
-statistical software programs. As a result, analysts interested in
-practicing reproducibility cannot easily access recommendations tailored
-specifically to their programming language or software of choice.
+statistical software programs. Since all statistical software programs
+operate differently, generalized recommendations on reproducibility
+often cannot go in depth. While they provide some useful guidelines,
+they can often be relatively unhelpful, particularly to new analysts who
+are unsure how to apply such recommendations within the software
+programs they are using.
 
-Additionally, all statistical software programs operate differently.
-Since analysts do not all share the same software preferences, there can
-be no single set of recommendations or guidelines for reproducibility.
-Instead, recommendations must be specific to certain software programs.
+In order to be most effective, reproducibility recommendations must be
+tailored to specific software programs. However, there has been very
+little work on reproducibility that is tailored to this level of
+specificity. In this paper, attempt to remedy that, focusing on
+reproducibility in the R programming language with a concentration on
+the RStudio integrated development environment.
 
-In this paper, we focus on reproducibility in the R programming
-language, with a specific concentration on the RStudio integrated
-development environment, due to its popularity for statistical analyses
-and the ease with which analysts can download and begin using the
-software.
+R is an ideal candidate for reproducibility recommendations due to the
+language's popularity for statistical analyses and the ease with which
+analysts can download and begin using the software.
 

### 2020-01-02 fertile_sdss.Rnw
added SDSS template and coverted abstract; #1

--- 2019-12-08 fertile.tex
@@
-% !TeX program = pdfLaTeX
-\documentclass[12pt]{article}
-\usepackage{amsmath}
-\usepackage{graphicx,psfrag,epsf}
-\usepackage{enumerate}
-\usepackage{natbib}
-\usepackage{textcomp}
-\usepackage[hyphens]{url} % not crucial - just used below for the URL
-\usepackage{hyperref}
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+% File SDSS2020_SampleExtendedAbstract.tex
+\documentclass[10pt]{article}
+\usepackage{sdss2020} % Uses Times Roman font (either newtx or times package)
+\usepackage{url}
+\usepackage{latexsym}
+\usepackage{amsmath, amsthm, amsfonts}
+\usepackage{algorithm, algorithmic}  
+\usepackage{graphicx}
+\usepackage{xspace}
 
-%\pdfminorversion=4
-% NOTE: To produce blinded version, replace "0" with "1" below.
-\newcommand{\blind}{0}
+% custom macros
 
-% DON'T change margins - should be 1 inch all around.
-\addtolength{\oddsidemargin}{-.5in}%
-\addtolength{\evensidemargin}{-.5in}%
-\addtolength{\textwidth}{1in}%
-\addtolength{\textheight}{1.3in}%
-\addtolength{\topmargin}{-.8in}%
+\newcommand{\R}{\textsf{R}\xspace}
+\newcommand{\RStudio}{\textsf{RStudio}\xspace}
+\newcommand{\cmd}[1]{\texttt{#1}}
+\newcommand{\pkg}[1]{\texttt{#1}}
 
-%% load any required packages here
+\title{Creating optimal conditions for reproducible data analysis in R with \pkg{fertile}}
+
+\author{
+  Audrey M. Bertin \\
+  Statistical and Data Sciences \\ 
+  Smith College \\
+  {\tt abertin@smith.edu} \\\And
+  Benjamin S. Baumer \\
+  Statistical and Data Sciences \\ 
+  Smith College \\
+  {\tt bbaumer@smith.edu}
+}
+
+\date{}
 
 \begin{document}
+\maketitle
+%\begin{abstract}
+%The advancement of scientific knowledge increasingly depends on ensuring that data-driven research is reproducible: that two people with the same data obtain the same results.
+However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation, and no clear consensus on standards of what constitutes reproducibility in published research.
+We focus on a series of common mistakes programmers make while conducting data science projects in R, primarily through the RStudio integrated development environment.
+\pkg{fertile} is an R package that operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that is already written for potential problems).
+Furthermore, \pkg{fertile} is designed to educate the user about why the mistakes are problematic, and how to fix them.
+We discuss experimental results from testing \pkg{fertile} in an introductory data science course.
+%\end{abstract}
 
-\def\spacingset#1{\renewcommand{\baselinestretch}%
-{#1}\small\normalsize} \spacingset{1}
+{\bf Keywords:} reproducibility, statistical software, workflow, collaboration
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
 
-\if0\blind
-{
-  \title{\bf Creating optimal conditions for reproducible data analysis in R with
-\texttt{fertile}}
+In the field of data science, an analyst has created \emph{reproducible} work if: 1) their code and data files have been made publicly available, and 2) these files produce identical results when run by another analyst.
 
-  \author{
-        Audrey Bertin \thanks{The authors gratefully acknowledge contributions from Hadley Wickham,
-Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the
-\texttt{tidyverse} team.} \\
-    Program in Statistical and Data Sciences, Smith College\\
-     and \\     Benjamin S. Baumer \\
-    Program in Statistical and Data Sciences, Smith College\\
-      }
-  \maketitle
-} \fi
-
-\if1\blind
-{
-  \bigskip
-  \bigskip
-  \bigskip
-  \begin{center}
-    {\LARGE\bf Creating optimal conditions for reproducible data analysis in R with
-\texttt{fertile}}
-  \end{center}
-  \medskip
-} \fi
-
-\bigskip
-\begin{abstract}
-The advancement of scientific knowledge increasingly depends on ensuring
-that data-driven research is reproducible: that two people with the same
-data obtain the same results. However, while the necessity of
-reproducibility is clear, there are significant behavioral and technical
-challenges that impede its widespread implementation, and no clear
-consensus on standards of what constitutes reproducibility in published
-research. We focus on a series of common mistakes programmers make while
-conducting data science projects in R, primarily through the RStudio
-integrated development environment. \texttt{fertile} is an R package
-that operates in two modes: proactively (to prevent reproducibility
-mistakes from happening in the first place), and retroactively
-(analyzing code that is already written for potential problems).
-Furthermore, \texttt{fertile} is designed to educate the user about why
-the mistakes are problematic, and how to fix them. We discuss
-experimental results from testing \texttt{fertile} in an introductory
-data science course.
-\end{abstract}
-
-\noindent%
-{\it Keywords:} reproducibility, statistical software, workflow, collaboration
-\vfill
-
-\newpage
-\spacingset{1.45} % DON'T change the spacing!
-
-\section{Extended Abstract}\label{extended-abstract}
-
-In the field of data science, an analyst has created \emph{reproducible}
-work if: 1) their code and data files have been made publicly available,
-and 2) these files produce identical results when run by another
-analyst.
-
-In the age of modern computer technology, where knowledge can be
-instantly shared worldwide, reproducibility is becoming incredibly
-important to the advancement of scientific knowledge.
+In the age of modern computer technology, where knowledge can be instantly shared worldwide, reproducibility is becoming incredibly important to the advancement of scientific knowledge.
 
 The first reason for this is that reproducible research is transparent.
-When researchers provide the code and data used for their work in a well
-organized and reproducible format, readers are able to understand the
-steps taken to generate findings from raw data and determine whether or
-not they believe the results to be reliable or trustworthy. Without
-reproducibility, research results must be blindly trusted.
+When researchers provide the code and data used for their work in a well organized and reproducible format, readers are able to understand the steps taken to generate findings from raw data and determine whether or not they believe the results to be reliable or trustworthy.
+Without reproducibility, research results must be blindly trusted.
 
-The second reason is that reproducibility allows for collaboration and
-extended research. The creators of reproducible research can easily
-receive feedback and recommendations on their work, and others
-interested in the research topic can see the code and easily apply the
-methods and ideas used in one project to their own future work with
-minimal effort.
+The second reason is that reproducibility allows for collaboration and extended research.
+The creators of reproducible research can easily receive feedback and recommendations on their work, and others interested in the research topic can see the code and easily apply the methods and ideas used in one project to their own future work with minimal effort.
 
-However, while the necessity of reproducibility is clear, there are
-significant behavioral and technical challenges that impede its
-widespread implementation, and no clear consensus on standards of what
-constitutes reproducibility in published research.
+However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation, and no clear consensus on standards of what constitutes reproducibility in published research.
 
-In defining the steps toward creating reproducible analyses, different
-researchers place their emphasis on different areas, including file
-structure, documentation, the use of file paths in code, and
-accessibility. Within these categories, the recommendations for what
-steps to take in ensuring reproducibility also vary.
+In defining the steps toward creating reproducible analyses, different researchers place their emphasis on different areas, including file structure, documentation, the use of file paths in code, and accessibility.
+Within these categories, the recommendations for what steps to take in ensuring reproducibility also vary.
 
-Much of the discussion around reproducibility is also generalized,
-written in a way applicable to users working with a variety of
-statistical software programs. Since all statistical software programs
-operate differently, generalized recommendations on reproducibility
-often cannot go in depth. While they provide some useful guidelines,
-they can often be relatively unhelpful, particularly to new analysts who
-are unsure how to apply such recommendations within the software
-programs they are using.
+Much of the discussion around reproducibility is also generalized, written in a way applicable to users working with a variety of statistical software programs.
+Since all statistical software programs operate differently, generalized recommendations on reproducibility often cannot go in depth.
+While they provide some useful guidelines, they can often be relatively unhelpful, particularly to new analysts who are unsure how to apply such recommendations within the software programs they are using.
 
-In order to be most effective, reproducibility recommendations must be
-tailored to specific software programs. However, there has been very
-little work on reproducibility that is tailored to this level of
-specificity. In this paper, attempt to remedy that, focusing on
-reproducibility in the R programming language with a concentration on
-the RStudio integrated development environment.
+In order to be most effective, reproducibility recommendations must be tailored to specific software programs.
+However, there has been very little work on reproducibility that is tailored to this level of specificity.
+In this paper, attempt to remedy that, focusing on reproducibility in the R programming language with a concentration on the RStudio integrated development environment.
 
-R is an ideal candidate for reproducibility recommendations due to the
-language's popularity for statistical analyses and the ease with which
-analysts can download and begin using the software.
+R is an ideal candidate for reproducibility recommendations due to the language's popularity for statistical analyses and the ease with which analysts can download and begin using the software.
 
-We propose an R package called \texttt{fertile}, which focuses on a
-series of common reproducibility-harming mistakes programmers make while
-conducting data science projects in R, warning users of their errors and
-providing recommendations for how to correct them.
+We propose an R package called \pkg{fertile}, which focuses on a series of common reproducibility-harming mistakes programmers make while conducting data science projects in R, warning users of their errors and providing recommendations for how to correct them.
 
-\texttt{fertile} operates in two modes: proactively (to prevent
-reproducibility mistakes from happening in the first place), and
-retroactively (analyzing code that is already written for potential
-problems).
+\pkg{fertile} operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that is already written for potential problems).
 
-Retroactively, \texttt{fertile} is designed to be run on an R Project
-folder, analyzing the project structure and the analyst's use of file
-paths and considering randomness, among other areas of interest.
-\texttt{fertile} creates reproducibility reports, identifying mistakes
-that users have made and providing recommendations for remedies. For
-example, fertile might identify an analyst's use of randomness in code
-and recommend setting a seed.
+Retroactively, \pkg{fertile} is designed to be run on an R Project folder, analyzing the project structure and the analyst's use of file paths and considering randomness, among other areas of interest.
+\pkg{fertile} creates reproducibility reports, identifying mistakes that users have made and providing recommendations for remedies.
+For example, fertile might identify an analyst's use of randomness in code and recommend setting a seed.
 
-Proactively, the package works similarly, identifying mistakes as they
-are made by the user and outputting an educational warning message
-identifying, and providing a solution for, the mistake. In its proactive
-warning system, \texttt{fertile} focuses primarily on the use of file
-paths. As users execute code, \texttt{fertile} looks for file paths
-passed to functions, identifying when absolute paths are provided or
-when paths point to a location outside of the project directory.
+Proactively, the package works similarly, identifying mistakes as they are made by the user and outputting an educational warning message identifying, and providing a solution for, the mistake.
+In its proactive warning system, \pkg{fertile} focuses primarily on the use of file paths.
+As users execute code, \pkg{fertile} looks for file paths passed to functions, identifying when absolute paths are provided or when paths point to a location outside of the project directory.
 
-\texttt{fertile} is designed to educate the user about why the mistakes
-are problematic, and how to fix them. It is written for a wide audience,
-simple enough to be used by students in an introductory data science
-course.
+\pkg{fertile} is designed to educate the user about why the mistakes are problematic, and how to fix them.
+It is written for a wide audience, simple enough to be used by students in an introductory data science course.
 
-\section{Introduction}\label{introduction}
-
-\section{The Importance of Ensuring
-Reproducibility}\label{the-importance-of-ensuring-reproducibility}
-
-One reason for this is that reproducible research encourages
-transparency. Providing one's audience with well organized, reproducible
-code and results allows readers to understand the steps taken to
-generate findings from raw data and determine for themselves whether or
-not they believe the results to be reliable or trustworthy.
-
-Additionally, reproducibility encourages collaboration and extended
-research, allows others to easily apply the methods used in one project
-to their own work with minimal effort. \citet{bray2014five}
-
-Fertile is designed to make reproducibility simple, providing fast and
-easy methods to test an R project for reproducibility. The package is
-intended to be usable by introductory data science students in their
-first semester of R.
-
---\textgreater{} make sure this info is included somewhere:
-
-From ``Five Concrete Reasons Your Students Should Be Learning to Analyze
-Data in the Reproducible Paradigm'' (\citet{bray2014five})
-
-\begin{itemize}
-\tightlist
-\item
-  Helps with transparency: make clear data cleaning steps between raw
-  data and final data
-\item
-  Helps with collaboration: easier to share code when it takes very few
-  steps to run on a different computer
-\end{itemize}
-
-The other big ideas on the importance/use of this package (not from
-sources):
-
-\begin{itemize}
-\tightlist
-\item
-  Make it easier for professors to grade students' code
-\item
-  Should be usable by intro level data science students
-\item
-  Possible use by reviewers of journal articles (and by those writing
-  the articles)
-\end{itemize}
-
-\section{What Defines Reproducibility in Data
-Science?}\label{what-defines-reproducibility-in-data-science}
-
-\section{\texorpdfstring{Creating Comprehensive Reproducibility Reports
-With
-``fertile''}{Creating Comprehensive Reproducibility Reports With fertile}}\label{creating-comprehensive-reproducibility-reports-with-fertile}
-
-Main points of different sources, as well as info about how they might
-be used for the paper.
-
-Here are some sources we might use for motivation behind the project:
-
-\subsubsection{The Reproducibility
-Crisis}\label{the-reproducibility-crisis}
-
-Big idea: most scientific fields are facing a reproducibility crisis and
-poor statistical use is considered one of the important reasons behind
-this.
-
-Not sure how useful this is due to the fact that it does not necessarily
-focus on the same kind of reproducibility we are looking at, which is
-code reproducibility rather than experimental reproducibility.
-
-\citet{baker20161}
-
-\subsubsection{Why is reproducibility
-important?}\label{why-is-reproducibility-important}
-
-From Popper in \texttt{The\ Logic\ of\ Scientific\ Discovery}:
-``non-reproducible single occurrences are of no significance to
-science.''
-
-\citet{popper2005logic}
-
-\subsubsection{What does reproducibility mean in data
-science?}\label{what-does-reproducibility-mean-in-data-science}
-
-``The ability of a researcher to duplicate the results of a prior study
-using the same materials as were used by the original investigator. That
-is, a second researcher might use the same raw data to build the same
-analysis files and implement the same statistical analysis in an attempt
-to yield the same results.''
-
-\citet{goodman2016does}
-
-From ``The Practice of Reproducibile Research'': Can all of the
-figures/calculations related to the result be reproduced in a single
-button press or at least with a reasonably small effort?
-
-\citet{kitzes2017practice}
-
-\subsubsection{What makes a data science project
-reproducibile?}\label{what-makes-a-data-science-project-reproducibile}
-
-Another way to think about this is: what features does fertile have that
-match what different sources think needs to be checked to ensure
-reproducibility?
-
-Some big ideas from the OpenSci discussion page:
-
-\begin{itemize}
-\tightlist
-\item
-  code should use tidy style
-\item
-  project should have short vignette files -- most written code should
-  be in an R/ directory
-\item
-  code and data should be stored in separate folders
-\item
-  a run environment and dependencies should be specified
-\item
-  there should be a README file
-\item
-  there should be a data dictionary
-\item
-  one master script
-\item
-  code should only use relative paths
-\item
-  code should be well commented and all variables should be defined
-\end{itemize}
-
-Big ideas from ``Packaging data analytical work reproducibly using R''
-
-\begin{itemize}
-\tightlist
-\item
-  research should be organized like an R package!
-\item
-  clear separation of data, method, and output
-\item
-  specify the computational environment that was used for the original
-  analysis (typically in a plain text file)
-\item
-  there should be a README describing the project and where to get
-  started
-\item
-  script files with reusable functions should go in an R/ directory
-\item
-  raw data files should be kept in a data/ directory
-\item
-  analysis scripts and report files should go in an analysis/ directory
-\item
-  for simpe projects, scripts should be given ascending names. For more
-  complicated ones, the use of some sort of makefile is recommended
-\item
-  there should also be a DESCRIPTION file with information about the
-  authors, project license, and software dependencies
-\end{itemize}
-
-\citet{marwick2018packaging}
-
-From ``The Practice of Reproducible Research''
-
-\begin{itemize}
-\item
-  Are the data openly accessible? If hosted online, is the web address
-  reliable long-term?
-\item
-  Are they in a commonly used and well-documented file format? Avoid
-  spreadsheets and instead use plain text data if possible!
-\item
-  Is the raw data available? Is sufficient metadata provided?
-\item
-  Are dependencies described properly?
-\item
-  Is full history of source code available through a public version
-  history
-\item
-  Is there a README?
-\item
-  Are functions documented?
-\item
-  Is there narrative documentation explaining how the different pieces
-  work together?
-\item
-  Are there usage examples?
-\end{itemize}
-
-Folder setup might look like:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Raw Data
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Data
-  \item
-    README
-  \end{itemize}
-\item
-  Clean Data
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Data
-  \end{itemize}
-\item
-  Results
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Results file
-  \end{itemize}
-\item
-  Src
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Analysis script
-  \item
-    Script to clean data
-  \end{itemize}
-\end{enumerate}
-
-\citet{kitzes2017practice}
-
-From R OpenSci's Reproducibility Guide:
-
-\url{http://ropensci.github.io/reproducibility-guide/}
-
-\begin{itemize}
-\tightlist
-\item
-  Is it clear where to begin?
-\item
-  Can you determinine which files were used as input to create output
-  files?
-\item
-  Is there documentation about every result?
-\item
-  Are exact versions of external applications noted?
-\item
-  If using randomness, are seeds noted?
-\item
-  Have you specified a license or noted licenses if you used other
-  people's content?
-\item
-  Are files easy to find?
-\item
-  Is it clear what the most recent file is?
-\item
-  Are there any folders that could be deleted?
-\item
-  Is analysis output done hierarchically?
-\item
-  Are there lots of manual data manipulation steps?
-\end{itemize}
-
-From ``A Guide to Reproducible Code in Ecology and Evolution'' (these
-ideas are pretty universal, though):
-
-A basic project structure:
-
-\begin{itemize}
-\item
-  The data folder contains all input data (and metadata) used in the
-  analysis.
-\item
-  The doc folder contains the manuscript
-\item
-  The figs directory contains figures generated by the analysis
-\item
-  The output folder contains any type of intermediate or output files
-  (e.g.~simulation outputs, models, processed datasets, etc.). You might
-  separate this and also have a cleaned-data folder.
-\item
-  The R directory contains R scripts with function definitions.
-\item
-  The reports folder contains RMarkdown files that document the analysis
-  or report on results
-\item
-  Consistent, ordered naming of scripts
-\item
-  Use portable paths
-\item
-  Write unit tests (only for advanced coding)
-\item
-  Show the packages you used
-\item
-  Record dependencies and versions of outside things you use
-\end{itemize}
-
-\citet{cooper2017guide}
-
-\section{Karl Broman's Suggestions}\label{karl-bromans-suggestions}
-
-\begin{itemize}
-\tightlist
-\item
-  Encapsulate everything within one directory
-\item
-  Separate raw data from derived data
-\item
-  Separate data from code
-\item
-  Use relative paths
-\item
-  Choose filenames correctly
-\item
-  Write README files
-\end{itemize}
-
-\url{http://kbroman.org/steps2rr/pages/organize.html}
-
-\section{Why focus on R?}\label{why-focus-on-r}
-
-\begin{itemize}
-\tightlist
-\item
-  R is the most popular language for statistical programming and is
-  specifically designed for statistics
-\item
-  R is great for reprodubility because the code is readable by users,
-  and RMarkdown is a great way to show/explain processes
-\item
-  R is easy to install and begin using
-\end{itemize}
-
---\textgreater{} from
-\url{https://openresearchsoftware.metajnl.com/articles/10.5334/jors.bu/print/}
-
-\section{Similar R Packages}\label{similar-r-packages}
-
-\subsubsection{rrtools}\label{rrtools}
-
-\url{https://github.com/benmarwick/rrtools}
-
-\begin{itemize}
-\tightlist
-\item
-  Creates a basic R package named after your research topic
-\item
-  Generates a license file
-\item
-  Connects to GitHub and creates a repository
-\item
-  Generates a README
-\item
-  Generates a reproducible directory structure
-\item
-  Creates a dockerfile
-\item
-  Creates a minimal travis file
-\item
-  Sets up testthat
-\end{itemize}
-
-\section{Msc sources to look at:}\label{msc-sources-to-look-at}
-
-\subsubsection{\texorpdfstring{Victoria Stodden's ``Implementing
-Reproducible Research''
-book}{Victoria Stodden's Implementing Reproducible Research book}}\label{victoria-stoddens-implementing-reproducible-research-book}
-
-\url{https://books.google.com/books?hl=en\&lr}=\&id=JcmSAwAAQBAJ\&oi=fnd\&pg=PP1\&dq=Victoria+stodden+reproducibility\&ots=ym\_gvSxMNH\&sig=B5N6FlkoocVjIprIZevtvrL33qQ\#v=onepage\&q=coombes\&f=false
-
-\bibliographystyle{agsm}
-\bibliography{bibliography.bib}
+\bibliographystyle{sdss2020} % Please do not change the bibliography style
+\bibliography{bibliography}
 

### 2020-01-02 fertile_sdss.Rnw
actually made edits; #1

@@
 \usepackage{sdss2020} % Uses Times Roman font (either newtx or times package)
-\usepackage{url}
+\usepackage{hyperref}
 \usepackage{latexsym}
@@
 
-In the field of data science, an analyst has created \emph{reproducible} work if: 1) their code and data files have been made publicly available, and 2) these files produce identical results when run by another analyst.
+Data-based research cannot be fully \emph{reproducible} unless the requisite code and data files produce identical results when run by another analyst. 
+As research in increasingly data-driven, and because knowledge can be shared worldwide so rapidly, reproducibility is critical to the advancement of scientific knowledge.
 
-In the age of modern computer technology, where knowledge can be instantly shared worldwide, reproducibility is becoming incredibly important to the advancement of scientific knowledge.
+When researchers provide the code and data used for their work in a well-organized and reproducible format, readers are more easily able to determine the veracity of any findings by following the steps from raw data to conclusions.
+The creators of reproducible research can more easily receive more specific feedback (including bug fixes) on their work.
+Moreover, others interested in the research topic can use the code to apply the methods and ideas used in one project to their own work with minimal effort.
 
-The first reason for this is that reproducible research is transparent.
-When researchers provide the code and data used for their work in a well organized and reproducible format, readers are able to understand the steps taken to generate findings from raw data and determine whether or not they believe the results to be reliable or trustworthy.
-Without reproducibility, research results must be blindly trusted.
+However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation, and no clear consensus on standards of what constitutes reproducibility in published research \cite{peng2009bio}.
+Not only are the \emph{components} of reproducible research up for discussion (e.g., need the software be open source?), but the corresponding \emph{recommendations} for ensuring reproducibility also vary (e.g., should raw and processed data files be in separate directories?).
 
-The second reason is that reproducibility allows for collaboration and extended research.
-The creators of reproducible research can easily receive feedback and recommendations on their work, and others interested in the research topic can see the code and easily apply the methods and ideas used in one project to their own future work with minimal effort.
-
-However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation, and no clear consensus on standards of what constitutes reproducibility in published research.
-
-In defining the steps toward creating reproducible analyses, different researchers place their emphasis on different areas, including file structure, documentation, the use of file paths in code, and accessibility.
-Within these categories, the recommendations for what steps to take in ensuring reproducibility also vary.
-
-Much of the discussion around reproducibility is also generalized, written in a way applicable to users working with a variety of statistical software programs.
-Since all statistical software programs operate differently, generalized recommendations on reproducibility often cannot go in depth.
-While they provide some useful guidelines, they can often be relatively unhelpful, particularly to new analysts who are unsure how to apply such recommendations within the software programs they are using.
-
-In order to be most effective, reproducibility recommendations must be tailored to specific software programs.
-However, there has been very little work on reproducibility that is tailored to this level of specificity.
-In this paper, attempt to remedy that, focusing on reproducibility in the R programming language with a concentration on the RStudio integrated development environment.
-
+Much of the discussion around reproducibility is also generalized---it is written to be applicable to users working with a variety of statistical software programs.
+Since all statistical software programs operate differently, generalized recommendations on reproducibility are often shallow and unspecific.
+While they provide useful guidelines, they can often be difficult to implement, particularly to new analysts who are unsure how to apply such recommendations within the software programs they are using.
+Thus, reproducibility recommendations tailored to specific software programs are more likely to be adopted. 
+In this paper, we focus on reproducibility in the R programming language with a concentration on projects that use the RStudio integrated development environment.
 R is an ideal candidate for reproducibility recommendations due to the language's popularity for statistical analyses and the ease with which analysts can download and begin using the software.
 
-We propose an R package called \pkg{fertile}, which focuses on a series of common reproducibility-harming mistakes programmers make while conducting data science projects in R, warning users of their errors and providing recommendations for how to correct them.
-
+We present an R package called \href{https://github.com/baumer-lab/fertile}{\pkg{fertile}}, which focuses on a series of common mistakes programmers make while conducting data science research in R.
 \pkg{fertile} operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that is already written for potential problems).
 
-Retroactively, \pkg{fertile} is designed to be run on an R Project folder, analyzing the project structure and the analyst's use of file paths and considering randomness, among other areas of interest.
-\pkg{fertile} creates reproducibility reports, identifying mistakes that users have made and providing recommendations for remedies.
-For example, fertile might identify an analyst's use of randomness in code and recommend setting a seed.
+Retroactively, \pkg{fertile} analyzes potential obstacles to reproducibility in an RStudio Project (i.e., a directory that contains an \cmd{.Rproj} file), including the directory structure, the analyst's use of file paths, randomness, etc.
+\pkg{fertile} creates reproducibility reports that identify potential mistakes and provide recommendations for remedies.
+For example, fertile might identify the use of randomness in code and recommend setting a seed.
 
-Proactively, the package works similarly, identifying mistakes as they are made by the user and outputting an educational warning message identifying, and providing a solution for, the mistake.
-In its proactive warning system, \pkg{fertile} focuses primarily on the use of file paths.
-As users execute code, \pkg{fertile} looks for file paths passed to functions, identifying when absolute paths are provided or when paths point to a location outside of the project directory.
+Proactively, the package identifies potential mistakes as they are made by the user and outputs an informative message as well as a recommended a solution.
+For example, \pkg{fertile} catches when a user passes a potentially problematic file path---such as an absolute path, or a path that points to a location outside of the project directory---to a variety of common input/output functions.
+\pkg{fertile} is even more aggressive with functions (like \cmd{setwd()}) that are almost certain to break reproducibility---it causes them to throw errors.
 
 \pkg{fertile} is designed to educate the user about why the mistakes are problematic, and how to fix them.
-It is written for a wide audience, simple enough to be used by students in an introductory data science course.
+It is written for a wide audience: simple enough to be used by students in an introductory data science course, but still helpful to experienced analysts.
+
+We also share preliminary results from a randomized, controlled experiment conducted on an undergraduate introductory data science course\footnote{This study is approved by Smith College IRB, Protocol \#19-032}.
+The purpose of the study is to determine whether \pkg{fertile} helps students produce data science research that is more likely to be reproducible.
 

### 2020-01-03 fertile_sdss.Rnw

@@
 \maketitle
-%\begin{abstract}
-%The advancement of scientific knowledge increasingly depends on ensuring that data-driven research is reproducible: that two people with the same data obtain the same results.
+\begin{abstract}
+The advancement of scientific knowledge increasingly depends on ensuring that data-driven research is reproducible: that two people with the same data obtain the same results.
 However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation, and no clear consensus on standards of what constitutes reproducibility in published research.
@@
 We discuss experimental results from testing \pkg{fertile} in an introductory data science course.
-%\end{abstract}
+\end{abstract}
 
@@
 
-We present an R package called \href{https://github.com/baumer-lab/fertile}{\pkg{fertile}}, which focuses on a series of common mistakes programmers make while conducting data science research in R.
+We present an R package called \href{https://github.com/baumer-lab/fertile}{\pkg{fertile}}\footnote{The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the `tidyverse` team.}, which focuses on a series of common mistakes programmers make while conducting data science research in R.
 \pkg{fertile} operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that is already written for potential problems).

### 2020-01-03 fertile_sdss.Rnw
additions to abstract

@@
 Data-based research cannot be fully \emph{reproducible} unless the requisite code and data files produce identical results when run by another analyst. 
-As research in increasingly data-driven, and because knowledge can be shared worldwide so rapidly, reproducibility is critical to the advancement of scientific knowledge.
+As research is becoming increasingly data-driven, and because knowledge can be shared worldwide so rapidly, reproducibility is critical to the advancement of scientific knowledge.
 
@@
 We present an R package called \href{https://github.com/baumer-lab/fertile}{\pkg{fertile}}\footnote{The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the `tidyverse` team.}, which focuses on a series of common mistakes programmers make while conducting data science research in R.
-\pkg{fertile} operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that is already written for potential problems).
+\pkg{fertile} operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that has already been written for potential problems).
 
@@
 
-Proactively, the package identifies potential mistakes as they are made by the user and outputs an informative message as well as a recommended a solution.
+Proactively, the package identifies potential mistakes as they are made by the user and outputs an informative message as well as a recommended solution.
 For example, \pkg{fertile} catches when a user passes a potentially problematic file path---such as an absolute path, or a path that points to a location outside of the project directory---to a variety of common input/output functions.
@@
 
-\pkg{fertile} is designed to educate the user about why the mistakes are problematic, and how to fix them.
-It is written for a wide audience: simple enough to be used by students in an introductory data science course, but still helpful to experienced analysts.
+The features and functions available in \pkg{fertile} are easily accessible to users.
+The retroactive reproducibility checks, also available individually, are packaged neatly into several primary functions that users of \pkg{fertile} can run quickly, in any order, by executing minimal lines of code.
 
-We also share preliminary results from a randomized, controlled experiment conducted on an undergraduate introductory data science course\footnote{This study is approved by Smith College IRB, Protocol \#19-032}.
+With two primary functions, users can access the majority of \pkg{fertile}'s retroactive features all at once:
+
+The \cmd{check()} function runs fifteen different reproducibility tests, noting which ones passed, which ones failed, the reason for failure, a recommended solution, and a guide to where to look for help.
+These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that to not have a call to \cmd{set.seed()} present.
+
+The \cmd{proj_analyze()}function creates a report documenting the structure of a data analysis project.
+This report contains informations about all packages referenced in code; the files present in the directory, and their types; suggestions for moving files to create a more organized structure; and a list of reproducibility-breaking file paths used in code.
+
+The proactive features are even more accessible, active immediately after loading the \pkg{fertile} package and requiring no additional code to access.
+
+In addition to the primary reproducability analsis functions present in \cmd{check}, \cmd{proj_analyze}, and the interactive warning system, several useful utility functions are made available to \pkg{fertile} users.
+Among other things, these include functions to check the type of a file and a way to create a copy of a project in a temp directory.
+
+\pkg{fertile} is designed to educate the user about why the mistakes are problematic, and how to fix them, promoting a greater understanding of reproducibility concepts in its users.
+
+It is also written for a wide audience, simple enough to be used by students in an introductory data science course, but still helpful to experienced analysts.
+
+The benefits and features of \pkg{fertile} can be accessed almost effortlessly, providing very little barrier to entry and making it easy for data analysists of all skill levels and backgrounds to gain a better understanding of how to make their work reproducible.
+
+In an effort to understand the package's effectiveness, we also share preliminary results from a randomized, controlled experiment conducted on an undergraduate introductory data science course\footnote{This study is approved by Smith College IRB, Protocol \#19-032}.
 The purpose of the study is to determine whether \pkg{fertile} helps students produce data science research that is more likely to be reproducible.

### 2020-01-03 fertile_sdss.Rnw
fixed LaTeX error

@@
 
-The \cmd{proj_analyze()}function creates a report documenting the structure of a data analysis project.
+The \cmd{proj\_analyze()} function creates a report documenting the structure of a data analysis project.
 This report contains informations about all packages referenced in code; the files present in the directory, and their types; suggestions for moving files to create a more organized structure; and a list of reproducibility-breaking file paths used in code.
@@
 
-In addition to the primary reproducability analsis functions present in \cmd{check}, \cmd{proj_analyze}, and the interactive warning system, several useful utility functions are made available to \pkg{fertile} users.
+In addition to the primary reproducability analsis functions present in \cmd{check}, \cmd{proj\_analyze()}, and the interactive warning system, several useful utility functions are made available to \pkg{fertile} users.
 Among other things, these include functions to check the type of a file and a way to create a copy of a project in a temp directory.

### 2020-01-03 fertile_sdss.Rnw
added examples; #1

@@
 Thus, reproducibility recommendations tailored to specific software programs are more likely to be adopted.
+
 In this paper, we focus on reproducibility in the R programming language with a concentration on projects that use the RStudio integrated development environment.
 R is an ideal candidate for reproducibility recommendations due to the language's popularity for statistical analyses and the ease with which analysts can download and begin using the software.
+
+\section{Methods}
 
@@
 \pkg{fertile} is even more aggressive with functions (like \cmd{setwd()}) that are almost certain to break reproducibility---it causes them to throw errors.
+
+\pkg{fertile} is available for download at \url{https://github.com/baumer-lab/fertile}.
+
+\section{Results}
 
@@
 
+<<message=FALSE>>=
+library(tidyverse)
+library(fertile)
+is_path_portable(".")
+is_path_portable("~")
+@
+
 \pkg{fertile} is designed to educate the user about why the mistakes are problematic, and how to fix them, promoting a greater understanding of reproducibility concepts in its users.
 
+<<include=FALSE>>=
+readr::write_csv(mtcars, "~/Desktop/my_data.csv")
+@
+
+<<>>=
+file.exists("~/Desktop/my_data.csv")
+read.csv("~/Desktop/my_data.csv")
+@
+
 It is also written for a wide audience, simple enough to be used by students in an introductory data science course, but still helpful to experienced analysts.
+
+<<message=FALSE>>=
+check_some(".", contains("paths"))
+@
 
@@
 
+\section{Conclusion}
+
+\pkg{fertile} is an R package that lowers barriers to reproducible data analysis projects in R.
+
 \bibliographystyle{sdss2020} % Please do not change the bibliography style

### 2020-01-04 fertile_sdss.Rnw
edited abstract examples

@@
 For example, \pkg{fertile} catches when a user passes a potentially problematic file path---such as an absolute path, or a path that points to a location outside of the project directory---to a variety of common input/output functions.
+
+<<include=FALSE>>=
+readr::write_csv(mtcars, "~/Desktop/my_data.csv")
+@
+
+<<message = FALSE>>=
+library(fertile)
+file.exists("~/Desktop/my_data.csv")
+read.csv("~/Desktop/my_data.csv")
+@
+
 \pkg{fertile} is even more aggressive with functions (like \cmd{setwd()}) that are almost certain to break reproducibility---it causes them to throw errors.
+
+<<message=FALSE>>=
+setwd("~/Desktop")
+@
 
@@
 library(fertile)
+
 is_path_portable(".")
 is_path_portable("~")
+
+is_data_file("~/Desktop/my_data.csv")
+is_image_file("~/Desktop/my_data.csv")
 @
@@
 \pkg{fertile} is designed to educate the user about why the mistakes are problematic, and how to fix them, promoting a greater understanding of reproducibility concepts in its users.
-
-<<include=FALSE>>=
-readr::write_csv(mtcars, "~/Desktop/my_data.csv")
-@
-
-<<>>=
-file.exists("~/Desktop/my_data.csv")
-read.csv("~/Desktop/my_data.csv")
-@
 

### 2020-01-06 fertile_sdss.Rnw
finalize two-page abstract; #1

--- 2020-01-03 fertile_sdss.Rnw
@@
 \usepackage{xspace}
+%\usepackage[T1]{fontenc}
 
@@
 \pkg{fertile} operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that has already been written for potential problems).
+\pkg{fertile} is available for download at \url{https://github.com/baumer-lab/fertile}.
 
-Retroactively, \pkg{fertile} analyzes potential obstacles to reproducibility in an RStudio Project (i.e., a directory that contains an \cmd{.Rproj} file), including the directory structure, the analyst's use of file paths, randomness, etc.
-\pkg{fertile} creates reproducibility reports that identify potential mistakes and provide recommendations for remedies.
-For example, fertile might identify the use of randomness in code and recommend setting a seed.
+\subsection{Proactive use}
 
@@
 For example, \pkg{fertile} catches when a user passes a potentially problematic file path---such as an absolute path, or a path that points to a location outside of the project directory---to a variety of common input/output functions.
-\pkg{fertile} is even more aggressive with functions (like \cmd{setwd()}) that are almost certain to break reproducibility---it causes them to throw errors.
-
-\pkg{fertile} is available for download at \url{https://github.com/baumer-lab/fertile}.
-
-\section{Results}
-
-The features and functions available in \pkg{fertile} are easily accessible to users.
-The retroactive reproducibility checks, also available individually, are packaged neatly into several primary functions that users of \pkg{fertile} can run quickly, in any order, by executing minimal lines of code.
-
-With two primary functions, users can access the majority of \pkg{fertile}'s retroactive features all at once:
-
-The \cmd{check()} function runs fifteen different reproducibility tests, noting which ones passed, which ones failed, the reason for failure, a recommended solution, and a guide to where to look for help.
-These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that to not have a call to \cmd{set.seed()} present.
-
-The \cmd{proj\_analyze()} function creates a report documenting the structure of a data analysis project.
-This report contains informations about all packages referenced in code; the files present in the directory, and their types; suggestions for moving files to create a more organized structure; and a list of reproducibility-breaking file paths used in code.
-
-The proactive features are even more accessible, active immediately after loading the \pkg{fertile} package and requiring no additional code to access.
-
-In addition to the primary reproducability analsis functions present in \cmd{check}, \cmd{proj\_analyze()}, and the interactive warning system, several useful utility functions are made available to \pkg{fertile} users.
-Among other things, these include functions to check the type of a file and a way to create a copy of a project in a temp directory.
-
-<<message=FALSE>>=
-library(tidyverse)
-library(fertile)
-is_path_portable(".")
-is_path_portable("~")
-@
-
-\pkg{fertile} is designed to educate the user about why the mistakes are problematic, and how to fix them, promoting a greater understanding of reproducibility concepts in its users.
 
@@
 
-<<>>=
+<<message = FALSE>>=
+library(tidyverse)
+library(fertile)
 file.exists("~/Desktop/my_data.csv")
@@
 
-It is also written for a wide audience, simple enough to be used by students in an introductory data science course, but still helpful to experienced analysts.
+\pkg{fertile} is even more aggressive with functions (like \cmd{setwd()}) that are almost certain to break reproducibility---it causes them to throw errors.
+
+<<message=FALSE>>=
+setwd("~/Desktop")
+@
+
+The proactive features are activated immediately after loading the \pkg{fertile} package and require no additional effort by the user.
+
+In addition to the interactive warning system, \pkg{fertile} provides several useful utility functions.
+Among other things, these include functions to check the type of a file and a way to create a copy of a project in a temporary directory.
+
+<<message=FALSE>>=
+is_path_portable(".")
+is_path_portable("~")
+
+is_data_file("~/Desktop/my_data.csv")
+@
+
+\subsection{Retroactive use}
+
+Retroactively, \pkg{fertile} analyzes potential obstacles to reproducibility in an RStudio Project (i.e., a directory that contains an \cmd{.Rproj} file), including the directory structure, the analyst's use of file paths, randomness, etc.
+\pkg{fertile} creates reproducibility reports that identify potential mistakes and provide recommendations for remedies.
+For example, \pkg{fertile} might identify the use of randomness in code and recommend setting a seed.
+
+Users can access the majority of \pkg{fertile}'s retroactive features through two primary functions.
+
+The \cmd{check()} function runs fifteen different reproducibility tests, noting which ones passed, which ones failed, the reason for failure, a recommended solution, and a guide to where to look for help.
+These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that to not have a call to \cmd{set.seed()} present.
+Subsets of the fifteen tests can be invoked using the select helper functions from \pkg{dplyr}\cite{R-dplyr} and the \cmd{check\_some()} function.
+
+<<eval=FALSE, include=FALSE>>=
+knitr::write_bib("dplyr", file = "pkgs.bib")
+@
 
@@
 
-The benefits and features of \pkg{fertile} can be accessed almost effortlessly, providing very little barrier to entry and making it easy for data analysists of all skill levels and backgrounds to gain a better understanding of how to make their work reproducible.
+The \cmd{proj\_analyze()} function creates a report documenting the structure of a data analysis project.
+This report contains informations about all packages referenced in code, the files present in the directory and their types, suggestions for moving files to create a more organized structure, and a list of reproducibility-breaking file paths used in code.
+
+\section{Results}
 
@@
 \pkg{fertile} is an R package that lowers barriers to reproducible data analysis projects in R. 
+The features of \pkg{fertile} can be accessed almost effortlessly, lowering barriers to entry and making it easy for data analysts of all skill levels and backgrounds to gain a better understanding of how to make their work reproducible.
+\pkg{fertile} is designed to educate the user about why the mistakes are problematic and how to fix them, promoting a greater understanding of reproducibility concepts in its users.
+It is written for a wide audience, simple enough to be used by students in an introductory data science course, but still helpful to experienced analysts.
 
 \bibliographystyle{sdss2020} % Please do not change the bibliography style
-\bibliography{bibliography}
+\bibliography{bibliography,pkgs}
 

### 2020-01-08 fertile_sdss.Rnw
human aspect of reproducibility

@@
 <<message = FALSE>>=
-library(tidyverse)
 library(fertile)
@@
 <<message=FALSE>>=
-is_path_portable(".")
 is_path_portable("~")
@@
 These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that to not have a call to \cmd{set.seed()} present.
-Subsets of the fifteen tests can be invoked using the select helper functions from \pkg{dplyr}\cite{R-dplyr} and the \cmd{check\_some()} function.
+Subsets of the fifteen tests can be invoked using the select helper functions from \pkg{dplyr} and the \cmd{check\_some()} function.
 
@@
 <<message=FALSE>>=
+library(tidyverse)
 check_some(".", contains("paths"))
@@
 \pkg{fertile} is an R package that lowers barriers to reproducible data analysis projects in R. 
-The features of \pkg{fertile} can be accessed almost effortlessly, lowering barriers to entry and making it easy for data analysts of all skill levels and backgrounds to gain a better understanding of how to make their work reproducible.
+The features of \pkg{fertile} can be accessed almost effortlessly, making it easy for data analysts of all skill levels and backgrounds to gain a better understanding of how to make their work reproducible.
 \pkg{fertile} is designed to educate the user about why the mistakes are problematic and how to fix them, promoting a greater understanding of reproducibility concepts in its users.
 It is written for a wide audience, simple enough to be used by students in an introductory data science course, but still helpful to experienced analysts.
+\pkg{fertile} also addresses a human challenge of reproducibility.
+In the moment, it often feels easiest to take a shortcut---to use an absolute path or change a working directory.
+However, when considering the long term path of a project, spending the extra time to improve reproducibility is worthwhile.
+\pkg{fertile}'s user-friendly features can help data analysts avoid these harmful shortcuts with minimal effort.
 

### 2020-01-09 fertile_sdss.Rnw
updated to proj_check()

@@
 
-The \cmd{check()} function runs fifteen different reproducibility tests, noting which ones passed, which ones failed, the reason for failure, a recommended solution, and a guide to where to look for help.
+The \cmd{proj\_check()} function runs fifteen different reproducibility tests, noting which ones passed, which ones failed, the reason for failure, a recommended solution, and a guide to where to look for help.
 These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that to not have a call to \cmd{set.seed()} present.

### 2020-01-09 fertile_sdss.Rnw
check_some() --> proj_check_some()
check_some() -> proj_check_some() in text

@@
 These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that to not have a call to \cmd{set.seed()} present.
-Subsets of the fifteen tests can be invoked using the select helper functions from \pkg{dplyr} and the \cmd{check\_some()} function.
+Subsets of the fifteen tests can be invoked using the select helper functions from \pkg{dplyr} and the \cmd{proj\_check\_some()} function.
 
@@
 library(tidyverse)
-check_some(".", contains("paths"))
+proj_check_some(".", contains("paths"))
 @

### 2020-01-12 fertile.Rmd
resolving issue #15/17 (references)

@@
 
+R OpenSci also has a discussion of code review practices where users commented. Not too much that was helpful from here (just a few odd comments):
+https://discuss.ropensci.org/t/how-do-you-review-code-that-accompanies-a-research-project-or-paper-help-ropensci-plan-a-community-call/1321/43
+
 From "A Guide to Reproducible Code in Ecology and Evolution" (these ideas are pretty universal, though):
@@
 
+### represtools
+
+https://pirategrunt.com/represtools/
+
+Has basic utility functions to support reproducibile research
+• Gather
+• Cook
+• Analyze
+• Present
+
+### orderly
+
+https://github.com/vimc/orderly
+
+Focuses on automating a series of basic steps in the prodcess of writing analyses.
+• Tracks inputs
+• Stores multiple versions of an analysis (unclear whether this is different stages?)
+• Track outputs
+• Create analyses that depend on the outputs of previous analyses
+• You have to declare what you're going to use and create BEFORE you start the analysis
+
+### workflowr
+
+https://f1000research.com/articles/8-1749
+
+Aims to instill a particular workflow that helps make projects more reproducible.
+
+• Version control (git)
+• Literate programming (Rmd)
+• Automatic checks & safeguards (seemingly related to whether different files in a project have been published on git, and not actually to do with code practices)
+• Sharing code / results via browsable website
+
+### checkers
+
+https://github.com/ropenscilabs/checkers
+
+• A framework for reviewing analysis projects
+• Seems to be something run after-the-fact
+• Currently a work in progress and not passing checks (as of Jan 2019)
+• Can create custom checks
+• Checks include:
+ - Comments
+ - Good coding style (lintr)
+ - Use of version control
+ - Whether preferred packages are used
+
 # Msc sources to look at:

### 2020-01-13 fertile_sdss_presentation.Rmd
presentation files

---
title: "Creating optimal conditions for reproducible data analysis in R with 'fertile'"
subtitle: 
author: |
  Audrey Bertin<br/>
  <span style = 'font-size: 70%;'>
  [Statistical & Data Sciences](http://www.smith.edu/sds), Smith College<br/>
  `r icon::fa("envelope")` &nbsp;[abertin@smith.edu](mailto:abertin@smith.edu) | 
  `r icon::fa("twitter")` @[ambertin99](http://www.twitter.com/ambertin99) |
  `r icon::fa("github")` [ambertin](http://www.github.com/ambertin)
  </span>
date: |
  SDSS 2020 <br/>
  Date <br/>
  `r sds::img_logo(width = 64)`
output:
  sds::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false

---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE}
# sds::duo_smith()
sds::mono_light_smith()
```

# Reproducibility

.pull-left[

• When code + data files produce **identical** results when run by another analyst

### Why this is important:
  - Research is increasingly data-driven
  - Knowledge can be shared worldwide rapidly
  - Transparency
  - Specific feedback
  - Sharing of knowledge / future work

]

.pull-right[

![Diagram by Stephanie Hicks and Roger Peng](images/reproducibility.png)

]

---

# Reproducibility challenges

• No consensus on standards of reproducibility in research
  - **components** (e.g. need it be open source?)
  - **recommendations** (e.g. should raw / processed data be stored separately?)

• Discussion is very generalized
  - For any software
  - Difficult to implement (particularly for new analysts)

---

# fertile addresses this

• `fertile`: An R package that focuses on common mistakes data analysts make in R

### Two modes

• Proactive 
  - Stops mistakes as they are happening

• Retroactive
  - Analyzes already-written code for potential issues

---
# Proactive use

• Identifies potential mistakes as they are made

• Outputs an informative message and a recommended solution

.pull-left[

• Ex: detecting problematic file paths (absolute or pointing outside a directory)

![](images/absolute.png)

]

.pull-right[

• Ex: setting working directories

![](images/setwd.png)

]

---
# Retroactive use

• Analyze an RStudio project:
  - Directory structure
  - File path use
  - Randomness

• Provide recommendations for remedies

• Features easily accessed w/ two simple functions

---
# Retroactive use

### 1. `proj_check()`

• Runs 15 reproducibility checks
  - Clear build chain
  - Clutter-free project root level
  - No unused files
  - Randomness without `set.seed()`

• Notes which pass/fail

• Gives recommendations for addressing issues

• Gives a guide of where to look for help

---

![](images/proj_check.png)

---
# Retroactive use

### 2. `proj_analyze()`

• Creates a report focused primarily on the structure of a data analysis project

• Contains:
  - Packages referenced
  - Files / types
  - Suggestions for reorganization
  - Problematic file paths

---

![](images/proj_analyze.png)

---
# Why `fertile` matters

• `fertile` lowers barriers to reproducibile data analysis in R

• Provides non-generalized recommendations

• Near-effortless accessibility

• Good for analysts of all skill levels / backgrounds

--- 
---

# `fertile` is different

• There are already R packages (some still in development) trying to address this issue:
  -rrtools
  -represtools
  -orderly
  -workflowr
  -checkers

• However, many of these are:
  -Complicated and difficult for beginners
  -Focused on just one aspect of reproducibility

• `fertile` is incredibly simple to use and covers a wide variety of topics

---

# Testing `fertile`

• Is `fertile` actually effective for teaching reproducibility?
• Experimented in a Smith College introductory data science course

• Findings:

---

class: center, middle
# Thank you!

---
class: center, middle
## Acknowledgements

The authors gratefully acknowledge contributions from:
Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the `tidyverse` team.

---

`r icon::fa("envelope")` [abertin@smith.edu](mailto:abertin@smith.edu) |
`r icon::fa("twitter")` [ambertin99](http://www.twitter.com/ambertin99) |
`r icon::fa("github")` [ambertin](http://www.github.com/ambertin)

### 2020-02-09 fertile.Rmd
literature review updates

@@
 https://books.google.com/books?hl=en&lr=&id=JcmSAwAAQBAJ&oi=fnd&pg=PP1&dq=Victoria+stodden+reproducibility&ots=ym_gvSxMNH&sig=B5N6FlkoocVjIprIZevtvrL33qQ#v=onepage&q=coombes&f=false
+
+# Literature Review
+
+## Papers and Books
+
+*What does research reproducibility mean? -- Goodman et al*
+
+•The language and conceptual framework of research reproducibility vary across the sciences. There are no clear standards agreed upon by all scientific researchers.
+
+*The practice of reproducible research: case studies and lessons from the data-intensive sciences -- Kitzes et al*
+
+• Reproducibility examples from a variety of fields all vary significantly but they share key principles / practices in common: 1.
+Use clear separation, labeling, and documentation.
+2.
+Automate documentation.
+3.
+Design a workflow as a sequence of small steps glued together with intermediate outputs from one step feeding into the next.
+
+*Packaging data analytical work reproducibly using {R} (and friends) -- Marwick et al*
+
+• The R package can be a suitable template for organizing files into a research compendium to enhance the reproducibility of research.
+
+*A guide to reproducible code in ecology and evolution -- Cooper et al*
+
+• Reproducible work should include a clear and well-separated file system, well-named files, a list of package dependencies, good documentation.
+Preferably, RMarkdown and version control should be used.
+
+*Reproducible research and Biostatistics -- Peng*
+
+• To meet the "reproducible" criterion in the biostatistics journal you must submit: a main script and any complementary side scripts, data and any other necessary files, a target file containing results.
+
+# Other articles
+
+*Karl Broman's suggestions:*
+
+• Keep the project contained in one directory, separate the components (raw data, derived data, code), use relative paths, write a README, and name things clearly.
+
+*R OpenSci's suggestions:*
+
+• To be reproducible, it should be easy to find files, be clear what each file is, and be easy to know what order to run files.
+There should be no extraneous files.
+Everything should be documented, dependencies noted, and seeds used for randomness.
+Automation should be used when possible.
+
+## Packages
+
+rrtools
+• Complementary to Ben Marwick's paper
+• Creates a basic R package structure for your research and institutes some testthat check functionality
+
+orderly
+• Based on the step-by-step analysis process. 
+• Build components that take inputs from other components
+• Have to declare project structure at the beginning.
+
+workflowr
+• Focused a lot on version control and making code available
+
+checkers
+• Allows you to create custom checks to look over your work
+
+*Notes: Almost all of these are things you need to do before you start your project.
+They each focus on a small subset of reproducibility components.
+They are somewhat complicated to use -- commands are complex and there are often a lot of them.

### 2020-02-09 fertile.Rmd
added use case + start of lit review

@@
 
-# Extended Abstract
+# Introduction
 
-In the field of data science, an analyst has created *reproducible* work if: 1) their code and data files have been made publicly available, and 2) these files produce identical results when run by another analyst.
+Data-based research cannot be fully *reproducible* unless the requisite code and data files produce identical results when run by another analyst.
 
-In the age of modern computer technology, where knowledge can be instantly shared worldwide, reproducibility is becoming incredibly important to the advancement of scientific knowledge.
+As research is becoming increasingly data-driven, and because knowledge can be shared worldwide so rapidly, reproducibility is critical to the advancement of scientific knowledge.
 
-The first reason for this is that reproducible research is transparent.
-When researchers provide the code and data used for their work in a well organized and reproducible format, readers are able to understand the steps taken to generate findings from raw data and determine whether or not they believe the results to be reliable or trustworthy.
-Without reproducibility, research results must be blindly trusted.
+When researchers provide the code and data used for their work in a well-organized and reproducible format, readers are more easily able to determine the veracity of any findings by following the steps from raw data to conclusions.
 
-The second reason is that reproducibility allows for collaboration and extended research.
-The creators of reproducible research can easily receive feedback and recommendations on their work, and others interested in the research topic can see the code and easily apply the methods and ideas used in one project to their own future work with minimal effort.
+The creators of reproducible research can more easily receive more specific feedback (including bug fixes) on their work.
+Moreover, others interested in the research topic can use the code to apply the methods and ideas used in one project to their own work with minimal effort.
 
-However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation, and no clear consensus on standards of what constitutes reproducibility in published research.
+However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation, and no clear consensus on standards of what constitutes reproducibility in published research (@peng2009bio).
+Not only are the *components* of reproducible research up for discussion (e.g., need the software be open source?), but the corresponding *recommendations* for ensuring reproducibility also vary (e.g., should raw and processed data files be in separate directories?).
 
-In defining the steps toward creating reproducible analyses, different researchers place their emphasis on different areas, including file structure, documentation, the use of file paths in code, and accessibility.
-Within these categories, the recommendations for what steps to take in ensuring reproducibility also vary.
+Much of the discussion around reproducibility is also generalized---it is written to be applicable to users working with a variety of statistical software programs.
+Since all statistical software programs operate differently, generalized recommendations on reproducibility are often shallow and unspecific.
+While they provide useful guidelines, they can often be difficult to implement, particularly to new analysts who are unsure how to apply such recommendations within the software programs they are using.
 
-Much of the discussion around reproducibility is also generalized, written in a way applicable to users working with a variety of statistical software programs.
-Since all statistical software programs operate differently, generalized recommendations on reproducibility often cannot go in depth.
-While they provide some useful guidelines, they can often be relatively unhelpful, particularly to new analysts who are unsure how to apply such recommendations within the software programs they are using.
+Thus, reproducibility recommendations tailored to specific software programs are more likely to be adopted.
 
-In order to be most effective, reproducibility recommendations must be tailored to specific software programs.
-However, there has been very little work on reproducibility that is tailored to this level of specificity.
-In this paper, attempt to remedy that, focusing on reproducibility in the R programming language with a concentration on the RStudio integrated development environment.
+In this paper, we focus on reproducibility in the R programming language with a concentration on projects that use the RStudio integrated development environment.
 
 R is an ideal candidate for reproducibility recommendations due to the language's popularity for statistical analyses and the ease with which analysts can download and begin using the software.  
+Several researchers and R users have recognized this and worked to publish a small body of papers and R packages focusing on increasing reproducibility in the R community.
 
-We propose an R package called `fertile`, which focuses on a series of common reproducibility-harming mistakes programmers make while conducting data science projects in R, warning users of their errors and providing recommendations for how to correct them.
+Much of this work is narrowly tailored, with each package effectively addressising a small component of reproducibility--file structure, modularization of code, version control, etc. 
+Many existing reproducibility packages, due to their focused nature, succeed at their area of focus well, but at a cost.
+They are often semi-complex to learn and operate, providing a barrier to entry for less-experienced data analysts.
 
-`fertile` operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that is already written for potential problems).
+There appear to be no existing tools that cater to analysts looking for an easy-to-learn, easy-to-implement, fast way to obtain a broad overview of their projects' reproducibility.
+To address this, we present an R package called [`fertile`](https://github.com/baumer-lab/fertile) [^1]:{The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the `tidyverse` team.}, a low barrier-to-entry package which focuses on a series of common mistakes programmers make while conducting data science research in R.
 
-Retroactively, `fertile` is designed to be run on an R Project folder, analyzing the project structure and the analyst's use of file paths and considering randomness, among other areas of interest.
-`fertile` creates reproducibility reports, identifying mistakes that users have made and providing recommendations for remedies.
-For example, fertile might identify an analyst's use of randomness in code and recommend setting a seed.
+# Literature Review
 
-Proactively, the package works similarly, identifying mistakes as they are made by the user and outputting an educational warning message identifying, and providing a solution for, the mistake.
-In its proactive warning system, `fertile` focuses primarily on the use of file paths.
-As users execute code, `fertile` looks for file paths passed to functions, identifying when absolute paths are provided or when paths point to a location outside of the project directory.
+## Academic Research
 
-`fertile` is designed to educate the user about why the mistakes are problematic, and how to fix them.
-It is written for a wide audience, simple enough to be used by students in an introductory data science course.
+*What does research reproducibility mean? -- Goodman et al*
+
+•The language and conceptual framework of research reproducibility vary across the sciences. There are no clear standards agreed upon by all scientific researchers.
+
+*The practice of reproducible research: case studies and lessons from the data-intensive sciences -- Kitzes et al*
+
+• Reproducibility examples from a variety of fields all vary significantly but they share key principles / practices in common: 1.
+Use clear separation, labeling, and documentation.
+2.
+Automate documentation.
+3.
+Design a workflow as a sequence of small steps glued together with intermediate outputs from one step feeding into the next.
+
+*Packaging data analytical work reproducibly using {R} (and friends) -- Marwick et al*
+
+• The R package can be a suitable template for organizing files into a research compendium to enhance the reproducibility of research.
+
+*A guide to reproducible code in ecology and evolution -- Cooper et al*
+
+• Reproducible work should include a clear and well-separated file system, well-named files, a list of package dependencies, good documentation.
+Preferably, RMarkdown and version control should be used.
+
+*Reproducible research and Biostatistics -- Peng*
+
+• To meet the "reproducible" criterion in the biostatistics journal you must submit: a main script and any complementary side scripts, data and any other necessary files, a target file containing results.
+
+## Other Articles
+
+*Karl Broman's suggestions:*
+
+• Keep the project contained in one directory, separate the components (raw data, derived data, code), use relative paths, write a README, and name things clearly.
+
+*R OpenSci's suggestions:*
+
+• To be reproducible, it should be easy to find files, be clear what each file is, and be easy to know what order to run files.
+There should be no extraneous files.
+Everything should be documented, dependencies noted, and seeds used for randomness.
+Automation should be used when possible.
+
+## R Packages
+
+rrtools
+• Complementary to Ben Marwick's paper
+• Creates a basic R package structure for your research and institutes some testthat check functionality
+
+orderly
+• Based on the step-by-step analysis process. 
+• Build components that take inputs from other components
+• Have to declare project structure at the beginning.
+
+workflowr
+• Focused a lot on version control and making code available
+
+checkers
+• Allows you to create custom checks to look over your work
+
+*Notes: Almost all of these are things you need to do before you start your project.
+They each focus on a small subset of reproducibility components.
+They are somewhat complicated to use -- commands are complex and there are often a lot of them.
+
+# Methods
+
+`fertile` operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that has already been written for potential problems).
+`fertile` is available for download at [https://github.com/baumer-lab/fertile](https://github.com/baumer-lab/fertile)
+
+## Proactive Use
+
+Proactively, the package identifies potential mistakes as they are made by the user and outputs an informative message as well as a recommended solution.
+For example, `fertile` catches when a user passes a potentially problematic file path---such as an absolute path, or a path that points to a location outside of the project directory---to a variety of common input/output functions.
+
+```{r setup, cache = F, include = FALSE}
+knitr::opts_chunk$set(error = TRUE)
+```
+
+```{r, include = FALSE}
+readr::write_csv(mtcars, "~/Desktop/my_data.csv")
+library(fertile)
+```
+
+```{r, message = FALSE}
+file.exists("~/Desktop/my_data.csv")
+read.csv("~/Desktop/my_data.csv")
+```
+
+`fertile` is even more aggressive with functions (like `setwd()`) that are almost certain to break reproducibility---it causes them to throw errors.
+
+```{r, message = FALSE}
+setwd("~/Desktop")
+```
+
+The proactive features are activated immediately after loading the `fertile` package and require no additional effort by the user.
+
+In addition to the interactive warning system, `fertile` provides several useful utility functions.
+Among other things, these include functions to check the type of a file and a way to create a copy of a project in a temporary directory.
+
+```{r}
+is_path_portable("~")
+
+is_data_file("~/Desktop/my_data.csv")
+```
+
+## Retroactive Use
+
+Retroactively, `fertile` analyzes potential obstacles to reproducibility in an RStudio Project (i.e., a directory that contains an `.Rproj` file), including the directory structure, the analyst's use of file paths, randomness, etc.
+`fertile}` creates reproducibility reports that identify potential mistakes and provide recommendations for remedies.
+For example, `fertile` might identify the use of randomness in code and recommend setting a seed.
+
+Users can access the majority of `fertile`'s retroactive features through two primary functions.
+
+The `proj_check()` function runs fifteen different reproducibility tests, noting which ones passed, which ones failed, the reason for failure, a recommended solution, and a guide to where to look for help.
+These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that to not have a call to `set.seed()` present.
+Subsets of the fifteen tests can be invoked using the select helper functions from `dplyr` and the `proj_check_some()` function.
+
+```{r, eval = FALSE, include = FALSE}
+knitr::write_bib("dplyr", file = "pkgs.bib")
+```
+
+```{r, message = FALSE}
+library(tidyverse)
+proj_check_some(".", contains("paths"))
+```
+
+The `proj_analyze()` function creates a report documenting the structure of a data analysis project.
+This report contains informations about all packages referenced in code, the files present in the directory and their types, suggestions for moving files to create a more organized structure, and a list of reproducibility-breaking file paths used in code.
+
+## Sample Use Cases
+
+`fertile` is designed in such a way as to be useful to users of all backgrounds and experiences.
+In this section, we provide two sample use cases demonstrating situations where users could benefit significantly from using the package.
+
+### Introductory Data Science Student
+
+Susan is taking an introductory data science course.
+This is her first time learning how to code, and she has not yet been exposed to ideas of research reproducibility.
+Her professor has assigned a data analysis project which must be completed in R Markdown.
+The project requires her to read in a data file located on her computer and use it to produce a graph.
+
+She reads in the data, makes the graph, and knits her .Rmd file.
+It compiles successfully, so she submits the assignment.
+The next day, she receives an email from her professor saying that her assignment failed to compile and that she needs to make changes and try again.
+
+Susan doesn't understand why it didn't work on the professor's computer when it did on her own.
+The professor recommends that she install `fertile` and run `proj_check()` on her assignment.
+She does this and gets a message informing her she used an absolute path to open her dataset, and she should use a relative path.
+She looks up what this means and then uses the new information to update her assignment.
+She resubmits and her second attempt is successful.
+
+On future projects, she always loads and runs fertile to make sure her work is okay before submitting.
+
+### 2. Experience R User
+
+# Results
+
+In an effort to understand the package's effectiveness, we also share preliminary results from a randomized, controlled experiment conducted on an undergraduate introductory data science course [^2]:{This study is approved by Smith College IRB, Protocol #19-032}.
+The purpose of the study is to determine whether `fertile` helps students produce data science research that is more likely to be reproducible.
+
+# Conclusion
+
+`fertile` is an R package that lowers barriers to reproducible data analysis projects in R. 
+The features of `fertile` can be accessed almost effortlessly, making it easy for data analysts of all skill levels and backgrounds to gain a better understanding of how to make their work reproducible.
+`fertile` is designed to educate the user about why the mistakes are problematic and how to fix them, promoting a greater understanding of reproducibility concepts in its users.
+It is written for a wide audience, simple enough to be used by students in an introductory data science course, but still helpful to experienced analysts.
+`fertile` also addresses a human challenge of reproducibility.
+In the moment, it often feels easiest to take a shortcut---to use an absolute path or change a working directory.
+However, when considering the long term path of a project, spending the extra time to improve reproducibility is worthwhile.
+`fertile`'s user-friendly features can help data analysts avoid these harmful shortcuts with minimal effort.
 
@@
 https://books.google.com/books?hl=en&lr=&id=JcmSAwAAQBAJ&oi=fnd&pg=PP1&dq=Victoria+stodden+reproducibility&ots=ym_gvSxMNH&sig=B5N6FlkoocVjIprIZevtvrL33qQ#v=onepage&q=coombes&f=false
-
-# Literature Review
-
-## Papers and Books
-
-*What does research reproducibility mean? -- Goodman et al*
-
-•The language and conceptual framework of research reproducibility vary across the sciences. There are no clear standards agreed upon by all scientific researchers.
-
-*The practice of reproducible research: case studies and lessons from the data-intensive sciences -- Kitzes et al*
-
-• Reproducibility examples from a variety of fields all vary significantly but they share key principles / practices in common: 1.
-Use clear separation, labeling, and documentation.
-2.
-Automate documentation.
-3.
-Design a workflow as a sequence of small steps glued together with intermediate outputs from one step feeding into the next.
-
-*Packaging data analytical work reproducibly using {R} (and friends) -- Marwick et al*
-
-• The R package can be a suitable template for organizing files into a research compendium to enhance the reproducibility of research.
-
-*A guide to reproducible code in ecology and evolution -- Cooper et al*
-
-• Reproducible work should include a clear and well-separated file system, well-named files, a list of package dependencies, good documentation.
-Preferably, RMarkdown and version control should be used.
-
-*Reproducible research and Biostatistics -- Peng*
-
-• To meet the "reproducible" criterion in the biostatistics journal you must submit: a main script and any complementary side scripts, data and any other necessary files, a target file containing results.
-
-# Other articles
-
-*Karl Broman's suggestions:*
-
-• Keep the project contained in one directory, separate the components (raw data, derived data, code), use relative paths, write a README, and name things clearly.
-
-*R OpenSci's suggestions:*
-
-• To be reproducible, it should be easy to find files, be clear what each file is, and be easy to know what order to run files.
-There should be no extraneous files.
-Everything should be documented, dependencies noted, and seeds used for randomness.
-Automation should be used when possible.
-
-## Packages
-
-rrtools
-• Complementary to Ben Marwick's paper
-• Creates a basic R package structure for your research and institutes some testthat check functionality
-
-orderly
-• Based on the step-by-step analysis process. 
-• Build components that take inputs from other components
-• Have to declare project structure at the beginning.
-
-workflowr
-• Focused a lot on version control and making code available
-
-checkers
-• Allows you to create custom checks to look over your work
-
-*Notes: Almost all of these are things you need to do before you start your project.
-They each focus on a small subset of reproducibility components.
-They are somewhat complicated to use -- commands are complex and there are often a lot of them.

### 2020-02-11 fertile.Rmd
use case #2

@@
 
-### 2. Experience R User
+### 2. Experienced R User
+
+Emma is a post-doc, with several years of R experience.
+She is familiar with some basic rules of reproducibility--file paths should always be relative and randomness should always be associated with a seed--but has never needed to pass any sort of reproducibility check before because her professors never emphasized that.
+
+She has just finished a research project and is looking to submit her work to a journal.
+When researching the journal she is interested in submitting to, she discovers that it has high standards for research reproducibility and a dedicated editor focusing on that aspect of submissions.
+She goes online and finds the journal's guidelines for reproducibility.
+They are more complete than any guidelines she has previously been required to conform to.
+In addition to notes about file paths and randomness, the journal requires a clean, well-organized folder structure, broken down by file category and cleaned of files that do not serve a purpose.
+
+Unsure of the best way to achieve this structure, Emma goes online to find help.
+In her search, she comes across `fertile`.
+She downloads the package, and in only a handful of commands, she identifies and removes excess files in her directory and automatically organizes her files into a structure reminiscent of an R package.
+She now meets the guidelines for the journal and can submit her research.
 

### 2020-02-16 fertile.Rmd
literature review additions

@@
 
-## Academic Research
+As is becoming increasingly clear, reproducibility is critically important to the advancement of knowledge in all fields of scientific research.
+Researchers appear to be acknowledging this, as publications focusing on reproducibility seem to have increased in frequency over the last several years.
+
+Much of the available literature is focused on the methods for achieving reproducibility in specific disciplines, though much of this information is generalizable to all areas of scientific research.
+There is also a small body of work focusing on reproducibility within the R community specifically.
+Some of this work is in the form of academic papers, while other literature comes from the blog posts and websites of several highly-regarded R-focused websites and researchers.
+In this section, we will discuss the reproducibility recommendations from both the formal and informal publications, which together influenced which features were included in `fertile`.
+
+There is also a small existing body of R packages focused on providing tools to assist researchers with achieving reproducibility.
+We will consider the functionality of these packages, identifying `fertile`'s unique role and functionality.
+
+## Published Literature
+
+As Goodman, Fanelli, and Ionnidis argue in their 2016 paper, *What does research reproducibilty mean?*, the language and conceptual framework of research reproducibility vary across the sciences.
+There are no clear standards agreed upon across fields.
+
+The variety of recommendations and techniques for achieving reproducibility can be clearly seen Kitzes, Turek, and Deniz' book *The Practice of Reproducible Research*, a collection of case studies on reproducibility practices from across the data-intensive sciences.
+Although the book idenfies no consensus on exact standards of reproducibility, several common trends and principles emerge from the case studies:
+
+1. Use clear separation, labeling and documentation.
+2. Automate processes when possible.
+3. Design the data analysis workflow as a sequence of small steps glued together, with outputs from one step serving as inputs into the next.
+
+*A guide to reproducible code in ecology and evolution*, focused on data analysis in R, identifies similar important reproducibility components, reiterating the need to clearly labled, well-documented, and well-separated files.
+In addition, the paper recommends that a list of dependencies be published and version control be used.
+
+Karl Broman, Professor of Biostatistics and Medical Informatics at the University of Wisconsin, Madison, reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a README.
+
+R OpenSci's reproducibility recommendations share similar components to those discussed previously.
+They focus on a need for a well-developed file system, with no extraneous files and clear labeling.
+They also reiterate the need to note dependencies and use automation when possible, while making clear a suggestion not present in the previously-discussed literature: the need to use seeds when running code involving randomness.
+
+When considered in combination, these sources provide a well-rounded picture of the components important to research reproducibility.
+Using this literature as a guideline, we identify several key features of reproducible work to focus on:
+
+• A well-designed file structure:
+  - Separate folders for different file types.
+  - No extraneous files.
+  - Minimal clutter.
+
+• Good documentation:
+  - Files are clearly named, preferably in a way where the order in which they should be run is clear.
+  - A README is present.
+  - Dependencies are noted.
+
+• Reproducible file paths:
+  - No absolute paths, or paths leading to locations outside of a project's directory, are used in code. Only portable (relative) paths.
+
+• Randomness is accounted for:
+  - If randomness is used in code, a seed must also be set.
+
+• Code conforms to tidyverse style
+
+*Packaging data analytical work reproducibly using {R} (and friends) -- Marwick et al*
+
+• The R package can be a suitable template for organizing files into a research compendium to enhance the reproducibility of research.
+
+*Reproducible research and Biostatistics -- Peng*
+
+• To meet the "reproducible" criterion in the biostatistics journal you must submit: a main script and any complementary side scripts, data and any other necessary files, a target file containing results.
+
+*R OpenSci's suggestions:*
+
+• To be reproducible, it should be easy to find files, be clear what each file is, and be easy to know what order to run files.
+There should be no extraneous files.
+Everything should be documented, dependencies noted, and seeds used for randomness.
+Automation should be used when possible.
+
+*A guide to reproducible code in ecology and evolution -- Cooper et al*
+
+• Reproducible work should include a clear and well-separated file system, well-named files, a list of package dependencies, good documentation.
+Preferably, RMarkdown and version control should be used.
+
+*Karl Broman's suggestions:*
+
+• Keep the project contained in one directory, separate the components (raw data, derived data, code), use relative paths, write a README, and name things clearly.
 
@@
 Design a workflow as a sequence of small steps glued together with intermediate outputs from one step feeding into the next.
-
-*Packaging data analytical work reproducibly using {R} (and friends) -- Marwick et al*
-
-• The R package can be a suitable template for organizing files into a research compendium to enhance the reproducibility of research.
-
-*A guide to reproducible code in ecology and evolution -- Cooper et al*
-
-• Reproducible work should include a clear and well-separated file system, well-named files, a list of package dependencies, good documentation.
-Preferably, RMarkdown and version control should be used.
-
-*Reproducible research and Biostatistics -- Peng*
-
-• To meet the "reproducible" criterion in the biostatistics journal you must submit: a main script and any complementary side scripts, data and any other necessary files, a target file containing results.
-
-## Other Articles
-
-*Karl Broman's suggestions:*
-
-• Keep the project contained in one directory, separate the components (raw data, derived data, code), use relative paths, write a README, and name things clearly.
-
-*R OpenSci's suggestions:*
-
-• To be reproducible, it should be easy to find files, be clear what each file is, and be easy to know what order to run files.
-There should be no extraneous files.
-Everything should be documented, dependencies noted, and seeds used for randomness.
-Automation should be used when possible.
 
@@
 
+As we have seen, many existing R packages focusing on reproducibility are designed in a way such that they can only be used prior to the start of a project, often for the purposes of creating an organized file structure or ordered file hierarchy.
+`fertile`, on the other hand, can be used at any stage in the data analysis workflow.
+It is designed to be used on data analyses organized as R Projects (i.e.
+folders containing a `.Rproj` file).
+Once an R Project is created, `fertile` can provide benefits at any time, both while coding as well as after the fact.
 `fertile` operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that has already been written for potential problems).
@@
 ```{r setup, cache = F, include = FALSE}
-knitr::opts_chunk$set(error = TRUE)
+knitr::opts_chunk$set(error = TRUE, collapse = TRUE)
 ```
 
-```{r, include = FALSE}
+```{r, include = FALSE, message = FALSE}
 readr::write_csv(mtcars, "~/Desktop/my_data.csv")
 library(fertile)
+library(tidyverse)
 ```
@@
 read.csv("~/Desktop/my_data.csv")
+```
+
+```{r}
+read.table("~/Desktop/my_data.csv")
 ```
@@
 is_data_file("~/Desktop/my_data.csv")
+is_image_file("~/Desktop/my_data.csv")
+```
+
+```{r}
+dir <- getwd()
+new_dir <- sandbox(dir)
+```
+
+```{r}
+dir
+fs::dir_ls(dir) %>% head()
+```
+
+```{r}
+new_dir
+fs::dir_ls(new_dir) %>% head()
 ```

### 2020-02-19 fertile.Rmd
small lit review addition

@@
 
-*Packaging data analytical work reproducibly using {R} (and friends) -- Marwick et al*
-
-• The R package can be a suitable template for organizing files into a research compendium to enhance the reproducibility of research.
-
-*Reproducible research and Biostatistics -- Peng*
-
-• To meet the "reproducible" criterion in the biostatistics journal you must submit: a main script and any complementary side scripts, data and any other necessary files, a target file containing results.
-
-*R OpenSci's suggestions:*
-
-• To be reproducible, it should be easy to find files, be clear what each file is, and be easy to know what order to run files.
-There should be no extraneous files.
-Everything should be documented, dependencies noted, and seeds used for randomness.
-Automation should be used when possible.
-
-*A guide to reproducible code in ecology and evolution -- Cooper et al*
-
-• Reproducible work should include a clear and well-separated file system, well-named files, a list of package dependencies, good documentation.
-Preferably, RMarkdown and version control should be used.
-
-*Karl Broman's suggestions:*
-
-• Keep the project contained in one directory, separate the components (raw data, derived data, code), use relative paths, write a README, and name things clearly.
-
-*What does research reproducibility mean? -- Goodman et al*
-
-•The language and conceptual framework of research reproducibility vary across the sciences. There are no clear standards agreed upon by all scientific researchers.
-
-*The practice of reproducible research: case studies and lessons from the data-intensive sciences -- Kitzes et al*
-
-• Reproducibility examples from a variety of fields all vary significantly but they share key principles / practices in common: 1.
-Use clear separation, labeling, and documentation.
-2.
-Automate documentation.
-3.
-Design a workflow as a sequence of small steps glued together with intermediate outputs from one step feeding into the next.
+Much of the available literature focuses on file structure, organization, and naming, and `fertile`'s features are consistent with this.
+The *ideal* file structure is not agreed upon by academics, though there are some recommendations focused on R that are available.
+*Packaging data analytical work reproducibility using {R} (and friends)* by Marwick, Boettiger, and Mullen provides the framework for file structure that `fertile` is based on: a structure similar to that of an R package, with an `R` folder, as well as `data`, `data-raw`, `inst` and `vignettes`.
 
@@
 https://books.google.com/books?hl=en&lr=&id=JcmSAwAAQBAJ&oi=fnd&pg=PP1&dq=Victoria+stodden+reproducibility&ots=ym_gvSxMNH&sig=B5N6FlkoocVjIprIZevtvrL33qQ#v=onepage&q=coombes&f=false
+
+*Packaging data analytical work reproducibly using {R} (and friends) -- Marwick et al*
+
+• The R package can be a suitable template for organizing files into a research compendium to enhance the reproducibility of research.
+
+*Reproducible research and Biostatistics -- Peng*
+
+• To meet the "reproducible" criterion in the biostatistics journal you must submit: a main script and any complementary side scripts, data and any other necessary files, a target file containing results.
+
+*R OpenSci's suggestions:*
+
+• To be reproducible, it should be easy to find files, be clear what each file is, and be easy to know what order to run files.
+There should be no extraneous files.
+Everything should be documented, dependencies noted, and seeds used for randomness.
+Automation should be used when possible.
+
+*A guide to reproducible code in ecology and evolution -- Cooper et al*
+
+• Reproducible work should include a clear and well-separated file system, well-named files, a list of package dependencies, good documentation.
+Preferably, RMarkdown and version control should be used.
+
+*Karl Broman's suggestions:*
+
+• Keep the project contained in one directory, separate the components (raw data, derived data, code), use relative paths, write a README, and name things clearly.
+
+*What does research reproducibility mean? -- Goodman et al*
+
+•The language and conceptual framework of research reproducibility vary across the sciences. There are no clear standards agreed upon by all scientific researchers.
+
+*The practice of reproducible research: case studies and lessons from the data-intensive sciences -- Kitzes et al*
+
+• Reproducibility examples from a variety of fields all vary significantly but they share key principles / practices in common: 1.
+Use clear separation, labeling, and documentation.
+2.
+Automate documentation.
+3.
+Design a workflow as a sequence of small steps glued together with intermediate outputs from one step feeding into the next.

### 2020-02-19 fertile.Rmd
added citations

--- 2020-02-16 fertile.Rmd
@@
 
-bibliography: bibliography.bib
+bibliography: [bibliography.bib, pkgs.bib]
 output: rticles::asa_article
 ---
+
+```{r, include=FALSE}
+knitr::write_bib(c("dplyr", "drake"), file = "pkgs.bib")
+```
 
@@
 
-As Goodman, Fanelli, and Ionnidis argue in their 2016 paper, *What does research reproducibilty mean?*, the language and conceptual framework of research reproducibility vary across the sciences.
-There are no clear standards agreed upon across fields.
+As @goodman2016does argue, the language and conceptual framework of research reproducibility vary across the sciences. There are no clear standards agreed upon across fields.
 
@@
 
-*Packaging data analytical work reproducibly using {R} (and friends) -- Marwick et al*
-
-• The R package can be a suitable template for organizing files into a research compendium to enhance the reproducibility of research.
-
-*Reproducible research and Biostatistics -- Peng*
-
-• To meet the "reproducible" criterion in the biostatistics journal you must submit: a main script and any complementary side scripts, data and any other necessary files, a target file containing results.
-
-*R OpenSci's suggestions:*
-
-• To be reproducible, it should be easy to find files, be clear what each file is, and be easy to know what order to run files.
-There should be no extraneous files.
-Everything should be documented, dependencies noted, and seeds used for randomness.
-Automation should be used when possible.
-
-*A guide to reproducible code in ecology and evolution -- Cooper et al*
-
-• Reproducible work should include a clear and well-separated file system, well-named files, a list of package dependencies, good documentation.
-Preferably, RMarkdown and version control should be used.
-
-*Karl Broman's suggestions:*
-
-• Keep the project contained in one directory, separate the components (raw data, derived data, code), use relative paths, write a README, and name things clearly.
-
-*What does research reproducibility mean? -- Goodman et al*
-
-•The language and conceptual framework of research reproducibility vary across the sciences. There are no clear standards agreed upon by all scientific researchers.
-
-*The practice of reproducible research: case studies and lessons from the data-intensive sciences -- Kitzes et al*
-
-• Reproducibility examples from a variety of fields all vary significantly but they share key principles / practices in common: 1.
-Use clear separation, labeling, and documentation.
-2.
-Automate documentation.
-3.
-Design a workflow as a sequence of small steps glued together with intermediate outputs from one step feeding into the next.
+Much of the available literature focuses on file structure, organization, and naming, and `fertile`'s features are consistent with this.
+The *ideal* file structure is not agreed upon by academics, though there are some recommendations focused on R that are available.
+@marwick2018packaging provides the framework for file structure that `fertile` is based on: a structure similar to that of an R package, with an `R` folder, as well as `data`, `data-raw`, `inst` and `vignettes`.
 
 ## R Packages
+
+See also the **drake** package [@R-drake].
 

### 2020-02-22 fertile.Rmd
reformatting package names

--- 2020-02-16 fertile.Rmd
@@
 There appear to be no existing tools that cater to analysts looking for an easy-to-learn, easy-to-implement, fast way to obtain a broad overview of their projects' reproducibility.
-To address this, we present an R package called [`fertile`](https://github.com/baumer-lab/fertile) [^1]:{The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the `tidyverse` team.}, a low barrier-to-entry package which focuses on a series of common mistakes programmers make while conducting data science research in R.
+To address this, we present an R package called [**fertile**](https://github.com/baumer-lab/fertile) [^1]:{The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the `tidyverse` team.}, a low barrier-to-entry package which focuses on a series of common mistakes programmers make while conducting data science research in R.
 
@@
 Some of this work is in the form of academic papers, while other literature comes from the blog posts and websites of several highly-regarded R-focused websites and researchers.
-In this section, we will discuss the reproducibility recommendations from both the formal and informal publications, which together influenced which features were included in `fertile`.
+In this section, we will discuss the reproducibility recommendations from both the formal and informal publications, which together influenced which features were included in **fertile**.
 
 There is also a small existing body of R packages focused on providing tools to assist researchers with achieving reproducibility.
-We will consider the functionality of these packages, identifying `fertile`'s unique role and functionality.
+We will consider the functionality of these packages, identifying **fertile**'s unique role and functionality.
 
@@
 
-*Packaging data analytical work reproducibly using {R} (and friends) -- Marwick et al*
-
-• The R package can be a suitable template for organizing files into a research compendium to enhance the reproducibility of research.
-
-*Reproducible research and Biostatistics -- Peng*
-
-• To meet the "reproducible" criterion in the biostatistics journal you must submit: a main script and any complementary side scripts, data and any other necessary files, a target file containing results.
-
-*R OpenSci's suggestions:*
-
-• To be reproducible, it should be easy to find files, be clear what each file is, and be easy to know what order to run files.
-There should be no extraneous files.
-Everything should be documented, dependencies noted, and seeds used for randomness.
-Automation should be used when possible.
-
-*A guide to reproducible code in ecology and evolution -- Cooper et al*
-
-• Reproducible work should include a clear and well-separated file system, well-named files, a list of package dependencies, good documentation.
-Preferably, RMarkdown and version control should be used.
-
-*Karl Broman's suggestions:*
-
-• Keep the project contained in one directory, separate the components (raw data, derived data, code), use relative paths, write a README, and name things clearly.
-
-*What does research reproducibility mean? -- Goodman et al*
-
-•The language and conceptual framework of research reproducibility vary across the sciences. There are no clear standards agreed upon by all scientific researchers.
-
-*The practice of reproducible research: case studies and lessons from the data-intensive sciences -- Kitzes et al*
-
-• Reproducibility examples from a variety of fields all vary significantly but they share key principles / practices in common: 1.
-Use clear separation, labeling, and documentation.
-2.
-Automate documentation.
-3.
-Design a workflow as a sequence of small steps glued together with intermediate outputs from one step feeding into the next.
+Much of the available literature focuses on file structure, organization, and naming, and **fertile`**s features are consistent with this.
+The *ideal* file structure is not agreed upon by academics, though there are some recommendations focused on R that are available.
+*Packaging data analytical work reproducibility using {R} (and friends)* by Marwick, Boettiger, and Mullen provides the framework for file structure that **fertile** is based on: a structure similar to that of an R package, with an `R` folder, as well as `data`, `data-raw`, `inst` and `vignettes`.
 
@@
 As we have seen, many existing R packages focusing on reproducibility are designed in a way such that they can only be used prior to the start of a project, often for the purposes of creating an organized file structure or ordered file hierarchy.
-`fertile`, on the other hand, can be used at any stage in the data analysis workflow.
+**fertile**, on the other hand, can be used at any stage in the data analysis workflow.
 It is designed to be used on data analyses organized as R Projects (i.e.
 folders containing a `.Rproj` file).
-Once an R Project is created, `fertile` can provide benefits at any time, both while coding as well as after the fact.
-`fertile` operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that has already been written for potential problems).
+Once an R Project is created, **fertile** can provide benefits at any time, both while coding as well as after the fact.
+**fertile** operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that has already been written for potential problems).
 `fertile` is available for download at [https://github.com/baumer-lab/fertile](https://github.com/baumer-lab/fertile)
@@
 Proactively, the package identifies potential mistakes as they are made by the user and outputs an informative message as well as a recommended solution.
-For example, `fertile` catches when a user passes a potentially problematic file path---such as an absolute path, or a path that points to a location outside of the project directory---to a variety of common input/output functions.
+For example, **fertile** catches when a user passes a potentially problematic file path---such as an absolute path, or a path that points to a location outside of the project directory---to a variety of common input/output functions.
 
@@
 readr::write_csv(mtcars, "~/Desktop/my_data.csv")
+library(tidyverse)
 library(fertile)
-library(tidyverse)
 ```
@@
 file.exists("~/Desktop/my_data.csv")
-read.csv("~/Desktop/my_data.csv")
+read_csv("~/Desktop/my_data.csv")
 ```
@@
 
-`fertile` is even more aggressive with functions (like `setwd()`) that are almost certain to break reproducibility---it causes them to throw errors.
+**fertile** is even more aggressive with functions (like `setwd()`) that are almost certain to break reproducibility---it causes them to throw errors.
 
@@
 
-The proactive features are activated immediately after loading the `fertile` package and require no additional effort by the user.
+The proactive features are activated immediately after loading the **fertile** package and require no additional effort by the user.
 
-In addition to the interactive warning system, `fertile` provides several useful utility functions.
+In addition to the interactive warning system, **fertile** provides several useful utility functions.
 Among other things, these include functions to check the type of a file and a way to create a copy of a project in a temporary directory.
@@
 
-Retroactively, `fertile` analyzes potential obstacles to reproducibility in an RStudio Project (i.e., a directory that contains an `.Rproj` file), including the directory structure, the analyst's use of file paths, randomness, etc.
-`fertile}` creates reproducibility reports that identify potential mistakes and provide recommendations for remedies.
-For example, `fertile` might identify the use of randomness in code and recommend setting a seed.
+Retroactively, **fertile** analyzes potential obstacles to reproducibility in an RStudio Project (i.e., a directory that contains an `.Rproj` file), including the directory structure, the analyst's use of file paths, randomness, etc.
+**fertile** creates reproducibility reports that identify potential mistakes and provide recommendations for remedies.
+For example, **fertile** might identify the use of randomness in code and recommend setting a seed.
 
-Users can access the majority of `fertile`'s retroactive features through two primary functions.
+Users can access the majority of **fertile**'s retroactive features through two primary functions.
 
@@
 These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that to not have a call to `set.seed()` present.
-Subsets of the fifteen tests can be invoked using the select helper functions from `dplyr` and the `proj_check_some()` function.
+Subsets of the fifteen tests can be invoked using the select helper functions from **dplyr** and the `proj_check_some()` function.
 
@@
 
-`fertile` is designed in such a way as to be useful to users of all backgrounds and experiences.
+**fertile** is designed in such a way as to be useful to users of all backgrounds and experiences.
 In this section, we provide two sample use cases demonstrating situations where users could benefit significantly from using the package.
@@
 Susan doesn't understand why it didn't work on the professor's computer when it did on her own.
-The professor recommends that she install `fertile` and run `proj_check()` on her assignment.
+The professor recommends that she install **fertile** and run `proj_check()` on her assignment.
 She does this and gets a message informing her she used an absolute path to open her dataset, and she should use a relative path.
@@
 Unsure of the best way to achieve this structure, Emma goes online to find help.
-In her search, she comes across `fertile`.
+In her search, she comes across **fertile**.
 She downloads the package, and in only a handful of commands, she identifies and removes excess files in her directory and automatically organizes her files into a structure reminiscent of an R package.
@@
 In an effort to understand the package's effectiveness, we also share preliminary results from a randomized, controlled experiment conducted on an undergraduate introductory data science course [^2]:{This study is approved by Smith College IRB, Protocol #19-032}.
-The purpose of the study is to determine whether `fertile` helps students produce data science research that is more likely to be reproducible.
+The purpose of the study is to determine whether **fertile** helps students produce data science research that is more likely to be reproducible.
 
@@
 
-`fertile` is an R package that lowers barriers to reproducible data analysis projects in R. 
-The features of `fertile` can be accessed almost effortlessly, making it easy for data analysts of all skill levels and backgrounds to gain a better understanding of how to make their work reproducible.
-`fertile` is designed to educate the user about why the mistakes are problematic and how to fix them, promoting a greater understanding of reproducibility concepts in its users.
+**fertile** is an R package that lowers barriers to reproducible data analysis projects in R. 
+The features of **fertile** can be accessed almost effortlessly, making it easy for data analysts of all skill levels and backgrounds to gain a better understanding of how to make their work reproducible.
+**fertile** is designed to educate the user about why the mistakes are problematic and how to fix them, promoting a greater understanding of reproducibility concepts in its users.
 It is written for a wide audience, simple enough to be used by students in an introductory data science course, but still helpful to experienced analysts.
-`fertile` also addresses a human challenge of reproducibility.
+**fertile** also addresses a human challenge of reproducibility.
 In the moment, it often feels easiest to take a shortcut---to use an absolute path or change a working directory.
 However, when considering the long term path of a project, spending the extra time to improve reproducibility is worthwhile.
-`fertile`'s user-friendly features can help data analysts avoid these harmful shortcuts with minimal effort.
+**fertile**'s user-friendly features can help data analysts avoid these harmful shortcuts with minimal effort.
 

### 2020-02-22 fertile.Rmd
updated citations + removed notes from document

@@
 
-As Goodman, Fanelli, and Ionnidis argue in their 2016 paper, *What does research reproducibilty mean?*, the language and conceptual framework of research reproducibility vary across the sciences.
-There are no clear standards agreed upon across fields.
+As argued by @Goodman341ps12, the language and conceptual framework of research reproducibility vary across the sciences. There are no clear standards agreed upon across fields.
 
-The variety of recommendations and techniques for achieving reproducibility can be clearly seen Kitzes, Turek, and Deniz' book *The Practice of Reproducible Research*, a collection of case studies on reproducibility practices from across the data-intensive sciences.
+The variety of recommendations and techniques for achieving reproducibility can be clearly seen in @kitzes2017practice, a collection of case studies on reproducibility practices from across the data-intensive sciences.
 Although the book idenfies no consensus on exact standards of reproducibility, several common trends and principles emerge from the case studies:
@@
 
-*A guide to reproducible code in ecology and evolution*, focused on data analysis in R, identifies similar important reproducibility components, reiterating the need to clearly labled, well-documented, and well-separated files.
+@cooper2017guide, focused on data analysis in R, identifies similar important reproducibility components, reiterating the need to clearly labled, well-documented, and well-separated files.
 In addition, the paper recommends that a list of dependencies be published and version control be used.
@@
 The *ideal* file structure is not agreed upon by academics, though there are some recommendations focused on R that are available.
-*Packaging data analytical work reproducibility using {R} (and friends)* by Marwick, Boettiger, and Mullen provides the framework for file structure that **fertile** is based on: a structure similar to that of an R package, with an `R` folder, as well as `data`, `data-raw`, `inst` and `vignettes`.
+@marwick2018packaging provides the framework for file structure that **fertile** is based on: a structure similar to that of an R package, with an `R` folder, as well as `data`, `data-raw`, `inst` and `vignettes`.
 
@@
 **fertile** operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that has already been written for potential problems).
-`fertile` is available for download at [https://github.com/baumer-lab/fertile](https://github.com/baumer-lab/fertile)
+**fertile** is available for download at [https://github.com/baumer-lab/fertile](https://github.com/baumer-lab/fertile)
 
@@
 **fertile**'s user-friendly features can help data analysts avoid these harmful shortcuts with minimal effort.
-
-# Introduction
-
-# The Importance of Ensuring Reproducibility
-
-One reason for this is that reproducible research encourages transparency.
-Providing one's audience with well organized, reproducible code and results allows readers to understand the steps taken to generate findings from raw data and determine for themselves whether or not they believe the results to be reliable or trustworthy.
-
-Additionally, reproducibility encourages collaboration and extended research, allows others to easily apply the methods used in one project to their own work with minimal effort. @bray2014five
-
-Fertile is designed to make reproducibility simple, providing fast and easy methods to test an R project for reproducibility.
-The package is intended to be usable by introductory data science students in their first semester of R.
-
---> make sure this info is included somewhere:
-
-From "Five Concrete Reasons Your Students Should Be Learning to Analyze Data in the Reproducible Paradigm" (@bray2014five)
-
-* Helps with transparency: make clear data cleaning steps between raw data and final data
-* Helps with collaboration: easier to share code when it takes very few steps to run on a different computer
-
-The other big ideas on the importance/use of this package (not from sources):
-
-* Make it easier for professors to grade students' code
-* Should be usable by intro level data science students
-* Possible use by reviewers of journal articles (and by those writing the articles)
-
-# What Defines Reproducibility in Data Science?
-
-# Creating Comprehensive Reproducibility Reports With "fertile"
-
-Main points of different sources, as well as info about how they might be used for the paper.
-
-Here are some sources we might use for motivation behind the project:
-
-### The Reproducibility Crisis
-
-Big idea: most scientific fields are facing a reproducibility crisis and poor statistical use is considered one of the important reasons behind this.
-
-Not sure how useful this is due to the fact that it does not necessarily focus on the same kind of reproducibility we are looking at, which is code reproducibility rather than experimental reproducibility.
-
-@baker20161
-
-### Why is reproducibility important?
-
-From Popper in `The Logic of Scientific Discovery`: "non-reproducible single occurrences are of no significance to science."
-
-@popper2005logic
-
-### What does reproducibility mean in data science?
-
-"The ability of a researcher to duplicate the results of a prior study using the same materials as were used by the original investigator.
-That is, a second researcher might use the same raw data to build the same analysis files and implement the same statistical analysis in an attempt to yield the same results."
-
-@goodman2016does
-
-From "The Practice of Reproducibile Research": Can all of the figures/calculations related to the result be reproduced in a single button press or at least with a reasonably small effort?
-
-@kitzes2017practice
-
-### What makes a data science project reproducibile?
-
-Another way to think about this is: what features does fertile have that match what different sources think needs to be checked to ensure reproducibility?
-
-Some big ideas from the OpenSci discussion page:
-
-* code should use tidy style
-* project should have short vignette files -- most written code should be in an R/ directory
-* code and data should be stored in separate folders
-* a run environment and dependencies should be specified
-* there should be a README file
-* there should be a data dictionary
-* one master script
-* code should only use relative paths
-* code should be well commented and all variables should be defined
-
-Big ideas from "Packaging data analytical work reproducibly using R"
-
-* research should be organized like an R package!
-* clear separation of data, method, and output
-* specify the computational environment that was used for the original analysis (typically in a plain text file)
-* there should be a README describing the project and where to get started
-* script files with reusable functions should go in an R/ directory
-* raw data files should be kept in a data/ directory
-* analysis scripts and report files should go in an analysis/ directory
-* for simpe projects, scripts should be given ascending names. For more complicated ones, the use of some sort of makefile is recommended
-* there should also be a DESCRIPTION file with information about the authors, project license, and software dependencies
-
-@marwick2018packaging
-
-From "The Practice of Reproducible Research"
-
-* Are the data openly accessible? If hosted online, is the web address reliable long-term?
-* Are they in a commonly used and well-documented file format? Avoid spreadsheets and instead use plain text data if possible!
-* Is the raw data available? Is sufficient metadata provided?
-
-* Are dependencies described properly?
-* Is full history of source code available through a public version history
-
-* Is there a README?
-* Are functions documented?
-* Is there narrative documentation explaining how the different pieces work together?
-* Are there usage examples?
-
-Folder setup might look like:
-
-1. Raw Data
-    + Data
-    + README
-2. Clean Data
-    + Data
-3. Results
-    + Results file
-4. Src
-    + Analysis script
-    + Script to clean data
-
-@kitzes2017practice
-
-From R OpenSci's Reproducibility Guide:
-
-http://ropensci.github.io/reproducibility-guide/
-
-* Is it clear where to begin?
-* Can you determinine which files were used as input to create output files?
-* Is there documentation about every result?
-* Are exact versions of external applications noted?
-* If using randomness, are seeds noted?
-* Have you specified a license or noted licenses if you used other people's content?
-* Are files easy to find?
-* Is it clear what the most recent file is?
-* Are there any folders that could be deleted?
-* Is analysis output done hierarchically?
-* Are there lots of manual data manipulation steps?
-
-R OpenSci also has a discussion of code review practices where users commented. Not too much that was helpful from here (just a few odd comments):
-https://discuss.ropensci.org/t/how-do-you-review-code-that-accompanies-a-research-project-or-paper-help-ropensci-plan-a-community-call/1321/43
-
-From "A Guide to Reproducible Code in Ecology and Evolution" (these ideas are pretty universal, though):
-
-A basic project structure:
-
-* The data folder contains all input data (and metadata) used in the analysis.
-* The doc folder contains the manuscript
-* The figs directory contains figures generated by the analysis
-* The output folder contains any type of intermediate or output files (e.g. simulation outputs, models, processed datasets, etc.). You might separate this and also have a cleaned-data folder.
-* The R directory contains R scripts with function definitions.
-* The reports folder contains RMarkdown files that document the analysis or report on results
-
-* Consistent, ordered naming of scripts
-* Use portable paths
-* Write unit tests (only for advanced coding)
-* Show the packages you used
-* Record dependencies and versions of outside things you use
-
-@cooper2017guide
-
-# Karl Broman's Suggestions
-
-* Encapsulate everything within one directory
-* Separate raw data from derived data
-* Separate data from code
-* Use relative paths
-* Choose filenames correctly
-* Write README files
-
-http://kbroman.org/steps2rr/pages/organize.html
-
-# Why focus on R?
-
-* R is the most popular language for statistical programming and is specifically designed for statistics
-* R is great for reprodubility because the code is readable by users, and RMarkdown is a great way to show/explain processes
-* R is easy to install and begin using
-
---> from https://openresearchsoftware.metajnl.com/articles/10.5334/jors.bu/print/
-
-# Similar R Packages
-
-### rrtools
-
-https://github.com/benmarwick/rrtools
-
-* Creates a basic R package named after your research topic
-* Generates a license file
-* Connects to GitHub and creates a repository
-* Generates a README
-* Generates a reproducible directory structure
-* Creates a dockerfile
-* Creates a minimal travis file
-* Sets up testthat
-
-### represtools
-
-https://pirategrunt.com/represtools/
-
-Has basic utility functions to support reproducibile research
-• Gather
-• Cook
-• Analyze
-• Present
-
-### orderly
-
-https://github.com/vimc/orderly
-
-Focuses on automating a series of basic steps in the prodcess of writing analyses.
-• Tracks inputs
-• Stores multiple versions of an analysis (unclear whether this is different stages?)
-• Track outputs
-• Create analyses that depend on the outputs of previous analyses
-• You have to declare what you're going to use and create BEFORE you start the analysis
-
-### workflowr
-
-https://f1000research.com/articles/8-1749
-
-Aims to instill a particular workflow that helps make projects more reproducible.
-
-• Version control (git)
-• Literate programming (Rmd)
-• Automatic checks & safeguards (seemingly related to whether different files in a project have been published on git, and not actually to do with code practices)
-• Sharing code / results via browsable website
-
-### checkers
-
-https://github.com/ropenscilabs/checkers
-
-• A framework for reviewing analysis projects
-• Seems to be something run after-the-fact
-• Currently a work in progress and not passing checks (as of Jan 2019)
-• Can create custom checks
-• Checks include:
- - Comments
- - Good coding style (lintr)
- - Use of version control
- - Whether preferred packages are used
-
-# Msc sources to look at:
-
-### Victoria Stodden's "Implementing Reproducible Research" book
-https://books.google.com/books?hl=en&lr=&id=JcmSAwAAQBAJ&oi=fnd&pg=PP1&dq=Victoria+stodden+reproducibility&ots=ym_gvSxMNH&sig=B5N6FlkoocVjIprIZevtvrL33qQ#v=onepage&q=coombes&f=false
-
-*Packaging data analytical work reproducibly using {R} (and friends) -- Marwick et al*
-
-• The R package can be a suitable template for organizing files into a research compendium to enhance the reproducibility of research.
-
-*Reproducible research and Biostatistics -- Peng*
-
-• To meet the "reproducible" criterion in the biostatistics journal you must submit: a main script and any complementary side scripts, data and any other necessary files, a target file containing results.
-
-*R OpenSci's suggestions:*
-
-• To be reproducible, it should be easy to find files, be clear what each file is, and be easy to know what order to run files.
-There should be no extraneous files.
-Everything should be documented, dependencies noted, and seeds used for randomness.
-Automation should be used when possible.
-
-*A guide to reproducible code in ecology and evolution -- Cooper et al*
-
-• Reproducible work should include a clear and well-separated file system, well-named files, a list of package dependencies, good documentation.
-Preferably, RMarkdown and version control should be used.
-
-*Karl Broman's suggestions:*
-
-• Keep the project contained in one directory, separate the components (raw data, derived data, code), use relative paths, write a README, and name things clearly.
-
-*What does research reproducibility mean? -- Goodman et al*
-
-•The language and conceptual framework of research reproducibility vary across the sciences. There are no clear standards agreed upon by all scientific researchers.
-
-*The practice of reproducible research: case studies and lessons from the data-intensive sciences -- Kitzes et al*
-
-• Reproducibility examples from a variety of fields all vary significantly but they share key principles / practices in common: 1.
-Use clear separation, labeling, and documentation.
-2.
-Automate documentation.
-3.
-Design a workflow as a sequence of small steps glued together with intermediate outputs from one step feeding into the next.

### 2020-02-23 fertile.Rmd
more bibliography updates

@@
 
-bibliography: bibliography.bib
+bibliography: [bibliography.bib, pkgs.bib]
 output: rticles::asa_article
 ---
+
+```{r, include = FALSE}
+knitr::write_bib(c("dplyr", "rrtools", "tidyverse", "orderly", "workflowr", "testthat"),  file = "pkgs.bib")
+```
 
@@
 There appear to be no existing tools that cater to analysts looking for an easy-to-learn, easy-to-implement, fast way to obtain a broad overview of their projects' reproducibility.
-To address this, we present an R package called [**fertile**](https://github.com/baumer-lab/fertile) [^1]:{The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the `tidyverse` team.}, a low barrier-to-entry package which focuses on a series of common mistakes programmers make while conducting data science research in R.
+To address this, we present an R package called [**fertile**](https://github.com/baumer-lab/fertile) [^1].
+
+[^1]:{The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the **tidyverse** team.}, a low barrier-to-entry package which focuses on a series of common mistakes programmers make while conducting data science research in R.
 
@@
 
-Much of the available literature focuses on file structure, organization, and naming, and **fertile`**s features are consistent with this.
-The *ideal* file structure is not agreed upon by academics, though there are some recommendations focused on R that are available.
-@marwick2018packaging provides the framework for file structure that **fertile** is based on: a structure similar to that of an R package, with an `R` folder, as well as `data`, `data-raw`, `inst` and `vignettes`.
+Much of the available literature focuses on file structure, organization, and 
+naming, and **fertile**s features are consistent with this. The *ideal* file 
+structure is not agreed upon by academics, though there are some recommendations 
+focused on R that are available. @marwick2018packaging provides the framework for 
+file structure that **fertile** is based on: a structure similar to that of an R 
+package, with an `R` folder, as well as `data`, `data-raw`, `inst` and `vignettes`.
 
@@
 
-rrtools
-• Complementary to Ben Marwick's paper
-• Creates a basic R package structure for your research and institutes some testthat check functionality
+There also exists a small selection of R packages that work to address the issue of research reproducibility, though not all of them are available on `CRAN`.
+**rrtools** [@R-rrtools], developed to address some of the issues discussed in @marwick2018packaging, creates a basic R package structure for a data analysis project and implements a basic **testthat** `check` functionality.
 
@@
 
-```{r, eval = FALSE, include = FALSE}
-knitr::write_bib("dplyr", file = "pkgs.bib")
-```
-
 ```{r, message = FALSE}

### 2020-02-23 fertile.Rmd
more citations + lit review of packages

@@
 output: rticles::asa_article
+
 ---
@@
 There also exists a small selection of R packages that work to address the issue of research reproducibility, though not all of them are available on `CRAN`.
-**rrtools** [@R-rrtools], developed to address some of the issues discussed in @marwick2018packaging, creates a basic R package structure for a data analysis project and implements a basic **testthat** `check` functionality.
 
-orderly
-• Based on the step-by-step analysis process. 
-• Build components that take inputs from other components
-• Have to declare project structure at the beginning.
+**rrtools** [@R-rrtools], developed to address some of the issues discussed in @marwick2018packaging, creates a basic R package structure for a data analysis project and implements a basic **testthat**::`check` functionality.
+The **orderly** [@R-orderly] package also focuses on file structure, requiring the user to declare a desired project structure (typically a step-by-step structure, where outputs from one step are inputs into the next) at the beginning and then creating the files necessary to achieve that structure.
+**workflowr**'s [@R-workflowr] functionality is based around version control and making code easily available online.
+It works to generate a website containing time-stamped, versioned, and documented results.
+**checkers** [@R-checkers] allows you to create custom checks that look at different aspects of reproducibility.
+**packrat** [@R-packrat] is focused on dependencies, creating a packaged folder containing a project as well as all of its dependencies, so that projects dependent on lesser-used packages can be easily shared across computers.
+Lastly, the **reproducible** [@R-reproducible] package focuses on the concept of caching--saving information so that projects can be run faster each time they are re-completed from the start.
 
-workflowr
-• Focused a lot on version control and making code available
+Many of these packages focus on one specific area of reproducibility, and while they each succeed well at their intended goal, they are not necessarily the most practical option for users trying to address a wide variety of factors influencing reproducibility at once.
 
-checkers
-• Allows you to create custom checks to look over your work
-
-*Notes: Almost all of these are things you need to do before you start your project.
-They each focus on a small subset of reproducibility components.
-They are somewhat complicated to use -- commands are complex and there are often a lot of them.
+Additionally, with the focused nature of these packages comes added complexity.
+The functions in these packages are often quite complex to use, and many steps must be completed to achieve the required reproducibility goal.
+This complex nature means that most reproducibility packages currently available are not easily accessible to users near the beginning of their R journey, nor particularly useful to those looking for quick and easy reproducibility checks.
 
@@
 
+\tiny
 ```{r}
@@
 ```
+\normalsize
 

### 2020-02-27 fertile.Rmd
edits to methods section

@@
 
-As we have seen, many existing R packages focusing on reproducibility are designed in a way such that they can only be used prior to the start of a project, often for the purposes of creating an organized file structure or ordered file hierarchy.
-**fertile**, on the other hand, can be used at any stage in the data analysis workflow.
+**fertile** looks to address these gaps, providing a simple, easy-to-learn reproducibility package that, rather than focusing intensely on a specific area, provides some information about a wide variety of aspects influencing reproducibility.
+
+The package also provides flexibility--offering benefits to users at any stage in the data analysis workflow, unlike some other available packages which can only be used before the creation of a new project or after it is finished.
+
 It is designed to be used on data analyses organized as R Projects (i.e.
 folders containing a `.Rproj` file).
-Once an R Project is created, **fertile** can provide benefits at any time, both while coding as well as after the fact.
-**fertile** operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that has already been written for potential problems).
-**fertile** is available for download at [https://github.com/baumer-lab/fertile](https://github.com/baumer-lab/fertile)
+Once an R Project is created, **fertile** provides benefits throughout the data anlaysis process, both while coding as well as after the fact.
+
+**fertile** achieves this by operating in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that has already been written for potential problems).
 
@@
 Proactively, the package identifies potential mistakes as they are made by the user and outputs an informative message as well as a recommended solution.
-For example, **fertile** catches when a user passes a potentially problematic file path---such as an absolute path, or a path that points to a location outside of the project directory---to a variety of common input/output functions.
+For example, **fertile** catches when a user passes a potentially problematic file path---such as an absolute path, or a path that points to a location outside of the project directory---to a variety of common input/output functions operating on many different file types.
 
@@
 
-**fertile** is even more aggressive with functions (like `setwd()`) that are almost certain to break reproducibility---it causes them to throw errors.
+**fertile** is even more aggressive with functions (like `setwd()`) that are almost certain to break reproducibility, causing them to throw errors that prevent their execution.
 
@@
 
-The proactive features are activated immediately after loading the **fertile** package and require no additional effort by the user.
+The proactive features are activated immediately after loading the **fertile** package and require no additional effort by the user to access.
 
-In addition to the interactive warning system, **fertile** provides several useful utility functions.
-Among other things, these include functions to check the type of a file and a way to create a copy of a project in a temporary directory.
+## Retroactive Use
+
+Retroactively, **fertile** analyzes potential obstacles to reproducibility in an RStudio Project (i.e., a directory that contains an `.Rproj` file), including the directory structure, the analyst's use of file paths, randomness, etc.
+**fertile** creates reproducibility reports that identify potential mistakes and provide recommendations for remedies.
+For example, **fertile** might identify the use of randomness in code and recommend setting a seed.
+
+Users can access the majority of **fertile**'s retroactive features through two primary functions.
+
+The `proj_check()` function runs fifteen different reproducibility tests, noting which ones passed, which ones failed, the reason for failure, a recommended solution, and a guide to where to look for help.
+These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that to not have a call to `set.seed()` present.
+Subsets of the fifteen tests can be invoked using the select helper functions from **dplyr** and the `proj_check_some()` function.
+
+```{r, message = FALSE}
+library(tidyverse)
+proj_check_some(".", contains("paths"))
+```
+
+The `proj_analyze()` function creates a report documenting the structure of a data analysis project.
+This report contains informations about all packages referenced in code, the files present in the directory and their types, suggestions for moving files to create a more organized structure, and a list of reproducibility-breaking file paths used in code.
+
+\tiny
+```{r, message = FALSE}
+proj_analyze(".")
+```
+\normalsize
+
+## Utility Functions
+
+**fertile** also provides several useful utility functions. Among other things, these include functions to check the type of a file and a way to create a copy of a project in a temporary directory.
 
@@
 \normalsize
-
-## Retroactive Use
-
-Retroactively, **fertile** analyzes potential obstacles to reproducibility in an RStudio Project (i.e., a directory that contains an `.Rproj` file), including the directory structure, the analyst's use of file paths, randomness, etc.
-**fertile** creates reproducibility reports that identify potential mistakes and provide recommendations for remedies.
-For example, **fertile** might identify the use of randomness in code and recommend setting a seed.
-
-Users can access the majority of **fertile**'s retroactive features through two primary functions.
-
-The `proj_check()` function runs fifteen different reproducibility tests, noting which ones passed, which ones failed, the reason for failure, a recommended solution, and a guide to where to look for help.
-These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that to not have a call to `set.seed()` present.
-Subsets of the fifteen tests can be invoked using the select helper functions from **dplyr** and the `proj_check_some()` function.
-
-```{r, message = FALSE}
-library(tidyverse)
-proj_check_some(".", contains("paths"))
-```
-
-The `proj_analyze()` function creates a report documenting the structure of a data analysis project.
-This report contains informations about all packages referenced in code, the files present in the directory and their types, suggestions for moving files to create a more organized structure, and a list of reproducibility-breaking file paths used in code.
 

### 2020-02-27 fertile.Rmd
added mention of logging

@@
 
+**fertile** also contains a logging functionality, which records commands with the potential to affect reproducibility that are run in the console.
+The package focuses mostly on package loading and file opening, notes which function was used, the path or package it referenced, and the timestamp at which that event happened.
+
+Users can access the log recording their commands at any time via the `log_report()` function:
+
+```{r, include = FALSE}
+log_clear()
+```
+
+```{r, message = FALSE, results = "hide"}
+library(purrr)
+library(forcats)
+library(fertile)
+read.csv("mtcars.csv")
+```
+
+\tiny
+```{r}
+log_report()
+```
+\normalsize
+
+The `log_clear()` function allows for management of the log, enabling users to erase it and start over at any time.
+
+```{r, message = FALSE}
+log_clear()
+log_report()
+```
+
 ## Utility Functions

### 2020-02-27 fertile.Rmd
additions to methods and conclusion

--- 2020-02-27 fertile.Rmd
@@
 
-**fertile** is even more aggressive with functions (like `setwd()`) that are almost certain to break reproducibility, causing them to throw errors that prevent their execution.
+**fertile** is even more aggressive with functions (like `setwd()`) that are almost certain to break reproducibility, causing them to throw errors that prevent their execution and providing recommendations for better alternatives.
 
@@
 
-The proactive features are activated immediately after loading the **fertile** package and require no additional effort by the user to access.
+These proactive warning features are activated immediately after loading the **fertile** package and require no additional effort by the user to access.
 
@@
 
-Retroactively, **fertile** analyzes potential obstacles to reproducibility in an RStudio Project (i.e., a directory that contains an `.Rproj` file), including the directory structure, the analyst's use of file paths, randomness, etc.
-**fertile** creates reproducibility reports that identify potential mistakes and provide recommendations for remedies.
-For example, **fertile** might identify the use of randomness in code and recommend setting a seed.
+Retroactively, **fertile** analyzes potential obstacles to reproducibility in an RStudio Project (i.e., a directory that contains an `.Rproj` file).
+The package considers several different aspects of the project which may influence reproducibility, including the directory structure, the analyst's use of file paths, and whether randomness is used thoughtfully.
 
-Users can access the majority of **fertile**'s retroactive features through two primary functions.
+The end products of these analyses are reproducibility reports summarizing a project's adherence to reproducibility standards and recommending remedies for where the project falls short.
+For example, **fertile** might identify the use of randomness in code and recommend setting a seed if one is not present.
+
+Users can access the majority of **fertile**'s retroactive features through two primary functions, `proj_check()` and `proj_analyze()`.
 
@@
 These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that to not have a call to `set.seed()` present.
-Subsets of the fifteen tests can be invoked using the select helper functions from **dplyr** and the `proj_check_some()` function.
+A full list is provided below:
+
+```{r}
+list_checks()
+```
+
+Subsets of the fifteen tests can be invoked using the select helper functions from **dplyr** in combination with the `proj_check_some()` function.
 
@@
 
+**fertile** also contains a logging functionality, which records commands with the potential to affect reproducibility that are run in the console, enabling users to look at their past history at any time.
+The package focuses mostly on package loading and file opening, noting which function was used, the path or package it referenced, and the timestamp at which that event happened.
+
+Users can access the log recording their commands at any time via the `log_report()` function:
+
+```{r, include = FALSE}
+log_clear()
+```
+
+```{r, message = FALSE, results = "hide"}
+library(purrr)
+library(forcats)
+library(fertile)
+read.csv("mtcars.csv")
+```
+
+\tiny
+```{r}
+log_report()
+```
+\normalsize
+
+The log, if not managed, can grow very long over time. For users who do not desire such functionality, the `log_clear()` provides a way to erase the log and start it over.
+
+```{r, message = FALSE}
+log_clear()
+log_report()
+```
+
+In addition to the interactive log, users can access something called the render log, which contains much of the information used to create the output for `proj_analyze()` using the function `render_log_report()`.
+
 ## Utility Functions
 
-**fertile** also provides several useful utility functions. Among other things, these include functions to check the type of a file and a way to create a copy of a project in a temporary directory.
+**fertile** also provides several useful utility functions that may assist with the process of data analysis.
+Among other things, these include functions to check the type of a file and a way to create a copy of a project in a temporary directory.
 
@@
 
+# Results
+
 # Conclusion
 
-**fertile** is an R package that lowers barriers to reproducible data analysis projects in R. 
-The features of **fertile** can be accessed almost effortlessly, making it easy for data analysts of all skill levels and backgrounds to gain a better understanding of how to make their work reproducible.
+**fertile** is an R package that lowers barriers to reproducible data analysis projects in R, providing a wide array of checks and suggestions addressing many different aspects of project reproducibility.
+
+Unlike with many other R packages focused on reproducibility that are currently available, the features of **fertile** can be accessed almost effortlessly, making it easy for data analysts of all skill levels and backgrounds to gain a better understanding of how to make their work reproducible.
+Many of **fertile**'s features can be accessed in only two lines of code and using minimal arguments, and some of the proactive features can be accessed purely upon loading the package.
+As a result, it is beneficial to a wide audience--simple enough to be used by students in an introductory data science course, but still helpful to experienced analysts.
+
 **fertile** is designed to educate the user about why the mistakes are problematic and how to fix them, promoting a greater understanding of reproducibility concepts in its users.
-It is written for a wide audience, simple enough to be used by students in an introductory data science course, but still helpful to experienced analysts.
+
 **fertile** also addresses a human challenge of reproducibility.

### 2020-02-29 fertile.Rmd
edits to conclusion

@@
 
-Unlike with many other R packages focused on reproducibility that are currently available, the features of **fertile** can be accessed almost effortlessly, making it easy for data analysts of all skill levels and backgrounds to gain a better understanding of how to make their work reproducible.
-Many of **fertile**'s features can be accessed in only two lines of code and using minimal arguments, and some of the proactive features can be accessed purely upon loading the package.
-As a result, it is beneficial to a wide audience--simple enough to be used by students in an introductory data science course, but still helpful to experienced analysts.
+**fertile** is meant to be educational, providing informative error messages showing how the ways in which the mistakes the user is making are problematic, as well as sharing recommendations on how to fix them.
+The package is designed in this way so as to promote a greater understanding of reproducibility concepts in its users, with the goal of increasing the overall awareness and understanding of reproducibility in the R community.
 
-**fertile** is designed to educate the user about why the mistakes are problematic and how to fix them, promoting a greater understanding of reproducibility concepts in its users.
+The package has very low barriers to entry, making it accessible to users with various levels of background knowledge.
+Unlike with many other R packages focused on reproducibility that are currently available, the features of **fertile** can be accessed almost effortlessly.
+Many of the retroactive features can be accessed in only two lines of code requiring  minimal arguments, and some of the proactive features can be accessed with no additional effort beyond loading the package.
+This, in combination with the fact that **fertile** does not focus on one specific area of reproducibility, instead covering (in less detail) a wide variety of topics, means that **fertile** makes it easy for data analysts of all skill levels to quickly gain a better understanding of the reproducibility of the work.
 
-**fertile** also addresses a human challenge of reproducibility.
 In the moment, it often feels easiest to take a shortcut---to use an absolute path or change a working directory.

### 2020-03-12 fertile.Rmd
draft edits

@@
 
-```{r, include = FALSE}
-knitr::write_bib(c("dplyr", "rrtools", "tidyverse", "orderly", "workflowr", "testthat"),  file = "pkgs.bib")
-```
-
 # Introduction
-
-Data-based research cannot be fully *reproducible* unless the requisite code and data files produce identical results when run by another analyst.
 
@@
 
+Data-based research cannot be fully *reproducible* unless the requisite code and data files produce identical results when run by another analyst.
 When researchers provide the code and data used for their work in a well-organized and reproducible format, readers are more easily able to determine the veracity of any findings by following the steps from raw data to conclusions.
@@
 While they provide useful guidelines, they can often be difficult to implement, particularly to new analysts who are unsure how to apply such recommendations within the software programs they are using.
-
 Thus, reproducibility recommendations tailored to specific software programs are more likely to be adopted.
@@
 In this paper, we focus on reproducibility in the R programming language with a concentration on projects that use the RStudio integrated development environment.
+R is an ideal candidate for reproducibility recommendations due to the language's popularity for statistical analyses and the ease with which analysts can download and begin using the software.
 
-R is an ideal candidate for reproducibility recommendations due to the language's popularity for statistical analyses and the ease with which analysts can download and begin using the software.  
-Several researchers and R users have recognized this and worked to publish a small body of papers and R packages focusing on increasing reproducibility in the R community.
+A small body of papers and R packages focusing on reproducibility in the R community have been published.
 
-Much of this work is narrowly tailored, with each package effectively addressising a small component of reproducibility--file structure, modularization of code, version control, etc. 
-Many existing reproducibility packages, due to their focused nature, succeed at their area of focus well, but at a cost.
-They are often semi-complex to learn and operate, providing a barrier to entry for less-experienced data analysts.
+Much of this work is narrowly tailored, with each package effectively addressising a small component of reproducibility---file structure, modularization of code, version control, etc.
+Many existing reproducibility packages, due to their focused nature, succeed at their area of focus, but at a cost.
+They are often difficult to learn and operate, providing a barrier to entry for less-experienced data analysts.
 
-There appear to be no existing tools that cater to analysts looking for an easy-to-learn, easy-to-implement, fast way to obtain a broad overview of their projects' reproducibility.
+There appear to be no existing tools that cater to analysts looking for an easy-to-learn, easy-to-implement, fast, way to obtain a broad overview of their projects' reproducibility.
 To address this, we present an R package called [**fertile**](https://github.com/baumer-lab/fertile) [^1].
 
-[^1]:{The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the **tidyverse** team.}, a low barrier-to-entry package which focuses on a series of common mistakes programmers make while conducting data science research in R.
+[^1]:The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the **tidyverse** team., a low barrier-to-entry package which focuses on a series of common mistakes programmers make while conducting data science research in R.
 
@@
 
-As is becoming increasingly clear, reproducibility is critically important to the advancement of knowledge in all fields of scientific research.
-Researchers appear to be acknowledging this, as publications focusing on reproducibility seem to have increased in frequency over the last several years.
+Reproducibility is critically important to the advancement of knowledge in all fields of scientific research.
+Researchers are acknowledging this; publications focusing on reproducibility seem to have increased in frequency over the last several years [**CITATIONS HERE**]
 
 Much of the available literature is focused on the methods for achieving reproducibility in specific disciplines, though much of this information is generalizable to all areas of scientific research.
-There is also a small body of work focusing on reproducibility within the R community specifically.
-Some of this work is in the form of academic papers, while other literature comes from the blog posts and websites of several highly-regarded R-focused websites and researchers.
-In this section, we will discuss the reproducibility recommendations from both the formal and informal publications, which together influenced which features were included in **fertile**.
+There is also work focusing specifically on reproducibility within the R community.
+Some of this work is in the form of academic papers, while other literature comes from blog posts, websites, and R packages, some of which are compiled by notable R-focused developers and researchers.
+In this section, we will discuss the reproducibility recommendations from both the formal and informal publications, which together influence the development of **fertile**.
 
-There is also a small existing body of R packages focused on providing tools to assist researchers with achieving reproducibility.
-We will consider the functionality of these packages, identifying **fertile**'s unique role and functionality.
+## Previous Work
 
-## Published Literature
+@Goodman341ps12 argues that the language and conceptual framework of research reproducibility varies across the sciences. There are no clear standards agreed upon across fields.
 
-As argued by @Goodman341ps12, the language and conceptual framework of research reproducibility vary across the sciences. There are no clear standards agreed upon across fields.
-
-The variety of recommendations and techniques for achieving reproducibility can be clearly seen in @kitzes2017practice, a collection of case studies on reproducibility practices from across the data-intensive sciences.
-Although the book idenfies no consensus on exact standards of reproducibility, several common trends and principles emerge from the case studies:
+@kitzes2017practice presents a collection of case studies on reproducibility practices from across the data-intensive sciences, illustrating a variety of recommendations and techniques for achieving reproducibility.
+Although the book does not come to a consensus on the exact standards of reproducibility that should be followed, several common trends and principles emerge from the case studies:
 
@@
 2. Automate processes when possible.
-3. Design the data analysis workflow as a sequence of small steps glued together, with outputs from one step serving as inputs into the next.
+3. Design the data analysis workflow as a sequence of small steps glued together, with outputs from one step serving as inputs into the next. [**MENTION THE UNIX PHILOSOPHY**]
 
-@cooper2017guide, focused on data analysis in R, identifies similar important reproducibility components, reiterating the need to clearly labled, well-documented, and well-separated files.
-In addition, the paper recommends that a list of dependencies be published and version control be used.
+@cooper2017guide focuses on data analysis in R and identifies similar important reproducibility components, while reiterating the need for clearly labeled, well-documented, and well-separated files.
+In addition, they recommend publishing a list of dependencies and using version control.
 
-Karl Broman, Professor of Biostatistics and Medical Informatics at the University of Wisconsin, Madison, reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a README.
+Broman reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a README.
+[**ADD CITATION FOR BROMAN**]
 
@@
 • A well-designed file structure:
+
   - Separate folders for different file types.
+
   - No extraneous files.
+
   - Minimal clutter.
@@
 • Good documentation:
+
   - Files are clearly named, preferably in a way where the order in which they should be run is clear.
+
   - A README is present.
+
   - Dependencies are noted.
@@
 • Reproducible file paths:
+
   - No absolute paths, or paths leading to locations outside of a project's directory, are used in code. Only portable (relative) paths.
@@
 • Randomness is accounted for:
+
   - If randomness is used in code, a seed must also be set.
 
-• Code conforms to tidyverse style
+[** NOTE ABOUT TIDYVERSE STYLE AND WHY WE ADD THAT**]
 

### 2020-03-15 fertile.Rmd
more draft edits

@@
 There appear to be no existing tools that cater to analysts looking for an easy-to-learn, easy-to-implement, fast, way to obtain a broad overview of their projects' reproducibility.
-To address this, we present an R package called [**fertile**](https://github.com/baumer-lab/fertile) [^1].
+To address this, we present an R package called [**fertile**](https://github.com/baumer-lab/fertile) [^1], a low barrier-to-entry package which focuses on a series of common mistakes programmers make while conducting data science research in R.
 
-[^1]:The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the **tidyverse** team., a low barrier-to-entry package which focuses on a series of common mistakes programmers make while conducting data science research in R.
+[^1]:The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the **tidyverse** team.
 
@@
 
-1. Use clear separation, labeling and documentation.
+1. Use clear separation, labeling, and documentation.
 2. Automate processes when possible.
@@
 
-R OpenSci's reproducibility recommendations share similar components to those discussed previously.
+R OpenSci's [**EXPLAIN WHO THEY ARE**] reproducibility recommendations share similar principles to those discussed previously.
 They focus on a need for a well-developed file system, with no extraneous files and clear labeling.
-They also reiterate the need to note dependencies and use automation when possible, while making clear a suggestion not present in the previously-discussed literature: the need to use seeds when running code involving randomness.
+They also reiterate the need to note dependencies and use automation when possible, while making clear a suggestion not present in the previously-discussed literature: the need to use seeds, which allow  for the saving and restoring of the random number generator state, when running code involving randomness.
 
@@
 
-Much of the available literature focuses on file structure, organization, and 
-naming, and **fertile**s features are consistent with this. The *ideal* file 
-structure is not agreed upon by academics, though there are some recommendations 
-focused on R that are available. @marwick2018packaging provides the framework for 
-file structure that **fertile** is based on: a structure similar to that of an R 
-package, with an `R` folder, as well as `data`, `data-raw`, `inst` and `vignettes`.
+Much of the available literature focuses on file structure, organization, 
+and naming, and **fertile**'s features are consistent with this. The 
+*ideal* file structure is not agreed upon by academics, though there are 
+some recommendations focused on R that are available.@marwick2018packaging
+provides the framework for file structure that **fertile** is based on: a 
+structure similar to that of an R package [**Cite Hadley's Writing R 
+Packages and the R Core Team's Writing R Extensions**], with an `R` 
+folder, as well as `data`, `data-raw`, `inst` and `vignettes`.
 
@@
 
-There also exists a small selection of R packages that work to address the issue of research reproducibility, though not all of them are available on `CRAN`.
+A small selection of R packages  work to address the issue of research reproducibility, although not all of them are available on `CRAN`.
 
-**rrtools** [@R-rrtools], developed to address some of the issues discussed in @marwick2018packaging, creates a basic R package structure for a data analysis project and implements a basic **testthat**::`check` functionality.
+**rrtools** [@R-rrtools] addresses some of the issues discussed in @marwick2018packaging by creating a basic R package structure for a data analysis project and implementing a basic **testthat**::`check` functionality.
 The **orderly** [@R-orderly] package also focuses on file structure, requiring the user to declare a desired project structure (typically a step-by-step structure, where outputs from one step are inputs into the next) at the beginning and then creating the files necessary to achieve that structure.
@@
 It works to generate a website containing time-stamped, versioned, and documented results.
-**checkers** [@R-checkers] allows you to create custom checks that look at different aspects of reproducibility.
+**checkers** [@R-checkers] allows you to create custom checks that examine different aspects of reproducibility.
 **packrat** [@R-packrat] is focused on dependencies, creating a packaged folder containing a project as well as all of its dependencies, so that projects dependent on lesser-used packages can be easily shared across computers.
@@
 
+[**Clarify why fertile comes in handy**]
 Additionally, with the focused nature of these packages comes added complexity.
 The functions in these packages are often quite complex to use, and many steps must be completed to achieve the required reproducibility goal.
-This complex nature means that most reproducibility packages currently available are not easily accessible to users near the beginning of their R journey, nor particularly useful to those looking for quick and easy reproducibility checks.
+This cumbersome nature means that most reproducibility packages currently available are not easily accessible to users near the beginning of their R journey, nor particularly useful to those looking for quick and easy reproducibility checks.
+[**Note that our recommendations are opinionated**]
 
@@
 
-**fertile** looks to address these gaps, providing a simple, easy-to-learn reproducibility package that, rather than focusing intensely on a specific area, provides some information about a wide variety of aspects influencing reproducibility.
+**fertile** addresses these gaps by providing a simple, easy-to-learn reproducibility package that, rather than focusing intensely on a specific area, provides some information about a wide variety of aspects influencing reproducibility.
 
@@
 
-It is designed to be used on data analyses organized as R Projects (i.e.
-folders containing a `.Rproj` file).
-Once an R Project is created, **fertile** provides benefits throughout the data anlaysis process, both while coding as well as after the fact.
+**fertile** is designed to be used on data analyses organized as R Projects (i.e.
+directories containing an `.Rproj` file).
+Once an R Project is created, **fertile** provides benefits throughout the data analysis process, both during development as well as after the fact.
 
@@
 library(tidyverse)
-library(fertile)
+
 ```
@@
 file.exists("~/Desktop/my_data.csv")
+library(fertile)
 read_csv("~/Desktop/my_data.csv")
@@
 
-These proactive warning features are activated immediately after loading the **fertile** package and require no additional effort by the user to access.
+These proactive warning features are activated immediately after attaching the **fertile** package and require no additional effort by the user.
 
@@
 Retroactively, **fertile** analyzes potential obstacles to reproducibility in an RStudio Project (i.e., a directory that contains an `.Rproj` file).
-The package considers several different aspects of the project which may influence reproducibility, including the directory structure, the analyst's use of file paths, and whether randomness is used thoughtfully.
+The package considers several different aspects of the project which may influence reproducibility, including the directory structure, file paths, and whether randomness is used thoughtfully.
 
@@
 
-Subsets of the fifteen tests can be invoked using the select helper functions from **dplyr** in combination with the `proj_check_some()` function.
+Subsets of the fifteen tests can be invoked using the **tidyselect** helper functions in combination with the more limited `proj_check_some()` function.
 
 ```{r, message = FALSE}
-library(tidyverse)
+library(tidyselect)
 proj_check_some(".", contains("paths"))
@@
 The `proj_analyze()` function creates a report documenting the structure of a data analysis project.
-This report contains informations about all packages referenced in code, the files present in the directory and their types, suggestions for moving files to create a more organized structure, and a list of reproducibility-breaking file paths used in code.
+This report contains information about all packages referenced in code, the files present in the directory and their types, suggestions for moving files to create a more organized structure, and a list of reproducibility-breaking file paths used in code.
 
@@
 
-**fertile** also contains a logging functionality, which records commands with the potential to affect reproducibility that are run in the console, enabling users to look at their past history at any time.
+**fertile** also contains  logging functionality, which records commands run in the console that have the potential to affect reproducibility, enabling users to look at their past history at any time.
 The package focuses mostly on package loading and file opening, noting which function was used, the path or package it referenced, and the timestamp at which that event happened.
@@
 
-The log, if not managed, can grow very long over time. For users who do not desire such functionality, the `log_clear()` provides a way to erase the log and start it over.
+The log, if not managed, can grow very long over time. For users who do not desire such functionality, the `log_clear()` provides a way to erase the log and start  over.
 
@@
 
-In addition to the interactive log, users can access something called the render log, which contains much of the information used to create the output for `proj_analyze()` using the function `render_log_report()`.
+In addition to the interactive log, users can access a render log, [**define this**] which contains much of the information used to create the output for `proj_analyze()` using the function `render_log_report()`.
 
@@
 **fertile** is designed in such a way as to be useful to users of all backgrounds and experiences.
-In this section, we provide two sample use cases demonstrating situations where users could benefit significantly from using the package.
+In this section, we provide two sample use cases demonstrating situations where users could benefit from using the package.
 
@@
 
-She reads in the data, makes the graph, and knits her .Rmd file.
+She reads in the data, makes the graph, and knits her `.Rmd` file.
 It compiles successfully, so she submits the assignment.
@@
 
-### 2. Experienced R User
+### Experienced R User
 
 Emma is a post-doc, with several years of R experience.
-She is familiar with some basic rules of reproducibility--file paths should always be relative and randomness should always be associated with a seed--but has never needed to pass any sort of reproducibility check before because her professors never emphasized that.
+She is familiar with some basic rules of reproducibility---file paths should always be relative and randomness should always be associated with a seed--but has never needed to pass any sort of reproducibility check before because her professors never emphasized that.
 
 She has just finished a research project and is looking to submit her work to a journal.
-When researching the journal she is interested in submitting to, she discovers that it has high standards for research reproducibility and a dedicated editor focusing on that aspect of submissions.
+When researching the journal to which she is interested in submitting, she discovers that it has high standards for research reproducibility and a dedicated editor focusing on that aspect of submission.
 She goes online and finds the journal's guidelines for reproducibility.
-They are more complete than any guidelines she has previously been required to conform to.
-In addition to notes about file paths and randomness, the journal requires a clean, well-organized folder structure, broken down by file category and cleaned of files that do not serve a purpose.
+They are more complete than any guidelines to which she has previously been required to conform.
+In addition to notes about file paths and randomness, the journal requires a clean, well-organized folder structure, broken down by file category and stripped of files that do not serve a purpose.
+[**add something about build chain and readme to requirements**]
 
@@
 
-In an effort to understand the package's effectiveness, we also share preliminary results from a randomized, controlled experiment conducted on an undergraduate introductory data science course [^2]:{This study is approved by Smith College IRB, Protocol #19-032}.
+In an effort to understand the package's effectiveness, we also share preliminary results from a randomized, controlled experiment conducted on an undergraduate introductory data science course [^2].
 The purpose of the study is to determine whether **fertile** helps students produce data science research that is more likely to be reproducible.
+
+[^2]: This study is approved by Smith College IRB, Protocol #19-032.
 
@@
 
-**fertile** is meant to be educational, providing informative error messages showing how the ways in which the mistakes the user is making are problematic, as well as sharing recommendations on how to fix them.
+**fertile** is meant to be educational, providing informative error messages that show how the ways in which the mistakes the user is making are problematic, as well as sharing recommendations on how to fix them.
 The package is designed in this way so as to promote a greater understanding of reproducibility concepts in its users, with the goal of increasing the overall awareness and understanding of reproducibility in the R community.
@@
 The package has very low barriers to entry, making it accessible to users with various levels of background knowledge.
-Unlike with many other R packages focused on reproducibility that are currently available, the features of **fertile** can be accessed almost effortlessly.
+Unlike many other R packages focused on reproducibility that are currently available, the features of **fertile** can be accessed almost effortlessly.
 Many of the retroactive features can be accessed in only two lines of code requiring  minimal arguments, and some of the proactive features can be accessed with no additional effort beyond loading the package.
-This, in combination with the fact that **fertile** does not focus on one specific area of reproducibility, instead covering (in less detail) a wide variety of topics, means that **fertile** makes it easy for data analysts of all skill levels to quickly gain a better understanding of the reproducibility of the work.
+This, in combination with the fact that **fertile** does not focus on one specific area of reproducibility, instead covering (albeit in less detail) a wide variety of topics, means that **fertile** makes it easy for data analysts of all skill levels to quickly gain a better understanding of the reproducibility of the work.
 

### 2020-03-15 fertile.Rmd
adding citations

@@
 
-Reproducibility is critically important to the advancement of knowledge in all fields of scientific research.
-Researchers are acknowledging this; publications focusing on reproducibility seem to have increased in frequency over the last several years [**CITATIONS HERE**]
+Reproducibility is critically important to the advancement of knowledge in
+all fields of scientific research. Researchers are acknowledging this; 
+publications and discussions focusing on reproducibility seem to have 
+increased in frequency over the last several years (@sep-scientific-reproducibility), (@plos-biology), (@bioessays-gosselin)
 
-Much of the available literature is focused on the methods for achieving reproducibility in specific disciplines, though much of this information is generalizable to all areas of scientific research.
-There is also work focusing specifically on reproducibility within the R community.
-Some of this work is in the form of academic papers, while other literature comes from blog posts, websites, and R packages, some of which are compiled by notable R-focused developers and researchers.
-In this section, we will discuss the reproducibility recommendations from both the formal and informal publications, which together influence the development of **fertile**.
+Much of the available literature is focused on the methods for achieving 
+reproducibility in specific disciplines, though much of this information 
+is generalizable to all areas of scientific research. There is also work 
+focusing specifically on reproducibility within the R community. Some of 
+this work is in the form of academic papers, while other literature comes 
+from blog posts, websites, and R packages, some of which are compiled by 
+notable R-focused developers and researchers.  In this section, we will 
+discuss the reproducibility recommendations from both the formal and 
+informal publications, which together influence the development of 
+**fertile**.
 
@@
 
-Broman reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a README.
-[**ADD CITATION FOR BROMAN**]
+Broman reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a README (@broman).
 
-R OpenSci's [**EXPLAIN WHO THEY ARE**] reproducibility recommendations share similar principles to those discussed previously.
+The reproducibility recommendations from R OpenSci, a non-profit initiative founded in 2011 to make scientific data retrieval reproducible, share similar principles to those discussed previously.
 They focus on a need for a well-developed file system, with no extraneous files and clear labeling.
-They also reiterate the need to note dependencies and use automation when possible, while making clear a suggestion not present in the previously-discussed literature: the need to use seeds, which allow  for the saving and restoring of the random number generator state, when running code involving randomness.
+They also reiterate the need to note dependencies and use automation when possible, while making clear a suggestion not present in the previously-discussed literature: the need to use seeds, which allow  for the saving and restoring of the random number generator state, when running code involving randomness @r-opensci.
 
@@
   - Separate folders for different file types.
-
   - No extraneous files.
-
   - Minimal clutter.
@@
   - Files are clearly named, preferably in a way where the order in which they should be run is clear.
-
   - A README is present.
-
   - Dependencies are noted.

### 2020-03-15 fertile.Rmd
changed examples to miceps

@@
 2. Automate processes when possible.
-3. Design the data analysis workflow as a sequence of small steps glued together, with outputs from one step serving as inputs into the next. [**MENTION THE UNIX PHILOSOPHY**]
+3.
+Design the data analysis workflow as a sequence of small steps glued together, with outputs from one step serving as inputs into the next.
+This is a common suggestion within the computing community, originating as part of the Unix philosophy (@unix).
 
@@
 
-• A well-designed file structure:
+1. A well-designed file structure:
 
@@
 
-• Good documentation:
+2. Good documentation:
 
@@
 
-• Reproducible file paths:
+3. Reproducible file paths:
 
@@
 
-• Randomness is accounted for:
+4. Randomness is accounted for:
 
@@
 
-[** NOTE ABOUT TIDYVERSE STYLE AND WHY WE ADD THAT**]
+5. Code conforms to tidyverse style:
+
+  - Although not discussed in any of the literature mentioned previously, we also believe **tidyverse** integration to be very important.
+Code conformation to **tidyverse** style helps ensure that analyses meeting **fertile**'s recommendations will be able to integrate seamlessly with packages from the **tidyverse**, an ever-expanding collection of packages designed to operate together with one another.
 
@@
 
-```{r, include = FALSE, message = FALSE}
-readr::write_csv(mtcars, "~/Desktop/my_data.csv")
-library(tidyverse)
-
-```
-
 ```{r, message = FALSE}
-file.exists("~/Desktop/my_data.csv")
 library(fertile)
-read_csv("~/Desktop/my_data.csv")
-```
-
-```{r}
-read.table("~/Desktop/my_data.csv")
+file.exists("project_miceps/mice.csv")
+abs_path <- fs::path_abs("project_miceps/mice.csv")
+read_csv(abs_path)
 ```
@@
 library(tidyselect)
-proj_check_some(".", contains("paths"))
+proj_check_some("project_miceps", contains("paths"))
 ```
@@
 ```{r, message = FALSE}
-proj_analyze(".")
+proj_analyze("project_miceps")
 ```
@@
 library(fertile)
-read.csv("mtcars.csv")
+read.csv("project_miceps/mice.csv")
 ```
@@
 
-is_data_file("~/Desktop/my_data.csv")
-is_image_file("~/Desktop/my_data.csv")
+is_data_file("project_miceps/mice.csv")
+is_image_file("project_miceps/mice.csv")
+is_image_file("project_miceps/proteins_v_time.png")
 ```
@@
 In addition to notes about file paths and randomness, the journal requires a clean, well-organized folder structure, broken down by file category and stripped of files that do not serve a purpose.
-[**add something about build chain and readme to requirements**]
+In order to be approved, submissions must also have a clear build chain and an informative README file.
 

### 2020-03-16 fertile.Rmd
more edits

@@
 When considered in combination, these sources provide a well-rounded picture of the components important to research reproducibility.
-Using this literature as a guideline, we identify several key features of reproducible work to focus on:
+Using this literature as a guideline, we identify several key features of reproducible work to focus on.
+These recommendations are opinionated---due to the lack of agreement on which components of reproducibility are most important, we select those that are mentioned most often, as well as some that are mentioned less but that we view as important.
 
@@
 *ideal* file structure is not agreed upon by academics, though there are 
-some recommendations focused on R that are available.@marwick2018packaging
+some recommendations focused on R that are available. @marwick2018packaging
 provides the framework for file structure that **fertile** is based on: a 
@@
 
-[**Clarify why fertile comes in handy**]
 Additionally, with the focused nature of these packages comes added complexity.
@@
 This cumbersome nature means that most reproducibility packages currently available are not easily accessible to users near the beginning of their R journey, nor particularly useful to those looking for quick and easy reproducibility checks.
-[**Note that our recommendations are opinionated**]
 
@@
 
-In addition to the interactive log, users can access a render log, [**define this**] which contains much of the information used to create the output for `proj_analyze()` using the function `render_log_report()`.
+In addition to the interactive log, users can access a render log, which is created by running all of the `.R` and `.Rmd` files within a project in order to collect information on packages, file paths, and file structure used to create the output for `proj_analyze()`.
+This log can be accessed with the function `render_log_report()`.
 

### 2020-03-16 fertile.Rmd
more citations and edits

@@
 publications and discussions focusing on reproducibility seem to have 
-increased in frequency over the last several years (@sep-scientific-reproducibility), (@plos-biology), (@bioessays-gosselin)
+increased in frequency over the last several years [@sep-scientific-reproducibility, @plos-biology, @bioessays-gosselin, @eisner-reproducibility, @engineering-reproducibility].
 
@@
 provides the framework for file structure that **fertile** is based on: a 
-structure similar to that of an R package [**Cite Hadley's Writing R 
-Packages and the R Core Team's Writing R Extensions**], with an `R` 
-folder, as well as `data`, `data-raw`, `inst` and `vignettes`.
+structure similar to that of an R package (@hadley-packages, @coreteam-extensions),
+with an `R` folder, as well as `data`, `data-raw`, `inst` and `vignettes`.
 

### 2020-07-10 STAT-paper-template/APA/wileyNJD-APA.tex
added missing .tex file

\documentclass[APA,LATO1COL]{WileyNJD-v2}

\articletype{Article Type}%

\usepackage{listings}
\usepackage{hyperref}
\usepackage{graphicx}
\graphicspath{{images/}}

\received{26 April 2016}
\revised{6 June 2016}
\accepted{6 June 2016}

\raggedbottom

\begin{document}

\title{Creating optimal conditions for reproducible data analysis in R with `fertile'\protect\thanks{The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the tidyverse team.}}

\author[1]{Audrey Bertin}

\author[2]{Benjamin S. Baumer}

\authormark{Bertin and Baumer, 2020}

\address[1]{\orgdiv{Program in Statistical and Data Sciences}, \orgname{Smith College}, \orgaddress{\city{Northampton}, \state{MA}}}

\address[2]{\orgdiv{Assistant Professor of Statistical and Data Sciences}, \orgname{Smith College}, \orgaddress{\city{Northampton}, \state{MA}}}

\corres{Audrey Bertin, Program in Statistical and Data Sciences, Smith College, Northampton, MA, 01063. \email{abertin@smith.edu}}

\abstract[Summary]{The advancement of scientific knowledge increasingly depends on ensuring that data-driven research is reproducible: that two people with the same data obtain the same results. However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation, and no clear consensus on standards of what constitutes reproducibility in published research. We focus on a series of common mistakes programmers make while conducting data science projects in R, primarily through the RStudio integrated development environment. \textbf{fertile} is an R package that operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that is already written for potential problems). Furthermore, \textbf{fertile} is designed to educate the user about why the mistakes are problematic, and how to fix them.}

\keywords{reproducibility, statistical software, workflow, collaboration}

\maketitle

\section{Introduction}\label{sec1}

As research is becoming increasingly data-driven, and because knowledge can be shared worldwide so rapidly, reproducibility is critical to the advancement of scientific knowledge. 
Data-based research cannot be fully reproducible unless the requisite code and data files
produce identical results when run by another analyst. When researchers provide the code and data used for their work in a well-organized and reproducible format, readers are more easily able to determine the veracity of any findings by following the steps from raw data to conclusions.

The creators of reproducible research can more easily receive more specific feedback (including bug fixes) on their work. Moreover, others interested in the research topic can use the code to apply the methods and ideas used in one project to their own work with minimal effort.

However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation, and no clear consensus on standards of what constitutes reproducibility in published research (\cite{peng2009bio}). Not only are the components of reproducible research up for discussion (e.g., need the software be open source?), but the corresponding recommendations for ensuring reproducibility also vary (e.g., should raw and processed data files be in separate directories?).

Much of the discussion around reproducibility is also generalized, written to be applicable to users working with a variety of statistical software programs. Since all statistical software programs operate differently, generalized recommendations on reproducibility are often shallow and unspecific. While they provide useful guidelines, they can often be difficult to implement, particularly to new analysts who are unsure how to apply such recommendations within the software programs they are using. Thus, reproducibility recommendations tailored to specific software programs are more likely to be adopted.

In this paper, we focus on reproducibility in the R programming language with a concentration on projects that use the RStudio integrated development environment. R is an ideal candidate for reproducibility recommendations due to the language's popularity for statistical analyses and the ease with which analysts can download and begin using the software.

A small body of papers and R packages focusing on reproducibility in the R community have been published. Much of this work is narrowly tailored, with each package effectively addressising a small component of reproducibility: file structure, modularization of code, version control, etc. Many existing reproducibility packages, due to their focused nature, succeed at their area of focus, but at a cost. They are often difficult to learn and operate, providing a barrier to entry for less-experienced data analysts.

There appear to be no existing tools that cater to analysts looking for an easy-to-learn, easy-to-implement, fast, way to obtain a broad overview of their projects' reproducibility. To address this, we present an R package called \textbf{fertile}, a low barrier-to-entry package which focuses on a series of common mistakes programmers make while conducting data science research in R.

\section{Literature Review}\label{sec2}

Reproducibility is critically important to the advancement of knowledge in all fields of scientific research. Researchers are acknowledging this; publications and discussions focusing on reproducibility seem to have increased in frequency over the last several years. (\cite{sep-scientific-reproducibility}, \cite{plos-biology}, \cite{bioessays-gosselin}, \cite{eisner-reproducibility}, \cite{engineering-reproducibility})

Much of the available literature is focused on the methods for achieving reproducibility in specific disciplines, though much of this information is generalizable to all areas of scientific research. There is also work focusing specifically on reproducibility within the R community. Some of this work is in the form of academic papers, while other literature comes from blog posts, websites, and R packages, some of which are compiled by notable R-focused developers and researchers. In this section, we will discuss the reproducibility recommendations from both the formal and informal publications, which together influence the development of \textbf{fertile}.

\subsection{Previous Work}

\cite{Goodman341ps12} argues that the language and conceptual framework of research reproducibility varies across the sciences. There are no clear standards agreed upon across fields.

\cite{kitzes2017practice} presents a collection of case studies on reproducibility practices from across the data-intensive sciences, illustrating a variety of recommendations and techniques for achieving reproducibility. Although the book does not come to a consensus on the exact standards of reproducibility that should be followed, several common trends and principles emerge from the case studies:

\begin{itemize}
\item[1] Use clear separation, labeling, and documentation.
\item[2] Automate processes when possible.
\item[3] Design the data analysis workflow as a sequence of small steps glued together, with
outputs from one step serving as inputs into the next. This is a common suggestion within the computing community, originating as part of the Unix philosophy (Gancarz (2002)).
\end{itemize}

\cite{cooper2017guide} focuses on data analysis in R and identifies similar important reproducibility components, while reiterating the need for clearly labeled, well-documented, and well-separated files. In addition, they recommend publishing a list of dependencies and using version control.

Broman reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a \textit{README} (\cite{broman}).

The reproducibility recommendations from R OpenSci, a non-profit initiative founded in 2011 to make scientific data retrieval reproducible, share similar principles to those discussed previously. They focus on a need for a well-developed file system, with no extraneous files and clear labeling. They also reiterate the need to note dependencies and use automation when possible, while making clear a suggestion not present in the previously-discussed literature: the need to use seeds, which allow for the saving and restoring of the random number generator state, when running code involving randomness (\cite{r-opensci}).

When considered in combination, these sources provide a well-rounded picture of the components important to research reproducibility. Using this literature as a guideline, we identify several key features of reproducible work to focus on. These recommendations are opinionated—due to the lack of agreement on which components of reproducibility are most important, we select those that are mentioned most often, as well as some that are mentioned less but that we view as important.

\vskip 0.5in

\begin{itemize}
\item[1] A well-designed file structure:
  \begin{itemize}
  \item Separate folders for different file types.
  \item No extraneous files.
  \item Minimal clutter.
  \end{itemize}
\item[2] Good documentation:
  \begin{itemize}
  \item Files are clearly named, preferably in a way where the order in which they should be run is clear.
  \item A README is present.
  \item Dependencies are noted.
  \end{itemize}
\item[3] Reproducible file paths:
  \begin{itemize}
  \item No absolute paths, or paths leading to locations outside of a project’s directory, are used in code. Only portable (relative) paths.
  \end{itemize}
\item[4] Randomness is accounted for:
  \begin{itemize}
  \item If randomness is used in code, a seed must also be set.
  \end{itemize}
\item[5] Code conforms to tidyverse style:
  \begin{itemize}
  \item Although not discussed in any of the literature mentioned previously, we also believe \textbf{tidyverse} integration to be very important. Code conformation to \textbf{tidyverse} style helps ensure that analyses meeting \textbf{fertile}'s recommendations will be able to integrate seamlessly with packages from the \textbf{tidyverse}, an ever-expanding collection of packages designed to operate together with one another.
  \end{itemize}
\end{itemize}

Much of the available literature focuses on file structure, organization, and naming, and \textbf{fertile}'s features are consistent with this. The ideal file structure is not agreed upon by academics, though there are some recommendations focused on R that are available. \cite{marwick2018packaging} provides the framework for file structure that \textbf{fertile} is based on: a structure similar to that of an R package (\cite{hadley-packages}, \cite{coreteam-extensions}), with an \textit{R} folder, as well as \textit{data}, \textit{data-raw}, \textit{inst} and \textit{vignettes}.

\subsection{R Packages}

A small selection of R packages work to address the issue of research reproducibility, although not all of them are available on CRAN.

\textbf{rrtools} (\cite{R-rrtools}) addresses some of the issues discussed in \cite{marwick2018packaging} by creating a basic R package structure for a data analysis project and implementing a basic \textit{testthat::check} functionality. The \textbf{orderly} (\cite{R-orderly}) package also focuses on file structure, requiring the user to declare a desired project structure (typically a step-by-step structure, where outputs from one step are inputs into the next) at the beginning and then creating the files necessary to achieve that structure. \textbf{workflowr}'s (\cite{R-workflowr}) functionality is based around version control and making code easily available online. It works to generate a website containing time-stamped, versioned, and documented results. \textbf{checkers} (\cite{R-checkers}) allows you to create custom checks that examine different aspects of reproducibility. \textbf{packrat} (\cite{R-packrat}) is focused on dependencies, creating a packaged folder containing a project as well as all of its dependencies, so that projects dependent on lesser-used packages can be easily shared across computers. \textbf{drake} (\cite{R-drake}) works to analyze workflows, skip steps where results are up to date, and provide evidence that results match the underlying code and data. Lastly, the \textbf{reproducible} (\cite{R-reproducible}) package focuses on the concept of caching: saving information so that projects can be run faster each time they are re-completed from the start.

Many of these packages focus on one specific area of reproducibility, and while they each succeed well at their intended goal, they are not necessarily the most practical option for users trying to address a wide variety of factors influencing reproducibility at once.

Additionally, with the focused nature of these packages comes added complexity. The functions in these packages are often quite complex to use, and many steps must be completed to achieve the required reproducibility goal. This cumbersome nature means that most reproducibility packages currently available are not easily accessible to users near the beginning of their R journey, nor particularly useful to those looking for quick and easy reproducibility checks.

\section{Methods}\label{sec3}

\textbf{fertile} addresses these gaps by providing a simple, easy-to-learn reproducibility package that, rather than focusing intensely on a specific area, provides some information about a wide variety of aspects influencing reproducibility. The package also provides flexibility, offering benefits to users at any stage in the data analysis workflow, unlike some other available packages which can only be used before the creation of a new project or after it is finished.

\textbf{fertile} is designed to be used on data analyses organized as R Projects (i.e. directories containing an \textit{.Rproj} file). Once an R Project is created, \textbf{fertile} provides benefits throughout the data analysis process, both during development as well as after the fact. \textbf{fertile} achieves this by operating in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that has already been written for potential problems).

\subsection{Proactive Use}

Proactively, the package identifies potential mistakes as they are made by the user and outputs an informative message as well as a recommended solution. For example, \textbf{fertile} catches when a user passes a potentially problematic file path--such as an absolute path, or a path that points to a location outside of the project directory--to a variety of common input/output functions operating on many different file types.

\vskip 0.25in

\centerline{\includegraphics[height=3pc,width=180mm]{fig-1}}
\centerline{\includegraphics[height=6pc,width=180mm]{fig-2}}

\vskip 0.25in

\textbf{fertile} is even more aggressive with functions (like \textit{setwd()}) that are almost certain to break reproducibility, causing them to throw errors that prevent their execution and providing recommendations for better alternatives.

\vskip 0.25in

\centerline{\includegraphics[height=3pc,width=180mm]{fig-3}}

\vskip 0.25in

These proactive warning features are activated immediately after attaching the \textbf{fertile} package and require no additional effort by the user.

\subsection{Retroactive Use}

Retroactively, \textbf{fertile} analyzes potential obstacles to reproducibility in an RStudio Project (i.e., a directory that contains an \textit{.Rproj} file). The package considers several different aspects of the project which may influence reproducibility, including the directory structure, file paths, and whether randomness is used thoughtfully.

The end products of these analyses are reproducibility reports summarizing a project's adherence to reproducibility standards and recommending remedies for where the project falls short. For example, \textbf{fertile} might identify the use of randomness in code and recommend setting a seed if one is not present.

Users can access the majority of \textbf{fertile}'s retroactive features through two primary functions, \textit{proj\_check()} and \textit{proj\_analyze()}.

The \textit{proj\_check()} function runs fifteen different reproducibility tests, noting which ones passed, which ones failed, the reason for failure, a recommended solution, and a guide to where to look for help. These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that to not have a call to \textit{set.seed()} present. A full list is provided below:

\vskip 0.25in

\centerline{\includegraphics[height=16pc,width=180mm]{fig-4}}

\vskip 0.25in

Subsets of the fifteen tests can be invoked using the tidyselect helper functions in combination with the more limited \textit{proj\_check\_some()} function.

\vskip 0.25in

\centerline{\includegraphics[height=12pc,width=180mm]{fig-5}}

\vskip 0.25in

The \textit{proj\_analyze()} function creates a report documenting the structure of a data analysis project. This report contains information about all packages referenced in code, the files present in the directory and their types, suggestions for moving files to create a more organized structure, and a list of reproducibility-breaking file paths used in code.

\vskip 0.25in

\centerline{\includegraphics[height=10pc,width=180mm]{fig-6}}
\centerline{\includegraphics[height=15pc,width=180mm]{fig-7}}

\vskip 0.25in

\textbf{fertile} also contains logging functionality, which records commands run in the console that have the potential to affect reproducibility, enabling users to look at their past history at any time. The package focuses mostly on package loading and file opening, noting which function was used, the path or package it referenced, and the timestamp at which that event happened.
Users can access the log recording their commands at any time via the \textit{log\_report()} function:

\vskip 0.25in

\centerline{\includegraphics[height=15pc,width=180mm]{fig-8}}

\vskip 0.25in

The log, if not managed, can grow very long over time. For users who do not desire such functionality, \textit{log\_clear()} provides a way to erase the log and start over.

\vskip 0.25in

\centerline{\includegraphics[height=5pc,width=180mm]{fig-9}}

\vskip 0.25in

\subsection{Utility Functions}

\textbf{fertile} also provides several useful utility functions that may assist with the process of data analysis.

\subsubsection{File Type Checks}

There are several functions that can be used to check the type of a file:

\vskip 0.25in

\centerline{\includegraphics[height=15pc,width=180mm]{fig-10}}

\vskip 0.25in

\subsubsection{Temporary Directories}

The \textit{sandbox()} function allows the user to make a copy of their project in a temporary directory. This can be useful for ensuring that projects run properlt when access to the local file system is removed.

\vskip 0.25in

\centerline{\includegraphics[height=22pc,width=180mm]{fig-11}}

\vskip 0.25in

\subsubsection{Managing Project Dependencies}

One of the challenges with ensuring that work is reproducible is the issue of dependencies. Many data analysis projects reference a variety of R packages in their code. When such projects are shared with other users who may not have the required packages downloaded, it can cause errors that prevent the project from running properly.

The \textit{proj\_pkg\_script()} function assists with this issue by making it simple and fast to download dependencies. When run on an R project directory, the function creates a \textit{.R} script file that contains the code needed to install all of the packages referenced in the project, differentiating between packages located on CRAN and those located on GitHub.

\vskip 0.25in

\centerline{\includegraphics[height=3pc,width=180mm]{fig-12}}

\vskip 0.25in

\centerline{\includegraphics[height=4pc,width=180mm]{fig-13}}

\vskip 0.25in

\subsection{Sample Use Cases}

\vskip 0.25in

\subsubsection{Introductory Data Science Student}

Susan is taking an introductory data science course. This is her first time learning how to code, and she has not yet been exposed to ideas of research reproducibility. Her professor has assigned a data analysis project which must be completed in R Markdown. The project requires her to read in a data file located on her computer and use it to produce a graph.

She reads in the data, makes the graph, and knits her \textit{.Rmd} file. It compiles successfully, so she submits the assignment. The next day, she receives an email from her professor saying that her assignment failed to compile and that she needs to make changes and try again.

Susan doesn't understand why it didn't work on the professor's computer when it did on her own. The professor recommends that she install \textbf{fertile} and run \textit{proj\_check()} on her assignment. She does this and gets a message informing her she used an absolute path to open her dataset, and she should use a relative path. She looks up what this means and then uses the new information to update her assignment. She resubmits and her second attempt is successful.

On future projects, she always loads and runs \textbf{fertile} to make sure her work is okay before submitting.

\subsubsection{Experienced R User}

Emma is a post-doc, with several years of R experience. She is familiar with some basic rules of reproducibility--file paths should always be relative and randomness should always be associated with a seed--but has never needed to pass any sort of reproducibility check before because her professors never emphasized that.

She has just finished a research project and is looking to submit her work to a journal. When researching the journal to which she is interested in submitting, she discovers that it has high standards for research reproducibility and a dedicated editor focusing on that aspect of submission. She goes online and finds the journal's guidelines for reproducibility. They are more complete than any guidelines to which she has previously been required to conform. In addition to notes about file paths and randomness, the journal requires a clean, well-organized folder structure, broken down by file category and stripped of files that do not serve a purpose. In order to be approved, submissions must also have a clear build chain and an informative \textit{README} file.

Unsure of the best way to achieve this structure, Emma goes online to find help. In her search, she comes across \textbf{fertile}. She downloads the package, and in only a handful of commands, she identifies and removes excess files in her directory and automatically organizes her files into a structure reminiscent of an R package. She now meets the guidelines for the journal and can submit her research.

\section{Conclusion}\label{sec4}

\textbf{fertile} is an R package that lowers barriers to reproducible data analysis projects in R, providing a wide array of checks and suggestions addressing many different aspects of project reproducibility, including file organization, file path usage, documentation, and dependencies.

\textbf{fertile} is meant to be educational, providing informative error messages that show how the ways in which the mistakes the user is making are problematic, as well as sharing recommendations on how to fix them. The package is designed in this way so as to promote a greater understanding of reproducibility concepts in its users, with the goal of increasing the overall awareness and understanding of reproducibility in the R community.

The package has very low barriers to entry, making it accessible to users with various levels of background knowledge. Unlike many other R packages focused on reproducibility that are currently available, the features of \textbf{fertile} can be accessed almost effortlessly. Many of the retroactive features can be accessed in only two lines of code requiring minimal arguments, and some of the proactive features can be accessed with no additional effort beyond loading the package. This, in combination with the fact that \textbf{fertile} does not focus on one specific area of reproducibility, instead covering (albeit in less detail) a wide variety of topics, means that \textbf{fertile} makes it easy for data analysts of all skill levels to quickly gain a better understanding of the reproducibility of the work.

In the moment, it often feels easiest to take a shortcut—to use an absolute path or change a working directory. However, when considering the long term path of a project, spending the extra time to improve reproducibility is worthwhile. \textbf{fertile}'s user-friendly features can help data analysts avoid these harmful shortcuts with minimal effort.

\section{Future Work}\label{sec5}

\subsection{Package Features}

The process of ensuring true reproducibility is highly complex and challenging. There are a significant number of components, many of which are described in \hyperref[sec2]{Section 2}, which must all be in place for work to be reproducibile. \textbf{fertile}, in its current version, addresses the vast majority of these in some way. However, there are several areas where further development to extend the available features of the package would be beneficial. These include the following:

\begin{itemize}
  \item Expanding dependency management features to include R session information and package version numbers in addition to package names.
  \item Expanding code and documentation style features to analyze whether code has been properly commented in addition to checking for a \textit{README} and tidy code style.
  \item Adding \textit{make}-like functionality that can analyze an R project structure and files and use this information to generate a Makefile. This Makefile would have information about target files and their prerequisites and would assist with making sure that re-running an analysis is done as quickly as possible by ensuring that only the necessary code and files that have been updated are run when rebuilding and re-running code.
\end{itemize}

\subsection{Testing Effectiveness}

\textbf{fertile} is designed to:

\begin{itemize}
\item[1] Be simple enough that users with minimal R experience can use the package without issue,
\item[2] Increase the reproducibility of work produced by its users, and
\item[3] Educate its users on why their work is or is not reproducibile and provide guidance on how to address any problems.
\end{itemize}

Anecdotal evidence and user feedback is useful, providing a baseline knowledge of whether these goals are being met. However, in order to truly test whether the package is successful in achieving its design goals, an experiment is necessary.

To test \textbf{fertile}'s effectiveness, we began an initial randomized control trial of the package on an introductory undergraduate data science course at Smith College in Spring 2020, comparing one group of students that received the package functions in their R environment to one that did not. Unfortunately, however, the trial had to be posponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing.

We hope to continue this experimentation and gather data on \textbf{fertile}'s effectiveness once the higher education environment returns to a relatively stable state.

\nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;
\bibliography{wileyNJD-APA}%

\section*{Author Biography}

\begin{biography}{\includegraphics[width=70pt,height=70pt]{audrey_bio}}{\textbf{Audrey Bertin} is an undergraduate student at Smith College intending on completing a Bachelor of arts in Statistical and Data Sciences and Public Policy in Spring 2021. Audrey is a recipient of Smith College's prestigious Zollman Scholarship and has participated in extensive research with faculty throughout her time at the college. Her research interests include data and science communication, data protection and privacy, algorithmic bias and injustice, and using data science for social good.

}
\end{biography}

\begin{biography}{\includegraphics[width=70pt,height=70pt]{ben_bio}}{\textbf{Benjamin S. Baumer} is an associate professor in the Statistical and Data Sciences program at Smith College. Ben is a co-author of The Sabermetric Revolution, Modern Data Science with R, and the second edition of Analyzing Baseball Data with R. Ben has received the Waller Education Award from the ASA Section on Statistics and Data Science Education, the Significant Contributor Award from the ASA Section on Statistics in Sports, and the Contemporary Baseball Analysis Award from the Society for American Baseball Research. His research interests include sports analytics, data science, statistics and data science education, statistical computing, and network science.}
\end{biography}

\end{document}

### 2020-07-14 STAT-paper-template/APA/fertile.Rnw
update for STAT

--- 2020-07-10 STAT-paper-template/APA/wileyNJD-APA.tex
@@
 \usepackage{graphicx}
+\usepackage{xspace}
 \graphicspath{{images/}}
+
+% custom macros
+
+\newcommand{\R}{\textsf{R}\xspace}
+\newcommand{\RStudio}{\textsf{RStudio}\xspace}
+\newcommand{\cmd}[1]{\textit{#1}}
+\newcommand{\pkg}[1]{\textbf{#1}}
+\newcommand{\func}[1]{\textit{#1}()}
 
@@
 
-\title{Creating optimal conditions for reproducible data analysis in R with `fertile'\protect\thanks{The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the tidyverse team.}}
+\title{Creating optimal conditions for reproducible data analysis in \R with `fertile'\protect\thanks{The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the \pkg{tidyverse} team.}}
 
-\author[1]{Audrey Bertin}
+\author[1]{Audrey M. Bertin}
 
@@
 
-\address[2]{\orgdiv{Assistant Professor of Statistical and Data Sciences}, \orgname{Smith College}, \orgaddress{\city{Northampton}, \state{MA}}}
+\address[2]{\orgdiv{Associate Professor of Statistical and Data Sciences}, \orgname{Smith College}, \orgaddress{\city{Northampton}, \state{MA}}}
 
@@
 We focus on a series of common mistakes programmers make while conducting data science projects in R, primarily through the RStudio integrated development environment.
-\textbf{fertile} is an R package that operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that is already written for potential problems).
-Furthermore, \textbf{fertile} is designed to educate the user about why the mistakes are problematic, and how to fix them.}
+\pkg{fertile} is an R package that operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that is already written for potential problems).
+Furthermore, \pkg{fertile} is designed to educate the user about why the mistakes are problematic, and how to fix them.}
 
@@
 There appear to be no existing tools that cater to analysts looking for an easy-to-learn, easy-to-implement, fast, way to obtain a broad overview of their projects' reproducibility.
-To address this, we present an R package called \textbf{fertile}, a low barrier-to-entry package which focuses on a series of common mistakes programmers make while conducting data science research in R.
+To address this, we present an R package called \pkg{fertile}, a low barrier-to-entry package which focuses on a series of common mistakes programmers make while conducting data science research in R.
 
@@
 Reproducibility is critically important to the advancement of knowledge in all fields of scientific research.
-Researchers are acknowledging this; publications and discussions focusing on reproducibility seem to have increased in frequency over the last several years.
-(\cite{sep-scientific-reproducibility}, \cite{plos-biology}, \cite{bioessays-gosselin}, \cite{eisner-reproducibility}, \cite{engineering-reproducibility})
+Researchers are acknowledging this; publications and discussions focusing on reproducibility seem to have increased in frequency over the last several years (\cite{sep-scientific-reproducibility}, \cite{plos-biology}, \cite{bioessays-gosselin}, \cite{eisner-reproducibility}, \cite{engineering-reproducibility}).
 
-Much of the available literature is focused on the methods for achieving reproducibility in specific disciplines, though much of this information is generalizable to all areas of scientific research.
+Much of the available literature is focused on methods for achieving reproducibility in specific disciplines, and some of their recommendations are generalizable to all areas of scientific research.
 There is also work focusing specifically on reproducibility within the R community.
 Some of this work is in the form of academic papers, while other literature comes from blog posts, websites, and R packages, some of which are compiled by notable R-focused developers and researchers.
-In this section, we will discuss the reproducibility recommendations from both the formal and informal publications, which together influence the development of \textbf{fertile}.
+In this section, we will discuss the reproducibility recommendations from both the formal and informal publications, which together influence the development of \pkg{fertile}.
 
@@
 
-\cite{Goodman341ps12} argues that the language and conceptual framework of research reproducibility varies across the sciences. There are no clear standards agreed upon across fields.
+\cite{Goodman341ps12} argue that the language and conceptual framework of research reproducibility varies across the sciences. There are no clear standards agreed upon across fields.
 
-\cite{kitzes2017practice} presents a collection of case studies on reproducibility practices from across the data-intensive sciences, illustrating a variety of recommendations and techniques for achieving reproducibility.
-Although the book does not come to a consensus on the exact standards of reproducibility that should be followed, several common trends and principles emerge from the case studies:
+\cite{kitzes2017practice} present a collection of case studies on reproducibility practices from across the data-intensive sciences, illustrating a variety of recommendations and techniques for achieving reproducibility.
+Although the book does not come to a consensus on the exact standards of reproducibility that should be followed, several common trends and principles emerge from their case studies:
 
@@
 
-Broman reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a \textit{README} (\cite{broman}).
+Broman reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a \cmd{README} (\cite{broman}).
 
@@
 Using this literature as a guideline, we identify several key features of reproducible work to focus on.
-These recommendations are opinionated—due to the lack of agreement on which components of reproducibility are most important, we select those that are mentioned most often, as well as some that are mentioned less but that we view as important.
-
-\vskip 0.5in
+These recommendations are opinionated---due to the lack of agreement on which components of reproducibility are most important, we select those that are mentioned most often, as well as some that are mentioned less but that we view as important.
 
@@
   \begin{itemize}
-  \item No absolute paths, or paths leading to locations outside of a project’s directory, are used in code. Only portable (relative) paths.
+  \item No absolute paths, or paths leading to locations outside of a project's directory, are used in code. Only portable (relative) paths.
   \end{itemize}
@@
   \end{itemize}
-\item[5] Code conforms to tidyverse style:
+\item[5] Code conforms to \pkg{tidyverse} style:
   \begin{itemize}
-  \item Although not discussed in any of the literature mentioned previously, we also believe \textbf{tidyverse} integration to be very important.
-Code conformation to \textbf{tidyverse} style helps ensure that analyses meeting \textbf{fertile}'s recommendations will be able to integrate seamlessly with packages from the \textbf{tidyverse}, an ever-expanding collection of packages designed to operate together with one another.
+  \item Although not discussed in any of the literature mentioned previously, we believe \pkg{tidyverse} integration to be very important.
+Code conformation to \pkg{tidyverse} style helps ensure that analyses meeting \pkg{fertile}'s recommendations will be able to integrate seamlessly with packages from the \pkg{tidyverse}, an ever-expanding collection of packages designed to operate together with one another.
   \end{itemize}
@@
 
-Much of the available literature focuses on file structure, organization, and naming, and \textbf{fertile}'s features are consistent with this.
+Much of the available literature focuses on file structure, organization, and naming, and \pkg{fertile}'s features are consistent with this.
 The ideal file structure is not agreed upon by academics, though there are some recommendations focused on R that are available.
-\cite{marwick2018packaging} provides the framework for file structure that \textbf{fertile} is based on: a structure similar to that of an R package (\cite{hadley-packages}, \cite{coreteam-extensions}), with an \textit{R} folder, as well as \textit{data}, \textit{data-raw}, \textit{inst} and \textit{vignettes}.
+\cite{marwick2018packaging} provides the framework for file structure that \pkg{fertile} is based on: a structure similar to that of an R package (\cite{hadley-packages}, \cite{coreteam-extensions}), with an \cmd{R} folder, as well as \cmd{data}, \cmd{data-raw}, \cmd{inst}, and \cmd{vignettes}.
 
@@
 
-\textbf{rrtools} (\cite{R-rrtools}) addresses some of the issues discussed in \cite{marwick2018packaging} by creating a basic R package structure for a data analysis project and implementing a basic \textit{testthat::check} functionality.
-The \textbf{orderly} (\cite{R-orderly}) package also focuses on file structure, requiring the user to declare a desired project structure (typically a step-by-step structure, where outputs from one step are inputs into the next) at the beginning and then creating the files necessary to achieve that structure.
-\textbf{workflowr}'s (\cite{R-workflowr}) functionality is based around version control and making code easily available online.
+\pkg{rrtools} (\cite{R-rrtools}) addresses some of the issues discussed in \cite{marwick2018packaging} by creating a basic R package structure for a data analysis project and implementing a basic \func{testthat::check} functionality.
+The \pkg{orderly} (\cite{R-orderly}) package also focuses on file structure, requiring the user to declare a desired project structure (typically a step-by-step structure, where outputs from one step are inputs into the next) at the beginning and then creating the files necessary to achieve that structure.
+\pkg{workflowr}'s (\cite{R-workflowr}) functionality is based around version control and making code easily available online.
 It works to generate a website containing time-stamped, versioned, and documented results.
-\textbf{checkers} (\cite{R-checkers}) allows you to create custom checks that examine different aspects of reproducibility.
-\textbf{packrat} (\cite{R-packrat}) is focused on dependencies, creating a packaged folder containing a project as well as all of its dependencies, so that projects dependent on lesser-used packages can be easily shared across computers.
-\textbf{drake} (\cite{R-drake}) works to analyze workflows, skip steps where results are up to date, and provide evidence that results match the underlying code and data.
-Lastly, the \textbf{reproducible} (\cite{R-reproducible}) package focuses on the concept of caching: saving information so that projects can be run faster each time they are re-completed from the start.
+\pkg{checkers} (\cite{R-checkers}) allows you to create custom checks that examine different aspects of reproducibility.
+\pkg{packrat} (\cite{R-packrat}) is focused on dependencies, creating a packaged folder containing a project as well as all of its dependencies, so that projects dependent on lesser-used packages can be easily shared across computers.
+\pkg{drake} (\cite{R-drake}) analyzes workflows, skip steps where results are up to date, and provides evidence that results match the underlying code and data.
+Lastly, the \pkg{reproducible} (\cite{R-reproducible}) package focuses on the concept of caching: saving information so that projects can be run faster each time they are re-completed from the start.
 
@@
 
-\textbf{fertile} addresses these gaps by providing a simple, easy-to-learn reproducibility package that, rather than focusing intensely on a specific area, provides some information about a wide variety of aspects influencing reproducibility.
+\pkg{fertile} addresses these gaps by providing a simple, easy-to-learn reproducibility package that, rather than focusing intensely on a specific area, provides some information about a wide variety of aspects influencing reproducibility.
 The package also provides flexibility, offering benefits to users at any stage in the data analysis workflow, unlike some other available packages which can only be used before the creation of a new project or after it is finished.
 
-\textbf{fertile} is designed to be used on data analyses organized as R Projects (i.e.
-directories containing an \textit{.Rproj} file).
-Once an R Project is created, \textbf{fertile} provides benefits throughout the data analysis process, both during development as well as after the fact.
-\textbf{fertile} achieves this by operating in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that has already been written for potential problems).
+\pkg{fertile} is designed to be used on data analyses organized as R Projects (i.e.
+directories containing an \cmd{.Rproj} file).
+Once an R Project is created, \pkg{fertile} provides benefits throughout the data analysis process, both during development as well as after the fact.
+\pkg{fertile} achieves this by operating in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that has already been written for potential problems).
 
@@
 Proactively, the package identifies potential mistakes as they are made by the user and outputs an informative message as well as a recommended solution.
-For example, \textbf{fertile} catches when a user passes a potentially problematic file path--such as an absolute path, or a path that points to a location outside of the project directory--to a variety of common input/output functions operating on many different file types.
+For example, \pkg{fertile} catches when a user passes a potentially problematic file path---such as an absolute path, or a path that points to a location outside of the project directory---to a variety of common input/output functions operating on many different file types.
 
-\vskip 0.25in
+<<write_csv, include=FALSE>>=
+readr::write_csv(mtcars, "~/Desktop/my_data.csv")
+@
 
-\centerline{\includegraphics[height=3pc,width=180mm]{fig-1}}
-\centerline{\includegraphics[height=6pc,width=180mm]{fig-2}}
+<<fertile, message = FALSE>>=
+library(fertile)
+search()
+file.exists("~/Desktop/my_data.csv")
+read.csv("~/Desktop/my_data.csv")
+@
 
-\vskip 0.25in
+\pkg{fertile} is even more aggressive with functions (like \func{setwd}) that are almost certain to break reproducibility, causing them to throw errors that prevent their execution and providing recommendations for better alternatives.
 
-\textbf{fertile} is even more aggressive with functions (like \textit{setwd()}) that are almost certain to break reproducibility, causing them to throw errors that prevent their execution and providing recommendations for better alternatives.
+<<setwd, message=FALSE>>=
+setwd("~/Desktop")
+@
 
-\vskip 0.25in
-
-\centerline{\includegraphics[height=3pc,width=180mm]{fig-3}}
-
-\vskip 0.25in
-
-These proactive warning features are activated immediately after attaching the \textbf{fertile} package and require no additional effort by the user.
+These proactive warning features are activated immediately after attaching the \pkg{fertile} package and require no additional effort by the user.
 
@@
 
-Retroactively, \textbf{fertile} analyzes potential obstacles to reproducibility in an RStudio Project (i.e., a directory that contains an \textit{.Rproj} file).
+Retroactively, \pkg{fertile} analyzes potential obstacles to reproducibility in an RStudio Project (i.e., a directory that contains an \cmd{.Rproj} file).
 The package considers several different aspects of the project which may influence reproducibility, including the directory structure, file paths, and whether randomness is used thoughtfully.
@@
 The end products of these analyses are reproducibility reports summarizing a project's adherence to reproducibility standards and recommending remedies for where the project falls short.
-For example, \textbf{fertile} might identify the use of randomness in code and recommend setting a seed if one is not present.
+For example, \pkg{fertile} might identify the use of randomness in code and recommend setting a seed if one is not present.
 
-Users can access the majority of \textbf{fertile}'s retroactive features through two primary functions, \textit{proj\_check()} and \textit{proj\_analyze()}.
+Users can access the majority of \pkg{fertile}'s retroactive features through two primary functions, \func{proj\_check} and \func{proj\_analyze}.
 
-The \textit{proj\_check()} function runs fifteen different reproducibility tests, noting which ones passed, which ones failed, the reason for failure, a recommended solution, and a guide to where to look for help.
-These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that to not have a call to \textit{set.seed()} present.
+The \func{proj\_check} function runs fifteen different reproducibility tests, noting which ones passed, which ones failed, the reason for failure, a recommended solution, and a guide to where to look for help.
+These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that to not have a call to \func{set.seed} present.
 A full list is provided below:
 
-\vskip 0.25in
+<<list-checks>>=
+list_checks()
+@
 
-\centerline{\includegraphics[height=16pc,width=180mm]{fig-4}}
+Subsets of the fifteen tests can be invoked using the \pkg{tidyselect} helper functions (\cite{R-tidyselect}) in combination with the more limited \func{proj\_check\_some} function.
 
-\vskip 0.25in
+<<set-project>>=
+proj_dir <- "~/Dropbox/git/baumer-lab/fertile/tests/testthat/project_miceps"
+@
 
-Subsets of the fifteen tests can be invoked using the tidyselect helper functions in combination with the more limited \textit{proj\_check\_some()} function.
+<<echo=FALSE>>=
+bad_file <- fs::path(proj_dir, "install_proj_packages.R")
+if (file.exists(bad_file)) {
+  fs::file_delete(bad_file)
+}
+@
 
-\vskip 0.25in
+<<proj_check_some>>=
+proj_check_some(proj_dir, contains("paths"))
+@
 
-\centerline{\includegraphics[height=12pc,width=180mm]{fig-5}}
-
-\vskip 0.25in
-
-The \textit{proj\_analyze()} function creates a report documenting the structure of a data analysis project.
+The \func{proj\_analyze} function creates a report documenting the structure of a data analysis project.
 This report contains information about all packages referenced in code, the files present in the directory and their types, suggestions for moving files to create a more organized structure, and a list of reproducibility-breaking file paths used in code.
 
-\vskip 0.25in
+<<proj_analyze>>=
+proj_analyze(proj_dir)
+@
 
-\centerline{\includegraphics[height=10pc,width=180mm]{fig-6}}
-\centerline{\includegraphics[height=15pc,width=180mm]{fig-7}}
+\pkg{fertile} also contains logging functionality, which records commands run in the console that have the potential to affect reproducibility, enabling users to look at their past history at any time.
+The package focuses mostly on package loading and file opening, noting which function was used, the path or package it referenced, and the timestamp at which that event happened.
+Users can access the log recording their commands at any time via the \func{log\_report} function:
 
-\vskip 0.25in
+<<read-csv, message=FALSE, include=FALSE>>=
+library(purrr)
+library(forcats)
+library(stringr)
+read_csv(fs::path(proj_dir, "mice.csv"))
+@
 
-\textbf{fertile} also contains logging functionality, which records commands run in the console that have the potential to affect reproducibility, enabling users to look at their past history at any time.
-The package focuses mostly on package loading and file opening, noting which function was used, the path or package it referenced, and the timestamp at which that event happened.
-Users can access the log recording their commands at any time via the \textit{log\_report()} function:
+<<log-report>>=
+log_report()
+@
 
-\vskip 0.25in
+The log, if not managed, can grow very long over time. For users who do not desire such functionality, \func{log\_clear} provides a way to erase the log and start over.
 
-\centerline{\includegraphics[height=15pc,width=180mm]{fig-8}}
-
-\vskip 0.25in
-
-The log, if not managed, can grow very long over time. For users who do not desire such functionality, \textit{log\_clear()} provides a way to erase the log and start over.
-
-\vskip 0.25in
-
-\centerline{\includegraphics[height=5pc,width=180mm]{fig-9}}
-
-\vskip 0.25in
+<<log-clear>>=
+log_clear()
+log_report()
+@
 
@@
 
-\textbf{fertile} also provides several useful utility functions that may assist with the process of data analysis.
+\pkg{fertile} also provides several useful utility functions that may assist with the process of data analysis.
 
@@
 
-\vskip 0.25in
+<<paths, message=FALSE>>=
+is_path_portable("~")
+@
 
-\centerline{\includegraphics[height=15pc,width=180mm]{fig-10}}
-
-\vskip 0.25in
+<<is_file>>=
+is_data_file(fs::path(proj_dir, "mice.csv"))
+is_image_file(fs::path(proj_dir, "mice.csv"))
+is_image_file(fs::path(proj_dir, "proteins_v_time.png"))
+@
 
@@
 
-The \textit{sandbox()} function allows the user to make a copy of their project in a temporary directory.
-This can be useful for ensuring that projects run properlt when access to the local file system is removed.
+The \func{sandbox} function allows the user to make a copy of their project in a temporary directory.
+This can be useful for ensuring that projects run properly when access to the local file system is removed.
 
-\vskip 0.25in
-
-\centerline{\includegraphics[height=22pc,width=180mm]{fig-11}}
-
-\vskip 0.25in
+<<sandbox>>=
+dir <- getwd()
+new_dir <- sandbox(dir)
+dir
+fs::dir_ls(dir) %>% head()
+new_dir
+fs::dir_ls(new_dir) %>% head()
+@
 
@@
 
-The \textit{proj\_pkg\_script()} function assists with this issue by making it simple and fast to download dependencies.
-When run on an R project directory, the function creates a \textit{.R} script file that contains the code needed to install all of the packages referenced in the project, differentiating between packages located on CRAN and those located on GitHub.
+The \func{proj\_pkg\_script} function assists with this issue by making it simple and fast to download dependencies.
+When run on an R project directory, the function creates a \cmd{.R} script file that contains the code needed to install all of the packages referenced in the project, differentiating between packages located on CRAN and those located on GitHub.
 
-\vskip 0.25in
-
-\centerline{\includegraphics[height=3pc,width=180mm]{fig-12}}
-
-\vskip 0.25in
-
-\centerline{\includegraphics[height=4pc,width=180mm]{fig-13}}
-
-\vskip 0.25in
+<<proj_pkg_script, eval=TRUE>>=
+install_script <- proj_pkg_script(proj_dir)
+cat(readChar(install_script, 1e5))
+@
 
 \subsection{Sample Use Cases}
-
-\vskip 0.25in
 
@@
 
-She reads in the data, makes the graph, and knits her \textit{.Rmd} file.
+She reads in the data, makes the graph, and knits her \cmd{.Rmd} file.
 It compiles successfully, so she submits the assignment.
@@
 Susan doesn't understand why it didn't work on the professor's computer when it did on her own.
-The professor recommends that she install \textbf{fertile} and run \textit{proj\_check()} on her assignment.
+The professor recommends that she install \pkg{fertile} and run \func{proj\_check} on her assignment.
 She does this and gets a message informing her she used an absolute path to open her dataset, and she should use a relative path.
@@
 
-On future projects, she always loads and runs \textbf{fertile} to make sure her work is okay before submitting.
+On future projects, she always loads and runs \pkg{fertile} to make sure her work is okay before submitting.
 
@@
 In addition to notes about file paths and randomness, the journal requires a clean, well-organized folder structure, broken down by file category and stripped of files that do not serve a purpose.
-In order to be approved, submissions must also have a clear build chain and an informative \textit{README} file.
+In order to be approved, submissions must also have a clear build chain and an informative \cmd{README} file.
 
 Unsure of the best way to achieve this structure, Emma goes online to find help.
-In her search, she comes across \textbf{fertile}.
+In her search, she comes across \pkg{fertile}.
 She downloads the package, and in only a handful of commands, she identifies and removes excess files in her directory and automatically organizes her files into a structure reminiscent of an R package.
@@
 
-\textbf{fertile} is an R package that lowers barriers to reproducible data analysis projects in R, providing a wide array of checks and suggestions addressing many different aspects of project reproducibility, including file organization, file path usage, documentation, and dependencies.
+\pkg{fertile} is an R package that lowers barriers to reproducible data analysis projects in R, providing a wide array of checks and suggestions addressing many different aspects of project reproducibility, including file organization, file path usage, documentation, and dependencies.
 
-\textbf{fertile} is meant to be educational, providing informative error messages that show how the ways in which the mistakes the user is making are problematic, as well as sharing recommendations on how to fix them.
+\pkg{fertile} is meant to be educational, providing informative error messages that show how the ways in which the mistakes the user is making are problematic, as well as sharing recommendations on how to fix them.
 The package is designed in this way so as to promote a greater understanding of reproducibility concepts in its users, with the goal of increasing the overall awareness and understanding of reproducibility in the R community.
@@
 The package has very low barriers to entry, making it accessible to users with various levels of background knowledge.
-Unlike many other R packages focused on reproducibility that are currently available, the features of \textbf{fertile} can be accessed almost effortlessly.
+Unlike many other R packages focused on reproducibility that are currently available, the features of \pkg{fertile} can be accessed almost effortlessly.
 Many of the retroactive features can be accessed in only two lines of code requiring minimal arguments, and some of the proactive features can be accessed with no additional effort beyond loading the package.
-This, in combination with the fact that \textbf{fertile} does not focus on one specific area of reproducibility, instead covering (albeit in less detail) a wide variety of topics, means that \textbf{fertile} makes it easy for data analysts of all skill levels to quickly gain a better understanding of the reproducibility of the work.
+This, in combination with the fact that \pkg{fertile} does not focus on one specific area of reproducibility, instead covering (albeit in less detail) a wide variety of topics, means that \pkg{fertile} makes it easy for data analysts of all skill levels to quickly gain a better understanding of the reproducibility of the work.
 
-In the moment, it often feels easiest to take a shortcut—to use an absolute path or change a working directory.
+In the moment, it often feels easiest to take a shortcut--to use an absolute path or change a working directory.
 However, when considering the long term path of a project, spending the extra time to improve reproducibility is worthwhile.
-\textbf{fertile}'s user-friendly features can help data analysts avoid these harmful shortcuts with minimal effort.
+\pkg{fertile}'s user-friendly features can help data analysts avoid these harmful shortcuts with minimal effort.
 
@@
 There are a significant number of components, many of which are described in \hyperref[sec2]{Section 2}, which must all be in place for work to be reproducibile.
-\textbf{fertile}, in its current version, addresses the vast majority of these in some way.
+\pkg{fertile}, in its current version, addresses the vast majority of these in some way.
 However, there are several areas where further development to extend the available features of the package would be beneficial.
@@
   \item Expanding dependency management features to include R session information and package version numbers in addition to package names.
-  \item Expanding code and documentation style features to analyze whether code has been properly commented in addition to checking for a \textit{README} and tidy code style.
-  \item Adding \textit{make}-like functionality that can analyze an R project structure and files and use this information to generate a Makefile.
+  \item Expanding code and documentation style features to analyze whether code has been properly commented in addition to checking for a \cmd{README} and tidy code style.
+  \item Adding \cmd{make}-like functionality that can analyze an R project structure and files and use this information to generate a Makefile.
 This Makefile would have information about target files and their prerequisites and would assist with making sure that re-running an analysis is done as quickly as possible by ensuring that only the necessary code and files that have been updated are run when rebuilding and re-running code.
@@
 
-\textbf{fertile} is designed to:
+\pkg{fertile} is designed to:
 
@@
 
-To test \textbf{fertile}'s effectiveness, we began an initial randomized control trial of the package on an introductory undergraduate data science course at Smith College in Spring 2020, comparing one group of students that received the package functions in their R environment to one that did not.
-Unfortunately, however, the trial had to be posponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing.
+To test \pkg{fertile}'s effectiveness, we began an initial randomized control trial of the package on an introductory undergraduate data science course at Smith College in Spring 2020, comparing one group of students that received the package functions in their R environment to one that did not.
+Unfortunately, however, the trial had to be postponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing.
 
-We hope to continue this experimentation and gather data on \textbf{fertile}'s effectiveness once the higher education environment returns to a relatively stable state.
+We hope to continue this experimentation and gather data on \pkg{fertile}'s effectiveness once the higher education environment returns to a relatively stable state.
 
 \nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;
-\bibliography{wileyNJD-APA}%
+\bibliography{refs}%
 
@@
 
-\begin{biography}{\includegraphics[width=70pt,height=70pt]{audrey_bio}}{\textbf{Audrey Bertin} is an undergraduate student at Smith College intending on completing a Bachelor of arts in Statistical and Data Sciences and Public Policy in Spring 2021.
+\begin{biography}{\includegraphics[width=70pt,height=70pt]{audrey_bio}}{\textbf{Audrey Bertin} is an undergraduate student at Smith College intending on completing a Bachelor of Arts in Statistical and Data Sciences and Public Policy in Spring 2021.
 Audrey is a recipient of Smith College's prestigious Zollman Scholarship and has participated in extensive research with faculty throughout her time at the college.
@@
 Baumer} is an associate professor in the Statistical and Data Sciences program at Smith College.
-Ben is a co-author of The Sabermetric Revolution, Modern Data Science with R, and the second edition of Analyzing Baseball Data with R.
+Ben is a co-author of \textit{The Sabermetric Revolution}, \textit{Modern Data Science with R}, and the second edition of \textit{Analyzing Baseball Data with R}.
 Ben has received the Waller Education Award from the ASA Section on Statistics and Data Science Education, the Significant Contributor Award from the ASA Section on Statistics in Sports, and the Contemporary Baseball Analysis Award from the Society for American Baseball Research.

### 2020-07-14 STAT-paper-template/APA/fertile.Rnw
added \R macros

@@
 We focus on a series of common mistakes programmers make while conducting data science projects in R, primarily through the RStudio integrated development environment.
-\pkg{fertile} is an R package that operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that is already written for potential problems).
+\pkg{fertile} is an \R package that operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that is already written for potential problems).
 Furthermore, \pkg{fertile} is designed to educate the user about why the mistakes are problematic, and how to fix them.}
@@
 
-In this paper, we focus on reproducibility in the R programming language with a concentration on projects that use the RStudio integrated development environment.
-R is an ideal candidate for reproducibility recommendations due to the language's popularity for statistical analyses and the ease with which analysts can download and begin using the software.
+In this paper, we focus on reproducibility in the \R programming language with a concentration on projects that use the RStudio integrated development environment.
+\R is an ideal candidate for reproducibility recommendations due to the language's popularity for statistical analyses and the ease with which analysts can download and begin using the software.
 
-A small body of papers and R packages focusing on reproducibility in the R community have been published.
+A small body of papers and \R packages focusing on reproducibility in the \R community have been published.
 Much of this work is narrowly tailored, with each package effectively addressising a small component of reproducibility: file structure, modularization of code, version control, etc.
@@
 There appear to be no existing tools that cater to analysts looking for an easy-to-learn, easy-to-implement, fast, way to obtain a broad overview of their projects' reproducibility.
-To address this, we present an R package called \pkg{fertile}, a low barrier-to-entry package which focuses on a series of common mistakes programmers make while conducting data science research in R.
+To address this, we present an \R package called \pkg{fertile}, a low barrier-to-entry package which focuses on a series of common mistakes programmers make while conducting data science research in R.
 
-\section{Literature Review}\label{sec2}
+\section{Literature Review}\label{sec:lit}
 
@@
 Much of the available literature is focused on methods for achieving reproducibility in specific disciplines, and some of their recommendations are generalizable to all areas of scientific research.
-There is also work focusing specifically on reproducibility within the R community.
-Some of this work is in the form of academic papers, while other literature comes from blog posts, websites, and R packages, some of which are compiled by notable R-focused developers and researchers.
+There is also work focusing specifically on reproducibility within the \R community.
+Some of this work is in the form of academic papers, while other literature comes from blog posts, websites, and \R packages, some of which are compiled by notable R-focused developers and researchers.
 In this section, we will discuss the reproducibility recommendations from both the formal and informal publications, which together influence the development of \pkg{fertile}.
@@
 
-\cite{cooper2017guide} focuses on data analysis in R and identifies similar important reproducibility components, while reiterating the need for clearly labeled, well-documented, and well-separated files.
+\cite{cooper2017guide} focuses on data analysis in \R and identifies similar important reproducibility components, while reiterating the need for clearly labeled, well-documented, and well-separated files.
 In addition, they recommend publishing a list of dependencies and using version control.
@@
 Much of the available literature focuses on file structure, organization, and naming, and \pkg{fertile}'s features are consistent with this.
-The ideal file structure is not agreed upon by academics, though there are some recommendations focused on R that are available.
-\cite{marwick2018packaging} provides the framework for file structure that \pkg{fertile} is based on: a structure similar to that of an R package (\cite{hadley-packages}, \cite{coreteam-extensions}), with an \cmd{R} folder, as well as \cmd{data}, \cmd{data-raw}, \cmd{inst}, and \cmd{vignettes}.
+The ideal file structure is not agreed upon by academics, though there are some recommendations focused on \R that are available.
+\cite{marwick2018packaging} provides the framework for file structure that \pkg{fertile} is based on: a structure similar to that of an \R package (\cite{hadley-packages}, \cite{coreteam-extensions}), with an \cmd{R} folder, as well as \cmd{data}, \cmd{data-raw}, \cmd{inst}, and \cmd{vignettes}.
 
@@
 
-A small selection of R packages work to address the issue of research reproducibility, although not all of them are available on CRAN.
+A small selection of \R packages work to address the issue of research reproducibility, although not all of them are available on CRAN.
 
-\pkg{rrtools} (\cite{R-rrtools}) addresses some of the issues discussed in \cite{marwick2018packaging} by creating a basic R package structure for a data analysis project and implementing a basic \func{testthat::check} functionality.
+\pkg{rrtools} (\cite{R-rrtools}) addresses some of the issues discussed in \cite{marwick2018packaging} by creating a basic \R package structure for a data analysis project and implementing a basic \func{testthat::check} functionality.
 The \pkg{orderly} (\cite{R-orderly}) package also focuses on file structure, requiring the user to declare a desired project structure (typically a step-by-step structure, where outputs from one step are inputs into the next) at the beginning and then creating the files necessary to achieve that structure.
@@
 The functions in these packages are often quite complex to use, and many steps must be completed to achieve the required reproducibility goal.
-This cumbersome nature means that most reproducibility packages currently available are not easily accessible to users near the beginning of their R journey, nor particularly useful to those looking for quick and easy reproducibility checks.
+This cumbersome nature means that most reproducibility packages currently available are not easily accessible to users near the beginning of their \R journey, nor particularly useful to those looking for quick and easy reproducibility checks.
 
@@
 
-\pkg{fertile} is designed to be used on data analyses organized as R Projects (i.e.
+\pkg{fertile} is designed to be used on data analyses organized as \R Projects (i.e.
 directories containing an \cmd{.Rproj} file).
-Once an R Project is created, \pkg{fertile} provides benefits throughout the data analysis process, both during development as well as after the fact.
+Once an \R Project is created, \pkg{fertile} provides benefits throughout the data analysis process, both during development as well as after the fact.
 \pkg{fertile} achieves this by operating in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that has already been written for potential problems).
@@
 One of the challenges with ensuring that work is reproducible is the issue of dependencies.
-Many data analysis projects reference a variety of R packages in their code.
+Many data analysis projects reference a variety of \R packages in their code.
 When such projects are shared with other users who may not have the required packages downloaded, it can cause errors that prevent the project from running properly.
@@
 The \func{proj\_pkg\_script} function assists with this issue by making it simple and fast to download dependencies.
-When run on an R project directory, the function creates a \cmd{.R} script file that contains the code needed to install all of the packages referenced in the project, differentiating between packages located on CRAN and those located on GitHub.
+When run on an \R project directory, the function creates a \cmd{.R} script file that contains the code needed to install all of the packages referenced in the project, differentiating between packages located on CRAN and those located on GitHub.
 
@@
 This is her first time learning how to code, and she has not yet been exposed to ideas of research reproducibility.
-Her professor has assigned a data analysis project which must be completed in R Markdown.
+Her professor has assigned a data analysis project which must be completed in \R Markdown.
 The project requires her to read in a data file located on her computer and use it to produce a graph.
@@
 
-\subsubsection{Experienced R User}
+\subsubsection{Experienced \R User}
 
-Emma is a post-doc, with several years of R experience.
+Emma is a post-doc, with several years of \R experience.
 She is familiar with some basic rules of reproducibility--file paths should always be relative and randomness should always be associated with a seed--but has never needed to pass any sort of reproducibility check before because her professors never emphasized that.
@@
 In her search, she comes across \pkg{fertile}.
-She downloads the package, and in only a handful of commands, she identifies and removes excess files in her directory and automatically organizes her files into a structure reminiscent of an R package.
+She downloads the package, and in only a handful of commands, she identifies and removes excess files in her directory and automatically organizes her files into a structure reminiscent of an \R package.
 She now meets the guidelines for the journal and can submit her research.
 
-\section{Conclusion}\label{sec4}
+\section{Results}
 
-\pkg{fertile} is an R package that lowers barriers to reproducible data analysis projects in R, providing a wide array of checks and suggestions addressing many different aspects of project reproducibility, including file organization, file path usage, documentation, and dependencies.
+\pkg{fertile} is designed to:
+
+\begin{itemize}
+\item[1] Be simple enough that users with minimal \R experience can use the package without issue,
+\item[2] Increase the reproducibility of work produced by its users, and
+\item[3] Educate its users on why their work is or is not reproducibile and provide guidance on how to address any problems.
+\end{itemize}
+
+Anecdotal evidence and user feedback is useful, providing a baseline knowledge of whether these goals are being met.
+However, in order to truly test whether the package is successful in achieving its design goals, an experiment is necessary.
+
+To test \pkg{fertile}'s effectiveness, we began an initial randomized control trial of the package on an introductory undergraduate data science course at Smith College in Spring 2020, comparing one group of students that received the package functions in their \R environment to one that did not.
+Unfortunately, however, the trial had to be postponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing.
+
+We hope to continue this experimentation and gather data on \pkg{fertile}'s effectiveness once the higher education environment returns to a relatively stable state.
+
+\section{Conclusion}\label{sec:conclusion}
+
+\pkg{fertile} is an \R package that lowers barriers to reproducible data analysis projects in \R, providing a wide array of checks and suggestions addressing many different aspects of project reproducibility, including file organization, file path usage, documentation, and dependencies.
 
 \pkg{fertile} is meant to be educational, providing informative error messages that show how the ways in which the mistakes the user is making are problematic, as well as sharing recommendations on how to fix them.
-The package is designed in this way so as to promote a greater understanding of reproducibility concepts in its users, with the goal of increasing the overall awareness and understanding of reproducibility in the R community.
+The package is designed in this way so as to promote a greater understanding of reproducibility concepts in its users, with the goal of increasing the overall awareness and understanding of reproducibility in the \R community.
 
 The package has very low barriers to entry, making it accessible to users with various levels of background knowledge.
-Unlike many other R packages focused on reproducibility that are currently available, the features of \pkg{fertile} can be accessed almost effortlessly.
+Unlike many other \R packages focused on reproducibility that are currently available, the features of \pkg{fertile} can be accessed almost effortlessly.
 Many of the retroactive features can be accessed in only two lines of code requiring minimal arguments, and some of the proactive features can be accessed with no additional effort beyond loading the package.
@@
 
-\section{Future Work}\label{sec5}
+\subsection{Future Work}\label{sec:future}
 
-\subsection{Package Features}
+\subsubsection{Package Features}
 
 The process of ensuring true reproducibility is highly complex and challenging.
-There are a significant number of components, many of which are described in \hyperref[sec2]{Section 2}, which must all be in place for work to be reproducibile.
+There are a significant number of components, many of which are described in \ref{sec:lit}, which must all be in place for work to be reproducibile.
 \pkg{fertile}, in its current version, addresses the vast majority of these in some way.
@@
 \begin{itemize}
-  \item Expanding dependency management features to include R session information and package version numbers in addition to package names.
+  \item Expanding dependency management features to include \R session information and package version numbers in addition to package names.
   \item Expanding code and documentation style features to analyze whether code has been properly commented in addition to checking for a \cmd{README} and tidy code style.
-  \item Adding \cmd{make}-like functionality that can analyze an R project structure and files and use this information to generate a Makefile.
+  \item Adding \cmd{make}-like functionality that can analyze an \R project structure and files and use this information to generate a Makefile.
 This Makefile would have information about target files and their prerequisites and would assist with making sure that re-running an analysis is done as quickly as possible by ensuring that only the necessary code and files that have been updated are run when rebuilding and re-running code.
 \end{itemize}
-
-\subsection{Testing Effectiveness}
-
-\pkg{fertile} is designed to:
-
-\begin{itemize}
-\item[1] Be simple enough that users with minimal R experience can use the package without issue,
-\item[2] Increase the reproducibility of work produced by its users, and
-\item[3] Educate its users on why their work is or is not reproducibile and provide guidance on how to address any problems.
-\end{itemize}
-
-Anecdotal evidence and user feedback is useful, providing a baseline knowledge of whether these goals are being met.
-However, in order to truly test whether the package is successful in achieving its design goals, an experiment is necessary.
-
-To test \pkg{fertile}'s effectiveness, we began an initial randomized control trial of the package on an introductory undergraduate data science course at Smith College in Spring 2020, comparing one group of students that received the package functions in their R environment to one that did not.
-Unfortunately, however, the trial had to be postponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing.
-
-We hope to continue this experimentation and gather data on \pkg{fertile}'s effectiveness once the higher education environment returns to a relatively stable state.
 

### 2020-07-15 STAT-paper-template/APA/fertile.Rnw
fixed non-reproducible path
switched from sweave to knitr

@@
 
-<<fertile, message = FALSE>>=
+<<fertile, message = FALSE, error = TRUE>>=
 library(fertile)
@@
 
-<<setwd, message=FALSE>>=
+<<setwd, message=FALSE, error = TRUE>>=
 setwd("~/Desktop")
@@
 <<set-project>>=
-proj_dir <- "~/Dropbox/git/baumer-lab/fertile/tests/testthat/project_miceps"
+proj_dir <- "project_miceps"
 @

### 2020-07-16 STAT-paper-template/APA/fertile.Rnw
addressing hadley comment

@@
 They are often difficult to learn and operate, providing a barrier to entry for less-experienced data analysts.
+This can make the goal of achieving reproducibility unnecessarily challenging, leading to users becoming discouraged and deciding that the effort necessary to succeed is not worth it.
 
-There appear to be no existing tools that cater to analysts looking for an easy-to-learn, easy-to-implement, fast, way to obtain a broad overview of their projects' reproducibility.
+A much more effective way to achieve widespread reproducibility is to make the process for doing so simple enough so that it takes next to no conscious effort to implement.
+However, there appear to be no existing tools that fit this model--catering to analysts looking for an easy-to-learn, easy-to-implement, fast, way to obtain a broad overview of their projects' reproducibility.
 To address this, we present an \R package called \pkg{fertile}, a low barrier-to-entry package which focuses on a series of common mistakes programmers make while conducting data science research in R.

### 2020-07-17 STAT-paper-template/APA/fertile.Rnw
added CI software to lit review

@@
 \item[3] Design the data analysis workflow as a sequence of small steps glued together, with
-outputs from one step serving as inputs into the next. This is a common suggestion within the computing community, originating as part of the Unix philosophy (Gancarz (2002)).
+outputs from one step serving as inputs into the next. This is a common suggestion within the computing community, originating as part of the Unix philosophy \cite{unix}.
 \end{itemize}
@@
 
-\subsection{R Packages}
+\subsection{R Packages and Other Software}
 
@@
 
+There are also several development tools external to R that focus on continuous integration and can be configured to address different aspects of reproducibility.
+\pkg{Wercker}, a command line tool that leverages Docker, enables users to test whether their projects will successfully compile when run on a variety of operating systems without access to the user's local hard drive \cite{wercker}.
+\pkg{Travis CI} is integrated with GitHub and can be configured to do similar checks on projects hosted in repositories \cite{travis}.
+\pkg{Circle CI}, another popular continuous integration tool, can also be used to check R code \cite{circle}.
+
+However, while these tools can be useful in some circumstances, they are generalized so as to be useful to the widest audience.
+As a result, their checks are not designed to be R-specific, which makes these tools sub-optimal for users looking to address reproducibility issues involving features specific to the R programming language, such as package installtion and seed setting.
+
 \section{Methods}\label{sec3}
@@
 \pkg{fertile} addresses these gaps by providing a simple, easy-to-learn reproducibility package that, rather than focusing intensely on a specific area, provides some information about a wide variety of aspects influencing reproducibility.
-The package also provides flexibility, offering benefits to users at any stage in the data analysis workflow, unlike some other available packages which can only be used before the creation of a new project or after it is finished.
+The package also provides flexibility, offering benefits to users at any stage in the data analysis workflow, and R-specific features, which address certain aspects of reproducibility that can be missed by external project development software.
 

### 2020-07-17 STAT-paper-template/APA/fertile.Rnw
added file path checks subsection

@@
 
-<<proj_check_some>>=
+<<proj_check_some, warning = FALSE>>=
 proj_check_some(proj_dir, contains("paths"))
@@
 
-\subsubsection{File Type Checks}
+\subsubsection{File Paths}
+
+For users focused on file paths, the \func{check\_path()} function analyzes a vector of paths (or a single path) to determine whether there are any absolute paths or paths that lead outside the project directory.
+
+<<path checks, error = TRUE>>=
+
+# Path inside the directory
+check_path("project_miceps")
+
+# Absolute path (current working directory)
+check_path(getwd())
+
+# Path outside the directory
+check_path("../fertile.Rmd")
+
+@
+
+\subsubsection{File Types}
 
 There are several functions that can be used to check the type of a file:
-
-<<paths, message=FALSE>>=
-is_path_portable("~")
-@
 
@@
 
-\subsubsection{Package Features}
-
 The process of ensuring true reproducibility is highly complex and challenging.

### 2020-07-17 STAT-paper-template/APA/fertile.Rnw
wording changes in intro

@@
 Much of the discussion around reproducibility is also generalized, written to be applicable to users working with a variety of statistical software programs.
-Since all statistical software programs operate differently, generalized recommendations on reproducibility are often shallow and unspecific.
-While they provide useful guidelines, they can often be difficult to implement, particularly to new analysts who are unsure how to apply such recommendations within the software programs they are using.
+Because all statistical software programs operate differently, generalized recommendations regarding reproducibility are often shallow and vague.
+While they provide useful guidelines, they can often be difficult to implement, particularly to new analysts who are unsure of how to apply such recommendations within the software programs they are using.
 Thus, reproducibility recommendations tailored to specific software programs are more likely to be adopted.
 
-In this paper, we focus on reproducibility in the \R programming language with a concentration on projects that use the RStudio integrated development environment.
-\R is an ideal candidate for reproducibility recommendations due to the language's popularity for statistical analyses and the ease with which analysts can download and begin using the software.
+In this paper, we focus on reproducibility in the \R programming language, concentrating specifically on projects that use the RStudio integrated development environment.
+\R is an ideal candidate for reproducibility recommendations because of its accessibility and popularity as a tool for statistical analysis.
 
-A small body of papers and \R packages focusing on reproducibility in the \R community have been published.
-Much of this work is narrowly tailored, with each package effectively addressising a small component of reproducibility: file structure, modularization of code, version control, etc.
-Many existing reproducibility packages, due to their focused nature, succeed at their area of focus, but at a cost.
+A small body of \R packages which focus on reproducibility have been published in recent years.
+Much of this work is narrowly tailored, with each package effectively addressing a small component of reproducibility: file structure, modularization of code, version control, etc.
+These packages often  succeed in their area of focus, but at the cost of accessibility to a wider audience.
 They are often difficult to learn and operate, providing a barrier to entry for less-experienced data analysts.
-This can make the goal of achieving reproducibility unnecessarily challenging, leading to users becoming discouraged and deciding that the effort necessary to succeed is not worth it.
+This can lead users to become discouraged and abandon attempts to achieve reproducibility.
 
-A much more effective way to achieve widespread reproducibility is to make the process for doing so simple enough so that it takes next to no conscious effort to implement.
-However, there appear to be no existing tools that fit this model--catering to analysts looking for an easy-to-learn, easy-to-implement, fast, way to obtain a broad overview of their projects' reproducibility.
-To address this, we present an \R package called \pkg{fertile}, a low barrier-to-entry package which focuses on a series of common mistakes programmers make while conducting data science research in R.
+A much more effective way of realizing widespread reproducibility is to make the process for doing so simple enough that it takes little to no conscious effort to implement.
+However, there appear to be no existing tools that fit this model--catering to analysts looking for a simple, easy-to-implement, and fast way to obtain a broad overview of their projects' reproducibility.
+To address this, we present \pkg{fertile}, a low barrier-to-entry package focusing on common mistakes programmers make while conducting data science research in R.
 
@@
 
-Reproducibility is critically important to the advancement of knowledge in all fields of scientific research.
-Researchers are acknowledging this; publications and discussions focusing on reproducibility seem to have increased in frequency over the last several years (\cite{sep-scientific-reproducibility}, \cite{plos-biology}, \cite{bioessays-gosselin}, \cite{eisner-reproducibility}, \cite{engineering-reproducibility}).
+Reproducibility is essential to the advancement of knowledge in all fields of scientific research.
+Researchers acknowledge this; publications and discussions focusing on reproducibility appear to have increased in frequency over the last several years (\cite{sep-scientific-reproducibility}, \cite{plos-biology}, \cite{bioessays-gosselin}, \cite{eisner-reproducibility}, \cite{engineering-reproducibility}).
 
-Much of the available literature is focused on methods for achieving reproducibility in specific disciplines, and some of their recommendations are generalizable to all areas of scientific research.
+Much of the available literature is focused on methods for achieving reproducibility in specific disciplines and some of their recommendations are generalizable to all areas of scientific research.
 There is also work focusing specifically on reproducibility within the \R community.
 Some of this work is in the form of academic papers, while other literature comes from blog posts, websites, and \R packages, some of which are compiled by notable R-focused developers and researchers.
-In this section, we will discuss the reproducibility recommendations from both the formal and informal publications, which together influence the development of \pkg{fertile}.
+In this section, we will discuss the reproducibility recommendations from both the formal and informal publications, which together have influenced the development of \pkg{fertile}.
 
@@
 
-\cite{cooper2017guide} focuses on data analysis in \R and identifies similar important reproducibility components, while reiterating the need for clearly labeled, well-documented, and well-separated files.
+\cite{cooper2017guide} focuses on data analysis in \R and identifies a similar list of important reproducibility components and reiterates the need for clearly labeled, well-documented, and well-separated files.
 In addition, they recommend publishing a list of dependencies and using version control.

### 2020-07-18 STAT-paper-template/APA/fertile.Rnw
rearranged first few sections

@@
 
+Data-based research is considered fully reproducible when the requisite code and data files
+produce identical results when run by another analyst.
+
 As research is becoming increasingly data-driven, and because knowledge can be shared worldwide so rapidly, reproducibility is critical to the advancement of scientific knowledge.
-Data-based research cannot be fully reproducible unless the requisite code and data files
-produce identical results when run by another analyst.
+Academics around the world have recognized this, and publications and discusions addressing reproducibility appear to have increased in the last several years (\cite{sep-scientific-reproducibility}, \cite{plos-biology}, \cite{bioessays-gosselin}, \cite{eisner-reproducibility}, \cite{engineering-reproducibility}).
+
+Reproducible research has a wide variety of benefits.
 When researchers provide the code and data used for their work in a well-organized and reproducible format, readers are more easily able to determine the veracity of any findings by following the steps from raw data to conclusions.
-
-The creators of reproducible research can more easily receive more specific feedback (including bug fixes) on their work.
+The creators of reproducible research can also more easily receive more specific feedback (including bug fixes) on their work.
 Moreover, others interested in the research topic can use the code to apply the methods and ideas used in one project to their own work with minimal effort.
@@
 
-Much of the discussion around reproducibility is also generalized, written to be applicable to users working with a variety of statistical software programs.
-Because all statistical software programs operate differently, generalized recommendations regarding reproducibility are often shallow and vague.
-While they provide useful guidelines, they can often be difficult to implement, particularly to new analysts who are unsure of how to apply such recommendations within the software programs they are using.
-Thus, reproducibility recommendations tailored to specific software programs are more likely to be adopted.
+Existing attempts to address reproducibility in data science are often either too generalized, resulting in shallow and vague recommendations that are challenging to implement, or too specific, approaching one aspect of reproducibility well but doing so in a highly technical way failes to capture the bigger picture and creates challenges for inexperienced users.
 
-In this paper, we focus on reproducibility in the \R programming language, concentrating specifically on projects that use the RStudio integrated development environment.
-\R is an ideal candidate for reproducibility recommendations because of its accessibility and popularity as a tool for statistical analysis.
-
-A small body of \R packages which focus on reproducibility have been published in recent years.
-Much of this work is narrowly tailored, with each package effectively addressing a small component of reproducibility: file structure, modularization of code, version control, etc.
-These packages often  succeed in their area of focus, but at the cost of accessibility to a wider audience.
-They are often difficult to learn and operate, providing a barrier to entry for less-experienced data analysts.
-This can lead users to become discouraged and abandon attempts to achieve reproducibility.
-
-A much more effective way of realizing widespread reproducibility is to make the process for doing so simple enough that it takes little to no conscious effort to implement.
-However, there appear to be no existing tools that fit this model--catering to analysts looking for a simple, easy-to-implement, and fast way to obtain a broad overview of their projects' reproducibility.
-To address this, we present \pkg{fertile}, a low barrier-to-entry package focusing on common mistakes programmers make while conducting data science research in R.
+In order to address these challenges, we present \pkg{fertile}, a low barrier-to-entry package focusing on common reproducibility mistakes programmers make while conducting data science research in R.
 
@@
 
-Reproducibility is essential to the advancement of knowledge in all fields of scientific research.
-Researchers acknowledge this; publications and discussions focusing on reproducibility appear to have increased in frequency over the last several years (\cite{sep-scientific-reproducibility}, \cite{plos-biology}, \cite{bioessays-gosselin}, \cite{eisner-reproducibility}, \cite{engineering-reproducibility}).
+\subsection{Written Work}\label{papers}
 
-Much of the available literature is focused on methods for achieving reproducibility in specific disciplines and some of their recommendations are generalizable to all areas of scientific research.
-There is also work focusing specifically on reproducibility within the \R community.
-Some of this work is in the form of academic papers, while other literature comes from blog posts, websites, and \R packages, some of which are compiled by notable R-focused developers and researchers.
-In this section, we will discuss the reproducibility recommendations from both the formal and informal publications, which together have influenced the development of \pkg{fertile}.
+Papers and other published works focusing on reproducibility can be found in all areas of scientific research.
+However, as \cite{Goodman341ps12} argue, the language and conceptual framework of research reproducibility varies significantly across the sciences, and there are no clear standards on reproducibility agreed upon by the scientific community as a whole.
 
-\subsection{Previous Work}
+\pkg{fertile} considers recommendations from a variety of fields, attempting to determine the key aspects of reproducibility faced by scientists in different disciplines and address as many as possible in a single piece of software.
 
-\cite{Goodman341ps12} argue that the language and conceptual framework of research reproducibility varies across the sciences. There are no clear standards agreed upon across fields.
+Although reproducibility recommendations vary, further investigation identifies several common ideas across publications.
 
@@
 
-A small selection of \R packages work to address the issue of research reproducibility, although not all of them are available on CRAN.
+Much of the written work discussed in Section \ref{papers} is highly generalized, written to be applicable to users working with a variety of statistical software programs.
+Because all statistical software programs operate differently, these recommendations are inherently vague and difficult to implement, particularly to new analysts who are relatively unfamiliar with their software.
+
+Thus, attempts to address reproducibility that are specific to certain software programs are more likely to be successful.
+\R, due to its accessibility and popularity as a tool for statistical analysis, is an ideal candidate such recommendations.
+
+Acknowledging this, researchers and members of the \R community have created a small body of \R packages that focus on research reproducibility, although not all of them are available on CRAN:
 
@@
 
-Many of these packages focus on one specific area of reproducibility, and while they each succeed well at their intended goal, they are not necessarily the most practical option for users trying to address a wide variety of factors influencing reproducibility at once.
-
-Additionally, with the focused nature of these packages comes added complexity.
-The functions in these packages are often quite complex to use, and many steps must be completed to achieve the required reproducibility goal.
+Many of these packages are narrowly tailored, with each effectively addressing a small component of reproducibility: file structure, modularization of code, version control, etc.
+These packages often  succeed in their area of focus, but at the cost of accessibility to a wider audience.
+Their functions are often quite complex to use, and many steps must be completed to achieve the required reproducibility goal.
 This cumbersome nature means that most reproducibility packages currently available are not easily accessible to users near the beginning of their \R journey, nor particularly useful to those looking for quick and easy reproducibility checks.
 
-There are also several development tools external to R that focus on continuous integration and can be configured to address different aspects of reproducibility.
-\pkg{Wercker}, a command line tool that leverages Docker, enables users to test whether their projects will successfully compile when run on a variety of operating systems without access to the user's local hard drive \cite{wercker}.
+A much more effective way of realizing widespread reproducibility is to make the process for doing so simple enough that it takes little to no conscious effort to implement.
+
+Several continuous integration code development tools external to R attempt to do this, providing a way to address different aspects of reproducibility with minimal code.
+For example, \pkg{Wercker}, a command line tool that leverages Docker, enables users to test whether their projects will successfully compile when run on a variety of operating systems without access to the user's local hard drive \cite{wercker}.
 \pkg{Travis CI} is integrated with GitHub and can be configured to do similar checks on projects hosted in repositories \cite{travis}.
@@
 However, while these tools can be useful in some circumstances, they are generalized so as to be useful to the widest audience.
-As a result, their checks are not designed to be R-specific, which makes these tools sub-optimal for users looking to address reproducibility issues involving features specific to the R programming language, such as package installtion and seed setting.
+As a result, their checks are not designed to be R-specific, which makes them sub-optimal for users looking to address reproducibility issues involving features specific to the R programming language, such as package installtion and seed setting.
 
@@
 
-\pkg{fertile} addresses these gaps by providing a simple, easy-to-learn reproducibility package that, rather than focusing intensely on a specific area, provides some information about a wide variety of aspects influencing reproducibility.
-The package also provides flexibility, offering benefits to users at any stage in the data analysis workflow, and R-specific features, which address certain aspects of reproducibility that can be missed by external project development software.
+\pkg{fertile} attempts to address these gaps in existing software by providing a simple, easy-to-learn reproducibility package that, rather than focusing intensely on a specific area, provides some information about a wide variety of aspects influencing reproducibility.
+\pkg{fertile} also allows for flexibility, offering benefits to users at any stage in the data analysis workflow, and provides R-specific features, which address certain aspects of reproducibility that can be missed by external project development software.
 
@@
 
-\begin{biography}{\includegraphics[width=70pt,height=70pt]{audrey_bio}}{\textbf{Audrey Bertin} is an undergraduate student at Smith College intending on completing a Bachelor of Arts in Statistical and Data Sciences and Public Policy in Spring 2021.
+\begin{biography}{\includegraphics[width=70pt,height=70pt]{audrey_bio}}{\textbf{Audrey M.
+Bertin} is an undergraduate student at Smith College intending on completing a Bachelor of Arts in Statistical and Data Sciences and Public Policy in Spring 2021.
 Audrey is a recipient of Smith College's prestigious Zollman Scholarship and has participated in extensive research with faculty throughout her time at the college.

### 2020-07-19 STAT-paper-template/APA/fertile.Rnw
added to sample use cases

@@
 
+\pkg{fertile} is wide in scope.
+The package's simplicity enables users of any background to take advantage of its features, and its big-picture design gives \pkg{fertile} the potential to provide benefits across a variety of disciplines.
+
+For example, professors could integrate \pkg{fertile} into their data science curricula, giving students an understanding and awareness of reproducibility early in their careers that can positively impact the reproducibility of their future work.
+It could also be used by experienced analysts working collaboratively who are looking to promote a smoother exchange of feedback and ideas.
+Journal reviewers may also find the package beneficial, allowing them to gain a fast overview of whether paper submissions meet reproducibility guidelines.
+
+The sample use cases in this section consider \pkg{fertile}'s applicability to some of these scenarios in detail.
+
 \subsubsection{Introductory Data Science Student}

### 2020-07-19 STAT-paper-template/APA/fertile.Rnw
added experiment description

@@
 
-To test \pkg{fertile}'s effectiveness, we began an initial randomized control trial of the package on an introductory undergraduate data science course at Smith College in Spring 2020, comparing one group of students that received the package functions in their \R environment to one that did not.
+To test \pkg{fertile}'s effectiveness, we began an initial randomized control trial of the package on an introductory undergraduate data science course at Smith College in Spring 2020.
 Unfortunately, however, the trial had to be postponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing.
+\\
 
-We hope to continue this experimentation and gather data on \pkg{fertile}'s effectiveness once the higher education environment returns to a relatively stable state.
+The experiment was structured as follows:
+
+\begin{itemize}
+
+\item Students are given a form at the start of the semester asking whether they consent to participate in a study on data science education.
+In order to successfully consent, they must provide their system username, collected through the command \cmd{Sys.getenv("LOGNAME")}.
+Since these individuals each username belongs to are often easily identifiable, each email address is then transformed into a hexadecimal string via a hashing algorithm in to maintain privacy.
+
+\item These hexadecimal strings are then randomly assigned into equally sized groups, one experimental group that receives the features of \pkg{fertile} and one group that receives a control.
+
+\item The students are then asked to download a package titled the same as the course number of the class: \pkg{sds192}, which was custom created for the purpose of this trial.
+It leverages an \cmd{.onAttach} function to scan the R environment and collect the username of the user who is loading the package and run it through the same hashing algorithm as used previously.
+It then identifies whether that user belongs to the experimental or the control group.
+Depending on the group they are in, they receive a different version of the package on loading:
+
+\item The experimental group receives the basic \pkg{sds192} package, which consists of some data sets and RMarkdown templates necessary for completing homework assignments and projects in the class, but they also have \pkg{fertile} installed and loaded silently in the background.
+The package's proactive features are enabled without their knowledge, and they can receive warning messages when they use absolute or non-portable paths or attempt to change their working directory.
+The control group receives only the basic \pkg{sds192} package, including its data sets and RMarkdown templates.
+They do not have \pkg{fertile} installed and loaded in the background and do not receive proactive warning messages.
+All students from both groups then use their version of the package throughout the semester on a variety of projects.
+
+\item In order to test \pkg{fertile}'s effectiveness, both groups are given a short quiz on different components of reproducibility that are intended to be taught by \pkg{fertile} at both the beginning and end of the semester.
+Their scores are then compared to see whether one group learned more than the other group or whether their scores were essentially equivalent.
+
+\end{itemize}
+
+Although the experiment was unsuccessful in its first attempt, we hope to run the same trial again and gather data on \pkg{fertile}'s effectiveness once the higher education environment returns to a relatively stable state.
 

### 2020-07-21 STAT-paper-template/APA/fertile.Rnw
added shimming section

@@
 
+\subsubsection{How It Works}
+
+In order to produce \pkg{fertile}'s warnings and reports, we take advantage of a function manipulation technique called \cmd{shimming}, which consists of the following steps:
+
+\begin{itemize}
+\item[1] We begin by choosing a function that exists in the R universe that we are interested in recording user interaction with.
+Popular functions linked to only one package, such as \cmd{read\_csv()} from \pkg{readr}, are ideal candidates.
+\item[2] Then, we creating a function with an identical name to this function within our own package.
+\item[3] We write this new function so that it calls the original function, captures the arguments that are passed in, and saves information of which function was called and which arguments were passed to it to a hidden log file.
+This process looks the same to the user as if they were calling the original function.
+\end{itemize}
+
+\pkg{fertile} shims many common functions, including those that read in a variety of data types, write data, and load packages.
+This allows us to track functions that may commonly be misused.
+For example, any time a user calls \cmd{read\_csv()} in their code, \pkg{fertile} captures this use, identifies the type of path the user supplied to the function, and uses that information to determine whether a warning is necessary.
+This works both proactively and retroactively, as the shimmed functions written in \pkg{fertile} are activated both when the user is coding interactively and when a file containing code is rendered.
+
+In order to ensure that the \pkg{fertile} versions of functions are always favored over their original forms when called, we manipulate the \R \cmd{search()} path so that \pkg{fertile} is always located in the first position using a shim of the \cmd{library()} function.
+In the \pkg{fertile} version of \cmd{library()}, we detach \pkg{fertile} from the search path, load the requested package, and then re-attach \pkg{fertile}.
+This makes certain that when a user executes a command, \R will check \pkg{fertile} for a matching function before considering other packages, enabling \pkg{fertile} to successfully record--and therefore report on--user behavior.
+
 \subsection{Utility Functions}

### 2020-08-04 STAT-paper-template/APA/fertile.Rnw
close #6

@@
 
-Several continuous integration code development tools external to R attempt to do this, providing a way to address different aspects of reproducibility with minimal code.
-For example, \pkg{Wercker}, a command line tool that leverages Docker, enables users to test whether their projects will successfully compile when run on a variety of operating systems without access to the user's local hard drive \cite{wercker}.
-\pkg{Travis CI} is integrated with GitHub and can be configured to do similar checks on projects hosted in repositories \cite{travis}.
-\pkg{Circle CI}, another popular continuous integration tool, can also be used to check R code \cite{circle}.
+Several continuous integration code development tools external to \R attempt to do this, providing a way to address different aspects of reproducibility with minimal code.
+For example, \pkg{wercker}---a command line tool that leverages Docker---enables users to test whether their projects will successfully compile when run on a variety of operating systems without access to the user's local hard drive \cite{wercker}.
+GitHub Actions is integrated with GitHub and can be configured to do similar checks on projects hosted in repositories.
+\href{https://travis-ci.com}{Travis CI} and \href{https://circleci.com/product}{Circle CI} are popular continuous integration tools which can also be used to check \R code.
 
 However, while these tools can be useful in some circumstances, they are generalized so as to be useful to the widest audience.
-As a result, their checks are not designed to be R-specific, which makes them sub-optimal for users looking to address reproducibility issues involving features specific to the R programming language, such as package installtion and seed setting.
+As a result, their checks are not designed to be \R-specific, which makes them sub-optimal for users looking to address reproducibility issues involving features specific to the \R programming language, such as package installation and seed setting.
 

### 2020-08-04 STAT-paper-template/APA/fertile.Rnw
close #5

@@
 
-A much more effective way of realizing widespread reproducibility is to make the process for doing so simple enough that it takes little to no conscious effort to implement.
+A more effective way of realizing widespread reproducibility is to make the process for doing so simple enough that it takes little to no conscious effort to implement.
+You want users to "fall into a hole"\footnote{We paraphrase Hadley Wickham.} of good practice.
 

### 2020-08-04 STAT-paper-template/APA/fertile.Rnw
various edits

@@
 
-In order to produce \pkg{fertile}'s warnings and reports, we take advantage of a function manipulation technique called \cmd{shimming}, which consists of the following steps:
+Much of the functionality in \pkg{fertile} is achieved by writing \href{https://en.wikipedia.org/wiki/Shim_(computing)}{shims}.
+\pkg{fertile}'s shimmed functions intercept the user's commands and perform various logging and checking tasks before executing the desired function.
+Our process is:
 
-\begin{itemize}
-\item[1] We begin by choosing a function that exists in the R universe that we are interested in recording user interaction with.
-Popular functions linked to only one package, such as \cmd{read\_csv()} from \pkg{readr}, are ideal candidates.
-\item[2] Then, we creating a function with an identical name to this function within our own package.
-\item[3] We write this new function so that it calls the original function, captures the arguments that are passed in, and saves information of which function was called and which arguments were passed to it to a hidden log file.
-This process looks the same to the user as if they were calling the original function.
-\end{itemize}
+\begin{enumerate}
+\item Identify an \R function is likely to be involved in operations that may break reproducibility.
+Popular functions associated with only one package (e.g., \func{read\_csv} from \pkg{readr}) are ideal candidates.
+\item Create a function in \pkg{fertile} with the same name, that takes the same arguments (and always the dots \cmd{...}).
+\item Write this new function so that it: 1) captures any arguments; 2) logs the name of the function called; 3) perform any checks on these arguments; and finally 4) calls the original function with the original arguments.
+Except where warranted, the execution looks the same to the user as if they were calling the original function.
+\end{enumerate}
+Most shims are quite simple and look something like what is shown below for \func{read\_csv}.
+
+<<>>=
+fertile::read_csv
+@
 
 \pkg{fertile} shims many common functions, including those that read in a variety of data types, write data, and load packages.
-This allows us to track functions that may commonly be misused.
-For example, any time a user calls \cmd{read\_csv()} in their code, \pkg{fertile} captures this use, identifies the type of path the user supplied to the function, and uses that information to determine whether a warning is necessary.
 This works both proactively and retroactively, as the shimmed functions written in \pkg{fertile} are activated both when the user is coding interactively and when a file containing code is rendered.
 
-In order to ensure that the \pkg{fertile} versions of functions are always favored over their original forms when called, we manipulate the \R \cmd{search()} path so that \pkg{fertile} is always located in the first position using a shim of the \cmd{library()} function.
-In the \pkg{fertile} version of \cmd{library()}, we detach \pkg{fertile} from the search path, load the requested package, and then re-attach \pkg{fertile}.
-This makes certain that when a user executes a command, \R will check \pkg{fertile} for a matching function before considering other packages, enabling \pkg{fertile} to successfully record--and therefore report on--user behavior.
+In order to ensure that the \pkg{fertile} versions of functions ("shims") are always favored over their original namesakes ("mask") when called, \pkg{fertile} uses its own shims of the \func{library} and \func{require} functions to  manipulate the \R \func{search} path so that it is always located in the first position.
+In the \pkg{fertile} version of \func{library}, we detach \pkg{fertile} from the search path, load the requested package, and then re-attach \pkg{fertile}.
+This ensures that when a user executes a command, \R will check \pkg{fertile} for a matching function before considering other packages.
+
+While it is possible that this shifty behavior could lead to unintended consequences, our goal is to catch a good deal of problems before they become problematic.
+Users can easily disable \pkg{fertile} by detaching it, or not loading it in the first place.
 
@@
 
-For users focused on file paths, the \func{check\_path()} function analyzes a vector of paths (or a single path) to determine whether there are any absolute paths or paths that lead outside the project directory.
+For users focused on file paths, the \func{check\_path} function analyzes a vector of paths (or a single path) to determine whether there are any absolute paths or paths that lead outside the project directory.
 
@@
 
-\begin{itemize}
-\item[1] Be simple enough that users with minimal \R experience can use the package without issue,
-\item[2] Increase the reproducibility of work produced by its users, and
-\item[3] Educate its users on why their work is or is not reproducibile and provide guidance on how to address any problems.
-\end{itemize}
+\begin{enumerate}
+\item Be simple enough that users with minimal \R experience can use the package without issue,
+\item Increase the reproducibility of work produced by its users, and
+\item Educate its users on why their work is or is not reproducible and provide guidance on how to address any problems.
+\end{enumerate}
 
@@
 Unfortunately, however, the trial had to be postponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing.
-\\
 
@@
 In order to successfully consent, they must provide their system username, collected through the command \cmd{Sys.getenv("LOGNAME")}.
-Since these individuals each username belongs to are often easily identifiable, each email address is then transformed into a hexadecimal string via a hashing algorithm in to maintain privacy.
+To maintain privacy the results are then transformed into a hexadecimal string via the \func{md5} hashing function.
 
@@
 
-\item The students are then asked to download a package titled the same as the course number of the class: \pkg{sds192}, which was custom created for the purpose of this trial.
-It leverages an \cmd{.onAttach} function to scan the R environment and collect the username of the user who is loading the package and run it through the same hashing algorithm as used previously.
+\item The students are then asked to download a package called \pkg{sds192} (the course number and prefix), which was created for the purpose of this trial.
+It leverages an \func{.onAttach} function to scan the \R environment and collect the username of the user who is loading the package and run it through the same hashing algorithm as used previously.
 It then identifies whether that user belongs to the experimental or the control group.
-Depending on the group they are in, they receive a different version of the package on loading:
+Depending on the group they are in, they receive a different version of the package:
 
-\item The experimental group receives the basic \pkg{sds192} package, which consists of some data sets and RMarkdown templates necessary for completing homework assignments and projects in the class, but they also have \pkg{fertile} installed and loaded silently in the background.
-The package's proactive features are enabled without their knowledge, and they can receive warning messages when they use absolute or non-portable paths or attempt to change their working directory.
-The control group receives only the basic \pkg{sds192} package, including its data sets and RMarkdown templates.
-They do not have \pkg{fertile} installed and loaded in the background and do not receive proactive warning messages.
+\item The experimental group receives the basic \pkg{sds192} package, which consists of some data sets and \R Markdown templates necessary for completing homework assignments and projects in the class, but they also have \pkg{fertile} installed and loaded silently in the background.
+The package's proactive features are enabled, and thus, for example, users will receive warning messages when they use absolute or non-portable paths or attempt to change their working directory.
+The control group receives only the basic \pkg{sds192} package, including its data sets and \R Markdown templates.
 All students from both groups then use their version of the package throughout the semester on a variety of projects.
@@
 
-Although the experiment was unsuccessful in its first attempt, we hope to run the same trial again and gather data on \pkg{fertile}'s effectiveness once the higher education environment returns to a relatively stable state.
+Although the experiment was unsuccessful in its first attempt, we hope to run the same trial again and gather data on \pkg{fertile}'s effectiveness.
 

### 2020-08-04 STAT-paper-template/APA/fertile.Rnw
added IRB protocol #

@@
 
-To test \pkg{fertile}'s effectiveness, we began an initial randomized control trial of the package on an introductory undergraduate data science course at Smith College in Spring 2020.
+To test \pkg{fertile}'s effectiveness, we began an initial randomized control trial of the package on an introductory undergraduate data science course at Smith College in Spring 2020 \footnote{This study was approved by Smith College IRB, Protocol \#19-032}.
 Unfortunately, however, the trial had to be postponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing.

### 2020-08-05 STAT-paper-template/APA/fertile.Rnw
revision

@@
 produce identical results when run by another analyst. 
-
 As research is becoming increasingly data-driven, and because knowledge can be shared worldwide so rapidly, reproducibility is critical to the advancement of scientific knowledge.
-Academics around the world have recognized this, and publications and discusions addressing reproducibility appear to have increased in the last several years (\cite{sep-scientific-reproducibility}, \cite{plos-biology}, \cite{bioessays-gosselin}, \cite{eisner-reproducibility}, \cite{engineering-reproducibility}).
+Academics around the world have recognized this, and publications and discussions addressing reproducibility appear to have increased in the last several years (\cite{sep-scientific-reproducibility}, \cite{plos-biology}, \cite{bioessays-gosselin}, \cite{eisner-reproducibility}, \cite{engineering-reproducibility}).
 
@@
 
-Existing attempts to address reproducibility in data science are often either too generalized, resulting in shallow and vague recommendations that are challenging to implement, or too specific, approaching one aspect of reproducibility well but doing so in a highly technical way failes to capture the bigger picture and creates challenges for inexperienced users.
+Existing attempts to address reproducibility in data science are often either too generalized---resulting in shallow and vague recommendations that are challenging to implement---or too specific---approaching one aspect of reproducibility well but doing so in a highly technical way that fails to capture the bigger picture and creates challenges for inexperienced users.
 
@@
 
-\section{Literature Review}\label{sec:lit}
+\section{Related Work}\label{sec:lit}
 
-\subsection{Written Work}\label{papers}
+\subsection{Literature Review}\label{sec:papers}
 
-Papers and other published works focusing on reproducibility can be found in all areas of scientific research.
+Publications on reproducibility can be found in all areas of scientific research.
 However, as \cite{Goodman341ps12} argue, the language and conceptual framework of research reproducibility varies significantly across the sciences, and there are no clear standards on reproducibility agreed upon by the scientific community as a whole.
 
-\pkg{fertile} considers recommendations from a variety of fields, attempting to determine the key aspects of reproducibility faced by scientists in different disciplines and address as many as possible in a single piece of software.
-
-Although reproducibility recommendations vary, further investigation identifies several common ideas across publications.
+We consider recommendations from a variety of fields, and determine the key aspects of reproducibility faced by scientists in different disciplines.
 
@@
 
-\begin{itemize}
-\item[1] Use clear separation, labeling, and documentation.
-\item[2] Automate processes when possible.
-\item[3] Design the data analysis workflow as a sequence of small steps glued together, with
-outputs from one step serving as inputs into the next. This is a common suggestion within the computing community, originating as part of the Unix philosophy \cite{unix}.
-\end{itemize}
+\begin{enumerate}
+\item Use clear separation, labeling, and documentation.
+\item Automate processes when possible.
+\item Design the data analysis workflow as a sequence of small steps glued together, with
+outputs from one step serving as inputs into the next. This is a common suggestion within the computing community, originating as part of the Unix philosophy (\cite{unix}).
+\end{enumerate}
 
@@
 In addition, they recommend publishing a list of dependencies and using version control.
-
-Broman reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a \cmd{README} (\cite{broman}).
+\cite{broman} reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a \cmd{README}.
 
@@
 When considered in combination, these sources provide a well-rounded picture of the components important to research reproducibility.
-Using this literature as a guideline, we identify several key features of reproducible work to focus on.
+Using this literature as a guideline, we identify several key features of reproducible work.
 These recommendations are opinionated---due to the lack of agreement on which components of reproducibility are most important, we select those that are mentioned most often, as well as some that are mentioned less but that we view as important.
 
-\begin{itemize}
-\item[1] A well-designed file structure:
+\begin{enumerate}
+\item A well-designed file structure:
   \begin{itemize}
@@
   \end{itemize}
-\item[2] Good documentation:
+\item Good documentation:
   \begin{itemize}
@@
   \end{itemize}
-\item[3] Reproducible file paths:
+\item Reproducible file paths:
   \begin{itemize}
-  \item No absolute paths, or paths leading to locations outside of a project's directory, are used in code. Only portable (relative) paths.
+  \item No absolute paths, or paths leading to locations outside of a project's directory, are used in code---only portable (relative) paths.
   \end{itemize}
-\item[4] Randomness is accounted for:
+\item Randomness is accounted for:
   \begin{itemize}
@@
   \end{itemize}
-\item[5] Code conforms to \pkg{tidyverse} style:
+\item Readable, styled code:
   \begin{itemize}
-  \item Although not discussed in any of the literature mentioned previously, we believe \pkg{tidyverse} integration to be very important.
-Code conformation to \pkg{tidyverse} style helps ensure that analyses meeting \pkg{fertile}'s recommendations will be able to integrate seamlessly with packages from the \pkg{tidyverse}, an ever-expanding collection of packages designed to operate together with one another.
+  \item Code should be written in a coherent style.
+Code that conforms to a style guide or is written in a consistent dialect is easier to read (\cite{hermans2017programming}).
+We believe that the \pkg{tidyverse} provides the most accessible dialect of \R.
   \end{itemize}
-\end{itemize}
+\end{enumerate}
 
 Much of the available literature focuses on file structure, organization, and naming, and \pkg{fertile}'s features are consistent with this. 
-The ideal file structure is not agreed upon by academics, though there are some recommendations focused on \R that are available.
-\cite{marwick2018packaging} provides the framework for file structure that \pkg{fertile} is based on: a structure similar to that of an \R package (\cite{hadley-packages}, \cite{coreteam-extensions}), with an \cmd{R} folder, as well as \cmd{data}, \cmd{data-raw}, \cmd{inst}, and \cmd{vignettes}.
+\cite{marwick2018packaging} provide the framework for file structure that \pkg{fertile} is based on: a structure similar to that of an \R package (\cite{hadley-packages}, \cite{coreteam-extensions}), with an \cmd{R} folder, as well as \cmd{data}, \cmd{data-raw}, \cmd{inst}, and \cmd{vignettes}.
 
@@
 
-Much of the written work discussed in Section \ref{papers} is highly generalized, written to be applicable to users working with a variety of statistical software programs.
+Much of the work discussed in Section \ref{sec:papers} is highly generalized, written to be applicable to users working with a variety of statistical software programs.
 Because all statistical software programs operate differently, these recommendations are inherently vague and difficult to implement, particularly to new analysts who are relatively unfamiliar with their software.
 
-Thus, attempts to address reproducibility that are specific to certain software programs are more likely to be successful.
-\R, due to its accessibility and popularity as a tool for statistical analysis, is an ideal candidate such recommendations.
+Focused attempts to address reproducibility in specific certain software programs are more likely to be successful. 
+We focus on \R, due to its open-source nature, accessibility, and popularity as a tool for statistical analysis.
 
-Acknowledging this, researchers and members of the \R community have created a small body of \R packages that focus on research reproducibility, although not all of them are available on CRAN:
+A small body of \R packages focus on research reproducibility.
 
@@
 
-Many of these packages are narrowly tailored, with each effectively addressing a small component of reproducibility: file structure, modularization of code, version control, etc.
+Many of these packages are narrow, with each effectively addressing a small component of reproducibility: file structure, modularization of code, version control, etc.
 These packages often  succeed in their area of focus, but at the cost of accessibility to a wider audience.
@@
 This cumbersome nature means that most reproducibility packages currently available are not easily accessible to users near the beginning of their \R journey, nor particularly useful to those looking for quick and easy reproducibility checks.
-
 A more effective way of realizing widespread reproducibility is to make the process for doing so simple enough that it takes little to no conscious effort to implement.
@@
 
-Several continuous integration code development tools external to \R attempt to do this, providing a way to address different aspects of reproducibility with minimal code.
-For example, \pkg{wercker}---a command line tool that leverages Docker---enables users to test whether their projects will successfully compile when run on a variety of operating systems without access to the user's local hard drive \cite{wercker}.
-GitHub Actions is integrated with GitHub and can be configured to do similar checks on projects hosted in repositories.
+\emph{Continuous integration} tools provide more general approaches to automated checking; which can enhance reproducibility with minimal code.
+For example, \pkg{wercker}---a command line tool that leverages Docker---enables users to test whether their projects will successfully compile when run on a variety of operating systems without access to the user's local hard drive (\cite{wercker}).
+GitHub Actions is integrated into GitHub and can be configured to do similar checks on projects hosted in repositories.
 \href{https://travis-ci.com}{Travis CI} and \href{https://circleci.com/product}{Circle CI} are popular continuous integration tools which can also be used to check \R code.
 
-However, while these tools can be useful in some circumstances, they are generalized so as to be useful to the widest audience.
+However, while these tools can be useful, they are generalized so as to be useful to the widest audience.
 As a result, their checks are not designed to be \R-specific, which makes them sub-optimal for users looking to address reproducibility issues involving features specific to the \R programming language, such as package installation and seed setting.
 
-\section{Methods}\label{sec3}
+\subsection{Our contribution}
 
 \pkg{fertile} attempts to address these gaps in existing software by providing a simple, easy-to-learn reproducibility package that, rather than focusing intensely on a specific area, provides some information about a wide variety of aspects influencing reproducibility.
-\pkg{fertile} also allows for flexibility, offering benefits to users at any stage in the data analysis workflow, and provides R-specific features, which address certain aspects of reproducibility that can be missed by external project development software.
+\pkg{fertile} is flexible, offering benefits to users at any stage in the data analysis workflow, and provides \R-specific features, which address certain aspects of reproducibility that can be missed by external project development software.
 
@@
 \pkg{fertile} achieves this by operating in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that has already been written for potential problems).
+
+\section{Methods}\label{sec3}
 
@@
 library(fertile)
-search()
 file.exists("~/Desktop/my_data.csv")
 read.csv("~/Desktop/my_data.csv")
+read.csv("../../../Desktop/my_data.csv")
 @
@@
 Users can access the majority of \pkg{fertile}'s retroactive features through two primary functions, \func{proj\_check} and \func{proj\_analyze}.
-
 The \func{proj\_check} function runs fifteen different reproducibility tests, noting which ones passed, which ones failed, the reason for failure, a recommended solution, and a guide to where to look for help.
-These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that to not have a call to \func{set.seed} present.
+These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that do not have a call to \func{set.seed} present.
 A full list is provided below:
@@
 
+\subsection{Logging}
+
 \pkg{fertile} also contains logging functionality, which records commands run in the console that have the potential to affect reproducibility, enabling users to look at their past history at any time.
@@
 
-<<log-clear>>=
+<<log-clear, include=FALSE>>=
 log_clear()
@@
 
-\subsubsection{How It Works}
+\subsection{How It Works}
 
@@
 \begin{enumerate}
-\item Identify an \R function is likely to be involved in operations that may break reproducibility.
+\item Identify an \R function that is likely to be involved in operations that may break reproducibility.
 Popular functions associated with only one package (e.g., \func{read\_csv} from \pkg{readr}) are ideal candidates.
-\item Create a function in \pkg{fertile} with the same name, that takes the same arguments (and always the dots \cmd{...}).
+\item Create a function in \pkg{fertile} with the same name that takes the same arguments (and always the dots \cmd{...}).
 \item Write this new function so that it: 1) captures any arguments; 2) logs the name of the function called; 3) perform any checks on these arguments; and finally 4) calls the original function with the original arguments.
@@
 
-In order to ensure that the \pkg{fertile} versions of functions ("shims") are always favored over their original namesakes ("mask") when called, \pkg{fertile} uses its own shims of the \func{library} and \func{require} functions to  manipulate the \R \func{search} path so that it is always located in the first position.
+In order to ensure that the \pkg{fertile} versions of functions ("shims") are always favored ("mask") over their original namesakes when called, \pkg{fertile} uses its own shims of the \func{library} and \func{require} functions to  manipulate the \R \func{search} path so that it is always located in the first position.
 In the \pkg{fertile} version of \func{library}, we detach \pkg{fertile} from the search path, load the requested package, and then re-attach \pkg{fertile}.
@@
 
-For users focused on file paths, the \func{check\_path} function analyzes a vector of paths (or a single path) to determine whether there are any absolute paths or paths that lead outside the project directory.
+The \func{check\_path} function analyzes a vector of paths (or a single path) to determine whether there are any absolute paths or paths that lead outside the project directory.
 
@@
 dir
-fs::dir_ls(dir) %>% head()
+fs::dir_ls(dir) %>% head(3)
 new_dir
-fs::dir_ls(new_dir) %>% head()
+fs::dir_ls(new_dir) %>% head(3)
 @
@@
 
-\pkg{fertile} is wide in scope.
-The package's simplicity enables users of any background to take advantage of its features, and its big-picture design gives \pkg{fertile} the potential to provide benefits across a variety of disciplines.
+\pkg{fertile}'s simplicity enables users of any background to take advantage of its features, and its big-picture design gives \pkg{fertile} the potential to provide benefits across a variety of disciplines.
 
@@
 The next day, she receives an email from her professor saying that her assignment failed to compile and that she needs to make changes and try again.
-
 Susan doesn't understand why it didn't work on the professor's computer when it did on her own.
@@
 Emma is a post-doc, with several years of \R experience.
-She is familiar with some basic rules of reproducibility--file paths should always be relative and randomness should always be associated with a seed--but has never needed to pass any sort of reproducibility check before because her professors never emphasized that.
+She is familiar with some basic rules of reproducibility---file paths should always be relative and randomness should always be associated with a seed---but has never needed to pass any sort of reproducibility check before because her professors never emphasized that.
 
@@
 
-\pkg{fertile} is designed to:
-
-\begin{enumerate}
-\item Be simple enough that users with minimal \R experience can use the package without issue,
-\item Increase the reproducibility of work produced by its users, and
-\item Educate its users on why their work is or is not reproducible and provide guidance on how to address any problems.
-\end{enumerate}
-
-Anecdotal evidence and user feedback is useful, providing a baseline knowledge of whether these goals are being met.
-However, in order to truly test whether the package is successful in achieving its design goals, an experiment is necessary.
+\pkg{fertile} is designed to: 1) be simple enough that users with minimal \R experience can use the package without issue; 2) increase the reproducibility of work produced by its users, and; 3) educate its users on why their work is or is not reproducible and provide guidance on how to address any problems.
 
 To test \pkg{fertile}'s effectiveness, we began an initial randomized control trial of the package on an introductory undergraduate data science course at Smith College in Spring 2020 \footnote{This study was approved by Smith College IRB, Protocol \#19-032}.
-Unfortunately, however, the trial had to be postponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing.
+Unfortunately, the trial had to be postponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing.
+Although the experiment was unsuccessful in its first attempt, we hope to run the same trial again and gather data on \pkg{fertile}'s effectiveness.
 
@@
 It then identifies whether that user belongs to the experimental or the control group.
-Depending on the group they are in, they receive a different version of the package:
+Depending on the group they are in, they receive a different version of the package.
 
@@
 
-Although the experiment was unsuccessful in its first attempt, we hope to run the same trial again and gather data on \pkg{fertile}'s effectiveness.
-
 \section{Conclusion}\label{sec:conclusion}
@@
 \pkg{fertile} is an \R package that lowers barriers to reproducible data analysis projects in \R, providing a wide array of checks and suggestions addressing many different aspects of project reproducibility, including file organization, file path usage, documentation, and dependencies.
-
 \pkg{fertile} is meant to be educational, providing informative error messages that show how the ways in which the mistakes the user is making are problematic, as well as sharing recommendations on how to fix them.
@@
 
-In the moment, it often feels easiest to take a shortcut--to use an absolute path or change a working directory.
+In the moment, it often feels easiest to take a shortcut---to use an absolute path or change a working directory.
 However, when considering the long term path of a project, spending the extra time to improve reproducibility is worthwhile.
@@
 
-The process of ensuring true reproducibility is highly complex and challenging.
-There are a significant number of components, many of which are described in \ref{sec:lit}, which must all be in place for work to be reproducibile.
-\pkg{fertile}, in its current version, addresses the vast majority of these in some way.
+\pkg{fertile}, in its current version, addresses the vast majority of the aspects of reproducibility identified in Section \ref{sec:papers} in some way.
 However, there are several areas where further development to extend the available features of the package would be beneficial.

### 2020-08-05 STAT-paper-template/APA/fertile.Rnw
expanded results + other sections

@@
 
+Each test can also be run individually by calling the function matching its check name:
+
+<<individual checks, warning = FALSE>>=
+
+has_tidy_raw_data(proj_dir)
+
+@
+
 The \func{proj\_analyze} function creates a report documenting the structure of a data analysis project.
@@
 \item Create a function in \pkg{fertile} with the same name that takes the same arguments (and always the dots \cmd{...}).
-\item Write this new function so that it: 1) captures any arguments; 2) logs the name of the function called; 3) perform any checks on these arguments; and finally 4) calls the original function with the original arguments.
-Except where warranted, the execution looks the same to the user as if they were calling the original function.
+\item Write this new function so that it: 
+\begin{enumerate}
+\item Captures any arguments
+\item Logs the name of the function called
+\item Performs any checks on these arguments, and, finally:
+\item Calls the original function with the original arguments. Except where warranted, the execution looks the same to the user as if they were calling the original function.
+\end{enumerate}
 \end{enumerate}
@@
 
-In order to ensure that the \pkg{fertile} versions of functions ("shims") are always favored ("mask") over their original namesakes when called, \pkg{fertile} uses its own shims of the \func{library} and \func{require} functions to  manipulate the \R \func{search} path so that it is always located in the first position.
+In order to ensure that the \pkg{fertile} versions of functions ("shims") are always favored ("masked") over their original namesakes when called, \pkg{fertile} uses its own shims of the \func{library} and \func{require} functions to  manipulate the \R \func{search} path so that it is always located in the first position.
 In the \pkg{fertile} version of \func{library}, we detach \pkg{fertile} from the search path, load the requested package, and then re-attach \pkg{fertile}.
@@
 is_image_file(fs::path(proj_dir, "proteins_v_time.png"))
+is_text_file(fs::path(proj_dir, "README.md"))
+is_r_file(fs::path(proj_dir, "analysis.Rmd"))
 @
@@
 new_dir <- sandbox(dir)
+
 dir
 fs::dir_ls(dir) %>% head(3)
+
 new_dir
@@
 
-<<proj_pkg_script, eval=TRUE>>=
+<<proj_pkg_script, eval=TRUE, cache=TRUE>>=
 install_script <- proj_pkg_script(proj_dir)
@@
 
-\pkg{fertile} is designed to: 1) be simple enough that users with minimal \R experience can use the package without issue; 2) increase the reproducibility of work produced by its users, and; 3) educate its users on why their work is or is not reproducible and provide guidance on how to address any problems.
+\pkg{fertile} is designed to: 
+\begin{enumerate}
+\item Be simple enough that users with minimal \R experience can use the package without issue.
+\item Increase the reproducibility of work produced by its users, and:
+\item Educate its users on why their work is or is not reproducible and provide guidance on how to address any problems.
+\end{enumerate}
 
 To test \pkg{fertile}'s effectiveness, we began an initial randomized control trial of the package on an introductory undergraduate data science course at Smith College in Spring 2020 \footnote{This study was approved by Smith College IRB, Protocol \#19-032}.
-Unfortunately, the trial had to be postponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing.
-Although the experiment was unsuccessful in its first attempt, we hope to run the same trial again and gather data on \pkg{fertile}'s effectiveness.
 
@@
 
-\item In order to test \pkg{fertile}'s effectiveness, both groups are given a short quiz on different components of reproducibility that are intended to be taught by \pkg{fertile} at both the beginning and end of the semester.
+Both groups are given a short quiz on different components of reproducibility that are intended to be taught by \pkg{fertile} at both the beginning and end of the semester.
 Their scores are then compared to see whether one group learned more than the other group or whether their scores were essentially equivalent.
+Additionally, for every homework assignment submitted, the professor takes note of whether or not the project compiles successfully.
 
 \end{itemize}
+
+Based on the results, we could determine whether \pkg{fertile} was successful at achieving its intended goals:
+
+\begin{enumerate}
+
+\item A lack of notable difference between the \func{experimental} and \func{control} groups in terms of the number of code-related questions asked throughout the semester would indicate that \pkg{fertile} achieved its goal of simplicity.
+\item A higher average for the \func{experimental} group in terms of the number of homework assignments that compiled successfully would indicate that \pkg{fertile} was successful in increasing reproducibility.
+\item A greater increase over the semester in the reproducibility quiz scores for students in the \func{experimental} group compared with the \func{control} group would indicate that \pkg{fertile} achieved its goal of educating users on reproducibility.
+
+\end{enumerate}
+
+Success in achieving these goals would provide valuable evidence showing \pkg{fertile}'s benefit as tool to help educators introduce reproducibility concepts in the classrom.
+
+ Unfortunately, we were unable to complete the analysis as intended as the trial had to be postponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing.
+Although the experiment was unsuccessful in its first attempt, we hope to run the same trial again and gather data on \pkg{fertile}'s effectiveness.
 

### 2020-08-09 STAT-paper-template/APA/fertile.Rnw
condensed spacing

@@
 \usepackage{xspace}
+\usepackage{enumitem}
+\usepackage{titlesec}
 \graphicspath{{images/}}
+\titlespacing{\section}{0pt}{10pt plus 2pt minus 1pt}{10pt plus 2pt minus 1pt}
+\titlespacing{\subsection}{0pt}{7pt plus 2pt minus 1pt}{7pt plus 2pt minus 1pt}
+\titlespacing{\subsubsection}{0pt}{7pt plus 2pt minus 1pt}{7pt plus 2pt minus 1pt}
 
@@
 
-Data-based research is considered fully reproducible when the requisite code and data files
-produce identical results when run by another analyst. 
+Data-based research is considered fully reproducible when the requisite code and data files produce identical results when run by another analyst.
 As research is becoming increasingly data-driven, and because knowledge can be shared worldwide so rapidly, reproducibility is critical to the advancement of scientific knowledge.
@@
 Not only are the components of reproducible research up for discussion (e.g., need the software be open source?), but the corresponding recommendations for ensuring reproducibility also vary (e.g., should raw and processed data files be in separate directories?).
-
 Existing attempts to address reproducibility in data science are often either too generalized---resulting in shallow and vague recommendations that are challenging to implement---or too specific---approaching one aspect of reproducibility well but doing so in a highly technical way that fails to capture the bigger picture and creates challenges for inexperienced users.
@@
 However, as \cite{Goodman341ps12} argue, the language and conceptual framework of research reproducibility varies significantly across the sciences, and there are no clear standards on reproducibility agreed upon by the scientific community as a whole.
-
 We consider recommendations from a variety of fields, and determine the key aspects of reproducibility faced by scientists in different disciplines.
@@
 \cite{kitzes2017practice} present a collection of case studies on reproducibility practices from across the data-intensive sciences, illustrating a variety of recommendations and techniques for achieving reproducibility.
-Although the book does not come to a consensus on the exact standards of reproducibility that should be followed, several common trends and principles emerge from their case studies:
-
-\begin{enumerate}
-\item Use clear separation, labeling, and documentation.
-\item Automate processes when possible.
-\item Design the data analysis workflow as a sequence of small steps glued together, with
-outputs from one step serving as inputs into the next. This is a common suggestion within the computing community, originating as part of the Unix philosophy (\cite{unix}).
-\end{enumerate}
+Although the book does not come to a consensus on the exact standards of reproducibility that should be followed, several common trends and principles emerge from their case studies: 1) Use clear separation, labeling, and documentation, 2) Automate processes when possible, and 3) Design the data analysis workflow as a sequence of small steps glued together, with outputs from one step serving as inputs into the next.
+This is a common suggestion within the computing community, originating as part of the Unix philosophy (\cite{unix}).
 
@@
 In addition, they recommend publishing a list of dependencies and using version control.
+
 \cite{broman} reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a \cmd{README}.
@@
 These recommendations are opinionated---due to the lack of agreement on which components of reproducibility are most important, we select those that are mentioned most often, as well as some that are mentioned less but that we view as important.
+\vspace{3mm}
 
-\begin{enumerate}
+\begin{enumerate} [nolistsep]
 \item A well-designed file structure:
-  \begin{itemize}
+  \begin{itemize}[noitemsep]
   \item Separate folders for different file types.
@@
 \item Good documentation:
-  \begin{itemize}
+  \begin{itemize} [noitemsep]
   \item Files are clearly named, preferably in a way where the order in which they should be run is clear.
@@
 Because all statistical software programs operate differently, these recommendations are inherently vague and difficult to implement, particularly to new analysts who are relatively unfamiliar with their software.
-
 Focused attempts to address reproducibility in specific certain software programs are more likely to be successful.
@@
 A small body of \R packages focus on research reproducibility.
-
 \pkg{rrtools} (\cite{R-rrtools}) addresses some of the issues discussed in \cite{marwick2018packaging} by creating a basic \R package structure for a data analysis project and implementing a basic \func{testthat::check} functionality.
@@
 Users can access the majority of \pkg{fertile}'s retroactive features through two primary functions, \func{proj\_check} and \func{proj\_analyze}.
+
 The \func{proj\_check} function runs fifteen different reproducibility tests, noting which ones passed, which ones failed, the reason for failure, a recommended solution, and a guide to where to look for help.
@@
 
-Each test can also be run individually by calling the function matching its check name:
-
-<<individual checks, warning = FALSE>>=
-
-has_tidy_raw_data(proj_dir)
-
-@
+Each test can also be run individually by calling the function matching its check name.
 
@@
 
-<<proj_analyze>>=
+<<proj_analyze, out.width = 180>>=
 proj_analyze(proj_dir)
@@
 library(forcats)
-library(stringr)
 read_csv(fs::path(proj_dir, "mice.csv"))
@@
 
-\begin{enumerate}
+\begin{enumerate}[noitemsep]
 \item Identify an \R function that is likely to be involved in operations that may break reproducibility.
@@
 \item Create a function in \pkg{fertile} with the same name that takes the same arguments (and always the dots \cmd{...}).
-\item Write this new function so that it: 
-\begin{enumerate}
-\item Captures any arguments
-\item Logs the name of the function called
-\item Performs any checks on these arguments, and, finally:
-\item Calls the original function with the original arguments. Except where warranted, the execution looks the same to the user as if they were calling the original function.
-\end{enumerate}
+\item Write this new function so that it: A) Captures any arguments, B) Logs the name of the function called, C) Performs any checks on these arguments, and D) Calls the original function with the original arguments.
+Except where warranted, the execution looks the same to the user as if they were calling the original function.
 \end{enumerate}
@@
 This ensures that when a user executes a command, \R will check \pkg{fertile} for a matching function before considering other packages.
-
 While it is possible that this shifty behavior could lead to unintended consequences, our goal is to catch a good deal of problems before they become problematic.
@@
 
-<<path checks, error = TRUE>>=
-
+<<path checks, error = TRUE, message = FALSE>>=
 # Path inside the directory
@@
 check_path("../fertile.Rmd")
-
 @
@@
 is_data_file(fs::path(proj_dir, "mice.csv"))
-is_image_file(fs::path(proj_dir, "mice.csv"))
 is_image_file(fs::path(proj_dir, "proteins_v_time.png"))
@@
 <<sandbox>>=
-dir <- getwd()
-new_dir <- sandbox(dir)
-
-dir
-fs::dir_ls(dir) %>% head(3)
-
-new_dir
-fs::dir_ls(new_dir) %>% head(3)
+proj_dir
+fs::dir_ls(proj_dir) %>% head(1)
+temp_dir <- sandbox(proj_dir)
+temp_dir
+fs::dir_ls(temp_dir) %>% head(1)
 @
@@
 
-\pkg{fertile} is designed to: 
-\begin{enumerate}
-\item Be simple enough that users with minimal \R experience can use the package without issue.
-\item Increase the reproducibility of work produced by its users, and:
-\item Educate its users on why their work is or is not reproducible and provide guidance on how to address any problems.
-\end{enumerate}
+\pkg{fertile} is designed to: 1) Be simple enough that users with minimal \R experience can use the package without issue, 2) Increase the reproducibility of work produced by its users, and 3) Educate its users on why their work is or is not reproducible and provide guidance on how to address any problems.
 
@@
 
-\begin{itemize}
+\begin{itemize}[noitemsep]
 
@@
 
-Both groups are given a short quiz on different components of reproducibility that are intended to be taught by \pkg{fertile} at both the beginning and end of the semester.
+\item Both groups are given a short quiz on different components of reproducibility that are intended to be taught by \pkg{fertile} at both the beginning and end of the semester.
 Their scores are then compared to see whether one group learned more than the other group or whether their scores were essentially equivalent.
@@
 
-Based on the results, we could determine whether \pkg{fertile} was successful at achieving its intended goals:
-
-\begin{enumerate}
-
-\item A lack of notable difference between the \func{experimental} and \func{control} groups in terms of the number of code-related questions asked throughout the semester would indicate that \pkg{fertile} achieved its goal of simplicity.
-\item A higher average for the \func{experimental} group in terms of the number of homework assignments that compiled successfully would indicate that \pkg{fertile} was successful in increasing reproducibility.
-\item A greater increase over the semester in the reproducibility quiz scores for students in the \func{experimental} group compared with the \func{control} group would indicate that \pkg{fertile} achieved its goal of educating users on reproducibility.
-
-\end{enumerate}
-
+Based on the results, we could determine whether \pkg{fertile} was successful at achieving its intended goals: 1) A lack of notable difference between the \func{experimental} and \func{control} groups in terms of the number of code-related questions asked throughout the semester would indicate that \pkg{fertile} achieved its goal of simplicity, 2) A higher average for the \func{experimental} group in terms of the number of homework assignments that compiled successfully would indicate that \pkg{fertile} was successful in increasing reproducibility, and 3) A greater increase over the semester in the reproducibility quiz scores for students in the \func{experimental} group compared with the \func{control} group would indicate that \pkg{fertile} achieved its goal of educating users on reproducibility.
 Success in achieving these goals would provide valuable evidence showing \pkg{fertile}'s benefit as tool to help educators introduce reproducibility concepts in the classrom.
@@
 
-\begin{itemize}
+\begin{itemize} [noitemsep]
   \item Expanding dependency management features to include \R session information and package version numbers in addition to package names.

### 2020-08-09 STAT-paper-template/APA/fertile.tex
added missing .tex file

--- 2020-08-09 STAT-paper-template/APA/fertile.Rnw
@@
-\documentclass[APA,LATO1COL]{WileyNJD-v2}
+\documentclass[APA,LATO1COL]{WileyNJD-v2}\usepackage[]{graphicx}\usepackage[]{color}
+% maxwidth is the original width if it is less than linewidth
+% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
 \raggedbottom
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 
-<<write_csv, include=FALSE>>=
-readr::write_csv(mtcars, "~/Desktop/my_data.csv")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(fertile)}
+\hlkwd{file.exists}\hlstd{(}\hlstr{"~/Desktop/my_data.csv"}\hlstd{)}
+\end{alltt}
+\begin{verbatim}
+## [1] TRUE
+\end{verbatim}
+\begin{alltt}
+\hlkwd{read.csv}\hlstd{(}\hlstr{"~/Desktop/my_data.csv"}\hlstd{)}
+\end{alltt}
 
-<<fertile, message = FALSE, error = TRUE>>=
-library(fertile)
-file.exists("~/Desktop/my_data.csv")
-read.csv("~/Desktop/my_data.csv")
-read.csv("../../../Desktop/my_data.csv")
-@
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Detected absolute paths}}\begin{alltt}
+\hlkwd{read.csv}\hlstd{(}\hlstr{"../../../Desktop/my_data.csv"}\hlstd{)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Detected paths that lead outside the project directory}}\end{kframe}
+\end{knitrout}
 
@@
 
-<<setwd, message=FALSE, error = TRUE>>=
-setwd("~/Desktop")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{setwd}\hlstd{(}\hlstr{"~/Desktop"}\hlstd{)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: setwd() is likely to break reproducibility. Use here::here() instead.}}\end{kframe}
+\end{knitrout}
 
@@
 
-<<list-checks>>=
-list_checks()
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{list_checks}\hlstd{()}
+\end{alltt}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- The available checks in `fertile` are as follows: ---------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+##  [1] "has_tidy_media"          "has_tidy_images"        
+##  [3] "has_tidy_code"           "has_tidy_raw_data"      
+##  [5] "has_tidy_data"           "has_tidy_scripts"       
+##  [7] "has_readme"              "has_no_lint"            
+##  [9] "has_proj_root"           "has_no_nested_proj_root"
+## [11] "has_only_used_files"     "has_clear_build_chain"  
+## [13] "has_no_absolute_paths"   "has_only_portable_paths"
+## [15] "has_no_randomness"
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<set-project>>=
-proj_dir <- "project_miceps"
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{proj_dir} \hlkwb{<-} \hlstr{"project_miceps"}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
-<<echo=FALSE>>=
-bad_file <- fs::path(proj_dir, "install_proj_packages.R")
-if (file.exists(bad_file)) {
-  fs::file_delete(bad_file)
-}
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{proj_check_some}\hlstd{(proj_dir,} \hlkwd{contains}\hlstd{(}\hlstr{"paths"}\hlstd{))}
+\end{alltt}
 
-<<proj_check_some, warning = FALSE>>=
-proj_check_some(proj_dir, contains("paths"))
-@
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... ----------------------------------------------------- fertile 0.0.0.9027 --}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ------------------------------------------- fertile 0.0.0.9027 --}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks ----------------------------------- fertile 0.0.0.9027 --}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# v Checking for no absolute paths}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# v Checking for only portable paths}}\end{kframe}
+\begin{kframe}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks ---------------------------------------- fertile 0.0.0.9027 --}}\end{kframe}
+\begin{kframe}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# v Reproducibility checks passed: 2}}\end{kframe}
+\end{knitrout}
 
@@
 
-<<proj_analyze, out.width = 180>>=
-proj_analyze(proj_dir)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{proj_analyze}\hlstd{(proj_dir)}
+\end{alltt}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps ------------------- fertile 0.0.0.9027 --}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code ----------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+## # A tibble: 9 x 3
+##   package       N used_in                    
+##   <chr>     <int> <chr>                      
+## 1 broom         1 project_miceps/analysis.Rmd
+## 2 dplyr         1 project_miceps/analysis.Rmd
+## 3 ggplot2       1 project_miceps/analysis.Rmd
+## 4 purrr         1 project_miceps/analysis.Rmd
+## 5 readr         1 project_miceps/analysis.Rmd
+## 6 rmarkdown     1 project_miceps/analysis.Rmd
+## 7 skimr         1 project_miceps/analysis.Rmd
+## 8 stargazer     1 project_miceps/analysis.Rmd
+## 9 tidyr         1 project_miceps/analysis.Rmd
+\end{verbatim}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory ------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+##                      file   ext    size
+## 1 Estrogen_Receptors.docx  docx  10.97K
+## 2      citrate_v_time.png   png 188.08K
+## 3     proteins_v_time.png   png 378.45K
+## 4   Blot_data_updated.csv   csv  14.43K
+## 5      CS_data_redone.csv   csv   7.39K
+## 6                mice.csv   csv  14.33K
+## 7               README.md    md      39
+## 8            miceps.Rproj Rproj     204
+## 9            analysis.Rmd   Rmd   4.94K
+##                                                                      mime
+## 1 application/vnd.openxmlformats-officedocument.wordprocessingml.document
+## 2                                                               image/png
+## 3                                                               image/png
+## 4                                                                text/csv
+## 5                                                                text/csv
+## 6                                                                text/csv
+## 7                                                           text/markdown
+## 8                                                            text/rstudio
+## 9                                                         text/x-markdown
+\end{verbatim}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files ----------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+##                  path_rel    dir_rel
+## 1   Blot_data_updated.csv   data-raw
+## 2      CS_data_redone.csv   data-raw
+## 3 Estrogen_Receptors.docx inst/other
+## 4            analysis.Rmd  vignettes
+## 5      citrate_v_time.png inst/image
+## 6                mice.csv   data-raw
+## 7     proteins_v_time.png inst/image
+##                                                                                                cmd
+## 1     file_move('project_miceps/Blot_data_updated.csv', fs::dir_create('project_miceps/data-raw'))
+## 2        file_move('project_miceps/CS_data_redone.csv', fs::dir_create('project_miceps/data-raw'))
+## 3 file_move('project_miceps/Estrogen_Receptors.docx', fs::dir_create('project_miceps/inst/other'))
+## 4             file_move('project_miceps/analysis.Rmd', fs::dir_create('project_miceps/vignettes'))
+## 5      file_move('project_miceps/citrate_v_time.png', fs::dir_create('project_miceps/inst/image'))
+## 6                  file_move('project_miceps/mice.csv', fs::dir_create('project_miceps/data-raw'))
+## 7     file_move('project_miceps/proteins_v_time.png', fs::dir_create('project_miceps/inst/image'))
+\end{verbatim}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged --------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+## NULL
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<read-csv, message=FALSE, include=FALSE>>=
-library(purrr)
-library(forcats)
-read_csv(fs::path(proj_dir, "mice.csv"))
-@
-
-<<log-report>>=
-log_report()
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{log_report}\hlstd{()}
+\end{alltt}
+\begin{verbatim}
+## # A tibble: 3 x 4
+##   path         path_abs                             func     timestamp          
+##   <chr>        <chr>                                <chr>    <dttm>             
+## 1 package:pur~ <NA>                                 base::l~ 2020-08-09 15:36:09
+## 2 package:for~ <NA>                                 base::l~ 2020-08-09 15:36:09
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-08-09 15:36:09
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
 The log, if not managed, can grow very long over time. For users who do not desire such functionality, \func{log\_clear} provides a way to erase the log and start over.
-
-<<log-clear, include=FALSE>>=
-log_clear()
-log_report()
-@
 
@@
 
-<<>>=
-fertile::read_csv
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{fertile}\hlopt{::}\hlstd{read_csv}
+\end{alltt}
+\begin{verbatim}
+## function (file, ...) 
+## {
+##     if (interactive_log_on()) {
+##         log_push(file, "readr::read_csv")
+##         check_path_safe(file)
+##         readr::read_csv(file, ...)
+##     }
+## }
+## <bytecode: 0x7fc56efc82c8>
+## <environment: namespace:fertile>
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<path checks, error = TRUE, message = FALSE>>=
-# Path inside the directory
-check_path("project_miceps")
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{# Path inside the directory}
+\hlkwd{check_path}\hlstd{(}\hlstr{"project_miceps"}\hlstd{)}
+\end{alltt}
+\begin{verbatim}
+## # A tibble: 0 x 3
+## # ... with 3 variables: path <chr>, problem <chr>, solution <chr>
+\end{verbatim}
+\begin{alltt}
+\hlcom{# Absolute path (current working directory)}
+\hlkwd{check_path}\hlstd{(}\hlkwd{getwd}\hlstd{())}
+\end{alltt}
 
-# Absolute path (current working directory)
-check_path(getwd())
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Detected absolute paths}}\begin{alltt}
+\hlcom{# Path outside the directory}
+\hlkwd{check_path}\hlstd{(}\hlstr{"../fertile.Rmd"}\hlstd{)}
+\end{alltt}
 
-# Path outside the directory
-check_path("../fertile.Rmd")
-@
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Detected paths that lead outside the project directory}}\end{kframe}
+\end{knitrout}
 
@@
 
-<<is_file>>=
-is_data_file(fs::path(proj_dir, "mice.csv"))
-is_image_file(fs::path(proj_dir, "proteins_v_time.png"))
-is_text_file(fs::path(proj_dir, "README.md"))
-is_r_file(fs::path(proj_dir, "analysis.Rmd"))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{is_data_file}\hlstd{(fs}\hlopt{::}\hlkwd{path}\hlstd{(proj_dir,} \hlstr{"mice.csv"}\hlstd{))}
+\end{alltt}
+\begin{verbatim}
+## [1] TRUE
+\end{verbatim}
+\begin{alltt}
+\hlkwd{is_image_file}\hlstd{(fs}\hlopt{::}\hlkwd{path}\hlstd{(proj_dir,} \hlstr{"proteins_v_time.png"}\hlstd{))}
+\end{alltt}
+\begin{verbatim}
+## [1] TRUE
+\end{verbatim}
+\begin{alltt}
+\hlkwd{is_text_file}\hlstd{(fs}\hlopt{::}\hlkwd{path}\hlstd{(proj_dir,} \hlstr{"README.md"}\hlstd{))}
+\end{alltt}
+\begin{verbatim}
+## [1] TRUE
+\end{verbatim}
+\begin{alltt}
+\hlkwd{is_r_file}\hlstd{(fs}\hlopt{::}\hlkwd{path}\hlstd{(proj_dir,} \hlstr{"analysis.Rmd"}\hlstd{))}
+\end{alltt}
+\begin{verbatim}
+## [1] TRUE
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<sandbox>>=
-proj_dir
-fs::dir_ls(proj_dir) %>% head(1)
-temp_dir <- sandbox(proj_dir)
-temp_dir
-fs::dir_ls(temp_dir) %>% head(1)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{proj_dir}
+\end{alltt}
+\begin{verbatim}
+## [1] "project_miceps"
+\end{verbatim}
+\begin{alltt}
+\hlstd{fs}\hlopt{::}\hlkwd{dir_ls}\hlstd{(proj_dir)} \hlopt{%>%} \hlkwd{head}\hlstd{(}\hlnum{1}\hlstd{)}
+\end{alltt}
+\begin{verbatim}
+## project_miceps/Blot_data_updated.csv
+\end{verbatim}
+\begin{alltt}
+\hlstd{temp_dir} \hlkwb{<-} \hlkwd{sandbox}\hlstd{(proj_dir)}
+\hlstd{temp_dir}
+\end{alltt}
+\begin{verbatim}
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmp0zg9IT/project_miceps
+\end{verbatim}
+\begin{alltt}
+\hlstd{fs}\hlopt{::}\hlkwd{dir_ls}\hlstd{(temp_dir)} \hlopt{%>%} \hlkwd{head}\hlstd{(}\hlnum{1}\hlstd{)}
+\end{alltt}
+\begin{verbatim}
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmp0zg9IT/project_miceps/Blot_data_updated.csv
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<proj_pkg_script, eval=TRUE, cache=TRUE>>=
-install_script <- proj_pkg_script(proj_dir)
-cat(readChar(install_script, 1e5))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{install_script} \hlkwb{<-} \hlkwd{proj_pkg_script}\hlstd{(proj_dir)}
+\hlkwd{cat}\hlstd{(}\hlkwd{readChar}\hlstd{(install_script,} \hlnum{1e5}\hlstd{))}
+\end{alltt}
+\begin{verbatim}
+## # Run this script to install the required packages for this R project.
+## # Packages hosted on CRAN...
+## install.packages(c( 'broom', 'dplyr', 'ggplot2', 'purrr', 'readr', 'rmarkdown', 'skimr', 'stargazer', 'tidyr' ))
+## # Packages hosted on GitHub...
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 

### 2020-08-09 STAT-paper-template/APA/fertile.tex
deleted unnecessary "empty" files

@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-08-09 15:36:09
-## 2 package:for~ <NA>                                 base::l~ 2020-08-09 15:36:09
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-08-09 15:36:09
+## 1 package:pur~ <NA>                                 base::l~ 2020-08-09 15:41:03
+## 2 package:for~ <NA>                                 base::l~ 2020-08-09 15:41:03
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-08-09 15:41:03
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x7fc56efc82c8>
+## <bytecode: 0x7fa5dc340718>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmp0zg9IT/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpJ7VyPP/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmp0zg9IT/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpJ7VyPP/project_miceps/Blot_data_updated.csv
 \end{verbatim}

### 2020-08-09 STAT-paper-template/APA/fertile.Rnw
small edits to results format

@@
 
-Based on the results, we could determine whether \pkg{fertile} was successful at achieving its intended goals: 1) A lack of notable difference between the \func{experimental} and \func{control} groups in terms of the number of code-related questions asked throughout the semester would indicate that \pkg{fertile} achieved its goal of simplicity, 2) A higher average for the \func{experimental} group in terms of the number of homework assignments that compiled successfully would indicate that \pkg{fertile} was successful in increasing reproducibility, and 3) A greater increase over the semester in the reproducibility quiz scores for students in the \func{experimental} group compared with the \func{control} group would indicate that \pkg{fertile} achieved its goal of educating users on reproducibility.
-Success in achieving these goals would provide valuable evidence showing \pkg{fertile}'s benefit as tool to help educators introduce reproducibility concepts in the classrom.
+Based on the results, we could determine whether \pkg{fertile} was successful at achieving its intended goals.
+A lack of notable difference between the \emph{experimental} and \emph{control} groups in terms of the number of code-related questions asked throughout the semester would indicate that \pkg{fertile} achieved its goal of simplicity.
+A higher average for the \emph{experimental} group in terms of the number of homework assignments that compiled successfully would indicate that \pkg{fertile} was successful in increasing reproducibility.
+A greater increase over the semester in the reproducibility quiz scores for students in the \emph{experimental} group compared with the \emph{control} group would indicate that \pkg{fertile} achieved its goal of educating users on reproducibility.
+Success according to these guidelines would provide valuable evidence showing \pkg{fertile}'s benefit as tool to help educators introduce reproducibility concepts in the classrom.
 

### 2020-08-09 STAT-paper-template/APA/fertile.tex
small edits to results format

@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-08-09 15:41:03
-## 2 package:for~ <NA>                                 base::l~ 2020-08-09 15:41:03
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-08-09 15:41:03
+## 1 package:pur~ <NA>                                 base::l~ 2020-08-09 15:50:04
+## 2 package:for~ <NA>                                 base::l~ 2020-08-09 15:50:04
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-08-09 15:50:04
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x7fa5dc340718>
+## <bytecode: 0x7fa9df1d8a48>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpJ7VyPP/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpMRbavu/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpJ7VyPP/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpMRbavu/project_miceps/Blot_data_updated.csv
 \end{verbatim}

### 2020-08-11 STAT-paper-template/APA/fertile.Rnw
small edits

@@
 \cite{kitzes2017practice} present a collection of case studies on reproducibility practices from across the data-intensive sciences, illustrating a variety of recommendations and techniques for achieving reproducibility.
-Although the book does not come to a consensus on the exact standards of reproducibility that should be followed, several common trends and principles emerge from their case studies: 1) Use clear separation, labeling, and documentation, 2) Automate processes when possible, and 3) Design the data analysis workflow as a sequence of small steps glued together, with outputs from one step serving as inputs into the next.
+Although the book does not come to a consensus on the exact standards of reproducibility that should be followed, several common trends and principles emerge from their case studies: 1) use clear separation, labeling, and documentation, 2) automate processes when possible, and 3) design the data analysis workflow as a sequence of small steps glued together, with outputs from one step serving as inputs into the next.
 This is a common suggestion within the computing community, originating as part of the Unix philosophy (\cite{unix}).
@@
 In addition, they recommend publishing a list of dependencies and using version control.
-
 \cite{broman} reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a \cmd{README}.
@@
 The package considers several different aspects of the project which may influence reproducibility, including the directory structure, file paths, and whether randomness is used thoughtfully.
-
 The end products of these analyses are reproducibility reports summarizing a project's adherence to reproducibility standards and recommending remedies for where the project falls short.
@@
 \item Create a function in \pkg{fertile} with the same name that takes the same arguments (and always the dots \cmd{...}).
-\item Write this new function so that it: A) Captures any arguments, B) Logs the name of the function called, C) Performs any checks on these arguments, and D) Calls the original function with the original arguments.
+\item Write this new function so that it: a) captures any arguments, b) logs the name of the function called, c) performs any checks on these arguments, and d) calls the original function with the original arguments.
 Except where warranted, the execution looks the same to the user as if they were calling the original function.
@@
 
-In order to ensure that the \pkg{fertile} versions of functions ("shims") are always favored ("masked") over their original namesakes when called, \pkg{fertile} uses its own shims of the \func{library} and \func{require} functions to  manipulate the \R \func{search} path so that it is always located in the first position.
+In order to ensure that the \pkg{fertile} versions of functions ("shims") always supersede ("mask") their original namesakes when called, \pkg{fertile} uses its own shims of the \func{library} and \func{require} functions to  manipulate the \R \func{search} path so that it is always located in the first position.
 In the \pkg{fertile} version of \func{library}, we detach \pkg{fertile} from the search path, load the requested package, and then re-attach \pkg{fertile}.
@@
 proj_dir
-fs::dir_ls(proj_dir) %>% head(1)
+fs::dir_ls(proj_dir) %>% head(3)
 temp_dir <- sandbox(proj_dir)
 temp_dir
-fs::dir_ls(temp_dir) %>% head(1)
+fs::dir_ls(temp_dir) %>% head(3)
 @
@@
 
-\pkg{fertile} is designed to: 1) Be simple enough that users with minimal \R experience can use the package without issue, 2) Increase the reproducibility of work produced by its users, and 3) Educate its users on why their work is or is not reproducible and provide guidance on how to address any problems.
+\pkg{fertile} is designed to: 1) be simple enough that users with minimal \R experience can use the package without issue, 2) increase the reproducibility of work produced by its users, and 3) educate its users on why their work is or is not reproducible and provide guidance on how to address any problems.
 
@@
 
-Based on the results, we could determine whether \pkg{fertile} was successful at achieving its intended goals.
+Based on the results, we hope to determine whether \pkg{fertile} was successful at achieving its intended goals.
 A lack of notable difference between the \emph{experimental} and \emph{control} groups in terms of the number of code-related questions asked throughout the semester would indicate that \pkg{fertile} achieved its goal of simplicity.
@@
 A greater increase over the semester in the reproducibility quiz scores for students in the \emph{experimental} group compared with the \emph{control} group would indicate that \pkg{fertile} achieved its goal of educating users on reproducibility.
-Success according to these guidelines would provide valuable evidence showing \pkg{fertile}'s benefit as tool to help educators introduce reproducibility concepts in the classrom.
+Success according to these metrics would provide evidence showing \pkg{fertile}'s benefit as tool to help educators introduce reproducibility concepts in the classroom.
 

### 2020-08-11 STAT-paper-template/APA/fertile.tex
small edits

@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- The available checks in `fertile` are as follows: ---------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- The available checks in `fertile` are as follows: -------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ##  [1] "has_tidy_media"          "has_tidy_images"        
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... ----------------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... --------------------------------------------------------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ----------------------------------------------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks ----------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks --------------------------------------------------------- fertile 0.0.0.9027 --}}
 
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks ---------------------------------------- fertile 0.0.0.9027 --}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks -------------------------------------------------------------- fertile 0.0.0.9027 --}}\end{kframe}
 \begin{kframe}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps ------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps ----------------------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code ----------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code --------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 9 x 3
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory ------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
-##                      file   ext    size
-## 1 Estrogen_Receptors.docx  docx  10.97K
-## 2      citrate_v_time.png   png 188.08K
-## 3     proteins_v_time.png   png 378.45K
-## 4   Blot_data_updated.csv   csv  14.43K
-## 5      CS_data_redone.csv   csv   7.39K
-## 6                mice.csv   csv  14.33K
-## 7               README.md    md      39
-## 8            miceps.Rproj Rproj     204
-## 9            analysis.Rmd   Rmd   4.94K
-##                                                                      mime
-## 1 application/vnd.openxmlformats-officedocument.wordprocessingml.document
-## 2                                                               image/png
-## 3                                                               image/png
-## 4                                                                text/csv
-## 5                                                                text/csv
-## 6                                                                text/csv
-## 7                                                           text/markdown
-## 8                                                            text/rstudio
-## 9                                                         text/x-markdown
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory ----------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+## # A tibble: 9 x 4
+##   file               ext        size mime                                       
+##   <fs::path>         <chr> <fs::byt> <chr>                                      
+## 1 Estrogen_Receptor~ docx     10.97K application/vnd.openxmlformats-officedocum~
+## 2 citrate_v_time.png png     188.29K image/png                                  
+## 3 proteins_v_time.p~ png     377.95K image/png                                  
+## 4 Blot_data_updated~ csv      14.43K text/csv                                   
+## 5 CS_data_redone.csv csv       7.39K text/csv                                   
+## 6 mice.csv           csv      14.33K text/csv                                   
+## 7 README.md          md           39 text/markdown                              
+## 8 miceps.Rproj       Rproj       204 text/rstudio                               
+## 9 analysis.Rmd       Rmd       4.94K text/x-markdown
 \end{verbatim}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files ----------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
-##                  path_rel    dir_rel
-## 1   Blot_data_updated.csv   data-raw
-## 2      CS_data_redone.csv   data-raw
-## 3 Estrogen_Receptors.docx inst/other
-## 4            analysis.Rmd  vignettes
-## 5      citrate_v_time.png inst/image
-## 6                mice.csv   data-raw
-## 7     proteins_v_time.png inst/image
-##                                                                                                cmd
-## 1     file_move('project_miceps/Blot_data_updated.csv', fs::dir_create('project_miceps/data-raw'))
-## 2        file_move('project_miceps/CS_data_redone.csv', fs::dir_create('project_miceps/data-raw'))
-## 3 file_move('project_miceps/Estrogen_Receptors.docx', fs::dir_create('project_miceps/inst/other'))
-## 4             file_move('project_miceps/analysis.Rmd', fs::dir_create('project_miceps/vignettes'))
-## 5      file_move('project_miceps/citrate_v_time.png', fs::dir_create('project_miceps/inst/image'))
-## 6                  file_move('project_miceps/mice.csv', fs::dir_create('project_miceps/data-raw'))
-## 7     file_move('project_miceps/proteins_v_time.png', fs::dir_create('project_miceps/inst/image'))
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files --------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+## # A tibble: 7 x 3
+##   path_rel           dir_rel    cmd                                             
+##   <fs::path>         <fs::path> <chr>                                           
+## 1 Blot_data_updated~ data-raw   file_move('project_miceps/Blot_data_updated.csv~
+## 2 CS_data_redone.csv data-raw   file_move('project_miceps/CS_data_redone.csv', ~
+## 3 Estrogen_Receptor~ inst/other file_move('project_miceps/Estrogen_Receptors.do~
+## 4 analysis.Rmd       vignettes  file_move('project_miceps/analysis.Rmd', fs::di~
+## 5 citrate_v_time.png inst/image file_move('project_miceps/citrate_v_time.png', ~
+## 6 mice.csv           data-raw   file_move('project_miceps/mice.csv', fs::dir_cr~
+## 7 proteins_v_time.p~ inst/image file_move('project_miceps/proteins_v_time.png',~
 \end{verbatim}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged --------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged ------------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## NULL
@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-08-09 15:50:04
-## 2 package:for~ <NA>                                 base::l~ 2020-08-09 15:50:04
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-08-09 15:50:04
+## 1 package:pur~ <NA>                                 base::l~ 2020-08-11 16:52:15
+## 2 package:for~ <NA>                                 base::l~ 2020-08-11 16:52:15
+## 3 project_mic~ /home/bbaumer/Dropbox/git/fertile-p~ readr::~ 2020-08-11 16:52:16
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x7fa9df1d8a48>
+## <bytecode: 0x55d182f14268>
 ## <environment: namespace:fertile>
@@
 \begin{alltt}
-\hlstd{fs}\hlopt{::}\hlkwd{dir_ls}\hlstd{(proj_dir)} \hlopt{%>%} \hlkwd{head}\hlstd{(}\hlnum{1}\hlstd{)}
+\hlstd{fs}\hlopt{::}\hlkwd{dir_ls}\hlstd{(proj_dir)} \hlopt{%>%} \hlkwd{head}\hlstd{(}\hlnum{3}\hlstd{)}
 \end{alltt}
 \begin{verbatim}
-## project_miceps/Blot_data_updated.csv
+## project_miceps/Blot_data_updated.csv   project_miceps/CS_data_redone.csv      
+## project_miceps/Estrogen_Receptors.docx
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpMRbavu/project_miceps
+## /tmp/RtmpK9UftW/project_miceps
 \end{verbatim}
 \begin{alltt}
-\hlstd{fs}\hlopt{::}\hlkwd{dir_ls}\hlstd{(temp_dir)} \hlopt{%>%} \hlkwd{head}\hlstd{(}\hlnum{1}\hlstd{)}
+\hlstd{fs}\hlopt{::}\hlkwd{dir_ls}\hlstd{(temp_dir)} \hlopt{%>%} \hlkwd{head}\hlstd{(}\hlnum{3}\hlstd{)}
 \end{alltt}
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpMRbavu/project_miceps/Blot_data_updated.csv
+## /tmp/RtmpK9UftW/project_miceps/Blot_data_updated.csv
+## /tmp/RtmpK9UftW/project_miceps/CS_data_redone.csv
+## /tmp/RtmpK9UftW/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}

### 2020-08-11 STAT-paper-template/APA/fertile.Rnw
additional grammar edits

@@
 
-\corres{Audrey Bertin, Program in Statistical and Data Sciences, Smith College, Northampton, MA, 01063. \email{abertin@smith.edu}}
+\corres{Audrey Bertin, Program in Statistical and Data Sciences, Smith College, Northampton, MA. \email{abertin@smith.edu}}
 
@@
 However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation, and no clear consensus on standards of what constitutes reproducibility in published research.
-We focus on a series of common mistakes programmers make while conducting data science projects in R, primarily through the RStudio integrated development environment.
-\pkg{fertile} is an \R package that operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that is already written for potential problems).
-Furthermore, \pkg{fertile} is designed to educate the user about why the mistakes are problematic, and how to fix them.}
+We present \pkg{fertile}, an R package that focuses on a series of common mistakes programmers make while conducting data science projects in R, primarily through the RStudio integrated development environment.
+\pkg{fertile} operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that is already written for potential problems).
+Furthermore, \pkg{fertile} is designed to educate users on why their mistakes are problematic and how to fix them.}
 
-\keywords{reproducibility, statistical software, workflow, collaboration}
+\keywords{reproducibility, statistical software, workflow, collaboration, education}
 
@@
 As research is becoming increasingly data-driven, and because knowledge can be shared worldwide so rapidly, reproducibility is critical to the advancement of scientific knowledge.
-Academics around the world have recognized this, and publications and discussions addressing reproducibility appear to have increased in the last several years (\cite{sep-scientific-reproducibility}, \cite{plos-biology}, \cite{bioessays-gosselin}, \cite{eisner-reproducibility}, \cite{engineering-reproducibility}).
+Academics around the world have recognized this, and publications and discussions addressing reproducibility appear to have increased in the last several years (\cite{eisner-reproducibility}; \cite{sep-scientific-reproducibility}; \cite{bioessays-gosselin}; \cite{engineering-reproducibility}; \cite{plos-biology}).
 
@@
 
-However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation, and no clear consensus on standards of what constitutes reproducibility in published research (\cite{peng2009bio}).
+However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation and no clear consensus on standards of what constitutes reproducibility in published research (\cite{peng2009bio}).
 Not only are the components of reproducible research up for discussion (e.g., need the software be open source?), but the corresponding recommendations for ensuring reproducibility also vary (e.g., should raw and processed data files be in separate directories?).
@@
 However, as \cite{Goodman341ps12} argue, the language and conceptual framework of research reproducibility varies significantly across the sciences, and there are no clear standards on reproducibility agreed upon by the scientific community as a whole.
-We consider recommendations from a variety of fields, and determine the key aspects of reproducibility faced by scientists in different disciplines.
+We consider recommendations from a variety of fields and determine the key aspects of reproducibility faced by scientists in different disciplines.
 
 \cite{kitzes2017practice} present a collection of case studies on reproducibility practices from across the data-intensive sciences, illustrating a variety of recommendations and techniques for achieving reproducibility.
-Although the book does not come to a consensus on the exact standards of reproducibility that should be followed, several common trends and principles emerge from their case studies: 1) use clear separation, labeling, and documentation, 2) automate processes when possible, and 3) design the data analysis workflow as a sequence of small steps glued together, with outputs from one step serving as inputs into the next.
+Although their work does not come to a consensus on the exact standards of reproducibility that should be followed, several common trends and principles emerge from their case studies: 1) use clear separation, labeling, and documentation, 2) automate processes when possible, and 3) design the data analysis workflow as a sequence of small steps glued together, with outputs from one step serving as inputs into the next.
 This is a common suggestion within the computing community, originating as part of the Unix philosophy (\cite{unix}).
 
-\cite{cooper2017guide} focuses on data analysis in \R and identifies a similar list of important reproducibility components and reiterates the need for clearly labeled, well-documented, and well-separated files.
+\cite{cooper2017guide} focus on data analysis in \R and identify a similar list of important reproducibility components, reinforcing the need for clearly labeled, well-documented, and well-separated files.
 In addition, they recommend publishing a list of dependencies and using version control.
@@
 Using this literature as a guideline, we identify several key features of reproducible work.
-These recommendations are opinionated---due to the lack of agreement on which components of reproducibility are most important, we select those that are mentioned most often, as well as some that are mentioned less but that we view as important.
+These recommendations are a matter of opinion---due to the lack of agreement on which components of reproducibility are most important, we select those that are mentioned most often, as well as some that are mentioned less but that we view as important.
 \vspace{3mm}
@@
 Much of the available literature focuses on file structure, organization, and naming, and \pkg{fertile}'s features are consistent with this.
-\cite{marwick2018packaging} provide the framework for file structure that \pkg{fertile} is based on: a structure similar to that of an \R package (\cite{hadley-packages}, \cite{coreteam-extensions}), with an \cmd{R} folder, as well as \cmd{data}, \cmd{data-raw}, \cmd{inst}, and \cmd{vignettes}.
+\cite{marwick2018packaging} provide the framework for file structure that \pkg{fertile} is based on: a structure similar to that of an \R package (\cite{coreteam-extensions}, \cite{hadley-packages}), with an \cmd{R} folder, as well as \cmd{data}, \cmd{data-raw}, \cmd{inst}, and \cmd{vignettes}.
 
@@
 
-A small body of \R packages focus on research reproducibility.
+A small body of \R packages focuses on research reproducibility.
 \pkg{rrtools} (\cite{R-rrtools}) addresses some of the issues discussed in \cite{marwick2018packaging} by creating a basic \R package structure for a data analysis project and implementing a basic \func{testthat::check} functionality.
@@
 \pkg{checkers} (\cite{R-checkers}) allows you to create custom checks that examine different aspects of reproducibility.
-\pkg{packrat} (\cite{R-packrat}) is focused on dependencies, creating a packaged folder containing a project as well as all of its dependencies, so that projects dependent on lesser-used packages can be easily shared across computers.
-\pkg{drake} (\cite{R-drake}) analyzes workflows, skip steps where results are up to date, and provides evidence that results match the underlying code and data.
+\pkg{packrat} (\cite{R-packrat}) is focused on dependencies, creating a packaged folder containing a project as well as all of its dependencies so that projects dependent on lesser-used packages can be easily shared across computers.
+\pkg{drake} (\cite{R-drake}) analyzes workflows, skips steps where results are up to date, and provides evidence that results match the underlying code and data.
 Lastly, the \pkg{reproducible} (\cite{R-reproducible}) package focuses on the concept of caching: saving information so that projects can be run faster each time they are re-completed from the start.
@@
 
-\emph{Continuous integration} tools provide more general approaches to automated checking; which can enhance reproducibility with minimal code.
+\emph{Continuous integration} tools provide more general approaches to automated checking, which can enhance reproducibility with minimal code.
 For example, \pkg{wercker}---a command line tool that leverages Docker---enables users to test whether their projects will successfully compile when run on a variety of operating systems without access to the user's local hard drive (\cite{wercker}).
-GitHub Actions is integrated into GitHub and can be configured to do similar checks on projects hosted in repositories.
-\href{https://travis-ci.com}{Travis CI} and \href{https://circleci.com/product}{Circle CI} are popular continuous integration tools which can also be used to check \R code.
+\pkg{GitHub Actions} is integrated into GitHub and can be configured to do similar checks on projects hosted in repositories.
+\href{https://travis-ci.com}{\pkg{Travis CI}} and \href{https://circleci.com/product}{\pkg{Circle CI}} are popular continuous integration tools that can also be used to check \R code.
 
@@
 
-\pkg{fertile}'s simplicity enables users of any background to take advantage of its features, and its big-picture design gives \pkg{fertile} the potential to provide benefits across a variety of disciplines.
+\pkg{fertile}'s simplicity enables users of any background to take advantage of its features and its big-picture design gives \pkg{fertile} the potential to provide benefits across a variety of disciplines.
 
@@
 Susan is taking an introductory data science course.
-This is her first time learning how to code, and she has not yet been exposed to ideas of research reproducibility.
-Her professor has assigned a data analysis project which must be completed in \R Markdown.
+This is her first time learning how to code and she has not yet been exposed to ideas of research reproducibility.
+Her professor has assigned a data analysis project that must be completed in \R Markdown.
 The project requires her to read in a data file located on her computer and use it to produce a graph.
@@
 The next day, she receives an email from her professor saying that her assignment failed to compile and that she needs to make changes and try again.
-Susan doesn't understand why it didn't work on the professor's computer when it did on her own.
+Susan does not understand why it did not work on the professor's computer when it did on her own.
 The professor recommends that she install \pkg{fertile} and run \func{proj\_check} on her assignment.
-She does this and gets a message informing her she used an absolute path to open her dataset, and she should use a relative path.
+She does this and gets a message informing her that she used an absolute path to open her dataset when she should have use a relative path instead.
 She looks up what this means and then uses the new information to update her assignment.
-She resubmits and her second attempt is successful.
+Her second submission compiles successfuly.
 
-On future projects, she always loads and runs \pkg{fertile} to make sure her work is okay before submitting.
+On future projects, she always loads and runs \pkg{fertile} before submitting.
 
@@
 
-Emma is a post-doc, with several years of \R experience.
+Emma is a post-doc with several years of \R experience.
 She is familiar with some basic rules of reproducibility---file paths should always be relative and randomness should always be associated with a seed---but has never needed to pass any sort of reproducibility check before because her professors never emphasized that.
@@
 
-\item The experimental group receives the basic \pkg{sds192} package, which consists of some data sets and \R Markdown templates necessary for completing homework assignments and projects in the class, but they also have \pkg{fertile} installed and loaded silently in the background.
-The package's proactive features are enabled, and thus, for example, users will receive warning messages when they use absolute or non-portable paths or attempt to change their working directory.
+\item The experimental group receives the basic \pkg{sds192} package, which consists of some data sets and \R Markdown templates necessary for completing homework assignments and projects in the class, but also has \pkg{fertile} installed and loaded silently in the background.
+The package's proactive features are enabled, and therefore users will receive warning messages when they use absolute or non-portable paths or attempt to change their working directory.
 The control group receives only the basic \pkg{sds192} package, including its data sets and \R Markdown templates.
@@
 \pkg{fertile} is an \R package that lowers barriers to reproducible data analysis projects in \R, providing a wide array of checks and suggestions addressing many different aspects of project reproducibility, including file organization, file path usage, documentation, and dependencies.
-\pkg{fertile} is meant to be educational, providing informative error messages that show how the ways in which the mistakes the user is making are problematic, as well as sharing recommendations on how to fix them.
+\pkg{fertile} is meant to be educational, providing informative error messages that indicate why users' mistakes are problematic and sharing recommendations on how to fix them.
 The package is designed in this way so as to promote a greater understanding of reproducibility concepts in its users, with the goal of increasing the overall awareness and understanding of reproducibility in the \R community.
@@
 Unlike many other \R packages focused on reproducibility that are currently available, the features of \pkg{fertile} can be accessed almost effortlessly.
-Many of the retroactive features can be accessed in only two lines of code requiring minimal arguments, and some of the proactive features can be accessed with no additional effort beyond loading the package.
+Many of the retroactive features can be accessed in only two lines of code requiring minimal arguments and some of the proactive features can be accessed with no additional effort beyond loading the package.
 This, in combination with the fact that \pkg{fertile} does not focus on one specific area of reproducibility, instead covering (albeit in less detail) a wide variety of topics, means that \pkg{fertile} makes it easy for data analysts of all skill levels to quickly gain a better understanding of the reproducibility of the work.

### 2020-08-11 STAT-paper-template/APA/fertile.tex
additional grammar edits

@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- The available checks in `fertile` are as follows: -------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- The available checks in `fertile` are as follows: ------------------------ fertile 0.0.0.9027 --}}\begin{verbatim}
 ##  [1] "has_tidy_media"          "has_tidy_images"        
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... --------------------------------------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... ------------------------------------------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ----------------------------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... --------------------------------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks --------------------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks ------------------------------------------- fertile 0.0.0.9027 --}}
 
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks -------------------------------------------------------------- fertile 0.0.0.9027 --}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks ------------------------------------------------ fertile 0.0.0.9027 --}}\end{kframe}
 \begin{kframe}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps ----------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps --------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code --------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code ------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 9 x 3
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory ----------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
-## # A tibble: 9 x 4
-##   file               ext        size mime                                       
-##   <fs::path>         <chr> <fs::byt> <chr>                                      
-## 1 Estrogen_Receptor~ docx     10.97K application/vnd.openxmlformats-officedocum~
-## 2 citrate_v_time.png png     188.29K image/png                                  
-## 3 proteins_v_time.p~ png     377.95K image/png                                  
-## 4 Blot_data_updated~ csv      14.43K text/csv                                   
-## 5 CS_data_redone.csv csv       7.39K text/csv                                   
-## 6 mice.csv           csv      14.33K text/csv                                   
-## 7 README.md          md           39 text/markdown                              
-## 8 miceps.Rproj       Rproj       204 text/rstudio                               
-## 9 analysis.Rmd       Rmd       4.94K text/x-markdown
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory --------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+##                      file   ext    size
+## 1 Estrogen_Receptors.docx  docx  10.97K
+## 2      citrate_v_time.png   png 188.08K
+## 3     proteins_v_time.png   png 378.45K
+## 4   Blot_data_updated.csv   csv  14.43K
+## 5      CS_data_redone.csv   csv   7.39K
+## 6                mice.csv   csv  14.33K
+## 7               README.md    md      39
+## 8            miceps.Rproj Rproj     204
+## 9            analysis.Rmd   Rmd   4.94K
+##                                                                      mime
+## 1 application/vnd.openxmlformats-officedocument.wordprocessingml.document
+## 2                                                               image/png
+## 3                                                               image/png
+## 4                                                                text/csv
+## 5                                                                text/csv
+## 6                                                                text/csv
+## 7                                                           text/markdown
+## 8                                                            text/rstudio
+## 9                                                         text/x-markdown
 \end{verbatim}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files --------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
-## # A tibble: 7 x 3
-##   path_rel           dir_rel    cmd                                             
-##   <fs::path>         <fs::path> <chr>                                           
-## 1 Blot_data_updated~ data-raw   file_move('project_miceps/Blot_data_updated.csv~
-## 2 CS_data_redone.csv data-raw   file_move('project_miceps/CS_data_redone.csv', ~
-## 3 Estrogen_Receptor~ inst/other file_move('project_miceps/Estrogen_Receptors.do~
-## 4 analysis.Rmd       vignettes  file_move('project_miceps/analysis.Rmd', fs::di~
-## 5 citrate_v_time.png inst/image file_move('project_miceps/citrate_v_time.png', ~
-## 6 mice.csv           data-raw   file_move('project_miceps/mice.csv', fs::dir_cr~
-## 7 proteins_v_time.p~ inst/image file_move('project_miceps/proteins_v_time.png',~
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files ------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+##                  path_rel    dir_rel
+## 1   Blot_data_updated.csv   data-raw
+## 2      CS_data_redone.csv   data-raw
+## 3 Estrogen_Receptors.docx inst/other
+## 4            analysis.Rmd  vignettes
+## 5      citrate_v_time.png inst/image
+## 6                mice.csv   data-raw
+## 7     proteins_v_time.png inst/image
+##                                                                                                cmd
+## 1     file_move('project_miceps/Blot_data_updated.csv', fs::dir_create('project_miceps/data-raw'))
+## 2        file_move('project_miceps/CS_data_redone.csv', fs::dir_create('project_miceps/data-raw'))
+## 3 file_move('project_miceps/Estrogen_Receptors.docx', fs::dir_create('project_miceps/inst/other'))
+## 4             file_move('project_miceps/analysis.Rmd', fs::dir_create('project_miceps/vignettes'))
+## 5      file_move('project_miceps/citrate_v_time.png', fs::dir_create('project_miceps/inst/image'))
+## 6                  file_move('project_miceps/mice.csv', fs::dir_create('project_miceps/data-raw'))
+## 7     file_move('project_miceps/proteins_v_time.png', fs::dir_create('project_miceps/inst/image'))
 \end{verbatim}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged ------------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged ----------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## NULL
@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-08-11 16:52:15
-## 2 package:for~ <NA>                                 base::l~ 2020-08-11 16:52:15
-## 3 project_mic~ /home/bbaumer/Dropbox/git/fertile-p~ readr::~ 2020-08-11 16:52:16
+## 1 package:pur~ <NA>                                 base::l~ 2020-08-11 18:41:48
+## 2 package:for~ <NA>                                 base::l~ 2020-08-11 18:41:48
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-08-11 18:41:48
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x55d182f14268>
+## <bytecode: 0x7fb540270948>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /tmp/RtmpK9UftW/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpuJZ1kZ/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /tmp/RtmpK9UftW/project_miceps/Blot_data_updated.csv
-## /tmp/RtmpK9UftW/project_miceps/CS_data_redone.csv
-## /tmp/RtmpK9UftW/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpuJZ1kZ/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpuJZ1kZ/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpuJZ1kZ/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}

### 2020-08-11 STAT-paper-template/APA/fertile.tex
restructuring for submission

@@
 \usepackage{listings}
+\usepackage{NJDapacite}
 \usepackage{hyperref}
@@
 \usepackage{titlesec}
-\graphicspath{{images/}}
 \titlespacing{\section}{0pt}{10pt plus 2pt minus 1pt}{10pt plus 2pt minus 1pt}
@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-08-11 18:41:48
-## 2 package:for~ <NA>                                 base::l~ 2020-08-11 18:41:48
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-08-11 18:41:48
+## 1 package:pur~ <NA>                                 base::l~ 2020-08-11 21:28:23
+## 2 package:for~ <NA>                                 base::l~ 2020-08-11 21:28:23
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-08-11 21:28:23
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x7fb540270948>
+## <bytecode: 0x7f8f8c998438>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpuJZ1kZ/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpGDaHPo/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpuJZ1kZ/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpuJZ1kZ/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpuJZ1kZ/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpGDaHPo/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpGDaHPo/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpGDaHPo/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}
@@
 
+\section*{Data Availability Statement}
+The sample project ("project\_miceps") and package code associated with this paper can be found in the "R" and "tests" folders at \href{https://github.com/baumer-lab/fertile}{https://github.com/baumer-lab/fertile}.
+
 \nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;

### 2020-08-18 STAT-paper-template/APA/fertile.Rnw
restructuring for submission

@@
 
-In order to address these challenges, we present \pkg{fertile}, a low barrier-to-entry package focusing on common reproducibility mistakes programmers make while conducting data science research in R.
+In order to address these challenges, we present \pkg{fertile}(\cite{R-fertile}), a low barrier-to-entry package focusing on common reproducibility mistakes programmers make while conducting data science research in R.
 

### 2020-08-18 STAT-paper-template/APA/fertile.tex

@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-08-11 21:28:23
-## 2 package:for~ <NA>                                 base::l~ 2020-08-11 21:28:23
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-08-11 21:28:23
+## 1 package:pur~ <NA>                                 base::l~ 2020-08-11 21:44:30
+## 2 package:for~ <NA>                                 base::l~ 2020-08-11 21:44:30
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-08-11 21:44:30
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x7f8f8c998438>
+## <bytecode: 0x7ff9e192df58>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpGDaHPo/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpr51HxJ/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpGDaHPo/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpGDaHPo/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpGDaHPo/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpr51HxJ/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpr51HxJ/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpr51HxJ/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}

### 2020-10-06 STAT-paper-template/APA/fertile.Rnw
modifications to address comments

@@
 
+\let\nofiles\relax
 \usepackage{listings}
@@
 \usepackage{titlesec}
+\usepackage{soul}
+\usepackage{natbib}
 \titlespacing{\section}{0pt}{10pt plus 2pt minus 1pt}{10pt plus 2pt minus 1pt}
@@
 \abstract[Summary]{The advancement of scientific knowledge increasingly depends on ensuring that data-driven research is reproducible: that two people with the same data obtain the same results.
-However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation, and no clear consensus on standards of what constitutes reproducibility in published research.
+However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation and no clear consensus on standards of what constitutes reproducibility in published research.
 We present \pkg{fertile}, an R package that focuses on a series of common mistakes programmers make while conducting data science projects in R, primarily through the RStudio integrated development environment.
-\pkg{fertile} operates in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that is already written for potential problems).
+\pkg{fertile} operates in two modes: proactively, to prevent reproducibility mistakes from happening in the first place, and retroactively, analyzing code that is already written for potential problems.
 Furthermore, \pkg{fertile} is designed to educate users on why their mistakes are problematic and how to fix them.}
 
-\keywords{reproducibility, statistical software, workflow, collaboration, education}
+\keywords{Quality control, Statistical computing, Statistical process control, Teaching Statistics}
 
@@
 As research is becoming increasingly data-driven, and because knowledge can be shared worldwide so rapidly, reproducibility is critical to the advancement of scientific knowledge.
-Academics around the world have recognized this, and publications and discussions addressing reproducibility appear to have increased in the last several years (\cite{eisner-reproducibility}; \cite{sep-scientific-reproducibility}; \cite{bioessays-gosselin}; \cite{engineering-reproducibility}; \cite{plos-biology}).
+Academics around the world have recognized this, and publications and discussions addressing reproducibility appear to have increased in the last several years \citep{eisner-reproducibility, sep-scientific-reproducibility, bioessays-gosselin, engineering-reproducibility, plos-biology}.
 
@@
 
-However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation and no clear consensus on standards of what constitutes reproducibility in published research (\cite{peng2009bio}).
+However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation and no clear consensus on standards of what constitutes reproducibility in published research \citep{peng2009bio}.
 Not only are the components of reproducible research up for discussion (e.g., need the software be open source?), but the corresponding recommendations for ensuring reproducibility also vary (e.g., should raw and processed data files be in separate directories?).
@@
 
-In order to address these challenges, we present \pkg{fertile}(\cite{R-fertile}), a low barrier-to-entry package focusing on common reproducibility mistakes programmers make while conducting data science research in R.
+In order to address these challenges, we present \pkg{fertile}\citep{R-fertile}, a low barrier-to-entry package focusing on common reproducibility mistakes programmers make while conducting data science research in R \citep{r-software}.
 
@@
 Publications on reproducibility can be found in all areas of scientific research.
-However, as \cite{Goodman341ps12} argue, the language and conceptual framework of research reproducibility varies significantly across the sciences, and there are no clear standards on reproducibility agreed upon by the scientific community as a whole.
+However, as \citet{Goodman341ps12} argue, the language and conceptual framework of research reproducibility \st{varies} \textbf{vary} significantly across the sciences, and there are no clear standards on reproducibility agreed upon by the scientific community as a whole.
 We consider recommendations from a variety of fields and determine the key aspects of reproducibility faced by scientists in different disciplines.
 
-\cite{kitzes2017practice} present a collection of case studies on reproducibility practices from across the data-intensive sciences, illustrating a variety of recommendations and techniques for achieving reproducibility.
+\citet{kitzes2017practice} present a collection of case studies on reproducibility practices from across the data-intensive sciences, illustrating a variety of recommendations and techniques for achieving reproducibility.
 Although their work does not come to a consensus on the exact standards of reproducibility that should be followed, several common trends and principles emerge from their case studies: 1) use clear separation, labeling, and documentation, 2) automate processes when possible, and 3) design the data analysis workflow as a sequence of small steps glued together, with outputs from one step serving as inputs into the next.
-This is a common suggestion within the computing community, originating as part of the Unix philosophy (\cite{unix}).
+This is a common suggestion within the computing community, originating as part of the Unix philosophy \citep{unix}.
 
-\cite{cooper2017guide} focus on data analysis in \R and identify a similar list of important reproducibility components, reinforcing the need for clearly labeled, well-documented, and well-separated files.
+\citet{cooper2017guide} focus on data analysis in \R and identify a similar list of important reproducibility components, reinforcing the need for clearly labeled, well-documented, and well-separated files.
 In addition, they recommend publishing a list of dependencies and using version control.
-\cite{broman} reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a \cmd{README}.
+\citet{broman} reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a \cmd{README}.
 
@@
 They focus on a need for a well-developed file system, with no extraneous files and clear labeling.
-They also reiterate the need to note dependencies and use automation when possible, while making clear a suggestion not present in the previously-discussed literature: the need to use seeds, which allow for the saving and restoring of the random number generator state, when running code involving randomness (\cite{r-opensci}).
+They also reiterate the need to note dependencies and use automation when possible, while making clear a suggestion not present in the previously-discussed literature: the need to use seeds, which allow for the saving and restoring of the random number generator state, when running code involving randomness \citep{r-opensci}.
 
@@
   \item Code should be written in a coherent style.
-Code that conforms to a style guide or is written in a consistent dialect is easier to read (\cite{hermans2017programming}).
+Code that conforms to a style guide or is written in a consistent dialect is easier to read \citep{hermans2017programming}.
 We believe that the \pkg{tidyverse} provides the most accessible dialect of \R.
@@
 Much of the available literature focuses on file structure, organization, and naming, and \pkg{fertile}'s features are consistent with this.
-\cite{marwick2018packaging} provide the framework for file structure that \pkg{fertile} is based on: a structure similar to that of an \R package (\cite{coreteam-extensions}, \cite{hadley-packages}), with an \cmd{R} folder, as well as \cmd{data}, \cmd{data-raw}, \cmd{inst}, and \cmd{vignettes}.
+\citet{marwick2018packaging} provide the framework for file structure that \pkg{fertile} is based on: a structure similar to that of an \R package \citep{coreteam-extensions, hadley-packages}, with an \cmd{R} folder, as well as \cmd{data}, \cmd{data-raw}, \cmd{inst}, and \cmd{vignettes}.
 
@@
 A small body of \R packages focuses on research reproducibility.
-\pkg{rrtools} (\cite{R-rrtools}) addresses some of the issues discussed in \cite{marwick2018packaging} by creating a basic \R package structure for a data analysis project and implementing a basic \func{testthat::check} functionality.
-The \pkg{orderly} (\cite{R-orderly}) package also focuses on file structure, requiring the user to declare a desired project structure (typically a step-by-step structure, where outputs from one step are inputs into the next) at the beginning and then creating the files necessary to achieve that structure.
-\pkg{workflowr}'s (\cite{R-workflowr}) functionality is based around version control and making code easily available online.
+\pkg{rrtools} \citep{R-rrtools} addresses some of the issues discussed in \citet{marwick2018packaging} by creating a basic \R package structure for a data analysis project and implementing a basic \func{testthat::check} functionality.
+The \pkg{orderly} \citep{R-orderly} package also focuses on file structure, requiring the user to declare a desired project structure (typically a step-by-step structure, where outputs from one step are inputs into the next) at the beginning and then creating the files necessary to achieve that structure.
+\pkg{workflowr}'s \citep{R-workflowr} functionality is based around version control and making code easily available online.
 It works to generate a website containing time-stamped, versioned, and documented results.
-\pkg{checkers} (\cite{R-checkers}) allows you to create custom checks that examine different aspects of reproducibility.
-\pkg{packrat} (\cite{R-packrat}) is focused on dependencies, creating a packaged folder containing a project as well as all of its dependencies so that projects dependent on lesser-used packages can be easily shared across computers.
-\pkg{drake} (\cite{R-drake}) analyzes workflows, skips steps where results are up to date, and provides evidence that results match the underlying code and data.
-Lastly, the \pkg{reproducible} (\cite{R-reproducible}) package focuses on the concept of caching: saving information so that projects can be run faster each time they are re-completed from the start.
+\pkg{checkers} \citep{R-checkers} allows you to create custom checks that examine different aspects of reproducibility.
+\pkg{packrat} \citep{R-packrat} is focused on dependencies, creating a packaged folder containing a project as well as all of its dependencies so that projects dependent on lesser-used packages can be easily shared across computers.
+\pkg{drake} \citep{R-drake} analyzes workflows, skips steps where results are up to date, and provides evidence that results match the underlying code and data.
+Lastly, the \pkg{reproducible} \citep{R-reproducible} package focuses on the concept of caching: saving information so that projects can be run faster each time they are re-completed from the start.
 
@@
 \emph{Continuous integration} tools provide more general approaches to automated checking, which can enhance reproducibility with minimal code.
-For example, \pkg{wercker}---a command line tool that leverages Docker---enables users to test whether their projects will successfully compile when run on a variety of operating systems without access to the user's local hard drive (\cite{wercker}).
+For example, \pkg{wercker}---a command line tool that leverages Docker---enables users to test whether their projects will successfully compile when run on a variety of operating systems without access to the user's local hard drive \citep{wercker}.
 \pkg{GitHub Actions} is integrated into GitHub and can be configured to do similar checks on projects hosted in repositories.
@@
 Once an \R Project is created, \pkg{fertile} provides benefits throughout the data analysis process, both during development as well as after the fact.
-\pkg{fertile} achieves this by operating in two modes: proactively (to prevent reproducibility mistakes from happening in the first place), and retroactively (analyzing code that has already been written for potential problems).
+\pkg{fertile} achieves this by operating in two modes: proactively, to prevent reproducibility mistakes from happening in the first place, and retroactively, analyzing code that has already been written for potential problems.
 
@@
 
-<<fertile, message = FALSE, error = TRUE>>=
+<<fertile, message = FALSE, warning = FALSE, error = TRUE>>=
 library(fertile)
@@
 
-These proactive warning features are activated immediately after attaching the \pkg{fertile} package and require no additional effort by the user.
-
 \subsection{Retroactive Use}
@@
 
-The \func{proj\_check} function runs fifteen different reproducibility tests, noting which ones passed, which ones failed, the reason for failure, a recommended solution, and a guide to where to look for help.
+The \func{proj\_check} function runs sixteen different reproducibility tests, noting which ones passed, which ones failed, the reason for failure, a recommended solution, and a guide to where to look for help.
 These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that do not have a call to \func{set.seed} present.
@@
 
-<<list-checks>>=
-list_checks()
-@
+\vspace{2mm}
 
-Subsets of the fifteen tests can be invoked using the \pkg{tidyselect} helper functions (\cite{R-tidyselect}) in combination with the more limited \func{proj\_check\_some} function.
+\includegraphics[width=500pt,height=280pt]{list-checks}
+
+Subsets of the sixteen tests can be invoked using the \pkg{tidyselect} helper functions \citep{R-tidyselect} in combination with the more limited \func{proj\_check\_some} function.
+These helper functions, which include \func{starts\_with}, \func{ends\_with}, \func{contains}, and \func{matches}, allow users to check for matches between a string or regular expression and a list of variables--in this case, searching for checks from the list in \func{list\_checks} whose names match a certain condition (i.e.
+ending with "root").
 
@@
 @
+
+\textbf{Both the proactive and retroactive features available in \pkg{fertile} are enabled automatically upon the loading of the package.
+To prevent users from accidentally enabling or disabling features, \pkg{fertile} does not provide any option to control which features are active.
+However, those looking to only use the proactive features can do so by refraining from using the retroactive-specific functions while those looking to only use the retroactive features can wait to load \pkg{fertile} until they are ready to produce a report on their project.}
 
@@
 Much of the functionality in \pkg{fertile} is achieved by writing \href{https://en.wikipedia.org/wiki/Shim_(computing)}{shims}.
-\pkg{fertile}'s shimmed functions intercept the user's commands and perform various logging and checking tasks before executing the desired function.
+\textbf{In computer programming generally, a shim is a function that intercepts a user's call, changes the arguments passed to it, and/or modifies the execution of the function in some way.} In the case of \pkg{fertile} specifically, shims are used to intercept the user's commands and perform various logging and path-checking tasks before executing the desired function.
 Our process is:
@@
 Popular functions associated with only one package (e.g., \func{read\_csv} from \pkg{readr}) are ideal candidates.
-\item Create a function in \pkg{fertile} with the same name that takes the same arguments (and always the dots \cmd{...}).
+\item Create a function in \pkg{fertile} with the same name that takes the same arguments \st{(and always the dots ...)}.
 \item Write this new function so that it: a) captures any arguments, b) logs the name of the function called, c) performs any checks on these arguments, and d) calls the original function with the original arguments.
 Except where warranted, the execution looks the same to the user as if they were calling the original function.
+Since \pkg{fertile} only needs access to the file path argument, "file" is the only declared argument for most shims.
+The other arguments are managed using the dots (...), which capture and save them, allowing them to eventually be passed back to the original shimmed function (i.e.
+\func{readr::read\_csv}).
+
 \end{enumerate}
@@
 When run on an \R project directory, the function creates a \cmd{.R} script file that contains the code needed to install all of the packages referenced in the project, differentiating between packages located on CRAN and those located on GitHub.
+<<proj_pkg_script, eval=TRUE>>=
+install_script <- proj_pkg_script(proj_dir)
+@
 
-<<proj_pkg_script, eval=TRUE, cache=TRUE>>=
-install_script <- proj_pkg_script(proj_dir)
+The package installation script for project miceps looks as follows:
+
+<<read script, echo = FALSE>>=
 cat(readChar(install_script, 1e5))
+@
+
+\textbf{\pkg{fertile} also keeps track of dependencies in additional detail.
+Every time the code contained within a project is rendered by `fertile`, the package generates a file capturing the `sessionInfo()` just after the code was run.
+This file contains information about the R version in which the code was run, the list of packages that were loaded, and their specific versions.}
+
+\textbf{Users can access this file via the \func{proj\_dependency\_report} function. Its output appears as a new window in RStudio with contents that look like the text below:}
+
+<<dependency report text>>=
+##  The R project located at '/Users/audreybertin/Documents/fertile-paper2/STAT-paper-template
+##  /APA/project_miceps' was last run in the following software environment:
+
+##  R version 4.0.2 (2020-06-22)
+##  Platform: x86_64-apple-darwin17.0 (64-bit)
+##  Running under: macOS Catalina 10.15.5
+
+##  Matrix products: default
+##  BLAS:   /Library/Frameworks/R.framework/Versions/4.0/Resources/lib/libRblas.dylib
+##  LAPACK: /Library/Frameworks/R.framework/Versions/4.0/Resources/lib/libRlapack.dylib
+
+##  locale:
+##  en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8
+
+##  attached base packages:
+##  stats     graphics  grDevices utils     datasets  methods   base
+
+##  other attached packages:
+##  stargazer_5.2.2    skimr_2.1.2        purrr_0.3.4       
+##  ggplot2_3.3.2      tidyr_1.1.2        readr_1.3.1        dplyr_1.0.2
+
 @
@@
 
-\section{Results}
+\section{Experimental Testing}
 
@@
 Unfortunately, we were unable to complete the analysis as intended as the trial had to be postponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing.
-Although the experiment was unsuccessful in its first attempt, we hope to run the same trial again and gather data on \pkg{fertile}'s effectiveness.
+Although the experiment was unsuccessful in its first attempt, \st{we hope to run the same trial again and gather data on fertile's effectiveness} \textbf{conditions at Smith College have now stabilized and we are currently running a second trial in the new online-only classroom environment}.
 
@@
 \begin{itemize} [noitemsep]
-  \item Expanding dependency management features to include \R session information and package version numbers in addition to package names.
-  \item Expanding code and documentation style features to analyze whether code has been properly commented in addition to checking for a \cmd{README} and tidy code style.
+  \item \st{Expanding dependency management features to include R session information and package version numbers in addition to package names.}
+  \item \st{Expanding code and documentation style features to analyze whether code has been properly commented in addition to checking for a README and tidy code style}.
   \item Adding \cmd{make}-like functionality that can analyze an \R project structure and files and use this information to generate a Makefile.
 This Makefile would have information about target files and their prerequisites and would assist with making sure that re-running an analysis is done as quickly as possible by ensuring that only the necessary code and files that have been updated are run when rebuilding and re-running code.
+  \item \textbf{Adding a badge system to improve user experience and simplify the process of understanding how reproducible any given project is.
+Under this system, projects would receive badges from \pkg{fertile} based on the sets of reproducibility checks they pass.
+For example, projects that pass all of the checks on file paths would receive a badge on their project for "Reproducible File Paths" while those who pass checks involving file clutter and project structure might be awarded a badge for "Exellent File Organization."}
+  \item \textbf{Implementing a method through which users can add their own shims to the package. This would allow users to modify \pkg{fertile} to fit the functions they use most often. }
 \end{itemize}

### 2020-10-06 STAT-paper-template/APA/fertile.tex
modifications to address comments

--- 2020-08-11 STAT-paper-template/APA/fertile.tex
@@
 
-In order to address these challenges, we present \pkg{fertile}, a low barrier-to-entry package focusing on common reproducibility mistakes programmers make while conducting data science research in R.
+In order to address these challenges, we present \pkg{fertile}\citep{R-fertile}, a low barrier-to-entry package focusing on common reproducibility mistakes programmers make while conducting data science research in R \citep{r-software}.
 
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{list_checks}\hlstd{()}
-\end{alltt}
+\vspace{2mm}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- The available checks in `fertile` are as follows: ------------------------ fertile 0.0.0.9027 --}}\begin{verbatim}
-##  [1] "has_tidy_media"          "has_tidy_images"        
-##  [3] "has_tidy_code"           "has_tidy_raw_data"      
-##  [5] "has_tidy_data"           "has_tidy_scripts"       
-##  [7] "has_readme"              "has_no_lint"            
-##  [9] "has_proj_root"           "has_no_nested_proj_root"
-## [11] "has_only_used_files"     "has_clear_build_chain"  
-## [13] "has_no_absolute_paths"   "has_only_portable_paths"
-## [15] "has_no_randomness"
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
+\includegraphics[width=500pt,height=280pt]{list-checks}
 
-Subsets of the fifteen tests can be invoked using the \pkg{tidyselect} helper functions (\cite{R-tidyselect}) in combination with the more limited \func{proj\_check\_some} function.
+Subsets of the sixteen tests can be invoked using the \pkg{tidyselect} helper functions \citep{R-tidyselect} in combination with the more limited \func{proj\_check\_some} function.
+These helper functions, which include \func{starts\_with}, \func{ends\_with}, \func{contains}, and \func{matches}, allow users to check for matches between a string or regular expression and a list of variables--in this case, searching for checks from the list in \func{list\_checks} whose names match a certain condition (i.e.
+ending with "root").
 
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... ------------------------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... --------------------------------------------------------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... --------------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ----------------------------------------------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks ------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks --------------------------------------------------------- fertile 0.0.0.9027 --}}
 
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks ------------------------------------------------ fertile 0.0.0.9027 --}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks -------------------------------------------------------------- fertile 0.0.0.9027 --}}\end{kframe}
 \begin{kframe}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps --------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps ----------------------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code ------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code --------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 9 x 3
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory --------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
-##                      file   ext    size
-## 1 Estrogen_Receptors.docx  docx  10.97K
-## 2      citrate_v_time.png   png 188.08K
-## 3     proteins_v_time.png   png 378.45K
-## 4   Blot_data_updated.csv   csv  14.43K
-## 5      CS_data_redone.csv   csv   7.39K
-## 6                mice.csv   csv  14.33K
-## 7               README.md    md      39
-## 8            miceps.Rproj Rproj     204
-## 9            analysis.Rmd   Rmd   4.94K
-##                                                                      mime
-## 1 application/vnd.openxmlformats-officedocument.wordprocessingml.document
-## 2                                                               image/png
-## 3                                                               image/png
-## 4                                                                text/csv
-## 5                                                                text/csv
-## 6                                                                text/csv
-## 7                                                           text/markdown
-## 8                                                            text/rstudio
-## 9                                                         text/x-markdown
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory ----------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+## # A tibble: 9 x 4
+##   file               ext        size mime                                       
+##   <fs::path>         <chr> <fs::byt> <chr>                                      
+## 1 Estrogen_Receptor~ docx     10.97K application/vnd.openxmlformats-officedocum~
+## 2 citrate_v_time.png png     184.04K image/png                                  
+## 3 proteins_v_time.p~ png     377.02K image/png                                  
+## 4 Blot_data_updated~ csv      14.43K text/csv                                   
+## 5 CS_data_redone.csv csv       7.39K text/csv                                   
+## 6 mice.csv           csv      14.33K text/csv                                   
+## 7 README.md          md           39 text/markdown                              
+## 8 miceps.Rproj       Rproj       204 text/rstudio                               
+## 9 analysis.Rmd       Rmd       4.94K text/x-markdown
 \end{verbatim}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files ------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
-##                  path_rel    dir_rel
-## 1   Blot_data_updated.csv   data-raw
-## 2      CS_data_redone.csv   data-raw
-## 3 Estrogen_Receptors.docx inst/other
-## 4            analysis.Rmd  vignettes
-## 5      citrate_v_time.png inst/image
-## 6                mice.csv   data-raw
-## 7     proteins_v_time.png inst/image
-##                                                                                                cmd
-## 1     file_move('project_miceps/Blot_data_updated.csv', fs::dir_create('project_miceps/data-raw'))
-## 2        file_move('project_miceps/CS_data_redone.csv', fs::dir_create('project_miceps/data-raw'))
-## 3 file_move('project_miceps/Estrogen_Receptors.docx', fs::dir_create('project_miceps/inst/other'))
-## 4             file_move('project_miceps/analysis.Rmd', fs::dir_create('project_miceps/vignettes'))
-## 5      file_move('project_miceps/citrate_v_time.png', fs::dir_create('project_miceps/inst/image'))
-## 6                  file_move('project_miceps/mice.csv', fs::dir_create('project_miceps/data-raw'))
-## 7     file_move('project_miceps/proteins_v_time.png', fs::dir_create('project_miceps/inst/image'))
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files --------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+## # A tibble: 7 x 3
+##   path_rel           dir_rel    cmd                                             
+##   <fs::path>         <fs::path> <chr>                                           
+## 1 Blot_data_updated~ data-raw   file_move('project_miceps/Blot_data_updated.csv~
+## 2 CS_data_redone.csv data-raw   file_move('project_miceps/CS_data_redone.csv', ~
+## 3 Estrogen_Receptor~ inst/other file_move('project_miceps/Estrogen_Receptors.do~
+## 4 analysis.Rmd       vignettes  file_move('project_miceps/analysis.Rmd', fs::di~
+## 5 citrate_v_time.png inst/image file_move('project_miceps/citrate_v_time.png', ~
+## 6 mice.csv           data-raw   file_move('project_miceps/mice.csv', fs::dir_cr~
+## 7 proteins_v_time.p~ inst/image file_move('project_miceps/proteins_v_time.png',~
 \end{verbatim}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged ----------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged ------------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## NULL
@@
 \end{knitrout}
+
+\textbf{Both the proactive and retroactive features available in \pkg{fertile} are enabled automatically upon the loading of the package.
+To prevent users from accidentally enabling or disabling features, \pkg{fertile} does not provide any option to control which features are active.
+However, those looking to only use the proactive features can do so by refraining from using the retroactive-specific functions while those looking to only use the retroactive features can wait to load \pkg{fertile} until they are ready to produce a report on their project.}
 
@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-08-11 21:28:23
-## 2 package:for~ <NA>                                 base::l~ 2020-08-11 21:28:23
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-08-11 21:28:23
+## 1 package:pur~ <NA>                                 base::l~ 2020-10-06 21:01:26
+## 2 package:for~ <NA>                                 base::l~ 2020-10-06 21:01:26
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-06 21:01:26
 \end{verbatim}
@@
 \begin{verbatim}
-## function (file, ...) 
-## {
+## function(file, ...) {
 ##   if (interactive_log_on()) {
@@
 ## }
-## <bytecode: 0x7f8f8c998438>
+## <bytecode: 0x7fbd9847b260>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpGDaHPo/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmp54U9ou/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpGDaHPo/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpGDaHPo/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpGDaHPo/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmp54U9ou/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmp54U9ou/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmp54U9ou/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}
@@
 When run on an \R project directory, the function creates a \cmd{.R} script file that contains the code needed to install all of the packages referenced in the project, differentiating between packages located on CRAN and those located on GitHub.
-
 \begin{knitrout}
@@
 \hlstd{install_script} \hlkwb{<-} \hlkwd{proj_pkg_script}\hlstd{(proj_dir)}
-\hlkwd{cat}\hlstd{(}\hlkwd{readChar}\hlstd{(install_script,} \hlnum{1e5}\hlstd{))}
 \end{alltt}
+\end{kframe}
+\end{knitrout}
+
+The package installation script for project miceps looks as follows:
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 \begin{verbatim}
@@
 \end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+\textbf{\pkg{fertile} also keeps track of dependencies in additional detail.
+Every time the code contained within a project is rendered by `fertile`, the package generates a file capturing the `sessionInfo()` just after the code was run.
+This file contains information about the R version in which the code was run, the list of packages that were loaded, and their specific versions.}
+
+\textbf{Users can access this file via the \func{proj\_dependency\_report} function. Its output appears as a new window in RStudio with contents that look like the text below:}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{##  The R project located at '/Users/audreybertin/Documents/fertile-paper2/STAT-paper-template}
+\hlcom{##  /APA/project_miceps' was last run in the following software environment:}
+
+\hlcom{##  R version 4.0.2 (2020-06-22)}
+\hlcom{##  Platform: x86_64-apple-darwin17.0 (64-bit)}
+\hlcom{##  Running under: macOS Catalina 10.15.5}
+
+\hlcom{##  Matrix products: default}
+\hlcom{##  BLAS:   /Library/Frameworks/R.framework/Versions/4.0/Resources/lib/libRblas.dylib}
+\hlcom{##  LAPACK: /Library/Frameworks/R.framework/Versions/4.0/Resources/lib/libRlapack.dylib}
+
+\hlcom{##  locale:}
+\hlcom{##  en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8}
+
+\hlcom{##  attached base packages:}
+\hlcom{##  stats     graphics  grDevices utils     datasets  methods   base     }
+
+\hlcom{##  other attached packages:}
+\hlcom{##  stargazer_5.2.2    skimr_2.1.2        purrr_0.3.4       }
+\hlcom{##  ggplot2_3.3.2      tidyr_1.1.2        readr_1.3.1        dplyr_1.0.2}
+\end{alltt}
 \end{kframe}

### 2020-10-06 STAT-paper-template/APA/fertile_arxiv.Rnw

--- 2020-08-18 STAT-paper-template/APA/fertile.Rnw
@@
-\documentclass[APA,LATO1COL]{WileyNJD-v2}
+% !TeX program = pdfLaTeX
+\documentclass[12pt,notitlepage]{article}
+\usepackage{amsmath}
+\usepackage{graphicx,psfrag,epsf}
+\usepackage{enumerate}
+\usepackage{natbib}
+\usepackage{textcomp}
+\usepackage[hyphens]{url} % not crucial - just used below for the URL
+\usepackage{hyperref}
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
-\articletype{Article Type}%
+%\pdfminorversion=4
 
-\usepackage{listings}
-\usepackage{NJDapacite}
-\usepackage{hyperref}
-\usepackage{graphicx}
+% DON'T change margins - should be 1 inch all around.
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.3in}%
+\addtolength{\topmargin}{-.8in}%
+
+%% load any required packages here
 \usepackage{xspace}
 \usepackage{enumitem}
-\usepackage{titlesec}
-\titlespacing{\section}{0pt}{10pt plus 2pt minus 1pt}{10pt plus 2pt minus 1pt}
-\titlespacing{\subsection}{0pt}{7pt plus 2pt minus 1pt}{7pt plus 2pt minus 1pt}
-\titlespacing{\subsubsection}{0pt}{7pt plus 2pt minus 1pt}{7pt plus 2pt minus 1pt}
 
@@
 
-\received{26 April 2016}
-\revised{6 June 2016}
-\accepted{6 June 2016}
+\begin{document}
 
-\raggedbottom
-
-\begin{document}
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+{#1}\small\normalsize} \spacingset{1}
 
@@
 
-\author[1]{Audrey M. Bertin}
+\author{Audrey M. Bertin\\
+Statistical and Data Sciences\\
+Smith College\\
+\and
+Benjamin S. Baumer\\
+Statistical and Data Sciences\\
+Smith College\\
+}
 
-\author[2]{Benjamin S. Baumer}
+\maketitle
 
-\authormark{Bertin and Baumer, 2020}
-
-\address[1]{\orgdiv{Program in Statistical and Data Sciences}, \orgname{Smith College}, \orgaddress{\city{Northampton}, \state{MA}}}
-
-\address[2]{\orgdiv{Associate Professor of Statistical and Data Sciences}, \orgname{Smith College}, \orgaddress{\city{Northampton}, \state{MA}}}
-
-\corres{Audrey Bertin, Program in Statistical and Data Sciences, Smith College, Northampton, MA. \email{abertin@smith.edu}}
-
-\abstract[Summary]{The advancement of scientific knowledge increasingly depends on ensuring that data-driven research is reproducible: that two people with the same data obtain the same results.
+\abstract{The advancement of scientific knowledge increasingly depends on ensuring that data-driven research is reproducible: that two people with the same data obtain the same results.
 However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation, and no clear consensus on standards of what constitutes reproducibility in published research.
@@
 Furthermore, \pkg{fertile} is designed to educate users on why their mistakes are problematic and how to fix them.}
-
-\keywords{reproducibility, statistical software, workflow, collaboration, education}
-
-\maketitle
 
@@
 \section*{Data Availability Statement}
-The sample project ("project\_miceps") and package code associated with this paper can be found in the "R" and "tests" folders at \href{https://github.com/baumer-lab/fertile}{https://github.com/baumer-lab/fertile}.
+The sample project \cmd{project\_miceps} and package code associated with this paper can be found in the \cmd{R} and \cmd{tests} folders at \url{https://github.com/baumer-lab/fertile}.
 
-\nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;
+\bibliographystyle{agsm}
 \bibliography{refs}%
 
-\section*{Author Biography}
-
-\begin{biography}{\includegraphics[width=70pt,height=70pt]{audrey_bio}}{\textbf{Audrey M.
-Bertin} is an undergraduate student at Smith College intending on completing a Bachelor of Arts in Statistical and Data Sciences and Public Policy in Spring 2021.
-Audrey is a recipient of Smith College's prestigious Zollman Scholarship and has participated in extensive research with faculty throughout her time at the college.
-Her research interests include data and science communication, data protection and privacy, algorithmic bias and injustice, and using data science for social good.
-
-}
-\end{biography}
-
-\begin{biography}{\includegraphics[width=70pt,height=70pt]{ben_bio}}{\textbf{Benjamin S.
-Baumer} is an associate professor in the Statistical and Data Sciences program at Smith College.
-Ben is a co-author of \textit{The Sabermetric Revolution}, \textit{Modern Data Science with R}, and the second edition of \textit{Analyzing Baseball Data with R}.
-Ben has received the Waller Education Award from the ASA Section on Statistics and Data Science Education, the Significant Contributor Award from the ASA Section on Statistics in Sports, and the Contemporary Baseball Analysis Award from the Society for American Baseball Research.
-His research interests include sports analytics, data science, statistics and data science education, statistical computing, and network science.}
-\end{biography}
-
 \end{document}

### 2020-10-07 STAT-paper-template/APA/fertile.Rnw
improved clarity of change highlighting

@@
 \usepackage{natbib}
+\usepackage{xcolor}
 \titlespacing{\section}{0pt}{10pt plus 2pt minus 1pt}{10pt plus 2pt minus 1pt}
@@
 Publications on reproducibility can be found in all areas of scientific research.
-However, as \citet{Goodman341ps12} argue, the language and conceptual framework of research reproducibility \st{varies} \textbf{vary} significantly across the sciences, and there are no clear standards on reproducibility agreed upon by the scientific community as a whole.
+However, as \citet{Goodman341ps12} argue, the language and conceptual framework of research reproducibility \textcolor{red}{\st{varies}} \textcolor{blue}{vary} significantly across the sciences, and there are no clear standards on reproducibility agreed upon by the scientific community as a whole.
 We consider recommendations from a variety of fields and determine the key aspects of reproducibility faced by scientists in different disciplines.
@@
 Subsets of the sixteen tests can be invoked using the \pkg{tidyselect} helper functions \citep{R-tidyselect} in combination with the more limited \func{proj\_check\_some} function.
-These helper functions, which include \func{starts\_with}, \func{ends\_with}, \func{contains}, and \func{matches}, allow users to check for matches between a string or regular expression and a list of variables--in this case, searching for checks from the list in \func{list\_checks} whose names match a certain condition (i.e.
-ending with "root").
+\textcolor{blue}{These helper functions, which include \func{starts\_with}, \func{ends\_with}, \func{contains}, and \func{matches}, allow users to check for matches between a string or regular expression and a list of variables--in this case, searching for checks from the list in \func{list\_checks} whose names match a certain condition (i.e.
+ending with "root")}.
 
@@
 
-\textbf{Both the proactive and retroactive features available in \pkg{fertile} are enabled automatically upon the loading of the package.
+\textcolor{blue}{Both the proactive and retroactive features available in \pkg{fertile} are enabled automatically upon the loading of the package.
 To prevent users from accidentally enabling or disabling features, \pkg{fertile} does not provide any option to control which features are active.
@@
 Much of the functionality in \pkg{fertile} is achieved by writing \href{https://en.wikipedia.org/wiki/Shim_(computing)}{shims}.
-\textbf{In computer programming generally, a shim is a function that intercepts a user's call, changes the arguments passed to it, and/or modifies the execution of the function in some way.} In the case of \pkg{fertile} specifically, shims are used to intercept the user's commands and perform various logging and path-checking tasks before executing the desired function.
+\textcolor{blue}{{In computer programming generally, a shim is a function that intercepts a user's call, changes the arguments passed to it, and/or modifies the execution of the function in some way.}} In the case of \pkg{fertile} specifically, shims are used to intercept the user's commands and perform various logging and path-checking tasks before executing the desired function.
 Our process is:
@@
 Popular functions associated with only one package (e.g., \func{read\_csv} from \pkg{readr}) are ideal candidates.
-\item Create a function in \pkg{fertile} with the same name that takes the same arguments \st{(and always the dots ...)}.
+\item Create a function in \pkg{fertile} with the same name that takes the same arguments \textcolor{red}{\st{(and always the dots ...)}}.
 \item Write this new function so that it: a) captures any arguments, b) logs the name of the function called, c) performs any checks on these arguments, and d) calls the original function with the original arguments.
 Except where warranted, the execution looks the same to the user as if they were calling the original function.
-Since \pkg{fertile} only needs access to the file path argument, "file" is the only declared argument for most shims.
+\textcolor{blue}{Since \pkg{fertile} only needs access to the file path argument, "file" is the only declared argument for most shims.
 The other arguments are managed using the dots (...), which capture and save them, allowing them to eventually be passed back to the original shimmed function (i.e.
-\func{readr::read\_csv}).
+\func{readr::read\_csv})}.
 
@@
 
-\textbf{\pkg{fertile} also keeps track of dependencies in additional detail.
+\textcolor{blue}{\pkg{fertile} also keeps track of dependencies in additional detail.
 Every time the code contained within a project is rendered by `fertile`, the package generates a file capturing the `sessionInfo()` just after the code was run.
@@
 
-\textbf{Users can access this file via the \func{proj\_dependency\_report} function. Its output appears as a new window in RStudio with contents that look like the text below:}
+\textcolor{blue}{Users can access this file via the \func{proj\_dependency\_report} function. Its output appears as a new window in RStudio with contents that look like the text below:}
 
@@
 Unfortunately, we were unable to complete the analysis as intended as the trial had to be postponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing.
-Although the experiment was unsuccessful in its first attempt, \st{we hope to run the same trial again and gather data on fertile's effectiveness} \textbf{conditions at Smith College have now stabilized and we are currently running a second trial in the new online-only classroom environment}.
+Although the experiment was unsuccessful in its first attempt, \textcolor{red}{\st{we hope to run the same trial again and gather data on fertile's effectiveness}} \textcolor{blue}{conditions at Smith College have now stabilized and we are currently running a second trial in the new online-only classroom environment}.
 
@@
 \begin{itemize} [noitemsep]
-  \item \st{Expanding dependency management features to include R session information and package version numbers in addition to package names.}
-  \item \st{Expanding code and documentation style features to analyze whether code has been properly commented in addition to checking for a README and tidy code style}.
+  \item \textcolor{red}{\st{Expanding dependency management features to include R session information and package version numbers in addition to package names.}}
+  \item \textcolor{red}{\st{Expanding code and documentation style features to analyze whether code has been properly commented in addition to checking for a README and tidy code style}}.
   \item Adding \cmd{make}-like functionality that can analyze an \R project structure and files and use this information to generate a Makefile.
 This Makefile would have information about target files and their prerequisites and would assist with making sure that re-running an analysis is done as quickly as possible by ensuring that only the necessary code and files that have been updated are run when rebuilding and re-running code.
-  \item \textbf{Adding a badge system to improve user experience and simplify the process of understanding how reproducible any given project is.
+  \item \textcolor{blue}{Adding a badge system to improve user experience and simplify the process of understanding how reproducible any given project is.
 Under this system, projects would receive badges from \pkg{fertile} based on the sets of reproducibility checks they pass.
 For example, projects that pass all of the checks on file paths would receive a badge on their project for "Reproducible File Paths" while those who pass checks involving file clutter and project structure might be awarded a badge for "Exellent File Organization."}
-  \item \textbf{Implementing a method through which users can add their own shims to the package. This would allow users to modify \pkg{fertile} to fit the functions they use most often. }
+  \item \textcolor{blue}{Implementing a method through which users can add their own shims to the package. This would allow users to modify \pkg{fertile} to fit the functions they use most often. }
 \end{itemize}
@@
 \section*{Data Availability Statement}
-The sample project ("project\_miceps") and package code associated with this paper can be found in the "R" and "tests" folders at \href{https://github.com/baumer-lab/fertile}{https://github.com/baumer-lab/fertile}.
+The sample project \cmd{project\_miceps} and package code associated with this paper can be found in the \cmd{R} and \cmd{tests} folders at \url{https://github.com/baumer-lab/fertile}.
 

### 2020-10-07 STAT-paper-template/APA/fertile.tex
improved clarity of change highlighting

@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... --------------------------------------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... ---------------------------------------------------------------------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ----------------------------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ------------------------------------------------------------------------------ fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks --------------------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks ---------------------------------------------------------------------- fertile 0.0.0.9027 --}}
 
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks -------------------------------------------------------------- fertile 0.0.0.9027 --}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks --------------------------------------------------------------------------- fertile 0.0.0.9027 --}}\end{kframe}
 \begin{kframe}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps ----------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps ------------------------------------------------------ fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code --------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code ---------------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 9 x 3
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory ----------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory ------------------------------------------------------------------------ fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 9 x 4
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files --------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files ---------------------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 7 x 3
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged ------------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged -------------------------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## NULL
@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-10-06 21:01:26
-## 2 package:for~ <NA>                                 base::l~ 2020-10-06 21:01:26
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-06 21:01:26
+## 1 package:pur~ <NA>                                 base::l~ 2020-10-07 17:58:10
+## 2 package:for~ <NA>                                 base::l~ 2020-10-07 17:58:10
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-07 17:58:10
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x7fbd9847b260>
+## <bytecode: 0x7f81bb25c1d8>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmp54U9ou/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpKbQKHU/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmp54U9ou/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmp54U9ou/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmp54U9ou/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpKbQKHU/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpKbQKHU/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpKbQKHU/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}

### 2020-10-07 STAT-paper-template/APA/fertile.Rnw
added tibble list of checks

@@
 
+<<list of checks, echo = FALSE>>=
+check_name <- c("has_proj_root()",
+           "has_no_nested_proj_root()",
+           "has_clear_build_chain()",
+           "has_tidy_[media/images/code/raw_data/data/scripts]()",
+           "has_only_used_files()",
+           "has_readme()",
+           "has_well_commented_code()",
+           "has_no_absolute_paths()",
+           "has_only_portable_paths()",
+           "has_no_randomness()",
+           "has_no_lint()")
+
+description <- c(
+  "Checks to make sure a single .Rproj file is found in the root directory of the project.",
+  "Checks to make sure there are no nested .Rproj files in subfolders of the project.",
+  "Looks for indication of a clear order in which to run R scripts.",
+  "There are six different ”has_tidy_...” functions. Each checks to make sure that no files of its given type (images, code, data, etc.) are located in the root directory of the project",
+  "Ensures that all files located in the project directory are being referenced or created by other files in the project.",
+  "Checks to make sure there is a README file in the root directory.",
+  "Checks to ensure that all code files contain an adequate number of comments (>10% of lines).",
+  "Checks to make sure that no absolute paths are referenced in code.",
+  "Checks to make sure that no paths in code lead outside the project directory.",
+  "Checks to make sure code does not contain uncontrolled randomness.",
+  "Examines code to determine whether it conforms to ”tidy” style guide recommendations.")
+
+checks_list <- tidyr::tibble(check_name, description)
+@
+
 \includegraphics[width=500pt,height=280pt]{list-checks}
@@
 Subsets of the sixteen tests can be invoked using the \pkg{tidyselect} helper functions \citep{R-tidyselect} in combination with the more limited \func{proj\_check\_some} function.
-\textcolor{blue}{These helper functions, which include \func{starts\_with}, \func{ends\_with}, \func{contains}, and \func{matches}, allow users to check for matches between a string or regular expression and a list of variables--in this case, searching for checks from the list in \func{list\_checks} whose names match a certain condition (i.e.
+\textcolor{blue}{These helper functions, which include \func{starts\_with}, \func{ends\_with}, \func{contains}, and \func{matches}, allow users to check for matches between a string or regular expression and a list of variables--in this case, searching for checks from the list provided by \func{list\_checks} whose names match a certain condition (i.e.
 ending with "root")}.

### 2020-10-07 STAT-paper-template/APA/fertile.tex
added tibble list of checks

@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-10-07 17:58:10
-## 2 package:for~ <NA>                                 base::l~ 2020-10-07 17:58:10
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-07 17:58:10
+## 1 package:pur~ <NA>                                 base::l~ 2020-10-07 18:10:27
+## 2 package:for~ <NA>                                 base::l~ 2020-10-07 18:10:27
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-07 18:10:27
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x7f81bb25c1d8>
+## <bytecode: 0x7f98482ffa60>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpKbQKHU/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpxVNb0M/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpKbQKHU/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpKbQKHU/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpKbQKHU/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpxVNb0M/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpxVNb0M/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpxVNb0M/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}

### 2020-10-07 STAT-paper-template/APA/fertile.Rnw
addressing more comments

@@
 
-\pkg{fertile} attempts to address these gaps in existing software by providing a simple, easy-to-learn reproducibility package that, rather than focusing intensely on a specific area, provides some information about a wide variety of aspects influencing reproducibility.
-\pkg{fertile} is flexible, offering benefits to users at any stage in the data analysis workflow, and provides \R-specific features, which address certain aspects of reproducibility that can be missed by external project development software.
+\pkg{fertile} attempts to address these gaps in existing software by providing a simple, easy-to-learn reproducibility package that, rather than focusing intensely on a specific area \textcolor{blue}{(such as file structure)}, provides some information about a wide variety of aspects influencing reproducibility.
+\textcolor{blue}{For example, \pkg{fertile} addresses file structure in part but also considers the quality of project documentation, the use of file paths, and the behavior of random number generation}.
+\pkg{fertile} is flexible, \textcolor{red}{\st{offering benefits}} \textcolor{blue}{providing reports on reproducibility} to users at any stage in the data analysis workflow, and provides \R-specific features, \textcolor{blue}{such as tools to manage package dependencies}, which address certain aspects of reproducibility that can be missed by external project development software.
 
@@
 directories containing an \cmd{.Rproj} file).
-Once an \R Project is created, \pkg{fertile} provides benefits throughout the data analysis process, both during development as well as after the fact.
+Once an \R Project is created, \pkg{fertile} provides \textcolor{red}{\st{benefits}} \textcolor{blue} {the user with feedback on their adherence to reproducibility standards} throughout the data analysis process, both during development as well as after the fact.
 \pkg{fertile} achieves this by operating in two modes: proactively, to prevent reproducibility mistakes from happening in the first place, and retroactively, analyzing code that has already been written for potential problems.
@@
 Proactively, the package identifies potential mistakes as they are made by the user and outputs an informative message as well as a recommended solution.
-For example, \pkg{fertile} catches when a user passes a potentially problematic file path---such as an absolute path, or a path that points to a location outside of the project directory---to a variety of common input/output functions operating on many different file types.
+For example, \pkg{fertile} \textcolor{red}{\st{catches}} \textcolor{blue}{throws an error} when a user passes a potentially problematic file path---such as an absolute path, or a path that points to a location outside of the project directory---to a variety of common input/output functions operating on many different file types.
 
@@
 
-\pkg{fertile} is even more aggressive with functions (like \func{setwd}) that are almost certain to break reproducibility, causing them to throw errors that prevent their execution and providing recommendations for better alternatives.
+\textcolor{blue}{Users who want to override this behavior and execute their function anyway are provided with a explanation of how to do so.}
+
+\pkg{fertile} is even more aggressive with functions (like \func{setwd}) that are almost certain to break reproducibility \textcolor{blue}{no matter the type of path passed to them}.
+\textcolor{red}\st{, causing them to throw errors that prevent their execution and providing recommendations for better alternatives}.
+\textcolor{blue}{When these functions are called, users are not presented with an option to override the resulting error.
+Instead, the only way to achieve the desired action is for the user to execute an alternative function that produces the same behavior but in a reproducible way.}
 

### 2020-10-07 STAT-paper-template/APA/fertile.tex
addressing more comments

@@
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Detected absolute paths}}\begin{alltt}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Detected absolute paths.
+Absolute paths are not reproducible and will likely only work on your computer.
+If you would like to continue anyway, re-execute your command by typing the name of the package it originates from followed by :: and the function call, e.g.
+if you previously typed 'read\_csv(path)', instead use 'readr::read\_csv(path)'}}\begin{alltt}
 \hlkwd{read.csv}\hlstd{(}\hlstr{"../../../Desktop/my_data.csv"}\hlstd{)}
@@
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Detected paths that lead outside the project directory}}\end{kframe}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Detected paths that lead outside the project directory.
+Such paths are not reproducible and will likely only work on your computer.
+If you would like to continue anyway, re-execute your command by typing the name of the package it originates from followed by :: and the function call, e.g.
+if you previously typed 'read\_csv(path)', instead use 'readr::read\_csv(path)'}}\end{kframe}
 \end{knitrout}
 
-\pkg{fertile} is even more aggressive with functions (like \func{setwd}) that are almost certain to break reproducibility, causing them to throw errors that prevent their execution and providing recommendations for better alternatives.
+\textcolor{blue}{Users who want to override this behavior and execute their function anyway are provided with a explanation of how to do so.}
+
+\pkg{fertile} is even more aggressive with functions (like \func{setwd}) that are almost certain to break reproducibility \textcolor{blue}{no matter the type of path passed to them}.
+\textcolor{red}\st{, causing them to throw errors that prevent their execution and providing recommendations for better alternatives}.
+\textcolor{blue}{When these functions are called, users are not presented with an option to override the resulting error.
+Instead, the only way to achieve the desired action is for the user to execute an alternative function that produces the same behavior but in a reproducible way.}
 
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... ---------------------------------------------------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... ------------------------------------------------------------------------ fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ------------------------------------------------------------------------------ fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... -------------------------------------------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks ---------------------------------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks ------------------------------------------------------ fertile 0.0.0.9027 --}}
 
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks --------------------------------------------------------------------------- fertile 0.0.0.9027 --}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks ----------------------------------------------------------- fertile 0.0.0.9027 --}}\end{kframe}
 \begin{kframe}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps ------------------------------------------------------ fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps -------------------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code ---------------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code ------------------------------------------------ fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 9 x 3
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory ------------------------------------------------------------------------ fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory -------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 9 x 4
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files ---------------------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files ------------------------------------------------------ fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 7 x 3
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged -------------------------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged ---------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## NULL
@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-10-07 18:10:27
-## 2 package:for~ <NA>                                 base::l~ 2020-10-07 18:10:27
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-07 18:10:27
+## 1 package:pur~ <NA>                                 base::l~ 2020-10-07 20:12:53
+## 2 package:for~ <NA>                                 base::l~ 2020-10-07 20:12:53
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-07 20:12:53
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x7f98482ffa60>
+## <bytecode: 0x7fdf2658a2a8>
 ## <environment: namespace:fertile>
@@
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Detected absolute paths}}\begin{alltt}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Detected absolute paths.
+Absolute paths are not reproducible and will likely only work on your computer.
+If you would like to continue anyway, re-execute your command by typing the name of the package it originates from followed by :: and the function call, e.g.
+if you previously typed 'read\_csv(path)', instead use 'readr::read\_csv(path)'}}\begin{alltt}
 \hlcom{# Path outside the directory}
@@
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Detected paths that lead outside the project directory}}\end{kframe}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Detected paths that lead outside the project directory.
+Such paths are not reproducible and will likely only work on your computer.
+If you would like to continue anyway, re-execute your command by typing the name of the package it originates from followed by :: and the function call, e.g.
+if you previously typed 'read\_csv(path)', instead use 'readr::read\_csv(path)'}}\end{kframe}
 \end{knitrout}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpxVNb0M/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNwfTrr/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpxVNb0M/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpxVNb0M/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpxVNb0M/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNwfTrr/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNwfTrr/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNwfTrr/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}

### 2020-10-07 STAT-paper-template/APA/response-to-reviewers.md
finishing up most of comments

# Paper Revision Changelog

Thank you for the excellent feedback on the first submission of the "fertile" manuscript. Below are our responses to the comments made by the Associate Editor and Reviewer:

**Associate Editor’s Comments:**

*p. 1: I suggest removing the parentheses from the abstract and replacing with commas.*

Great suggestion. This has been changed. A similar piece of text later in the manuscript had the same parentheses and these have been removed as well.

*p. 1: Collapse the parenthetical citations starting with Eisner 2018 into a single \citep (not \cite) command, with citations separated by commas within the \citep command, and don't add an extra set of parentheses.  Make this change throughout the manuscript, noting the difference between \citep and \cite.*

The citations throughout the manuscript have now been changed. Parentheticals are now in \citep format, while in-text citations are \citet.

*p. 2: You might want to cite R itself.*

A citation for R has been added just before the literature review section.

*p. 2:  "language and conceptual framework ... varies" -> "language and conceptual framework ... vary"*

This grammatical mistake has been corrected.

*p. 6: In Section 3.4, it might help some readers if you provide a short explanation of what a shim is.*

Good point. To address this, I added a new sentence explaining the definition of a shim in computer science before proceeding to describe their purpose in fertile.

*p. 8: In Section 3.5.4, Do you want to address whether fertile has any way of checking that particular versions of any dependencies are in place? I presume that package versions that change over time can sometimes destroy reproducibility.*

This feature (version-specific dependency tracking) has been added to fertile since the original manuscript was submitted. A short section detailing the new functionality is now at the bottom of the existing section on project dependency management right after the code for proj_pkg_script().

*p. 10:  Too bad the pandemic interrupted your cool-sounding study!  It does seem that the study could be run in an online-only class environment though; is this true and is it worth mentioning?*

Yes, this is true! Since the submission of the original manuscript, we decided to conduct another trial in the classroom in fall 2020, now that Smith College is 100% online. The text has been updated to reflect this.

**Reviewer’s Comments:**

General:
*- the use of shims is likely unfamiliar to R users. Therefore, I have some concern about the ability of a user to control how aggressively this package performs. For example, if a user specifies an absolute path and receives an error message, it will be difficult for the user to determine what function was used to specify this error message and determine if there is a way to be less aggressive (to turn that into a warning, instead of an error, for example). My fear is that this functionality may limit use. However, I do like that the sandbox() function addresses this to an extent, allowing users to test things out and "get it right" before retuning to where the original error message was received.*

This is an excellent point! Part of the reason that fertile is so strict on this behavior is that in my own personal experience, many R coders (particularly those with little experience) tend to ignore warnings as long as their code executes properly. Having fertile’s path-related behavior be error based ensures that no one will skip over the messages that come up. However, my co-author and I do recognize that some users may still want to run these functions despite the warnings. In order to take this into account, we have added detail to the error messages produced by the shims. The new messages provide users with a method to override the error produced by fertile.

Specific minor suggestions (in order of manuscript):
*- A few suggested references to include in Intro and/or 2.1:* 
   - *reproducibility framework: Patil et al. (https://www.nature.com/articles/s41562-019-0629-z)*
   - *good enough practices: Wilson et al. (https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1005510)*
   - *best practices: Wilson et al (https://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.1001745)*

The first reference was an excellent fit for the introduction, and has been added there, while the second and third fit well into the literature review and have been mentioned accordingly.

- *Section 2.3: currently there are too many vague phrases "some information," "a wide variety of aspects," "offering benefits," "provides benefits," etc. - description here limits readers ability to truly understand what your work does to build on the shortcomings in 2.2. For a more specific example, Docker/CI is described in 2.2 and it's not until the end of the paper (3.5.4) that I have a good sense as to whether this work addresses those shortcomings. Would love more specifics in 2.2.*

Many of the vague phrases in 2.3 now have more specifics attached to them. This should help clarify the benefit of fertile compared with some of the shortcomings in 2.2.

- *Section 3.1: may be helpful to be very clear here that these are errors and not warnings. "catches" wasn't clear as to whether I would be allowed to specify an absolute path and would get a warning, or whether this would aggressively error. (Note: the code makes this clear, but the text may benefit from slight clarification).*

This has been clarified in the text. In addition, further explanation of the difference in response between the read.csv() and setwd() examples has been provided.

- *3.2 `list_checks()` is a nice function, but a table with the checks and a brief description of what that check does in the paper would be beneficial.*

This table has been added! Since list_checks() provides no additional benefit to the user than the information in this table, it is no longer mentioned.

- *in 3.1 and 3.2, it may be nice to describe explicitly whether it's possible to just use one or the other (proactive vs retroactive)....similar to how in 3.3 it's described how logging can be turned off (or reset). I could envision one wanting to develop without error and then check at the end (retroactive only). Philosophically, I could see that you would not want to build that in and that people could then just load your package at the end...but this may be something explicitly stated in the manuscript.*

A comment on this has been added to the end of the description of retroactive behavior.

- *Section 3.2: The average user may not know what the tidyselect helper functions are.  A bit more explanation here would likely help readers/users.*

A short description of tidy select helpers and their applicability here has been added just before the proj_check_some() example.

- *Section 3.4: The average reader may not understand what "and always the dots" means in point 2. Slightly more explanation may be helpful here.*

The original “and always the dots” mention has been removed and slightly more explanation has been added under item d.

- *Section 3.5.4: the line of code `cat(readChar(install_script, 1e5))` is not intuitive, and that line may benefit from a comment with a brief explanation within the code itself (or in the text above)*

Since the code is intuitive, it has now been hidden from view. Instead, there is now a sentence stating that we are looking into the contents of the file.

- *Section 4: Suggestion to rename to future directions or something other than results, given the lack of results here (no fault of the authors, here! simply beyond one's control.) But, I did read and was bummed to see no results here after the section header was Results. Excited to see these results in the future, regardless!*

Agreed on this point. No results are discussed so that warrants a name change. Since the section is about an experiment, however, I have chosen to replace it with “Experimental Testing”. We also decided to run a second trial this semester, so future directions is no longer an applicable name as it is happening currently!

*Very minor* (feel free to ignore)
- *I haven't dug into the code fully so I may have missed it, but I could imagine that a user may want the ability to add their own shim to make a check for something as new packages/functions come out or for something specific to their own uses. Is there functionality for one to create/add their own shim to their installation of fertile?*

There is not currently a functionality for this, but it is definitely something of interest for the future. As a result, we have added it to the future work section!

**Additional Changes**

*The first two bullets in the future work section have been completed since the submission of the original manuscript. Due to this, those bullets have been removed and replaced with new future steps.*

### 2020-10-07 STAT-paper-template/APA/fertile.Rnw
finishing up most of comments

@@
 
-Data-based research is considered fully reproducible when the requisite code and data files produce identical results when run by another analyst.
+Data-based research is considered fully reproducible when the requisite code and data files produce identical results when run by another analyst \citep{patil2019visual}.
 As research is becoming increasingly data-driven, and because knowledge can be shared worldwide so rapidly, reproducibility is critical to the advancement of scientific knowledge.
@@
 \citet{broman} reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a \cmd{README}.
+\textcolor{blue}{\citet{wilson2017good} argue that, to follow good practice, an analysis must be located in a well-organized directory, be supplemented by a file containing information about the project (i.e.
+a README), and a provide an explicit list of dependencies.}
+\textcolor{blue}{\citet{wilson2014best}} emphasize communication, recommending that code be human-readable and consistently styled for ease of understanding once shared.}
 

### 2020-10-07 STAT-paper-template/APA/fertile.tex
finishing up most of comments

@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... ------------------------------------------------------------------------ fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... ------------------------------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... -------------------------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... --------------------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks ------------------------------------------------------ fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks ------------------------------- fertile 0.0.0.9027 --}}
 
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks ----------------------------------------------------------- fertile 0.0.0.9027 --}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks ------------------------------------ fertile 0.0.0.9027 --}}\end{kframe}
 \begin{kframe}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps -------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps --------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code ------------------------------------------------ fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code ------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 9 x 3
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory -------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory --------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 9 x 4
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files ------------------------------------------------------ fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files ------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 7 x 3
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged ---------------------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged ----------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## NULL
@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-10-07 20:12:53
-## 2 package:for~ <NA>                                 base::l~ 2020-10-07 20:12:53
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-07 20:12:53
+## 1 package:pur~ <NA>                                 base::l~ 2020-10-07 20:50:29
+## 2 package:for~ <NA>                                 base::l~ 2020-10-07 20:50:29
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-07 20:50:30
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x7fdf2658a2a8>
+## <bytecode: 0x7fd40f5adbf0>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNwfTrr/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpFsW2WL/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNwfTrr/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNwfTrr/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNwfTrr/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpFsW2WL/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpFsW2WL/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpFsW2WL/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}

### 2020-10-07 STAT-paper-template/APA/fertile.tex
fixed bracket error

@@
 \pkg{fertile} is even more aggressive with functions (like \func{setwd}) that are almost certain to break reproducibility \textcolor{blue}{no matter the type of path passed to them}.
-\textcolor{red}\st{, causing them to throw errors that prevent their execution and providing recommendations for better alternatives}.
+\textcolor{red}{\st{, causing them to throw errors that prevent their execution and providing recommendations for better alternatives}}.
 \textcolor{blue}{When these functions are called, users are not presented with an option to override the resulting error.
@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-10-07 20:50:29
-## 2 package:for~ <NA>                                 base::l~ 2020-10-07 20:50:29
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-07 20:50:30
+## 1 package:pur~ <NA>                                 base::l~ 2020-10-07 21:00:13
+## 2 package:for~ <NA>                                 base::l~ 2020-10-07 21:00:13
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-07 21:00:13
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x7fd40f5adbf0>
+## <bytecode: 0x7fa3a7975308>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpFsW2WL/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpnSVnsZ/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpFsW2WL/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpFsW2WL/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpFsW2WL/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpnSVnsZ/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpnSVnsZ/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpnSVnsZ/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}

### 2020-10-07 STAT-paper-template/APA/fertile.Rnw
fixed bracket error
swapped i.e. for e.g.

@@
 Subsets of the sixteen tests can be invoked using the \pkg{tidyselect} helper functions \citep{R-tidyselect} in combination with the more limited \func{proj\_check\_some} function.
-\textcolor{blue}{These helper functions, which include \func{starts\_with}, \func{ends\_with}, \func{contains}, and \func{matches}, allow users to check for matches between a string or regular expression and a list of variables--in this case, searching for checks from the list provided by \func{list\_checks} whose names match a certain condition (i.e.
+\textcolor{blue}{These helper functions, which include \func{starts\_with}, \func{ends\_with}, \func{contains}, and \func{matches}, allow users to check for matches between a string or regular expression and a list of variables--in this case, searching for checks from the list provided by \func{list\_checks} whose names match a certain condition (e.g.
 ending with "root")}.
@@
 \textcolor{blue}{Since \pkg{fertile} only needs access to the file path argument, "file" is the only declared argument for most shims.
-The other arguments are managed using the dots (...), which capture and save them, allowing them to eventually be passed back to the original shimmed function (i.e.
+The other arguments are managed using the dots (...), which capture and save them, allowing them to eventually be passed back to the original shimmed function (e.g.
 \func{readr::read\_csv})}.

### 2020-10-07 STAT-paper-template/APA/fertile.tex
swapped i.e. for e.g.

@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-10-07 21:00:13
-## 2 package:for~ <NA>                                 base::l~ 2020-10-07 21:00:13
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-07 21:00:13
+## 1 package:pur~ <NA>                                 base::l~ 2020-10-07 22:16:42
+## 2 package:for~ <NA>                                 base::l~ 2020-10-07 22:16:42
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-07 22:16:42
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x7fa3a7975308>
+## <bytecode: 0x7fe560195598>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpnSVnsZ/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpnHOZCD/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpnSVnsZ/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpnSVnsZ/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpnSVnsZ/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpnHOZCD/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpnHOZCD/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpnHOZCD/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}

### 2020-10-08 STAT-paper-template/APA/fertile.Rnw
added missing bracket

@@
 a README), and a provide an explicit list of dependencies.}
-\textcolor{blue}{\citet{wilson2014best}} emphasize communication, recommending that code be human-readable and consistently styled for ease of understanding once shared.}
+\textcolor{blue}{{\citet{wilson2014best}} emphasize communication, recommending that code be human-readable and consistently styled for ease of understanding once shared.}
 

### 2020-10-08 STAT-paper-template/APA/fertile.tex
added missing bracket

@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... ------------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... -------------------------------------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... --------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ---------------------------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks ------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks -------------------------------------- fertile 0.0.0.9027 --}}
 
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks ------------------------------------ fertile 0.0.0.9027 --}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks ------------------------------------------- fertile 0.0.0.9027 --}}\end{kframe}
 \begin{kframe}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps --------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps ---------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code ------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code -------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 9 x 3
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory --------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory ---------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 9 x 4
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files ------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files -------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 7 x 3
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged ----------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged ------------------------------------------ fertile 0.0.0.9027 --}}\begin{verbatim}
 ## NULL
@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-10-07 22:16:42
-## 2 package:for~ <NA>                                 base::l~ 2020-10-07 22:16:42
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-07 22:16:42
+## 1 package:pur~ <NA>                                 base::l~ 2020-10-08 14:49:12
+## 2 package:for~ <NA>                                 base::l~ 2020-10-08 14:49:12
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-08 14:49:12
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x7fe560195598>
+## <bytecode: 0x7fc745046ad8>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpnHOZCD/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpWD3heO/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpnHOZCD/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpnHOZCD/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpnHOZCD/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpWD3heO/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpWD3heO/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpWD3heO/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}

### 2020-10-09 STAT-paper-template/APA/response-to-reviewers.md
added one more bullet point

@@
 
-*The first two bullets in the future work section have been completed since the submission of the original manuscript. Due to this, those bullets have been removed and replaced with new future steps.*
+The first two bullets in the future work section have been completed since the submission of the original manuscript. Due to this, those bullets have been removed and replaced with new future steps.
+
+Additionally, a new bullet point on code commenting (relevant to the has_well_commented_code() function as well as one of the added sources, has been added to the summary of reproducibility components.

### 2020-10-09 STAT-paper-template/APA/fertile.tex
added one more bullet point

@@
   \item Dependencies are noted.
+  \item \textcolor{blue}{Code files contain descriptive comments.}
   \end{itemize}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... -------------------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... ----------------------------------------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ---------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ------------------------------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks -------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks ----------------------------------------- fertile 0.0.0.9027 --}}
 
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks ------------------------------------------- fertile 0.0.0.9027 --}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks ---------------------------------------------- fertile 0.0.0.9027 --}}\end{kframe}
 \begin{kframe}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps ---------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps ------------------------- fertile 0.0.0.9027 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code -------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code ----------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 9 x 3
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory ---------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory ------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 9 x 4
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files -------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files ----------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## # A tibble: 7 x 3
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged ------------------------------------------ fertile 0.0.0.9027 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged --------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
 ## NULL
@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-10-08 14:49:12
-## 2 package:for~ <NA>                                 base::l~ 2020-10-08 14:49:12
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-08 14:49:12
+## 1 package:pur~ <NA>                                 base::l~ 2020-10-09 15:45:45
+## 2 package:for~ <NA>                                 base::l~ 2020-10-09 15:45:45
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-09 15:45:45
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x7fc745046ad8>
+## <bytecode: 0x7fb02d55d1b0>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpWD3heO/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpOuHj47/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpWD3heO/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpWD3heO/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpWD3heO/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpOuHj47/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpOuHj47/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpOuHj47/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}

### 2020-10-09 STAT-paper-template/APA/fertile.Rnw
added one more bullet point
added table; close #18

@@
 \usepackage{xcolor}
+% kableExtra packages: https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf
+\usepackage{booktabs}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+%\usepackage[normalem]{ulem}
+\usepackage{makecell}
 \titlespacing{\section}{0pt}{10pt plus 2pt minus 1pt}{10pt plus 2pt minus 1pt}
@@
 These tests include: looking for a clear build chain, checking to make sure the root level of the project is clear of clutter, confirming that there are no files present that are not being directly used by or created by the code, and looking for uses of randomness that do not have a call to \func{set.seed} present.
-A full list is provided below:
+A full list is provided in Table \ref{tab:list-of-checks}:
 
@@
 
-<<list of checks, echo = FALSE>>=
+<<list-of-checks, echo = FALSE>>=
 check_name <- c("has_proj_root()",
@@
   "Looks for indication of a clear order in which to run R scripts.",
-  "There are six different ”has_tidy_...” functions. Each checks to make sure that no files of its given type (images, code, data, etc.) are located in the root directory of the project",
+  "There are six different 'has_tidy_...' functions. Each checks to make sure that no files of its given type (images, code, data, etc.) are located in the root directory of the project",
   "Ensures that all files located in the project directory are being referenced or created by other files in the project.",
@@
   "Checks to make sure code does not contain uncontrolled randomness.",
-  "Examines code to determine whether it conforms to ”tidy” style guide recommendations.")
+  "Examines code to determine whether it conforms to 'tidy' style guide recommendations.")
 
-checks_list <- tidyr::tibble(check_name, description)
+library(magrittr)
+tibble::tibble("Check" = check_name, "Functionality" = description) %>%
+  kableExtra::kbl( 
+    caption = "Table illustrating the list of check available in fertile. ",
+    format = "latex",
+    booktabs = TRUE
+  ) %>%
+  kableExtra::kable_styling(latex_options = c("striped"), full_width = TRUE) %>%
+#  kableExtra::column_spec(1, width = "2in") %>%
+  kableExtra::column_spec(2, width = "4in")
 @
-
-\includegraphics[width=500pt,height=280pt]{list-checks}
 

### 2020-10-09 STAT-paper-template/APA/response-to-reviewers.md
updating w/ fertile

@@
 
-Additionally, a new bullet point on code commenting (relevant to the has_well_commented_code() function as well as one of the added sources, has been added to the summary of reproducibility components.
+Additionally, a new bullet point on code commenting (relevant to the has_well_commented_code() function as well as one of the added sources), has been added to the summary of reproducibility components.

### 2020-10-09 STAT-paper-template/APA/fertile.Rnw
updating w/ fertile

@@
 library(fertile)
+library(readr)
 file.exists("~/Desktop/my_data.csv")
 read.csv("~/Desktop/my_data.csv")
-read.csv("../../../Desktop/my_data.csv")
+read_csv("../../../Desktop/my_data.csv")
 @
 
-\textcolor{blue}{Users who want to override this behavior and execute their function anyway are provided with a explanation of how to do so.}
+\textcolor{blue}{Users who want to override this behavior and execute their function anyway are provided with a piece of code to execute that will achieve this for them.}
 

### 2020-10-09 STAT-paper-template/APA/fertile.tex
updating w/ fertile

@@
 \hlkwd{library}\hlstd{(fertile)}
+\hlkwd{library}\hlstd{(readr)}
 \hlkwd{file.exists}\hlstd{(}\hlstr{"~/Desktop/my_data.csv"}\hlstd{)}
@@
 Absolute paths are not reproducible and will likely only work on your computer.
-If you would like to continue anyway, re-execute your command by typing the name of the package it originates from followed by :: and the function call, e.g.
-if you previously typed 'read\_csv(path)', instead use 'readr::read\_csv(path)'}}\begin{alltt}
-\hlkwd{read.csv}\hlstd{(}\hlstr{"../../../Desktop/my_data.csv"}\hlstd{)}
+If you would like to continue anyway, please execute the following command: utils::read.csv('\textasciitilde{}/Desktop/my\_data.csv')}}\begin{alltt}
+\hlkwd{read_csv}\hlstd{(}\hlstr{"../../../Desktop/my_data.csv"}\hlstd{)}
 \end{alltt}
@@
 Such paths are not reproducible and will likely only work on your computer.
-If you would like to continue anyway, re-execute your command by typing the name of the package it originates from followed by :: and the function call, e.g.
-if you previously typed 'read\_csv(path)', instead use 'readr::read\_csv(path)'}}\end{kframe}
+If you would like to continue anyway, please execute the following command: readr::read\_csv('../../../Desktop/my\_data.csv')}}\end{kframe}
 \end{knitrout}
 
-\textcolor{blue}{Users who want to override this behavior and execute their function anyway are provided with a explanation of how to do so.}
+\textcolor{blue}{Users who want to override this behavior and execute their function anyway are provided with a piece of code to execute that will achieve this for them.}
 
@@
 
-\includegraphics[width=500pt,height=280pt]{list-checks}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{table}
+
+\caption{\label{tab:list-of-checks}Table illustrating the list of check available in fertile. }
+\centering
+\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright\arraybackslash}p{4in}}
+\toprule
+Check & Functionality\\
+\midrule
+\cellcolor{gray!6}{has\_proj\_root()} & \cellcolor{gray!6}{Checks to make sure a single .Rproj file is found in the root directory of the project.}\\
+has\_no\_nested\_proj\_root() & Checks to make sure there are no nested .Rproj files in subfolders of the project.\\
+\cellcolor{gray!6}{has\_clear\_build\_chain()} & \cellcolor{gray!6}{Looks for indication of a clear order in which to run R scripts.}\\
+has\_tidy\_[media/images/code/raw\_data/data/scripts]() & There are six different 'has\_tidy\_...' functions.
+Each checks to make sure that no files of its given type (images, code, data, etc.) are located in the root directory of the project\\
+\cellcolor{gray!6}{has\_only\_used\_files()} & \cellcolor{gray!6}{Ensures that all files located in the project directory are being referenced or created by other files in the project.}\\
+\addlinespace
+has\_readme() & Checks to make sure there is a README file in the root directory.\\
+\cellcolor{gray!6}{has\_well\_commented\_code()} & \cellcolor{gray!6}{Checks to ensure that all code files contain an adequate number of comments (>10\% of lines).}\\
+has\_no\_absolute\_paths() & Checks to make sure that no absolute paths are referenced in code.\\
+\cellcolor{gray!6}{has\_only\_portable\_paths()} & \cellcolor{gray!6}{Checks to make sure that no paths in code lead outside the project directory.}\\
+has\_no\_randomness() & Checks to make sure code does not contain uncontrolled randomness.\\
+\addlinespace
+\cellcolor{gray!6}{has\_no\_lint()} & \cellcolor{gray!6}{Examines code to determine whether it conforms to 'tidy' style guide recommendations.}\\
+\bottomrule
+\end{tabu}
+\end{table}
+
+\end{knitrout}
 
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... ----------------------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... --------------------------------------------------- fertile 0.0.0.9028 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ------------------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ----------------------------------------- fertile 0.0.0.9028 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks ----------------------------------------- fertile 0.0.0.9027 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Error in eval\_tidy(xs[[j]], mask): object '.Random.seed' not found}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# v Checking for no absolute paths}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks --------------------------------- fertile 0.0.0.9028 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# v Checking for only portable paths}}\end{kframe}
-\begin{kframe}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks ---------------------------------------------- fertile 0.0.0.9027 --}}\end{kframe}
-\begin{kframe}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# v Reproducibility checks passed: 2}}\end{kframe}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: arrange() failed at implicit mutate() step.
+\\\#\# x Could not create a temporary column for `..1`.\\\#\# i `..1` is `timestamp`.}}\end{kframe}
 \end{knitrout}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps ------------------------- fertile 0.0.0.9027 --}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code ----------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
-## # A tibble: 9 x 3
-##   package       N used_in                    
-##   <chr>     <int> <chr>                      
-## 1 broom         1 project_miceps/analysis.Rmd
-## 2 dplyr         1 project_miceps/analysis.Rmd
-## 3 ggplot2       1 project_miceps/analysis.Rmd
-## 4 purrr         1 project_miceps/analysis.Rmd
-## 5 readr         1 project_miceps/analysis.Rmd
-## 6 rmarkdown     1 project_miceps/analysis.Rmd
-## 7 skimr         1 project_miceps/analysis.Rmd
-## 8 stargazer     1 project_miceps/analysis.Rmd
-## 9 tidyr         1 project_miceps/analysis.Rmd
-\end{verbatim}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory ------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
-## # A tibble: 9 x 4
-##   file               ext        size mime                                       
-##   <fs::path>         <chr> <fs::byt> <chr>                                      
-## 1 Estrogen_Receptor~ docx     10.97K application/vnd.openxmlformats-officedocum~
-## 2 citrate_v_time.png png     184.04K image/png                                  
-## 3 proteins_v_time.p~ png     377.02K image/png                                  
-## 4 Blot_data_updated~ csv      14.43K text/csv                                   
-## 5 CS_data_redone.csv csv       7.39K text/csv                                   
-## 6 mice.csv           csv      14.33K text/csv                                   
-## 7 README.md          md           39 text/markdown                              
-## 8 miceps.Rproj       Rproj       204 text/rstudio                               
-## 9 analysis.Rmd       Rmd       4.94K text/x-markdown
-\end{verbatim}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files ----------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
-## # A tibble: 7 x 3
-##   path_rel           dir_rel    cmd                                             
-##   <fs::path>         <fs::path> <chr>                                           
-## 1 Blot_data_updated~ data-raw   file_move('project_miceps/Blot_data_updated.csv~
-## 2 CS_data_redone.csv data-raw   file_move('project_miceps/CS_data_redone.csv', ~
-## 3 Estrogen_Receptor~ inst/other file_move('project_miceps/Estrogen_Receptors.do~
-## 4 analysis.Rmd       vignettes  file_move('project_miceps/analysis.Rmd', fs::di~
-## 5 citrate_v_time.png inst/image file_move('project_miceps/citrate_v_time.png', ~
-## 6 mice.csv           data-raw   file_move('project_miceps/mice.csv', fs::dir_cr~
-## 7 proteins_v_time.p~ inst/image file_move('project_miceps/proteins_v_time.png',~
-\end{verbatim}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged --------------------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
-## NULL
-\end{verbatim}
-\end{kframe}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: arrange() failed at implicit mutate() step.
+\\\#\# x Could not create a temporary column for `..1`.\\\#\# i `..1` is `timestamp`.}}\end{kframe}
 \end{knitrout}
@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-10-09 15:45:45
-## 2 package:for~ <NA>                                 base::l~ 2020-10-09 15:45:45
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-09 15:45:45
+## 1 package:pur~ <NA>                                 base::l~ 2020-10-09 23:56:28
+## 2 package:for~ <NA>                                 base::l~ 2020-10-09 23:56:28
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-09 23:56:28
 \end{verbatim}
@@
 ##     log_push(file, "readr::read_csv")
-##     check_path_safe(file)
+##     check_path_safe(file, ... = "readr::read_csv")
 ##     readr::read_csv(file, ...)
@@
 ## }
-## <bytecode: 0x7fb02d55d1b0>
+## <bytecode: 0x7fe67a348478>
 ## <environment: namespace:fertile>
@@
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Detected absolute paths.
-Absolute paths are not reproducible and will likely only work on your computer.
-If you would like to continue anyway, re-execute your command by typing the name of the package it originates from followed by :: and the function call, e.g.
-if you previously typed 'read\_csv(path)', instead use 'readr::read\_csv(path)'}}\begin{alltt}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Detected absolute paths. Absolute paths are not reproducible and will likely only work on your computer.}}\begin{alltt}
 \hlcom{# Path outside the directory}
@@
 {\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Detected paths that lead outside the project directory.
-Such paths are not reproducible and will likely only work on your computer.
-If you would like to continue anyway, re-execute your command by typing the name of the package it originates from followed by :: and the function call, e.g.
-if you previously typed 'read\_csv(path)', instead use 'readr::read\_csv(path)'}}\end{kframe}
+Such paths are not reproducible and will likely only work on your computer.}}\end{kframe}
 \end{knitrout}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpOuHj47/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpisOoIV/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpOuHj47/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpOuHj47/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpOuHj47/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpisOoIV/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpisOoIV/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpisOoIV/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}

### 2020-10-10 STAT-paper-template/APA/fertile.tex
added log example

@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... --------------------------------------------------- fertile 0.0.0.9028 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... -------------------------- fertile 0.0.0.9028 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ----------------------------------------- fertile 0.0.0.9028 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ---------------- fertile 0.0.0.9028 --}}
 
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks --------------------------------- fertile 0.0.0.9028 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks -------- fertile 0.0.0.9028 --}}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: arrange() failed at implicit mutate() step.
-\\\#\# x Could not create a temporary column for `..1`.\\\#\# i `..1` is `timestamp`.}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ---------------- fertile 0.0.0.9028 --}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval\_tidy(xs[[j]], mask): object '.Random.seed' not found}}\end{kframe}
 \end{knitrout}
@@
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: arrange() failed at implicit mutate() step.
-\\\#\# x Could not create a temporary column for `..1`.\\\#\# i `..1` is `timestamp`.}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ---------------- fertile 0.0.0.9028 --}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval\_tidy(xs[[j]], mask): object '.Random.seed' not found}}\end{kframe}
 \end{knitrout}
@@
 \begin{alltt}
+\hlkwd{library}\hlstd{(purrr)}
+\hlkwd{library}\hlstd{(forcats)}
+\hlkwd{read_csv}\hlstd{(fs}\hlopt{::}\hlkwd{path}\hlstd{(proj_dir,} \hlstr{"mice.csv"}\hlstd{))}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
 \hlkwd{log_report}\hlstd{()}
@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-10-09 23:56:28
-## 2 package:for~ <NA>                                 base::l~ 2020-10-09 23:56:28
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-09 23:56:28
+## 1 package:pur~ <NA>                                 base::l~ 2020-10-10 00:39:23
+## 2 package:for~ <NA>                                 base::l~ 2020-10-10 00:39:23
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-10 00:39:23
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x7fe67a348478>
+## <bytecode: 0x7fcede54ad08>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpisOoIV/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNAq0UZ/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpisOoIV/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpisOoIV/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpisOoIV/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNAq0UZ/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNAq0UZ/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNAq0UZ/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}

### 2020-10-10 STAT-paper-template/APA/fertile.tex
hard-coded latex file

@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... -------------------------- fertile 0.0.0.9028 --}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ---------------- fertile 0.0.0.9028 --}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Error in eval\_tidy(xs[[j]], mask): object '.Random.seed' not found}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks -------- fertile 0.0.0.9028 --}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ---------------- fertile 0.0.0.9028 --}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval\_tidy(xs[[j]], mask): object '.Random.seed' not found}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... ---------------------------------------------- fertile 0.0.0.9027 --}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ------------------------------------ fertile 0.0.0.9027 --}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks ---------------------------- fertile 0.0.0.9027 --}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# v Checking for no absolute paths}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# v Checking for only portable paths}}\end{kframe}
+\begin{kframe}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks --------------------------------- fertile 0.0.0.9027 --}}\end{kframe}
+\begin{kframe}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# v Reproducibility checks passed: 2}}\end{kframe}
 \end{knitrout}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ---------------- fertile 0.0.0.9028 --}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval\_tidy(xs[[j]], mask): object '.Random.seed' not found}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps ------------ fertile 0.0.0.9027 --}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code ---------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+## # A tibble: 9 x 3
+##   package       N used_in                    
+##   <chr>     <int> <chr>                      
+## 1 broom         1 project_miceps/analysis.Rmd
+## 2 dplyr         1 project_miceps/analysis.Rmd
+## 3 ggplot2       1 project_miceps/analysis.Rmd
+## 4 purrr         1 project_miceps/analysis.Rmd
+## 5 readr         1 project_miceps/analysis.Rmd
+## 6 rmarkdown     1 project_miceps/analysis.Rmd
+## 7 skimr         1 project_miceps/analysis.Rmd
+## 8 stargazer     1 project_miceps/analysis.Rmd
+## 9 tidyr         1 project_miceps/analysis.Rmd
+\end{verbatim}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory ------------------------------ fertile 0.0.0.9027 --}}\begin{verbatim}
+##                      file   ext    size
+## 1 Estrogen_Receptors.docx  docx  10.97K
+## 2      citrate_v_time.png   png 188.08K
+## 3     proteins_v_time.png   png 378.45K
+## 4   Blot_data_updated.csv   csv  14.43K
+## 5      CS_data_redone.csv   csv   7.39K
+## 6                mice.csv   csv  14.33K
+## 7               README.md    md      39
+## 8            miceps.Rproj Rproj     204
+## 9            analysis.Rmd   Rmd   4.94K
+##                                                                      mime
+## 1 application/vnd.openxmlformats-officedocument.wordprocessingml.document
+## 2                                                               image/png
+## 3                                                               image/png
+## 4                                                                text/csv
+## 5                                                                text/csv
+## 6                                                                text/csv
+## 7                                                           text/markdown
+## 8                                                            text/rstudio
+## 9                                                         text/x-markdown
+\end{verbatim}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files ---------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+##                  path_rel    dir_rel
+## 1   Blot_data_updated.csv   data-raw
+## 2      CS_data_redone.csv   data-raw
+## 3 Estrogen_Receptors.docx inst/other
+## 4            analysis.Rmd  vignettes
+## 5      citrate_v_time.png inst/image
+## 6                mice.csv   data-raw
+## 7     proteins_v_time.png inst/image
+##                                                                                                cmd
+## 1     file_move('project_miceps/Blot_data_updated.csv', fs::dir_create('project_miceps/data-raw'))
+## 2        file_move('project_miceps/CS_data_redone.csv', fs::dir_create('project_miceps/data-raw'))
+## 3 file_move('project_miceps/Estrogen_Receptors.docx', fs::dir_create('project_miceps/inst/other'))
+## 4             file_move('project_miceps/analysis.Rmd', fs::dir_create('project_miceps/vignettes'))
+## 5      file_move('project_miceps/citrate_v_time.png', fs::dir_create('project_miceps/inst/image'))
+## 6                  file_move('project_miceps/mice.csv', fs::dir_create('project_miceps/data-raw'))
+## 7     file_move('project_miceps/proteins_v_time.png', fs::dir_create('project_miceps/inst/image'))
+\end{verbatim}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged -------------------------------- fertile 0.0.0.9027 --}}\begin{verbatim}
+## NULL
+\end{verbatim}
+\end{kframe}
 \end{knitrout}
@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-10-10 00:39:23
-## 2 package:for~ <NA>                                 base::l~ 2020-10-10 00:39:23
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-10 00:39:23
+## 1 package:pur~ <NA>                                 base::l~ 2020-10-10 13:33:50
+## 2 package:for~ <NA>                                 base::l~ 2020-10-10 13:33:50
+## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-10 13:33:50
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x7fcede54ad08>
+## <bytecode: 0x7fea6a6f2048>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNAq0UZ/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpxR3hjP/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNAq0UZ/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNAq0UZ/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNAq0UZ/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpxR3hjP/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpxR3hjP/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpxR3hjP/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}
@@
 \nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;
-\bibliography{refs}%
+%\bibliography{refs}%
+
+\begin{thebibliography}{}
+
+\bibitem [\protect \citeauthoryear {%
+Baker%
+}{%
+Baker%
+}{%
+{\protect \APACyear {2016}}%
+}]{%
+baker20161}
+\APACinsertmetastar {%
+baker20161}%
+\begin{APACrefauthors}%
+Baker, M.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2016}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {1,500 scientists lift the lid on reproducibility} {1,500
+  scientists lift the lid on reproducibility}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Nature News}{533}{7604}{452}.
+\newblock
+\begin{APACrefURL}
+  \url{https://www.nature.com/news/1-500-scientists-lift-the-lid-on-reproducibility-1.19970}
+  \end{APACrefURL}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Baumer%
+\ \BBA {} Bertin%
+}{%
+Baumer%
+\ \BBA {} Bertin%
+}{%
+{\protect \APACyear {2020}}%
+}]{%
+R-fertile}
+\APACinsertmetastar {%
+R-fertile}%
+\begin{APACrefauthors}%
+Baumer, B\BPBI S.%
+\BCBT {}\ \BBA {} Bertin, A\BPBI M.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2020}{}{}.
+\newblock
+\APACrefbtitle {fertile: creating optimal conditions for reproducibility.}
+  {fertile: creating optimal conditions for reproducibility.}
+\newblock
+\begin{APACrefURL} \url{https://github.com/baumer-lab/fertile} \end{APACrefURL}
+\newblock
+\APACrefnote{R package version 0.0.0.9027}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+J.~Blischak%
+, Carbonetto%
+\BCBL {}\ \BBA {} Stephens%
+}{%
+J.~Blischak%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2019}}%
+}]{%
+R-workflowr}
+\APACinsertmetastar {%
+R-workflowr}%
+\begin{APACrefauthors}%
+Blischak, J.%
+, Carbonetto, P.%
+\BCBL {}\ \BBA {} Stephens, M.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2019}{}{}.
+\newblock
+\APACrefbtitle {workflowr: A Framework for Reproducible and Collaborative Data
+  Science.} {workflowr: A framework for reproducible and collaborative data
+  science.}
+\newblock
+\begin{APACrefURL} \url{https://CRAN.R-project.org/package=workflowr}
+  \end{APACrefURL}
+\newblock
+\APACrefnote{R package version 1.6.0}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+J\BPBI D.~Blischak%
+, Carbonetto%
+\BCBL {}\ \BBA {} Stephens%
+}{%
+J\BPBI D.~Blischak%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2019}}%
+}]{%
+workflowr2019}
+\APACinsertmetastar {%
+workflowr2019}%
+\begin{APACrefauthors}%
+Blischak, J\BPBI D.%
+, Carbonetto, P.%
+\BCBL {}\ \BBA {} Stephens, M.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2019}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Creating and sharing reproducible research code the
+  workflowr way} {Creating and sharing reproducible research code the workflowr
+  way}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{F1000Research}{8}{1749}{}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.12688/f1000research.20843.1}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} 10.12688/f1000research.20843.1 \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Broman%
+}{%
+Broman%
+}{%
+{\protect \APACyear {2019}}%
+}]{%
+broman}
+\APACinsertmetastar {%
+broman}%
+\begin{APACrefauthors}%
+Broman, K.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2019}{}{}.
+\newblock
+\APACrefbtitle {initial steps toward reproducible research: organize your data
+  and code.} {initial steps toward reproducible research: organize your data
+  and code.}
+\newblock
+\begin{APACrefURL} \url{https://kbroman.org/steps2rr/pages/organize.html}
+  \end{APACrefURL}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Cooper%
+\ \protect \BOthers {.}}{%
+Cooper%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2017}}%
+}]{%
+cooper2017guide}
+\APACinsertmetastar {%
+cooper2017guide}%
+\begin{APACrefauthors}%
+Cooper, N.%
+, Hsing, P\BHBI Y.%
+, Croucher, M.%
+, Graham, L.%
+, James, T.%
+, Krystalli, A.%
+\BCBL {}\ \BBA {} Michonneau, F.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2017}{}{}.
+\newblock
+\APACrefbtitle {A guide to reproducible code in ecology and evolution.} {A
+  guide to reproducible code in ecology and evolution.}
+\newblock
+\begin{APACrefURL}
+  \url{https://www.britishecologicalsociety.org/wp-content/uploads/2017/12/guide-to-reproducible-code.pdf}
+  \end{APACrefURL}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Eisner%
+}{%
+Eisner%
+}{%
+{\protect \APACyear {2018}}%
+}]{%
+eisner-reproducibility}
+\APACinsertmetastar {%
+eisner-reproducibility}%
+\begin{APACrefauthors}%
+Eisner, D\BPBI A.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2018}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Reproducibility of science: Fraud, impact factors and
+  carelessness} {Reproducibility of science: Fraud, impact factors and
+  carelessness}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Journal of Molecular and Cellular
+  Cardiology}{114}{}{364-368}.
+\newblock
+\begin{APACrefURL}
+  \url{http://www.sciencedirect.com/science/article/pii/S0022282817303334}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} https://doi.org/10.1016/j.yjmcc.2017.10.009 \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Fidler%
+\ \BBA {} Wilcox%
+}{%
+Fidler%
+\ \BBA {} Wilcox%
+}{%
+{\protect \APACyear {2018}}%
+}]{%
+sep-scientific-reproducibility}
+\APACinsertmetastar {%
+sep-scientific-reproducibility}%
+\begin{APACrefauthors}%
+Fidler, F.%
+\BCBT {}\ \BBA {} Wilcox, J.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2018}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Reproducibility of Scientific Results} {Reproducibility
+  of scientific results}.{\BBCQ}
+\newblock
+\BIn{} E\BPBI N.~Zalta\ (\BED), \APACrefbtitle {The Stanford Encyclopedia of
+  Philosophy} {The stanford encyclopedia of philosophy}\ (\PrintOrdinal{Winter
+  2018}\ \BEd).
+\newblock
+\APACaddressPublisher{}{Metaphysics Research Lab, Stanford University}.
+\newblock
+\APAChowpublished
+  {\url{https://plato.stanford.edu/archives/win2018/entries/scientific-reproducibility/}}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+FitzJohn%
+\ \protect \BOthers {.}}{%
+FitzJohn%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2020}}%
+}]{%
+R-orderly}
+\APACinsertmetastar {%
+R-orderly}%
+\begin{APACrefauthors}%
+FitzJohn, R.%
+, Ashton, R.%
+, Hill, A.%
+, Eden, M.%
+, Hinsley, W.%
+, Russell, E.%
+\BCBL {}\ \BBA {} Thompson, J.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2020}{}{}.
+\newblock
+\APACrefbtitle {orderly: Lightweight Reproducible Reporting.} {orderly:
+  Lightweight reproducible reporting.}
+\newblock
+\begin{APACrefURL} \url{https://CRAN.R-project.org/package=orderly}
+  \end{APACrefURL}
+\newblock
+\APACrefnote{R package version 1.0.4}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Gancarz%
+}{%
+Gancarz%
+}{%
+{\protect \APACyear {2003}}%
+}]{%
+unix}
+\APACinsertmetastar {%
+unix}%
+\begin{APACrefauthors}%
+Gancarz, M.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYear{2003}.
+\newblock
+\APACrefbtitle {Linux and the Unix Philosophy} {Linux and the unix philosophy}\
+  (\PrintOrdinal{2nd}\ \BEd).
+\newblock
+\APACaddressPublisher{Woburn, MA}{Digital Press}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Goodman%
+, Fanelli%
+\BCBL {}\ \BBA {} Ioannidis%
+}{%
+Goodman%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2016}}%
+}]{%
+Goodman341ps12}
+\APACinsertmetastar {%
+Goodman341ps12}%
+\begin{APACrefauthors}%
+Goodman, S\BPBI N.%
+, Fanelli, D.%
+\BCBL {}\ \BBA {} Ioannidis, J\BPBI P\BPBI A.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2016}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {What does research reproducibility mean?} {What does
+  research reproducibility mean?}{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Science Translational Medicine}{8}{341}{1--6}.
+\newblock
+\begin{APACrefURL} \url{https://stm.sciencemag.org/content/8/341/341ps12}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} 10.1126/scitranslmed.aaf5027 \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Gosselin%
+}{%
+Gosselin%
+}{%
+{\protect \APACyear {2020}}%
+}]{%
+bioessays-gosselin}
+\APACinsertmetastar {%
+bioessays-gosselin}%
+\begin{APACrefauthors}%
+Gosselin, R\BHBI D.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2020}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Statistical Analysis Must Improve to Address the
+  Reproducibility Crisis: The ACcess to Transparent Statistics (ACTS) Call to
+  Action} {Statistical analysis must improve to address the reproducibility
+  crisis: The access to transparent statistics (acts) call to action}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{BioEssays}{42}{1}{1900189}.
+\newblock
+\begin{APACrefURL}
+  \url{https://onlinelibrary.wiley.com/doi/abs/10.1002/bies.201900189}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} 10.1002/bies.201900189 \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Henry%
+\ \BBA {} Wickham%
+}{%
+Henry%
+\ \BBA {} Wickham%
+}{%
+{\protect \APACyear {2020}}%
+}]{%
+R-tidyselect}
+\APACinsertmetastar {%
+R-tidyselect}%
+\begin{APACrefauthors}%
+Henry, L.%
+\BCBT {}\ \BBA {} Wickham, H.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2020}{}{}.
+\newblock
+\APACrefbtitle {tidyselect: Select from a Set of Strings.} {tidyselect: Select
+  from a set of strings.}
+\newblock
+\begin{APACrefURL} \url{https://CRAN.R-project.org/package=tidyselect}
+  \end{APACrefURL}
+\newblock
+\APACrefnote{R package version 1.1.0}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Hermans%
+\ \BBA {} Aldewereld%
+}{%
+Hermans%
+\ \BBA {} Aldewereld%
+}{%
+{\protect \APACyear {2017}}%
+}]{%
+hermans2017programming}
+\APACinsertmetastar {%
+hermans2017programming}%
+\begin{APACrefauthors}%
+Hermans, F.%
+\BCBT {}\ \BBA {} Aldewereld, M.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2017}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Programming is writing is programming} {Programming is
+  writing is programming}.{\BBCQ}
+\newblock
+\BIn{} \APACrefbtitle {Companion to the first International Conference on the
+  Art, Science and Engineering of Programming} {Companion to the first
+  international conference on the art, science and engineering of programming}\
+  (\BPGS\ 1--8).
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Kitzes%
+, Turek%
+\BCBL {}\ \BBA {} Deniz%
+}{%
+Kitzes%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2017}}%
+}]{%
+kitzes2017practice}
+\APACinsertmetastar {%
+kitzes2017practice}%
+\begin{APACrefauthors}%
+Kitzes, J.%
+, Turek, D.%
+\BCBL {}\ \BBA {} Deniz, F.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYear{2017}.
+\newblock
+\APACrefbtitle {The Practice of Reproducible Research: Case Studies and Lessons
+  from the Data-Intensive Sciences} {The practice of reproducible research:
+  Case studies and lessons from the data-intensive sciences}.
+\newblock
+\APACaddressPublisher{Berkeley, CA}{University of California Press}.
+\newblock
+\begin{APACrefURL} \url{https://www.practicereproducibleresearch.org}
+  \end{APACrefURL}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Martinez%
+\ \protect \BOthers {.}}{%
+Martinez%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2018}}%
+}]{%
+r-opensci}
+\APACinsertmetastar {%
+r-opensci}%
+\begin{APACrefauthors}%
+Martinez, C.%
+, Hollister, J.%
+, Marwick, B.%
+, Sz{\"o}cs, E.%
+, Zeitlin, S.%
+, Kinoshita, B\BPBI P.%
+\BDBL {}Meinke, B.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2018}{}{}.
+\newblock
+\APACrefbtitle {{Reproducibility in Science: A Guide to enhancing
+  reproducibility in scientific results and writing}.} {{Reproducibility in
+  Science: A Guide to enhancing reproducibility in scientific results and
+  writing}.}
+\newblock
+\begin{APACrefURL} \url{http://ropensci.github.io/reproducibility-guide/}
+  \end{APACrefURL}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Marwick%
+}{%
+Marwick%
+}{%
+{\protect \APACyear {2019}}%
+}]{%
+R-rrtools}
+\APACinsertmetastar {%
+R-rrtools}%
+\begin{APACrefauthors}%
+Marwick, B.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2019}{}{}.
+\newblock
+\APACrefbtitle {rrtools: Creates a Reproducible Research Compendium.} {rrtools:
+  Creates a reproducible research compendium.}
+\newblock
+\begin{APACrefURL} \url{https://github.com/benmarwick/rrtools} \end{APACrefURL}
+\newblock
+\APACrefnote{R package version 0.1.0}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Marwick%
+, Boettiger%
+\BCBL {}\ \BBA {} Mullen%
+}{%
+Marwick%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2018}}%
+}]{%
+marwick2018packaging}
+\APACinsertmetastar {%
+marwick2018packaging}%
+\begin{APACrefauthors}%
+Marwick, B.%
+, Boettiger, C.%
+\BCBL {}\ \BBA {} Mullen, L.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2018}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Packaging data analytical work reproducibly using {R}
+  (and friends)} {Packaging data analytical work reproducibly using {R} (and
+  friends)}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{The American Statistician}{72}{1}{80--88}.
+\newblock
+\begin{APACrefURL} \url{https://peerj.com/preprints/3192.pdf} \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} doi.org/10.1080/00031305.2017.1375986 \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+McArthur%
+}{%
+McArthur%
+}{%
+{\protect \APACyear {2019}}%
+}]{%
+engineering-reproducibility}
+\APACinsertmetastar {%
+engineering-reproducibility}%
+\begin{APACrefauthors}%
+McArthur, S\BPBI L.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2019}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Repeatability, Reproducibility, and Replicability:
+  Tackling the 3R challenge in biointerface science and engineering}
+  {Repeatability, reproducibility, and replicability: Tackling the 3r challenge
+  in biointerface science and engineering}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Biointerphases}{14}{2}{1--2}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1116/1.5093621} \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} 10.1116/1.5093621 \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+McIntire%
+\ \BBA {} Chubaty%
+}{%
+McIntire%
+\ \BBA {} Chubaty%
+}{%
+{\protect \APACyear {2020}}%
+}]{%
+R-reproducible}
+\APACinsertmetastar {%
+R-reproducible}%
+\begin{APACrefauthors}%
+McIntire, E\BPBI J\BPBI B.%
+\BCBT {}\ \BBA {} Chubaty, A\BPBI M.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2020}{}{}.
+\newblock
+\APACrefbtitle {reproducible: A Set of Tools that Enhance Reproducibility
+  Beyond Package Management.} {reproducible: A set of tools that enhance
+  reproducibility beyond package management.}
+\newblock
+\begin{APACrefURL} \url{https://CRAN.R-project.org/package=reproducible}
+  \end{APACrefURL}
+\newblock
+\APACrefnote{R package version 1.0.0}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+OpenSci%
+}{%
+OpenSci%
+}{%
+{\protect \APACyear {2020}}%
+}]{%
+R-drake}
+\APACinsertmetastar {%
+R-drake}%
+\begin{APACrefauthors}%
+OpenSci, R.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2020}{}{}.
+\newblock
+\APACrefbtitle {drake: A Pipeline Toolkit for Reproducible Computation at
+  Scale.} {drake: A pipeline toolkit for reproducible computation at scale.}
+\newblock
+\begin{APACrefURL} \url{https://cran.r-project.org/package=drake}
+  \end{APACrefURL}
+\newblock
+\APACrefnote{R package version 7.11.0}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+{Oracle Corporation}%
+}{%
+{Oracle Corporation}%
+}{%
+{\protect \APACyear {2019}}%
+}]{%
+wercker}
+\APACinsertmetastar {%
+wercker}%
+\begin{APACrefauthors}%
+{Oracle Corporation}.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2019}{}{}.
+\newblock
+\APACrefbtitle {Wercker.} {Wercker.}
+\newblock
+\begin{APACrefURL} \url{https://github.com/wercker/wercker} \end{APACrefURL}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Patil%
+, Peng%
+\BCBL {}\ \BBA {} Leek%
+}{%
+Patil%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2019}}%
+}]{%
+patil2019visual}
+\APACinsertmetastar {%
+patil2019visual}%
+\begin{APACrefauthors}%
+Patil, P.%
+, Peng, R\BPBI D.%
+\BCBL {}\ \BBA {} Leek, J\BPBI T.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2019}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {A visual tool for defining reproducibility and
+  replicability} {A visual tool for defining reproducibility and
+  replicability}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Nature human behaviour}{3}{7}{650--652}.
+\newblock
+\begin{APACrefURL} \url{https://www.nature.com/articles/s41562-019-0629-z}
+  \end{APACrefURL}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Peng%
+}{%
+Peng%
+}{%
+{\protect \APACyear {2009}}%
+}]{%
+peng2009bio}
+\APACinsertmetastar {%
+peng2009bio}%
+\begin{APACrefauthors}%
+Peng, R\BPBI D.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2009}{07}{}.
+\newblock
+{\BBOQ}\APACrefatitle {{Reproducible research and Biostatistics}}
+  {{Reproducible research and Biostatistics}}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Biostatistics}{10}{3}{405-408}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1093/biostatistics/kxp014}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} 10.1093/biostatistics/kxp014 \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Popper%
+}{%
+Popper%
+}{%
+{\protect \APACyear {2005}}%
+}]{%
+popper2005logic}
+\APACinsertmetastar {%
+popper2005logic}%
+\begin{APACrefauthors}%
+Popper, K.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYear{2005}.
+\newblock
+\APACrefbtitle {The logic of scientific discovery} {The logic of scientific
+  discovery}.
+\newblock
+\APACaddressPublisher{New York, NY}{Routledge}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+{R Core Team}%
+}{%
+{R Core Team}%
+}{%
+{\protect \APACyear {2020}}%
+}]{%
+r-software}
+\APACinsertmetastar {%
+r-software}%
+\begin{APACrefauthors}%
+{R Core Team}.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2020}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {R: A Language and Environment for Statistical Computing}
+  {R: A language and environment for statistical computing}{\BBCQ}\
+  [\bibcomputersoftwaremanual].
+\newblock
+\APACaddressPublisher{Vienna, Austria}{}.
+\newblock
+\begin{APACrefURL} \url{https://www.R-project.org/} \end{APACrefURL}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+R-Core-Team%
+}{%
+R-Core-Team%
+}{%
+{\protect \APACyear {2020}}%
+}]{%
+coreteam-extensions}
+\APACinsertmetastar {%
+coreteam-extensions}%
+\begin{APACrefauthors}%
+R-Core-Team.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2020}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Writing R Extensions} {Writing r extensions}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{R Foundation for Statistical Computing}{}{}{}.
+\newblock
+\begin{APACrefURL}
+  \url{http://cran.stat.unipd.it/doc/manuals/r-release/R-exts.pdf}
+  \end{APACrefURL}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Ross%
+, DeCicco%
+\BCBL {}\ \BBA {} Randhawa%
+}{%
+Ross%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2018}}%
+}]{%
+R-checkers}
+\APACinsertmetastar {%
+R-checkers}%
+\begin{APACrefauthors}%
+Ross, N.%
+, DeCicco, L.%
+\BCBL {}\ \BBA {} Randhawa, N.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2018}{}{}.
+\newblock
+\APACrefbtitle {checkers: Automated checking of best practices for research
+  compendia.} {checkers: Automated checking of best practices for research
+  compendia.}
+\newblock
+\begin{APACrefURL}
+  \url{https://github.com/ropenscilabs/checkers/blob/master/DESCRIPTIONr}
+  \end{APACrefURL}
+\newblock
+\APACrefnote{R package version 0.1.0}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Ushey%
+, McPherson%
+, Cheng%
+, Atkins%
+\BCBL {}\ \BBA {} Allaire%
+}{%
+Ushey%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2018}}%
+}]{%
+R-packrat}
+\APACinsertmetastar {%
+R-packrat}%
+\begin{APACrefauthors}%
+Ushey, K.%
+, McPherson, J.%
+, Cheng, J.%
+, Atkins, A.%
+\BCBL {}\ \BBA {} Allaire, J.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2018}{}{}.
+\newblock
+\APACrefbtitle {packrat: A Dependency Management System for Projects and their
+  R Package Dependencies.} {packrat: A dependency management system for
+  projects and their r package dependencies.}
+\newblock
+\begin{APACrefURL} \url{https://CRAN.R-project.org/package=packrat}
+  \end{APACrefURL}
+\newblock
+\APACrefnote{R package version 0.5.0}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Wallach%
+, Boyack%
+\BCBL {}\ \BBA {} Ioannidis%
+}{%
+Wallach%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2018}}%
+}]{%
+plos-biology}
+\APACinsertmetastar {%
+plos-biology}%
+\begin{APACrefauthors}%
+Wallach, J\BPBI D.%
+, Boyack, K\BPBI W.%
+\BCBL {}\ \BBA {} Ioannidis, J\BPBI P\BPBI A.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2018}{11}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Reproducible research practices, transparency, and open
+  access data in the biomedical literature, 2015-2017} {Reproducible research
+  practices, transparency, and open access data in the biomedical literature,
+  2015-2017}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{PLOS Biology}{16}{11}{1-20}.
+\newblock
+\begin{APACrefURL} \url{https://doi.org/10.1371/journal.pbio.2006930}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} 10.1371/journal.pbio.2006930 \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Wickham%
+}{%
+Wickham%
+}{%
+{\protect \APACyear {2011}}%
+}]{%
+testthat2011}
+\APACinsertmetastar {%
+testthat2011}%
+\begin{APACrefauthors}%
+Wickham, H.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2011}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {testthat: Get Started with Testing} {testthat: Get
+  started with testing}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{The R Journal}{3}{}{5--10}.
+\newblock
+\begin{APACrefURL}
+  \url{https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf}
+  \end{APACrefURL}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Wickham%
+}{%
+Wickham%
+}{%
+{\protect \APACyear {2015}}%
+}]{%
+hadley-packages}
+\APACinsertmetastar {%
+hadley-packages}%
+\begin{APACrefauthors}%
+Wickham, H.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYear{2015}.
+\newblock
+\APACrefbtitle {R Packages} {R packages}\ (\PrintOrdinal{1st}\ \BEd).
+\newblock
+\APACaddressPublisher{Sebastopol, CA}{O'Reilly Media, Inc.}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Wickham%
+}{%
+Wickham%
+}{%
+{\protect \APACyear {2019}}%
+{\protect \APACexlab {{\protect \BCnt {1}}}}}]{%
+R-testthat}
+\APACinsertmetastar {%
+R-testthat}%
+\begin{APACrefauthors}%
+Wickham, H.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2019{\protect \BCnt {1}}}{}{}.
+\newblock
+\APACrefbtitle {testthat: Unit Testing for {R}.} {testthat: Unit testing for
+  {R}.}
+\newblock
+\begin{APACrefURL} \url{https://CRAN.R-project.org/package=testthat}
+  \end{APACrefURL}
+\newblock
+\APACrefnote{R package version 2.3.1}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Wickham%
+}{%
+Wickham%
+}{%
+{\protect \APACyear {2019}}%
+{\protect \APACexlab {{\protect \BCnt {2}}}}}]{%
+R-tidyverse}
+\APACinsertmetastar {%
+R-tidyverse}%
+\begin{APACrefauthors}%
+Wickham, H.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2019{\protect \BCnt {2}}}{}{}.
+\newblock
+\APACrefbtitle {tidyverse: Easily Install and Load the 'Tidyverse'.}
+  {tidyverse: Easily install and load the 'tidyverse'.}
+\newblock
+\begin{APACrefURL} \url{https://CRAN.R-project.org/package=tidyverse}
+  \end{APACrefURL}
+\newblock
+\APACrefnote{R package version 1.3.0}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Wickham%
+, Averick%
+\BCBL {}\ \protect \BOthers {.}}{%
+Wickham%
+, Averick%
+\BCBL {}\ \protect \BOthers {.}}{%
+{\protect \APACyear {2019}}%
+}]{%
+tidyverse2019}
+\APACinsertmetastar {%
+tidyverse2019}%
+\begin{APACrefauthors}%
+Wickham, H.%
+, Averick, M.%
+, Bryan, J.%
+, Chang, W.%
+, McGowan, L\BPBI D.%
+, Fran{\c c}ois, R.%
+\BDBL {}Yutani, H.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2019}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Welcome to the {tidyverse}} {Welcome to the
+  {tidyverse}}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{Journal of Open Source Software}{4}{43}{1686}.
+\newblock
+\begin{APACrefURL} \url{https://tidyverse.tidyverse.org/articles/paper.html}
+  \end{APACrefURL}
+\newblock
+\begin{APACrefDOI} 10.21105/joss.01686 \end{APACrefDOI}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Wickham%
+, Fran{\c c}ois%
+, Henry%
+\BCBL {}\ \BBA {} M{\"u}ller%
+}{%
+Wickham%
+, Fran{\c c}ois%
+\BCBL {}\ \protect \BOthers {.}}{%
+{\protect \APACyear {2019}}%
+}]{%
+R-dplyr}
+\APACinsertmetastar {%
+R-dplyr}%
+\begin{APACrefauthors}%
+Wickham, H.%
+, Fran{\c c}ois, R.%
+, Henry, L.%
+\BCBL {}\ \BBA {} M{\"u}ller, K.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2019}{}{}.
+\newblock
+\APACrefbtitle {dplyr: A Grammar of Data Manipulation.} {dplyr: A grammar of
+  data manipulation.}
+\newblock
+\begin{APACrefURL} \url{https://CRAN.R-project.org/package=dplyr}
+  \end{APACrefURL}
+\newblock
+\APACrefnote{R package version 0.8.3}
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Wilson%
+\ \protect \BOthers {.}}{%
+Wilson%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2014}}%
+}]{%
+wilson2014best}
+\APACinsertmetastar {%
+wilson2014best}%
+\begin{APACrefauthors}%
+Wilson, G.%
+, Aruliah, D\BPBI A.%
+, Brown, C\BPBI T.%
+, Hong, N\BPBI P\BPBI C.%
+, Davis, M.%
+, Guy, R\BPBI T.%
+\BDBL {}others%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2014}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Best practices for scientific computing} {Best practices
+  for scientific computing}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{PLOS Biology}{12}{1}{e1001745}.
+\PrintBackRefs{\CurrentBib}
+
+\bibitem [\protect \citeauthoryear {%
+Wilson%
+\ \protect \BOthers {.}}{%
+Wilson%
+\ \protect \BOthers {.}}{%
+{\protect \APACyear {2017}}%
+}]{%
+wilson2017good}
+\APACinsertmetastar {%
+wilson2017good}%
+\begin{APACrefauthors}%
+Wilson, G.%
+, Bryan, J.%
+, Cranston, K.%
+, Kitzes, J.%
+, Nederbragt, L.%
+\BCBL {}\ \BBA {} Teal, T\BPBI K.%
+\end{APACrefauthors}%
+\unskip\
+\newblock
+\APACrefYearMonthDay{2017}{}{}.
+\newblock
+{\BBOQ}\APACrefatitle {Good enough practices in scientific computing} {Good
+  enough practices in scientific computing}.{\BBCQ}
+\newblock
+\APACjournalVolNumPages{PLOS Biology}{13}{6}{e1005510}.
+\newblock
+\begin{APACrefURL}
+  \url{https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1005510}
+  \end{APACrefURL}
+\PrintBackRefs{\CurrentBib}
+
+\end{thebibliography}
 

### 2020-10-11 STAT-paper-template/APA/fertile.tex
back to original tex file (not hard coded)

--- 2020-10-10 STAT-paper-template/APA/fertile.tex
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... -------------------------- fertile 0.0.0.9028 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... ------------------------------------------------------------ fertile 0.0.0.9028 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ---------------- fertile 0.0.0.9028 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... -------------------------------------------------- fertile 0.0.0.9028 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Error in eval\_tidy(xs[[j]], mask): object '.Random.seed' not found}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks ------------------------------------------ fertile 0.0.0.9028 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks -------- fertile 0.0.0.9028 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# v Checking for no absolute paths}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ---------------- fertile 0.0.0.9028 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# v Checking for only portable paths}}\end{kframe}
+\begin{kframe}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval\_tidy(xs[[j]], mask): object '.Random.seed' not found}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks ----------------------------------------------- fertile 0.0.0.9028 --}}\end{kframe}
+\begin{kframe}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# v Reproducibility checks passed: 2}}\end{kframe}
 \end{knitrout}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ---------------- fertile 0.0.0.9028 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps -------------------------- fertile 0.0.0.9028 --}}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval\_tidy(xs[[j]], mask): object '.Random.seed' not found}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code ------------------------------------ fertile 0.0.0.9028 --}}\begin{verbatim}
+## # A tibble: 9 x 3
+##   package       N used_in                    
+##   <chr>     <int> <chr>                      
+## 1 broom         1 project_miceps/analysis.Rmd
+## 2 dplyr         1 project_miceps/analysis.Rmd
+## 3 ggplot2       1 project_miceps/analysis.Rmd
+## 4 purrr         1 project_miceps/analysis.Rmd
+## 5 readr         1 project_miceps/analysis.Rmd
+## 6 rmarkdown     1 project_miceps/analysis.Rmd
+## 7 skimr         1 project_miceps/analysis.Rmd
+## 8 stargazer     1 project_miceps/analysis.Rmd
+## 9 tidyr         1 project_miceps/analysis.Rmd
+\end{verbatim}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory -------------------------------------------- fertile 0.0.0.9028 --}}\begin{verbatim}
+## # A tibble: 9 x 4
+##   file               ext        size mime                                       
+##   <fs::path>         <chr> <fs::byt> <chr>                                      
+## 1 Estrogen_Receptor~ docx     10.97K application/vnd.openxmlformats-officedocum~
+## 2 citrate_v_time.png png     188.68K image/png                                  
+## 3 proteins_v_time.p~ png     379.22K image/png                                  
+## 4 Blot_data_updated~ csv      14.43K text/csv                                   
+## 5 CS_data_redone.csv csv       7.39K text/csv                                   
+## 6 mice.csv           csv      14.33K text/csv                                   
+## 7 README.md          md           39 text/markdown                              
+## 8 miceps.Rproj       Rproj       204 text/rstudio                               
+## 9 analysis.Rmd       Rmd       4.94K text/x-markdown
+\end{verbatim}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files ------------------------------------------ fertile 0.0.0.9028 --}}\begin{verbatim}
+## # A tibble: 7 x 3
+##   path_rel           dir_rel    cmd                                             
+##   <fs::path>         <fs::path> <chr>                                           
+## 1 Blot_data_updated~ data-raw   file_move('project_miceps/Blot_data_updated.csv~
+## 2 CS_data_redone.csv data-raw   file_move('project_miceps/CS_data_redone.csv', ~
+## 3 Estrogen_Receptor~ inst/other file_move('project_miceps/Estrogen_Receptors.do~
+## 4 analysis.Rmd       vignettes  file_move('project_miceps/analysis.Rmd', fs::di~
+## 5 citrate_v_time.png inst/image file_move('project_miceps/citrate_v_time.png', ~
+## 6 mice.csv           data-raw   file_move('project_miceps/mice.csv', fs::dir_cr~
+## 7 proteins_v_time.p~ inst/image file_move('project_miceps/proteins_v_time.png',~
+\end{verbatim}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged ---------------------------------------------- fertile 0.0.0.9028 --}}\begin{verbatim}
+## NULL
+\end{verbatim}
+\end{kframe}
 \end{knitrout}
@@
 \begin{verbatim}
-## # A tibble: 3 x 4
+## # A tibble: 5 x 4
 ##   path         path_abs                             func     timestamp          
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-10-10 00:39:23
-## 2 package:for~ <NA>                                 base::l~ 2020-10-10 00:39:23
-## 3 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-10 00:39:23
+## 1 package:pur~ <NA>                                 base::l~ 2020-10-11 15:10:40
+## 2 package:for~ <NA>                                 base::l~ 2020-10-11 15:10:40
+## 3 Seed Before  <NA>                                 14       2020-10-11 15:10:40
+## 4 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-11 15:10:40
+## 5 Seed After   <NA>                                 28       2020-10-11 15:10:40
 \end{verbatim}
@@
 ## function(file, ...) {
+## 
 ##   if (interactive_log_on()) {
+##     if (Sys.getenv("FERTILE_RENDER_MODE") == TRUE){
+##       log_push('Seed Before', .Random.seed[2])
+##     }
+## 
 ##     log_push(file, "readr::read_csv")
+## 
 ##     check_path_safe(file, ... = "readr::read_csv")
-##     readr::read_csv(file, ...)
+## 
+##     data <- readr::read_csv(file, ...)
+## 
+##     if (Sys.getenv("FERTILE_RENDER_MODE") == TRUE){
+##       log_push('Seed After', .Random.seed[2])
+##     }
+## 
+##     return (data)
 ##   }
 ## }
-## <bytecode: 0x7fcede54ad08>
+## <bytecode: 0x7f9a68cb9468>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNAq0UZ/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpz372mQ/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNAq0UZ/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNAq0UZ/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpNAq0UZ/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpz372mQ/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpz372mQ/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpz372mQ/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}

### 2020-10-11 STAT-paper-template/APA/response-to-reviewers.md

@@
 Specific minor suggestions (in order of manuscript):
-*- A few suggested references to include in Intro and/or 2.1:* 
+
+- *A few suggested references to include in Intro and/or 2.1:* 
     - *reproducibility framework: Patil et al. (https://www.nature.com/articles/s41562-019-0629-z)*
@@
 
-A comment on this has been added to the end of the description of retroactive behavior.
+Thank you for pointing this out. A comment to this effect has been added to the end of the description of retroactive behavior.
 
@@
 
-Since the code is intuitive, it has now been hidden from view. Instead, there is now a sentence stating that we are looking into the contents of the file.
+That code has been hidden from view. Instead, there is now a sentence stating that we are looking into the contents of the file.
 
@@
 
-There is not currently a functionality for this, but it is definitely something of interest for the future. As a result, we have added it to the future work section!
+Thank you for this excellent idea. There is not currently a functionality for this, but it is definitely something of interest for the future. As a result, we have added it to the future work section.
 

### 2020-10-11 STAT-paper-template/APA/fertile.Rnw
added log example

@@
 \citet{broman} reiterates the need for clear naming and file separation while sharing several additional suggestions: keep the project contained in one directory, use relative paths, and include a \cmd{README}.
-\textcolor{blue}{\citet{wilson2017good} argue that, to follow good practice, an analysis must be located in a well-organized directory, be supplemented by a file containing information about the project (i.e.
-a README), and a provide an explicit list of dependencies.}
+\textcolor{blue}{\citet{wilson2017good} argue that to follow good practice, an analysis must be located in a well-organized directory, be supplemented by a file containing information about the project (i.e., a README), and a provide an explicit list of dependencies.}
 \textcolor{blue}{{\citet{wilson2014best}} emphasize communication, recommending that code be human-readable and consistently styled for ease of understanding once shared.}
@@
 \textcolor{blue}{For example, \pkg{fertile} addresses file structure in part but also considers the quality of project documentation, the use of file paths, and the behavior of random number generation}.
-\pkg{fertile} is flexible, \textcolor{red}{\st{offering benefits}} \textcolor{blue}{providing reports on reproducibility} to users at any stage in the data analysis workflow, and provides \R-specific features, \textcolor{blue}{such as tools to manage package dependencies}, which address certain aspects of reproducibility that can be missed by external project development software.
+\pkg{fertile} is flexible, \textcolor{red}{\st{offering benefits}} \textcolor{blue}{providing reports on reproducibility} to users at any stage in the data analysis workflow.
+It also provides \R-specific features, \textcolor{blue}{such as tools to manage package dependencies}, which address certain aspects of reproducibility that can be missed by external project development software.
 
@@
 
-\textcolor{blue}{Users who want to override this behavior and execute their function anyway are provided with a piece of code to execute that will achieve this for them.}
+\textcolor{blue}{Users who want to override this behavior and execute their function anyway are provided with a snipped of code to execute that will achieve this for them.}
 
@@
 \textcolor{blue}{When these functions are called, users are not presented with an option to override the resulting error.
-Instead, the only way to achieve the desired action is for the user to execute an alternative function that produces the same behavior but in a reproducible way.}
+Instead, the only way to achieve the desired action is for the user to execute a different function that produces the same behavior but in a reproducible way.}
 
@@
   "Looks for indication of a clear order in which to run R scripts.",
-  "There are six different 'has_tidy_...' functions. Each checks to make sure that no files of its given type (images, code, data, etc.) are located in the root directory of the project",
+  "There are six different 'has_tidy_*()' functions. Each checks to make sure that no files of its given type (images, code, data, etc.) are located in the root directory of the project",
   "Ensures that all files located in the project directory are being referenced or created by other files in the project.",
@@
 Subsets of the sixteen tests can be invoked using the \pkg{tidyselect} helper functions \citep{R-tidyselect} in combination with the more limited \func{proj\_check\_some} function.
-\textcolor{blue}{These helper functions, which include \func{starts\_with}, \func{ends\_with}, \func{contains}, and \func{matches}, allow users to check for matches between a string or regular expression and a list of variables--in this case, searching for checks from the list provided by \func{list\_checks} whose names match a certain condition (e.g.
-ending with "root")}.
+\textcolor{blue}{These helper functions, which include \func{starts\_with}, \func{ends\_with}, \func{contains}, and \func{matches}, allow users to check for matches between a string or regular expression and a list of variables.
+In this case, we search for checks whose names match a certain condition (e.g., contain the string "paths")}.
 
@@
 To prevent users from accidentally enabling or disabling features, \pkg{fertile} does not provide any option to control which features are active.
-However, those looking to only use the proactive features can do so by refraining from using the retroactive-specific functions while those looking to only use the retroactive features can wait to load \pkg{fertile} until they are ready to produce a report on their project.}
+However, those who want to use the proactive features can simply  refrain from using the retroactive-specific functions.
+Conversely, those who want to use only the retroactive features can wait to load \pkg{fertile} until they are ready to produce a report on their project.}
 
@@
 
-<<read-csv, message=FALSE, include=FALSE>>=
+<<include = FALSE>>=
+log_clear()
+@
+
+<<read-csv, message=FALSE, results = 'hide'>>=
 library(purrr)
@@
 Except where warranted, the execution looks the same to the user as if they were calling the original function.
-\textcolor{blue}{Since \pkg{fertile} only needs access to the file path argument, "file" is the only declared argument for most shims.
+\textcolor{blue}{For most shims, "file" is the only argument of interest.
 The other arguments are managed using the dots (...), which capture and save them, allowing them to eventually be passed back to the original shimmed function (e.g.
@@
 \textcolor{blue}{\pkg{fertile} also keeps track of dependencies in additional detail.
-Every time the code contained within a project is rendered by `fertile`, the package generates a file capturing the `sessionInfo()` just after the code was run.
-This file contains information about the R version in which the code was run, the list of packages that were loaded, and their specific versions.}
+Every time the code contained within a project is rendered by \pkg{fertile}, the package generates a file capturing the \func{sessionInfo} just after the code was run.
+This file contains information about the \R version in which the code was run, the list of packages that were loaded, and their specific versions.}
 
@@
 Under this system, projects would receive badges from \pkg{fertile} based on the sets of reproducibility checks they pass.
-For example, projects that pass all of the checks on file paths would receive a badge on their project for "Reproducible File Paths" while those who pass checks involving file clutter and project structure might be awarded a badge for "Exellent File Organization."}
+For example, projects that pass all of the checks on file paths would receive a badge on their project for "Reproducible File Paths" while those who pass checks involving file clutter and project structure might be awarded a badge for "Excellent File Organization."}
   \item \textcolor{blue}{Implementing a method through which users can add their own shims to the package. This would allow users to modify \pkg{fertile} to fit the functions they use most often. }

### 2020-10-11 STAT-paper-template/APA/fertile.tex
back to original tex file (not hard coded)

@@
 \cellcolor{gray!6}{has\_clear\_build\_chain()} & \cellcolor{gray!6}{Looks for indication of a clear order in which to run R scripts.}\\
-has\_tidy\_[media/images/code/raw\_data/data/scripts]() & There are six different 'has\_tidy\_...' functions.
+has\_tidy\_[media/images/code/raw\_data/data/scripts]() & There are six different 'has\_tidy\_*()' functions.
 Each checks to make sure that no files of its given type (images, code, data, etc.) are located in the root directory of the project\\
@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-10-11 15:10:40
-## 2 package:for~ <NA>                                 base::l~ 2020-10-11 15:10:40
-## 3 Seed Before  <NA>                                 14       2020-10-11 15:10:40
-## 4 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-11 15:10:40
-## 5 Seed After   <NA>                                 28       2020-10-11 15:10:40
+## 1 package:pur~ <NA>                                 base::l~ 2020-10-11 15:12:26
+## 2 package:for~ <NA>                                 base::l~ 2020-10-11 15:12:26
+## 3 Seed Before  <NA>                                 21       2020-10-11 15:12:26
+## 4 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-11 15:12:26
+## 5 Seed After   <NA>                                 35       2020-10-11 15:12:26
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x7f9a68cb9468>
+## <bytecode: 0x7fb05e74f770>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpz372mQ/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpo24pt6/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpz372mQ/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpz372mQ/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpz372mQ/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpo24pt6/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpo24pt6/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpo24pt6/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}

### 2020-10-14 STAT-paper-template/APA/response-to-reviewers.md
changed sample shim to read.csv

@@
 Additionally, a new bullet point on code commenting (relevant to the has_well_commented_code() function as well as one of the added sources), has been added to the summary of reproducibility components.
+
+Finally, the sample shim (previously `read_csv()`) has been changed to `read.csv()`.
+This is because of a change to the behavior of `readr::read_csv` since the original submission of the manuscript which required the shim for that function to be adjusted---making it no longer representative of most shims in `fertile`.

### 2020-10-14 STAT-paper-template/APA/fertile.tex
changed sample shim to read.csv

@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... ------------------------------------------------------------ fertile 0.0.0.9028 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... -------------------------------------------------------- fertile 0.0.0.9028 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... -------------------------------------------------- fertile 0.0.0.9028 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ---------------------------------------------- fertile 0.0.0.9028 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks ------------------------------------------ fertile 0.0.0.9028 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks -------------------------------------- fertile 0.0.0.9028 --}}
 
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks ----------------------------------------------- fertile 0.0.0.9028 --}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks ------------------------------------------- fertile 0.0.0.9028 --}}\end{kframe}
 \begin{kframe}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps -------------------------- fertile 0.0.0.9028 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps ---------------------- fertile 0.0.0.9028 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code ------------------------------------ fertile 0.0.0.9028 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code -------------------------------- fertile 0.0.0.9028 --}}\begin{verbatim}
 ## # A tibble: 9 x 3
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory -------------------------------------------- fertile 0.0.0.9028 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory ---------------------------------------- fertile 0.0.0.9028 --}}\begin{verbatim}
 ## # A tibble: 9 x 4
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files ------------------------------------------ fertile 0.0.0.9028 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files -------------------------------------- fertile 0.0.0.9028 --}}\begin{verbatim}
 ## # A tibble: 7 x 3
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged ---------------------------------------------- fertile 0.0.0.9028 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged ------------------------------------------ fertile 0.0.0.9028 --}}\begin{verbatim}
 ## NULL
@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-10-11 15:12:26
-## 2 package:for~ <NA>                                 base::l~ 2020-10-11 15:12:26
-## 3 Seed Before  <NA>                                 21       2020-10-11 15:12:26
-## 4 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-11 15:12:26
-## 5 Seed After   <NA>                                 35       2020-10-11 15:12:26
+## 1 package:pur~ <NA>                                 base::l~ 2020-10-14 18:12:20
+## 2 package:for~ <NA>                                 base::l~ 2020-10-14 18:12:20
+## 3 Seed Before  <NA>                                 20       2020-10-14 18:12:20
+## 4 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-14 18:12:20
+## 5 Seed After   <NA>                                 34       2020-10-14 18:12:20
 \end{verbatim}
@@
 \end{enumerate}
-Most shims are quite simple and look something like what is shown below for \func{read\_csv}.
+Most shims are quite simple and look something like what is shown below for \textcolor{blue}{\func{read.csv}}.
 
@@
 \begin{alltt}
-\hlstd{fertile}\hlopt{::}\hlstd{read_csv}
+\hlstd{fertile}\hlopt{::}\hlstd{read.csv}
 \end{alltt}
@@
 ## function(file, ...) {
-## 
 ##   if (interactive_log_on()) {
-##     if (Sys.getenv("FERTILE_RENDER_MODE") == TRUE){
-##       log_push('Seed Before', .Random.seed[2])
-##     }
-## 
-##     log_push(file, "readr::read_csv")
-## 
-##     check_path_safe(file, ... = "readr::read_csv")
-## 
-##     data <- readr::read_csv(file, ...)
-## 
-##     if (Sys.getenv("FERTILE_RENDER_MODE") == TRUE){
-##       log_push('Seed After', .Random.seed[2])
-##     }
-## 
-##     return (data)
+##     log_push(file, "utils::read.csv")
+##     check_path_safe(file, ... = "utils::read.csv")
+##     utils::read.csv(file, ...)
 ##   }
 ## }
-## <bytecode: 0x7fb05e74f770>
+## <bytecode: 0x7fc7f433c548>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpo24pt6/project_miceps
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpyKNgZn/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpo24pt6/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpo24pt6/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/Rtmpo24pt6/project_miceps/Estrogen_Receptors.docx
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpyKNgZn/project_miceps/Blot_data_updated.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpyKNgZn/project_miceps/CS_data_redone.csv
+## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpyKNgZn/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}

### 2020-10-15 STAT-paper-template/APA/response-to-reviewers.md
added reference and edits

--- 2020-10-11 STAT-paper-template/APA/response-to-reviewers.md
@@
-# Paper Revision Changelog
+# Response to reviewers
 
@@
 
-**Associate Editor’s Comments:**
+## Associate Editor’s Comments
 
@@
 
-**Reviewer’s Comments:**
+## Reviewer’s Comments
 
 General:
+
 *- the use of shims is likely unfamiliar to R users.
@@
 This is an excellent point!
-Part of the reason that fertile is so strict on this behavior is that in my own personal experience, many R coders (particularly those with little experience) tend to ignore warnings as long as their code executes properly.
+Part of the reason that fertile is so strict on this behavior is that in our experience, many R coders (particularly those with little experience) tend to ignore warnings as long as their code executes properly.
 Having fertile’s path-related behavior be error based ensures that no one will skip over the messages that come up.
-However, my co-author and I do recognize that some users may still want to run these functions despite the warnings.
+However, we do recognize that some users may still want to run these functions despite the warnings.
 In order to take this into account, we have added detail to the error messages produced by the shims.
-The new messages provide users with a method to override the error produced by fertile.
+The new messages provide users with a strategy to override the error produced by fertile.
 
@@
 
-This table has been added! Since list_checks() provides no additional benefit to the user than the information in this table, it is no longer mentioned.
+This table has been added! Since `list_checks()` provides no additional benefit to the user than the information in this table, it is no longer mentioned.
 
@@
 
-A short description of tidy select helpers and their applicability here has been added just before the proj_check_some() example.
+A short description of tidy select helpers and their applicability here has been added just before the `proj_check_some()` example.
 
@@
 No results are discussed so that warrants a name change.
-Since the section is about an experiment, however, I have chosen to replace it with “Experimental Testing”.
-We also decided to run a second trial this semester, so future directions is no longer an applicable name as it is happening currently!
+Since the section is about an experiment, however, we have chosen to replace it with “Experimental Testing”.
+We also decided to run a second trial this semester, so "future directions" is no longer an applicable name as it is happening currently!
 
@@
 
-**Additional Changes**
+## Additional Changes
 

### 2020-10-15 STAT-paper-template/APA/fertile.Rnw
changed sample shim to read.csv
added reference and edits

@@
 
-\textcolor{blue}{Users who want to override this behavior and execute their function anyway are provided with a snipped of code to execute that will achieve this for them.}
+\textcolor{blue}{For users who want to override this behavior and execute their function anyway, we provide a snippet of code to bypass \pkg{fertile} and execute the original function call.}
 
@@
 \textcolor{blue}{When these functions are called, users are not presented with an option to override the resulting error.
-Instead, the only way to achieve the desired action is for the user to execute a different function that produces the same behavior but in a reproducible way.}
+Instead, the preferred way to achieve the desired action is for the user to execute a different function that produces the same behavior but in a reproducible way.
+For example, by using the \func{here} function from the \pkg{here} package \citep{R-here} instead of \func{setwd}.}
 
@@
 
-\textcolor{blue}{Both the proactive and retroactive features available in \pkg{fertile} are enabled automatically upon the loading of the package.
+\textcolor{blue}{Both the proactive and retroactive features available in \pkg{fertile} are enabled automatically upon loading the package.
 To prevent users from accidentally enabling or disabling features, \pkg{fertile} does not provide any option to control which features are active.

### 2020-10-15 STAT-paper-template/APA/fertile.Rnw
close #16; #17

@@
 \end{enumerate}
-Most shims are quite simple and look something like what is shown below for \func{read\_csv}.
+Most shims are quite simple and look something like what is shown below for \textcolor{blue}{\func{read.csv}}.
 
 <<>>=
-fertile::read_csv
+fertile::read.csv
 @
@@
 
-One of the challenges with ensuring that work is reproducible is the issue of dependencies.
+One of the challenges with ensuring that work is reproducible is \textcolor{blue}{\st{the issue of}} dependencies.
 Many data analysis projects reference a variety of \R packages in their code.

### 2020-10-15 STAT-paper-template/APA/fertile.tex
close #16; #17

@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling... -------------------------------------------------------- fertile 0.0.0.9028 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Compiling...
+------------------------------------------------------------------------------------------------------------------ fertile 0.0.0.9028 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts... ---------------------------------------------- fertile 0.0.0.9028 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Rendering R scripts...
+-------------------------------------------------------------------------------------------------------- fertile 0.0.0.9028 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks -------------------------------------- fertile 0.0.0.9028 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Running reproducibility checks ------------------------------------------------------------------------------------------------ fertile 0.0.0.9028 --}}
 
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks ------------------------------------------- fertile 0.0.0.9028 --}}\end{kframe}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Summary of fertile checks ----------------------------------------------------------------------------------------------------- fertile 0.0.0.9028 --}}\end{kframe}
 \begin{kframe}
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps ---------------------- fertile 0.0.0.9028 --}}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Analysis of reproducibility for project\_miceps -------------------------------------------------------------------------------- fertile 0.0.0.9028 --}}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code -------------------------------- fertile 0.0.0.9028 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Packages referenced in source code ------------------------------------------------------------------------------------------ fertile 0.0.0.9028 --}}\begin{verbatim}
 ## # A tibble: 9 x 3
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory ---------------------------------------- fertile 0.0.0.9028 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Files present in directory -------------------------------------------------------------------------------------------------- fertile 0.0.0.9028 --}}\begin{verbatim}
 ## # A tibble: 9 x 4
@@
 ## 1 Estrogen_Receptor~ docx     10.97K application/vnd.openxmlformats-officedocum~
-## 2 citrate_v_time.png png     188.68K image/png                                  
-## 3 proteins_v_time.p~ png     379.22K image/png                                  
+## 2 citrate_v_time.png png     188.29K image/png                                  
+## 3 proteins_v_time.p~ png     377.95K image/png                                  
 ## 4 Blot_data_updated~ csv      14.43K text/csv                                   
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files -------------------------------------- fertile 0.0.0.9028 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Suggestions for moving files ------------------------------------------------------------------------------------------------ fertile 0.0.0.9028 --}}\begin{verbatim}
 ## # A tibble: 7 x 3
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged ------------------------------------------ fertile 0.0.0.9028 --}}\begin{verbatim}
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# --\ \  Problematic paths logged ---------------------------------------------------------------------------------------------------- fertile 0.0.0.9028 --}}\begin{verbatim}
 ## NULL
@@
 ##   <chr>        <chr>                                <chr>    <dttm>             
-## 1 package:pur~ <NA>                                 base::l~ 2020-10-14 18:12:20
-## 2 package:for~ <NA>                                 base::l~ 2020-10-14 18:12:20
-## 3 Seed Before  <NA>                                 20       2020-10-14 18:12:20
-## 4 project_mic~ /Users/audreybertin/Documents/ferti~ readr::~ 2020-10-14 18:12:20
-## 5 Seed After   <NA>                                 34       2020-10-14 18:12:20
+## 1 package:pur~ <NA>                                 base::l~ 2020-10-15 17:20:28
+## 2 package:for~ <NA>                                 base::l~ 2020-10-15 17:20:28
+## 3 Seed Before  <NA>                                 20       2020-10-15 17:20:28
+## 4 project_mic~ /home/bbaumer/Dropbox/git/fertile-p~ readr::~ 2020-10-15 17:20:28
+## 5 Seed After   <NA>                                 35       2020-10-15 17:20:28
 \end{verbatim}
@@
 ## }
-## <bytecode: 0x7fc7f433c548>
+## <bytecode: 0x55e884da4358>
 ## <environment: namespace:fertile>
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpyKNgZn/project_miceps
+## /tmp/RtmpqsLI66/project_miceps
 \end{verbatim}
@@
 \begin{verbatim}
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpyKNgZn/project_miceps/Blot_data_updated.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpyKNgZn/project_miceps/CS_data_redone.csv
-## /var/folders/v6/f62qz88s0sd5n3yqw9d8sb300000gn/T/RtmpyKNgZn/project_miceps/Estrogen_Receptors.docx
+## /tmp/RtmpqsLI66/project_miceps/Blot_data_updated.csv
+## /tmp/RtmpqsLI66/project_miceps/CS_data_redone.csv
+## /tmp/RtmpqsLI66/project_miceps/Estrogen_Receptors.docx
 \end{verbatim}
